### 2012-09-19 process.tex
initial latex files with 2 reference papers
rbac-slr.tex

\section{Methodology and Process} \label{sec:process}

Reference can be Andrew Meneely's Validating Software Metrics: A Spectrum of Philosophies paper.

### 2012-09-27 process.tex
Introduction and RQs

@@
 
-Reference can be Andrew Meneely's Validating Software Metrics: A Spectrum of Philosophies paper.
+We collect papers on extended role-based access control~\cite{ferraiolo:rbac}.
+We select 27 papers out of XXX papers.
+
+\subsection{Research Questions} \label{subsec:research questions}
+We have following research questions.
+\begin{itemize}
+	\item RQ1. What are problems in current RBAC model to propose extended RBAC models?
+	\item RQ2. What are reasons to propose RBAC extensions? 
+	\item RQ3. Which extended features of RBAC model are proposed?
+	\item RQ4. How they provide evidence to show that their model work in practice? For example, they provide a prototype to run a real example.
+	\item RQ5. What is criteria for evaluation? We investigate how they evaluate their research work.
+
+\end{itemize}
+
+For RQ1-RQ3, we compare of proposed extended models.
+For RQ4, we compare quality of completeness of the proposed models.
+Beyond modeling of extended RBAC models, an prototype shows that their models work in practice and improves the quality
+of the completeness of research papers. For RQ5, we compare criteria for evaluation of extended RBAC models.
+
+\subsection{Categorization} \label{subsec:categorization}
+
+This section describes categorization of papers based on specific extended features as follows. 
+\begin{itemize}
+	\item Temporal-related-constraints~\cite{mossakowski03:temporal,aich07:STARBAC,kumar06:strbac,samuel07:spatio-temporal,ray07:spatio,chandran05:llt,aich09:role,bertino01:trbac, li08:fine, joshi05:generalized, chen08:spatio-temporal}
+	\item Inheritance~\cite{ren10:machine}
+	\item Location~\cite{chandran05:llt}
+	\item Delegation~\cite{hasebe10:capability}
+	\item General constraints~\cite{alam06:constraint,yao08:task}
+	\item Context~\cite{tzelepi01:flexible,haibo05:context,cholewka00:acontext-sensitive,huang06:pervasive,motta03:contextual,bao08:role}						
+	\item General extension~\cite{han08:extended,zhang06:collaborative}
+	\item Combination with other access control models such as Task-based access control~\cite{yao08:task,oh03:task,zhou07:network, oh00:task} and Agent-based access control~\cite{yamazaki04:designing}
+\end{itemize}

### 2013-02-03 latex/RBAC-SLR.tex
Moves original latex documents into old/ directory and adds a latex build directory for current documentation state.

% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\usepackage[utf8]{inputenc}
\DeclareUnicodeCharacter{00A0}{\nobreakspace}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{longtable}
\usepackage{sphinx}
\usepackage{multirow}

\title{RBAC-SLR Documentation}
\date{February 02, 2013}
\release{0.1}
\author{Eric D. Helms}
\newcommand{\sphinxlogo}{}
\renewcommand{\releasename}{Release}
\makeindex

\makeatletter
\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
    \let\PYG@ul=\relax \let\PYG@tc=\relax%
    \let\PYG@bc=\relax \let\PYG@ff=\relax}
\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
\def\PYG@toks#1+{\ifx\relax#1\empty\else%
    \PYG@tok{#1}\expandafter\PYG@toks\fi}
\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}

\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.25,0.82}{##1}}}
\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@cs\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{1.00,0.94,0.94}{\strut ##1}}}
\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.19,0.19,0.19}{##1}}}
\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.84,0.33,0.22}{##1}}}
\expandafter\def\csname PYG@tok@nl\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.13,0.44}{##1}}}
\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.38,0.68,0.84}{##1}}}
\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
\expandafter\def\csname PYG@tok@nd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
\expandafter\def\csname PYG@tok@ne\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.49}{##1}}}
\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.44,0.63,0.82}{##1}}}
\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.45}{##1}}}
\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.32,0.47,0.09}{##1}}}
\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.14,0.33,0.53}{##1}}}
\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.56,0.13,0.00}{##1}}}
\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}

\def\PYGZbs{\char`\\}
\def\PYGZus{\char`\_}
\def\PYGZob{\char`\{}
\def\PYGZcb{\char`\}}
\def\PYGZca{\char`\^}
\def\PYGZam{\char`\&}
\def\PYGZlt{\char`\<}
\def\PYGZgt{\char`\>}
\def\PYGZsh{\char`\#}
\def\PYGZpc{\char`\%}
\def\PYGZdl{\char`\$}
\def\PYGZti{\char`\~}
% for compatibility with earlier versions
\def\PYGZat{@}
\def\PYGZlb{[}
\def\PYGZrb{]}
\makeatother

\begin{document}

\maketitle
\tableofcontents
\phantomsection\label{index::doc}

Contents:

\chapter{1. Introduction}
\label{intro:introduction}\label{intro:systematic-literature-review-of-rbac-analysis-notes}\label{intro::doc}
Why is the base model of RBAC extended by newer models?

Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.  In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions of specific actions on resources to roles instead of individual users.  Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.

RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.

RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
Standard RBAC model considers only role-user association and role hierarchy.
Since standard RBAC model has limitations such as specifying environmental constraints or context information
Researchers developed extended models of RBAC to overcome the limitations.
However, as researchers often develop their own specialized extended models of RBAC,
their research cannot be generalized or compared with other research work appropriately.
As a result, researchers could take time on reinventing the wheel.
But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?

The goal of this work is to synthesize available research results on extended models of RBAC. We analyze their extended features and claimed research contributions to find limitations of current RBAC models and what extent of extended features by
comparing with similar research work.
We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
\begin{description}
\item[{Our research give benefits to a community as follows:}] \leavevmode\begin{itemize}
\item {} 
Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.

\item {} 
Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.

\item {} 
Our work shows a criteria in comparison among research results.

\item {} 
Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.

\end{itemize}

\end{description}

\chapter{2. Methodology and Process}
\label{process:methodology-and-process}\label{process::doc}
For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
\begin{itemize}
\item {} 
Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}

\item {} 
IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}

\item {} 
ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}

\item {} 
CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}

\end{itemize}

For each of the criteria below, a search was performed on each of the search engines for a total of 12 data sets.  The criteria used were:
\begin{itemize}
\item {} 
role based access control

\item {} 
RBAC

\item {} 
role-based access control

\end{itemize}

The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.  For each criteria for each search engine, the results were captured until a stopping criteria was met.  Each run was performed as follows:
\begin{enumerate}
\item {} 
Remote call to search engine with current search start position and the current search criteria.

\item {} 
Parse results and extract paper title, authors and year of publication.

\item {} 
Compare results against stopping criteria.

\end{enumerate}
\begin{itemize}
\item {} 
If stopping criteria met, stop search.

\item {} 
If stopping criteria not met, increase search position by number of results and return to step 1.

\end{itemize}

The stopping criteria used was either after the first 1000 results, a limitation imposed by some of the search engines, or if ten consecutive results did not contain the search criteria phrase within the title.  After gathering all 12 data sets, the data was combined into a master list by systematically comparing the bibliographic information for each.  After producing a master list, a series of assessment rounds were performed to narrow the paper list and identify primary sources.

There were a total of three elimination rounds.  The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.  Each round was performed as follows:
\begin{itemize}
\item {} 
Each reviewer independently classified papers as relevant, irrelevant or uncertain.

\item {} 
Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.

\item {} 
Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.  From this discussion, papers were either thrown out or kept until the next round of the review.  Ties were broken by an indepedent party.

\end{itemize}

The results of the searches is summarised below:

\begin{tabulary}{\linewidth}{|L|L|L|L|L|}
\hline
\textbf{
Search Engines
} & \textbf{
RBAC
} & \textbf{
role based access control
} & \textbf{
role-based access control
} & \textbf{
Total
}\\\hline

Google Scholar
 & 
651
 & 
213
 & 
435
 & 
1299
\\\hline

ACM Portal
 & 
500
 & 
20
 & 
720
 & 
1240
\\\hline

IEEExplore
 & 
200
 & 
40
 & 
230
 & 
470
\\\hline

CiteSeerX
 & 
100
 & 
100
 & 
150
 & 
350
\\\hline
 &  &  &  & \\\hline

Totals
 & 
1451
 & 
373
 & 
1535
 & 
3359
\\\hline

Combined
 &  &  &  & 
\textbf{1716}
\\\hline
\end{tabulary}

\section{Research Questions}
\label{process:research-questions}
Based on our intial intent for the systematic literature review, and notes taken during the third phase of elimination, we have identified the following research questions:
\begin{itemize}
\item {} 
RQ1. What are the deficiencies in current RBAC model to propose extended RBAC models?

\item {} 
RQ2. What are the motivations behind RBAC extensions?

\item {} 
RQ3. What are the categorizations of RBAC model proposed?

\item {} 
RQ4. For each categorization, what are the extended features of RBAC model proposed?

\item {} 
RQ5. Do these models have corresponding implementations in practice?

\item {} 
RQ6. How are extended RBAC models evaluated in theory and in practice?

\item {} 
RQ7. Are there any commonalties or generalizations across all categorizations?

\item {} 
RQ8. What domains or scenarios serve as inspiration for these extensions?

\end{itemize}

\chapter{RQ1: What are the deficiencies in current RBAC model?}
\label{research_questions/RQ_1:rq1-what-are-the-deficiencies-in-current-rbac-model}\label{research_questions/RQ_1::doc}

\section{Common Themes}
\label{research_questions/RQ_1:common-themes}\begin{itemize}
\item {} 
\end{itemize}

\section{By Category}
\label{research_questions/RQ_1:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
no support for partial inheritance

\item {} 
inability to support permission assignment constraints

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Missing ability to apply constraints between role and permissions and enforce

\item {} 
Missing context that defines if a role can be activated at execution time

\item {} 
Role activation on a per object or process execution rather than session based

\item {} 
No allowance for context to check permission at execution moment based on context

\item {} 
activation of roles based on current context rather than session intiation

\item {} 
conflict resolution

\item {} 
RBAC cannot handle dynamic authorization and entities

\item {} 
Role cannot be resource dependent

\item {} 
RBAC is too simple

\item {} 
Missing granularity

\item {} 
Inability to handle dynamic nature of multiple projects with multiple resources

\item {} 
Too simple

\item {} 
Lacks context

\item {} 
Handling authorization between applications and within can be different

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
RBAC cannot handle cross-organizational roles and access in an administratively easy way

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Privacy regulations are complex, with many entities, RBAC lacks ability to handle all of these

\item {} 
Context, conflict detection and resolution

\item {} 
RBAC lacks ability to define purpose

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
No support for multiple security domains

\item {} 
No support for same security, different applications

\item {} 
Lack of flexibility

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Lack of contextual information

\item {} 
Role needs a defined schema

\item {} 
Role activation based on context

\item {} 
Giving a permission an operation

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
Needs an entity to represent a task and connect a role and a permission through it

\item {} 
Need for team and tasks

\item {} 
Need for entity between permission and roles to hold constraint and allow for dynamic situations and entities

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
RBAC can't handle time based constraints for role enablement

\end{itemize}

\end{description}

\chapter{RQ2: What are the motivations behind RBAC extensions?}
\label{research_questions/RQ_2:rq2-what-are-the-motivations-behind-rbac-extensions}\label{research_questions/RQ_2::doc}

\section{Common Themes}
\label{research_questions/RQ_2:common-themes}\begin{itemize}
\item {} 
Roles, users and permissions need context associations

\item {} 
Location and time constraints as well as contexts needed

\item {} 
RBAC lacks ability to link in objects

\item {} 
RBAC does not handle single user base and multiple organizations/applications

\item {} 
Role enablement based on context

\item {} 
User granting/revoking of roles based on context

\end{itemize}

\section{By Category}
\label{research_questions/RQ_2:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No support for partial inheritance

\item {} 
Permission assignment constraints

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Relationship between object and viewer can't be represented

\item {} 
Object content matters

\item {} 
Location and time constraints on the user

\item {} 
Departmental context around access

\item {} 
RBAC does not handle many small services to wrap around them

\item {} 
RBAC lacks security principle enforcement

\item {} 
Lack of specifying context, and permission checks of dynamic evaluation time rules

\item {} 
RBAC is static and lacks ability to handle dynamic authorization and definitions

\item {} 
RBAC ignores context

\item {} 
Role granulatiry not well controlled

\item {} 
Roles should depend on the resources

\item {} 
RBAC cannot handle multiple projects with multiple tasks and differing roles in each system but the same user

\item {} 
Lack of context support

\item {} 
Too simple for dynamic environments

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
Support cross-domain delegation

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
Reduce administrative complexity of RBAC for multi-organizational

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Lack of components, purpose binding, conditions, obligations to handle privacy needs

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
Need to support multiple securtiy domains

\item {} 
Need to support same security with different applications

\item {} 
Role heirarchy adds complexity and reduces flexibility

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Lack of contextual information with roles, users and permissions

\item {} 
Role needs a schema to allow specification of attributes

\item {} 
Location based security scenarios

\item {} 
Role activation based on location

\item {} 
Support physical and logical locations

\item {} 
Attaching operations to permissions

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
RBAC doesn't inherently support location and time constraints

\item {} 
Spatial and temporal awareness of roles, inheritance and permissions

\item {} 
Context mechanisms needed

\item {} 
Physical and logical locations, hybrids

\item {} 
Role assignment based off time and space, as well as permissions

\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
RBAC can't handle large enterprises with multiple applications and single user base

\item {} 
Security requirements for enterprises (look up these)

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
Workflow based organizations need time base constraint for role enablement

\end{itemize}

\end{description}

\chapter{RQ3: What are the categorizations of proposed RBAC extensions?}
\label{research_questions/RQ_3:rq3-what-are-the-categorizations-of-proposed-rbac-extensions}\label{research_questions/RQ_3::doc}
Across the papers found, a number of higher level categories emerged that each proposed model fell in to.  Some of these categories had singular primary papers, while others had multiple extension models proposed that were deemed independent of one another.

For each category, the breakdown of the papers for each category is:

\begin{tabulary}{\linewidth}{|L|L|}
\hline
\textbf{
Paper
} & \textbf{
Category
}\\\hline

Alam, M. and Hafner, M. and Breu, R. 2006
 & 
Constraint
\\\hline

Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001
 & 
Context
\\\hline

Haibo, SHEN and Fan, HONG 2005
 & 
Context
\\\hline

Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000
 & 
Context
\\\hline

Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006
 & 
Context
\\\hline

Motta, G.H.M.B. and Furuie, S.S. 2003
 & 
Context
\\\hline

Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008
 & 
Context
\\\hline

Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004
 & 
Context
\\\hline

Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008
 & 
Context
\\\hline

Thein, N. and others 2011
 & 
Context
\\\hline

Zou, D. and He, L. and Jin, H. and Chen, X. 2009
 & 
Context
\\\hline

Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010
 & 
Delegation
\\\hline

Zhang, Z. and Zhang, X. and Sandhu, R. 2006
 & 
Organizational
\\\hline

Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007
 & 
Privacy
\\\hline

Masoumzadeh, A. and Joshi, J. 2008
 & 
Privacy
\\\hline

Zhao, Y. and Zhao, Y. and Lu, H. 2008
 & 
Resource
\\\hline

Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005
 & 
Spatial
\\\hline

Ray, I. and Kumar, M. and Yu, L. 2006
 & 
Spatial
\\\hline

Hansen, F. and Oleshchuk, V. 2003
 & 
Spatial
\\\hline

Aich, S. and Sural, S. and Majumdar, A. 2007
 & 
Spatio-Temporal
\\\hline

Chen, L. and Crampton, J. 2008
 & 
Spatio-Temporal
\\\hline

Samuel, A. and Ghafoor, A. and Bertino, E. 2007
 & 
Spatio-Temporal
\\\hline

Chandran, S. and Joshi, J. 2005
 & 
Spatio-Temporal
\\\hline

Ray, I. and Toahchoodee, M. 2007
 & 
Spatio-Temporal
\\\hline

Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009
 & 
Spatio-Temporal
\\\hline

Yao, L. and Kong, X. and Xu, Z. 2008
 & 
Task
\\\hline

ZHANG, S. and CHEN, X. and HOU, G. 2009
 & 
Task
\\\hline

Oh, S. and Park, S. 2003
 & 
Task
\\\hline

Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005
 & 
Temporal
\\\hline
\end{tabulary}

\section{By Category}
\label{research_questions/RQ_3:by-category}\begin{description}
\item[{Constraint (1)}] \leavevmode
a limitation or restriction

\item[{Context (8)}] \leavevmode
circumstances in which an event occurs

\item[{Delegation (1)}] \leavevmode
the act of empowering to act for another

\item[{Organizational (1)}] \leavevmode
of or relating to an organization

\item[{Privacy (1)}] \leavevmode
freedom from unauthorized intrusion

\item[{Resource (1)}] \leavevmode
an available supply than be drawn on when needed

\item[{Spatial (3)}] \leavevmode
of or relating to facility in percieving relations in space

\item[{Spatio/Temporal (6)}] \leavevmode
combining space and time

\item[{Task (3)}] \leavevmode
a usually assigned piece of work often to be finished within a certain time

\item[{Temporal (1)}] \leavevmode
of or relating to the sequence of time or to a particular time

\end{description}

{\hfill\scalebox{1.000000}{\includegraphics{categories_map.png}}\hfill}

\chapter{RQ4: For each categorization, what are the extended features of RBAC proposed?}
\label{research_questions/RQ_4::doc}\label{research_questions/RQ_4:rq4-for-each-categorization-what-are-the-extended-features-of-rbac-proposed}
Is this the same as RQ\_1 ?

\section{Common Themes}
\label{research_questions/RQ_4:common-themes}\begin{itemize}
\item {} 
\end{itemize}

\section{By Category}
\label{research_questions/RQ_4:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\end{description}

\chapter{RQ5. Do these models have corresponding implementations in practice?}
\label{research_questions/RQ_5::doc}\label{research_questions/RQ_5:rq5-do-these-models-have-corresponding-implementations-in-practice}
When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, bringing the model to life is strong evidence that the
proposed model can work in practice.  The concept of authorization, and access control
is rooted in a business need.  Thus, any access control model needs to be feasible
in the real world not just on paper.  We analyzed the primary sources to see how many
proposed models actually had implmenetations associated with them.  And quantified the
type of implementation.  Whether the implementation was for a real system, for a prototype
and/or used in a production environment.

\section{Common Themes}
\label{research_questions/RQ_5:common-themes}\begin{itemize}
\item {} 
21 No's

\item {} 
8 Yes's

\item {} 
Half the implementation's were prototypes

\end{itemize}

\section{By Category}
\label{research_questions/RQ_5:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
No - system architecture provided

\item {} 
No

\item {} 
Yes - prototype with VB6 and MS Access97

\item {} 
Yes - prototype system, system architecture, algorithms implemented, simulation scenarios

\item {} 
Yes - on an EHR system

\item {} 
Yes - prototype using XML container

\item {} 
Yes - web based API agent

\item {} 
No

\item {} 
No

\item {} 
No

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\item {} 
No

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
Yes - integrated with telemedicine application, web based

\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
Yes - mentions of systems that exist using the model

\item {} 
Yes - prototype system

\item {} 
No

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\end{description}

\chapter{RQ6. How are extended RBAC models evaluated in theory and in practice?}
\label{research_questions/RQ_6:rq6-how-are-extended-rbac-models-evaluated-in-theory-and-in-practice}\label{research_questions/RQ_6::doc}

\section{Common Themes}
\label{research_questions/RQ_6:common-themes}\begin{itemize}
\item {} 
12 had no evaluation whatsoever

\item {} 
8 present example scenarios and how their model would apply

\item {} 
6 presented performance and time complexity analysis

\end{itemize}

\section{By Category}
\label{research_questions/RQ_6:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
No evaluation

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Present hospital scenarios and explanation of how new model would handle

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
No

\item {} 
Scenarios with how model affects, performance test of the model vs. traditional RBAC

\item {} 
Description of a fake project management system that the model is designed for

\item {} 
Applied to conceptual situation

\item {} 
No

\item {} 
Example scenario comparing their model vs other models

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
Comparison caluclation for number of roles vs standard model, numerical size comparisons against other proposed models

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Evaluate the performance of permission assignment and authorization checks for various sizes and context variables

\item {} 
No

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
Model comparison from a qualitative standpoint

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Algorithm and Big-O performance evaluation

\item {} 
No except minor scenarios and how the model handles

\item {} 
No other than mathematical specification

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
Mathematical description

\item {} 
Mathematical analysis of role graph, complexity comparisons across different models

\item {} 
Pseudo example scenario

\item {} 
No

\item {} 
Detailed scenario with examination of how model handles situation

\item {} 
Discuess difference btween their model and other models, time-complexity analysis to show no performance impact with standard model

\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
No

\item {} 
No

\item {} 
No

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
Mathematical modeling and graphs to detail usage and interactions, explanation of example scenarios

\end{itemize}

\end{description}

\chapter{RQ7. Are there any commonalties or generalizations across all categorizations?}
\label{research_questions/RQ_7:rq7-are-there-any-commonalties-or-generalizations-across-all-categorizations}\label{research_questions/RQ_7::doc}

\section{Common Themes}
\label{research_questions/RQ_7:common-themes}\begin{itemize}
\item {} 
\end{itemize}

\section{By Category}
\label{research_questions/RQ_7:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
\end{itemize}

\end{description}

\chapter{RQ8. What domains or scenarios serve as inspiration for these extensions?}
\label{research_questions/RQ_8:rq8-what-domains-or-scenarios-serve-as-inspiration-for-these-extensions}\label{research_questions/RQ_8::doc}

\section{Common Themes}
\label{research_questions/RQ_8:common-themes}\begin{itemize}
\item {} 
Medical domain

\item {} 
Pervasive computing environments

\item {} 
Mobile devices

\item {} 
Large-scale organizations with many sub-departments

\item {} 
Enterprise, organization workflows

\end{itemize}

\section{By Category}
\label{research_questions/RQ_8:by-category}\begin{description}
\item[{Constraint}] \leavevmode\begin{itemize}
\item {} 
Medical domain

\item {} 
Service Oriented Architecture

\end{itemize}

\item[{Context}] \leavevmode\begin{itemize}
\item {} 
Medical imaging database

\item {} 
Handling an insurance claim workflow

\item {} 
Pervasive computing environments

\item {} 
Medical emergency overrides

\item {} 
Electronic health records

\item {} 
Supermarket

\item {} 
Project management system with large project, task and user base

\item {} 
Power load forecasting system (multi-level user environment)

\end{itemize}

\item[{Delegation}] \leavevmode\begin{itemize}
\item {} 
Large-scale electronic health record systems

\end{itemize}

\item[{Organizational}] \leavevmode\begin{itemize}
\item {} 
Online tutoring system

\item {} 
State education system with schools, districts, etc.

\end{itemize}

\item[{Privacy}] \leavevmode\begin{itemize}
\item {} 
Privacy within the medical domain, HIPAA

\end{itemize}

\item[{Resource}] \leavevmode\begin{itemize}
\item {} 
N/A

\end{itemize}

\item[{Spatial}] \leavevmode\begin{itemize}
\item {} 
Hospital with multi-floors, departments

\end{itemize}

\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
\item {} 
Pervasive computing environments

\item {} 
Hospital using mobile devices with multiple user types on a single floor, departments

\item {} 
Mobile, peer-to-peer

\item {} 
Doctor out in the world needing access during a random medical emergency

\item {} 
Students at a university

\end{itemize}

\item[{Task}] \leavevmode\begin{itemize}
\item {} 
Large enterprises with multiple applications and a single user base

\item {} 
Hospital needs

\item {} 
Complex enterprise environments

\end{itemize}

\item[{Temporal}] \leavevmode\begin{itemize}
\item {} 
General organizational needs for time constraints

\end{itemize}

\end{description}

\chapter{Constraint}
\label{categories/constraint::doc}\label{categories/constraint:constraint}\begin{description}
\item[{Definition}] \leavevmode
a limitation or restriction

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://dl.acm.org/citation.cfm?id=1501451}{A constraint based role based access control in the SECTET a model-driven approach}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{A constraint based role based access control in the SECTET a model-driven approach}
\begin{quote}
\begin{description}
\item[{Inspired by a real example?}] \leavevmode
SECTECT is a framework that provides workflow for Service oriented Architectures and is the primary target.

\item[{Scenarios or examples.}] \leavevmode
Presented generic scenarios from the medical domain as examples to describe how mechanisms worked.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
The claim is that with respect to Service Oriented Architecture, and the execution of web services, RBAC lacks needed features.

\item {} 
RBAC does not support partial inheritance

\item {} 
``total inheritance of permissions in RBAC is against the principle of least privilege''

\item {} 
ability for some sub-role permissions not be available to super role

\item {} 
``permission assignment constraints''

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No evidence of use in a real system.

\item[{Is there an evaluation? If yes, how did they do one? If not, why?}] \leavevmode
No.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Presents a modification to a core aspect of RBAC

\item {} 
Allow partial inheritance at the permission level by a superrole instead of treating the role as an atomic entity

\end{itemize}

\item[{Notes.}] \leavevmode
Apply a restricted status to a permission which prevents a superrole from inheriting that permission

\end{description}
\end{quote}

\chapter{Context}
\label{categories/context::doc}\label{categories/context:context}\begin{description}
\item[{Definition.}] \leavevmode
circumstances in which an event occurs

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://dl.acm.org/citation.cfm?id=1232473}{A flexible content and context-based access control model for multimedia medical image database systems}

\item {} 
\href{http://ieeexplore.ieee.org/xpl/login.jsp?tp=\&arnumber=1552897\&url=http\%3A\%2F\%2Fieeexplore.ieee.org\%2Fxpls\%2Fabs\_all.jsp\%3Farnumber\%3D1552897}{A Context-Aware Role-Based Access Control Model for Web Services}

\item {} 
\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.102.9126\&rep=rep1\&type=pdf}{A context-sensitive access control model and prototype implementation}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4079196}{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1229859}{A contextual role-based access control authorization model for electronic patient record}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4709140}{A Role and Context Based Access Control Model with UML}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1376833}{Designing an agent-based RBAC system for dynamic security policy}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4664701}{An extended RBAC model based on granular logic}

\item {} 
\href{http://ieeexplore.ieee.org.prox.lib.ncsu.edu/xpl/articleDetails.jsp?tp=\&arnumber=5967517\&contentType=Conference+Publications\&searchField\%3DSearch\_All\%26queryText\%3Dleveraging+access+control+mechanism+of+android+smartphone+using+context-related+role-based+access+control+model}{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}

\item {} 
\href{http://www.sciencedirect.com.prox.lib.ncsu.edu/science/article/pii/S1084804508000520}{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{A flexible content and context-based access control model for multimedia medical image database systems}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Real world inspiration based upon medical image database.  One simple scenario shown where the model is applicable.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
What the image is of is important to who can view

\item {} 
Domain governs who can see a particular image (e.g. departments in a hospital)

\item {} 
Location for determining if user is accessing from secure location or location of importance (e.g. emergency room)

\item {} 
Time constraints for validity periods

\item {} 
Relationship between images and viewer

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
Provided system architecture for entire medical imaging database with access control system. No implied implementation of model.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No evaluation.  Present proposed situation and hospital structure with scenario where new model would be applicable. No reason for lack of evaluation given.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Proposes adding of constraints between the Role-Permission relationship which must be satisfied in order for the holder of said role to be able to use those permissions

\end{itemize}

\item[{This appears to be an active mechanism during enforcement not assignment}] \leavevmode\begin{itemize}
\item {} 
Adds user attributes (e.g. company position, location, user name)

\item {} 
Adds a decision rule to role-permission relationship

\item {} 
Role-perm relationship contains 5 attributes: identifier, subject/role, action, target objects, constraints

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A Context-Aware Role-Based Access Control Model for Web Services}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
No.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
claim is that RBAC can not handle many small services that need roles/perms wrapped around them

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
services replace objects/actions

\item {} 
adds contexts that must be satisfied

\item {} 
splits roles into local and global, where local is reflective of actions within a service/domain and global roles are reflective of the permissions to access the service

\item {} 
roles are activated based on context at execution time

\end{itemize}

\item[{Notes.}] \leavevmode\begin{itemize}
\item {} 
context oriented, provides what appears to be a different spin or implementation of heirarchies

\end{itemize}

\end{description}
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Context-sensitive access control model and prototype implementation}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Initiates the paper with case scenario ``handle an insurance claim'' workflow.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
points out contrasts between base RBAC and improvements clearly

\item {} 
claims that RBAC supports security principles but does not enforce the use of

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
Prototype created using VB6 and MS Access97.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
largely using mechanisms prescribed by RBAC specifically SoD and LP see note about how RBAC provides for but not the use of the previous security practices

\item {} 
roles are activated whenever a tasks is chosen to be performed and checked if the user can have said role

\item {} 
provides more implementation of SoD and LP than additions to model

\end{itemize}

\item[{Notes.}] \leavevmode\begin{itemize}
\item {} 
context oriented, provides what appears to be a different spin or implementation of heirarchies

\end{itemize}

\end{description}
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Inspired by the needs of access control in pervasive computing environments

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
``static method of management user-role-permissions''

\end{itemize}

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
Created prototype system to verify model correctness and system architecture

\item {} 
Implemented the algorithms proposed by their model

\item {} 
Simulated scenarios but the details of the simulations are not in the paper, nor are the results in a consumable manner

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Claim to be an extension and follow the standard RBAC execution model except by adding two different checks where permissions are re-evaluated by the context of the current user according to the already specific rules

\item {} 
context aware rules that dynamically determine permissions

\item {} 
constraining the permissions based on a users location

\item {} 
assignment/removal of a role based on context information

\item {} 
connection of roles, granting a user certain role permissions when another user/role is present

\item {} 
Provide mechanism to deal with conflicts

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A contextual role-based access control authorization model for electronic patient record}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Traditional RBAC does not support emergency overrides based on the situation

\item {} 
Inspired by the needs of an EHR application in practice

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Claim RBAC is static and cannot handle the dynamic authorization and definitions needed by EHR application complexity

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
Implemented using Java and LDAP on InCor EHR system

\item {} 
They quote number of users and rules, but provide no evidence of comparison or any observation data of the system in use

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Showed some discussion evidence for the requirement for these extensions to RBAC, and they implemented it, however, they provided little evidence as to the effectiveness

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Claim to be a context based extension

\item {} 
Focused on the authorization aspect of permissions for a given role

\item {} 
Contains both positive and negative permission authorizations

\item {} 
Provides for overriding of authorizations based on inheritance

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A Role and Context Based Access Control Model with UML}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Describes a scenario whereby a supermarket could not effectively use RBAC to encapsulate all the various access scenarios they need

\item {} 
Propose a model that adds attributes and system context

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
RBAC ignores context

\item {} 
RBAC is too simple and unilateral and does not align with real world

\item {} 
Role granularity is not well controlled

\item {} 
Roles should depend on the resources not vice-versa

\item {} 
Roles should e assigned to one resource instead of resources being assigned to one role

\end{itemize}

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
Prototype system created using XML container to contain roles and conditionals

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
No evidence as to the comparison of the model in terms of the need over traditional RBAC other than conjecture around a scenario

\item {} 
The performance test is really only valid for their implementation and shows no grounded re-worldness

\item {} 
performed a performance analysis of traditional RBAC and C-RBAC based off their prototype

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode
Extension of the traditional model adding context and maps roles to context and contex to resources with authorization and verification mixed in

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Designing an agent-based RBAC system for dynamic security policy}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
users and application sbecoming larger

\item {} 
many applications have dynamic attributes and defining scenarios for all of them is difficult

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
For their project manage system, traditional RBAC cannot handle the dynamic nature of multiple projects with multiple tasks that different users may be managing in one and not in another

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
Talks about some weird web based API for the agent

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Discuss a fake project management system claiming this type of system is the target for this type of access control

\item[{Does this model extend the core model?}] \leavevmode
Defines model clearly
Extension that defines an Abstract Role, Context Rules and Context Information and the actual role is decided upon based off the context inputs and the rules

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{An extended RBAC model based on granular logic}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Inspired by a muli-level user environment with a complicated authorization management - power load forecasting system

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
traditional RBAC does not provide context

\item {} 
RBAC is too simple for large dyanmic environments

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No evidence, applied in theory to a conceptual situation, no comparison

\item[{Does this model extend the core model?}] \leavevmode
Additions to specification, and rules around whether a user is authorized to perform an action based on their role and the context within the role

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Access control needs of smart phones

\item {} 
Installation of third party applications that a user needs to trust

\item {} 
User must grant device privileges to the application

\item {} 
Parents want to limit the amount of time kids use phone

\item {} 
User might can to limit accessbility to friends and admins

\item {} 
Companies want to limit data access by employee phones

\item {} 
User loses company phone in admin mode, context to prevent leakage

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
smartphone is centralized, user-centric system where identities are known in advance

\item {} 
smartphone has lots of contextual info

\item {} 
\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No.

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Adds objects, environmental context, policies and decisions to model

\item {} 
Object is an accessible entity

\item {} 
a property of the system at the moment of interaction

\item {} 
policy is the formal specification of the access control

\end{itemize}

\item[{Notes.}] \leavevmode\begin{itemize}
\item {} 
precedential order of access privileges to prvent policy bugs

\end{itemize}

\end{description}
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
proliferation of distributed applications

\item {} 
example of limited disk space when a database read and update are performed to resolve which operation to give preference to

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
need flexible and dynamic authorization constraints

\item {} 
authorization differs between and within applications

\item {} 
Users may be granted access to the application and entities or may be granted access to objects with the application depending on context

\item {} 
role and permission constraints are also possible

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
analysis of example scenario, comparison of constraint vs other models,

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
adds object sets, entities, status and set of authorization attributes

\item {} 
permissions on objects with constraints

\item {} 
user and user constraints mapped to roles with permission attribute constraints

\end{itemize}

\item[{Notes.}] \leavevmode\begin{itemize}
\item {} 
classification of constraints as either users eligibility to to use a resource/service or constraints on the users actual use of a resource (limiting what they can do)

\end{itemize}

\end{description}
\end{quote}

\chapter{Historical}
\label{categories/historical::doc}\label{categories/historical:historical}\begin{description}
\item[{Definition.}] \leavevmode
based on or reconstructed from an event, custom, style, etc., in the past

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.182.8833\&rep=rep1\&type=pdf}{HRBAC: Historical Role-Based Access Control}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{HRBAC: Historical Role-Based Access Control}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
a need to incorporate information of the past to make access control decisions

\item {} 
for example, ensuring a user had a set of previous roles before being granted a higher responsibility role (e.g. ``chief'' role needs to have once had a manager role)

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
in order to store and express historical entities of each base entity in order to apply constraints of current assignments and roles and permissions based off the past

\end{itemize}

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
No

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
No, only examples of mechanisms

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Yes, extends each of the 4 levels of RBAC

\item {} 
incorporates new entities to represent historical version of each core entity

\item {} 
provides various historical constraints

\item {} 
provides algorithms for performing constraint resolution

\item {} 
highly defined model and entities

\end{itemize}

\end{description}

Notes.
\end{quote}

\chapter{Organizational}
\label{categories/organizational:organizational}\label{categories/organizational::doc}\begin{description}
\item[{Definition.}] \leavevmode
of or relating to an organization

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4207551}{ROBAC: Scalable Role and Organization Based Access Control Models}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{ROBAC: Scalable Role and Organization Based Access Control Models}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
inspired by context from business-to-business and business-to-consumer context

\item {} 
a theoritical web based report delivery system e.g. a set of educational professionals from schools, districts, and states

\item {} 
also provide example of an online tutoring system

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
claim that their model can significantly reduce administrative complexity for multi-organization

\item {} 
more succint and intuitive than RBAC

\item {} 
same expressive power

\end{itemize}

\end{description}

Did they implement the model?
\begin{description}
\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
present comparison calculations for the number of roles that must be generated between their model and standard

\item {} 
provide numerical size comparisons with other proposed models (Role Templates, Team-based Access Control, Organizational Units

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
extension of RBAC to consider organization and role (a.k.a. context)

\item {} 
an organization must own the object, and the role must give permission on the type for a user to access

\end{itemize}

\end{description}

Notes.
\end{quote}

\chapter{Privacy}
\label{categories/privacy::doc}\label{categories/privacy:privacy}\begin{description}
\item[{Definition.}] \leavevmode
freedom from unauthorized intrusion

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://dl.acm.org/citation.cfm?id=1266848}{Privacy-aware role-based access control}

\item {} 
\href{http://www.sis.pitt.edu/~amirreza/papers/masoumzadeh2008\_purbac.pdf}{PuRBAC: Purpose-Aware Role-Based Access Control}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{Privacy-aware role-based access control}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
inspired by conventional RBAC's inability to handle privacy

\item {} 
extra attributes are needed to support the intricacies of privacy

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}]
due to the lack of basic components required by privacy regulations, especially purpose binding (i.e., data collected for one purpose should not used for another purpose without user consent), conditions, and obligations.  Despite its limitations, existing access control technology can be used as a starting point for managing personal identifiable information

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Evaluated the performance of assigning permissions and checking authorization for various permission size bases as well as variety in number of context variables

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Defines additions to each level of the base RBAC model

\item {} 
Adds objects, purposes, actions, obligations and conditions to core

\item {} 
Includs context variables that can be very general, from space to time to other

\item {} 
Conflict detection and resolution mechanisms

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{PuRBAC: Purpose-Aware Role-Based Access Control}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Web and companies with privacy policies make it harder to follow said policies and protect user data

\end{itemize}

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
privacy policies include purposes under which collected data can be used

\item {} 
current RBAC model does not include a entity or relationships to incorporate purpose

\item {} 
authorization requests should include a purpose for performing the intended action

\end{itemize}

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
No.

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
No evaluation, conclusion includes note that further analysis needs to be done

\item {} 
Discusses entities added by this model, the way they were added and why this method is the best way to add purpose and support privacy

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
extends the base and heirarchy portion of the model, adds in a hybrid heirarchy

\item {} 
connects purposes to permissions

\end{itemize}

\item[{Notes.}] \leavevmode\begin{itemize}
\item {} 
claim to reduce policy complexity by associating a purpose with only a permission and not with roles and users as well

\item {} 
locks down the users ability to declare purpose, by putting authorization controls around the purposes

\end{itemize}

\end{description}
\end{quote}

\chapter{Resource}
\label{categories/resource:resource}\label{categories/resource::doc}\begin{description}
\item[{Definition.}] \leavevmode
an available supply that can be drawn on when needed

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4609645}{A Flexible Role- and Resource-Based Access Control Model}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{A Flexible Role- and Resource-Based Access Control Model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
No.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
RBAC cannot support multiple security domains

\item {} 
cannot support same security with different applications

\item {} 
lack of flexibility in authorization

\item {} 
discard role heirarchy for directory structure to reduce complexity and increase flexibility

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Comparison of the two models from a qualitative standpoint

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Not an extension

\item {} 
Re-imagining of the hierarchy aspect and adding a resources entity

\end{itemize}

\end{description}

Notes.
\end{quote}

\chapter{Spatial}
\label{categories/spatial::doc}\label{categories/spatial:spatial}\begin{description}
\item[{Definition.}] \leavevmode
of or relating to facility in perceiving relations in space

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://dl.acm.org/citation.cfm?id=1063985}{GEO-RBAC: a spatially aware RBAC}

\item {} 
\href{http://www.springerlink.com/index/4N6XK46321M574UH.pdf}{LRBAC: A location-aware role-based access control model}

\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1285394}{Spatial role-based access control model for wireless networks}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{GEO-RBAC: a spatially aware RBAC}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Hospital based examples of activating roles, allowing doctors to only see patient records while in the department of the patient
Extended example section that discusses how a hospital setting with some typical actors would perform simple patient access

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Lack of contextual information with roles, users and permissions
Role needs a schema which allows specification of more attributes of the role

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Shows that this model can be be represented by first-order logic and that they can be evaluated in PTIME ?

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Introduces the concept of a spatial role

\item {} 
Extends the base model at each of the 4 levels

\item {} 
Specify mutual exclusion constraint for separation of duty

\item {} 
Split model into a model without constraints and an extensive model with constraints

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{LRBAC: A location-aware role-based access control model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Some brief real world concepts with regards to intricate security scenarios based on location of user or object

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
Allowing for different security scenarios based on location

\item {} 
Role activation based on location of object or user

\item {} 
Add objects to define physical or logical with location data

\item {} 
Attaches an operation to a permission

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Describes in full definition each object and it's role in the graph
Describes a number of specific operations that would be performed and how the model handles them with respect to locations
Description of how the core RBAC is related to each piece of their model
No evaluation
Minor scenarios

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Extension adding operation, object and role enabling based on a constraint

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Spatial role-based access control model for wireless networks}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Provide an application scenario of where this would be needed and used

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
standard RBAC cannot handle location based constraints and attributes

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Mathematical specification of each aspect
Definitions of each element in their model

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
extension at every level of RBAC

\item {} 
includes separation of duty and heiarchry concerns specifically

\end{itemize}

\end{description}

Notes.
\end{quote}

\chapter{Spatio-Temporal}
\label{categories/spatio_temporal:spatio-temporal}\label{categories/spatio_temporal::doc}\begin{description}
\item[{Definition.}] \leavevmode
combining space and time

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://www.springerlink.com/index/tq15n88122v63754.pdf}{STARBAC: Spatiotemporal role based access control}

\item {} 
\href{http://dl.acm.org/citation.cfm?id=1368341}{On spatio-temporal constraints and inheritance in role-based access control}

\item {} 
\href{https://www.cerias.purdue.edu/tools\_and\_resources/bibtex\_archive/archive/2007-08.pdf}{A framework for specification and verification of generalized spatio-temporal role based access control model}

\item {} 
\href{http://www.springerlink.com/index/77j0j540004585wv.pdf}{LoT-RBAC: A Location and Time-Based RBAC Model}

\item {} 
\href{http://www.springerlink.com/index/C3T50338535HX115.pdf}{A spatio-temporal role-based access control model}

\item {} 
\href{http://www.springerlink.com/index/r1n6j114g5431174.pdf}{Role Based Access Control with Spatiotemporal Context for Mobile Applications}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{STARBAC: Spatiotemporal role based access control}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Similar reasoning as other location based models, No formal evaluation, scenarios or real world examples.

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Extension of RBAC and of SRBAC by combining spatial and temporal

\end{description}

Did they implement the model?
\begin{description}
\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Mathematical explanations, definitions
Includes descriptions and math behind the semantics

\end{description}

Does this model extend the core model?

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{On spatio-temporal constraints and inheritance in role-based access control}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Believe that previous models focus on syntax and not semantics

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
pervasive computing environments requirement for spatial and temporal awareness

\item {} 
previous spatio-temporal do nt address interaction between these constraints and inheritance

\item {} 
makes 4 claims about how existing models do not handle the above

\item {} 
trusted entities to override restrictions

\item {} 
attempt to show compatibility with standard model and claim others do not

\end{itemize}

\end{description}

Did they implement the model?
\begin{description}
\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
Examine previous models and point out deficiencies, use of graphs to show off relationships

\item {} 
Evaluation includes a mathematical examination of the role graphs, and ordering

\item {} 
authors note that there is added complexity in their method, but that is the price to pay to get user  and user-role application constraints

\item {} 
trade-off between complexity of policies and complexity of constraints

\item {} 
authors do comparisons of complexity for the various models

\item {} 
authors note that SoD should be considered in the future

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode
Additions to the specification for constraints with a specific look at how to do spatial and temporal

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A framework for specification and verification of generalized spatio-temporal role based access control model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Based off hospital example with mobile devices inspiring combining location and time based models
Provides an example hospital access control policy that requires space and time - based on the layout of the floor of a hospital and the various users and entitie sthat might exist such as nurse, doctor and surgeons

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Need to be able to handle add attributes that location and spatial specification requires
Add context mechanisms

\item[{Did they implement the model?}] \leavevmode
No

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
To some degree, an example, and some analysis of developing policy and how to deal with conflicts

\item[{Does this model extend the core model?}] \leavevmode
Builds on GTRBAC, extends at all 4 levels
Adds role activation/deactivation
Provides conflict resolution
Includes a GSTRBAC policy specification model using Alloy

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{LoT-RBAC: A Location and Time-Based RBAC Model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Inspired by growing needs of mobile, peer-to-peer devices with regards to activating and specifying permissions and roles based on time and location
Detailed scenario of a person having a medical emergency, and a doctor on site needing to request patient data, as well as the progression through the ambulance, to hospital and citing such location based clues to help know if the doctor should be granted privileges such as being in or near his own car

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Previous location and time based RBAC models did not incorporate location for both the user and role
In-depth, detailed location heirarchies and specifications are needed
Physical, logical, and hybrid location

\item[{Did they implement the model?}] \leavevmode
No

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No

\end{description}

Does this model extend the core model?

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{A Spatio-temporal Role-Based Access Control Model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Growth of wireless networks, mobile devices, the move towards pervasive computing

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Claim that previously examined models of spatial, temporal, or spatio-temporal lack needed aspects
Role assignment should be dependent on space and time
Permissions need to be dependent on space and time not just roles

\item[{Did they implement the model?}] \leavevmode
No. Authors acknowledge need to implement.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
No.  Best they do is provide a detailed scenario with an examination of how such a scenario would be represented in their model. Author acknowledge detailed analysis is needed.

\item[{Does this model extend the core model?}] \leavevmode
Defines for Space - physical location, logical location and mapping functions and relationship definitions
Defines for Time - time instant, time interval
Adds objects to Core model, object can be physical or logical (computer or files e.g.)
Defines a number of potential role hierarchy strategies
Examines impact on both static and dynamic separation of duties for new attributes and mechanisms

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Role Based Access Control with Spatiotemporal Context for Mobile Applications}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Need and desire to put spatio-temporal context and constraint on roles and permissions - such as students at a university or hospital facilities

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
Needs enhancements to be able to handle space and time attributes on roles and permissions

\item[{Did they implement the model?}] \leavevmode
Yes.  Integrated an access control system into iMedik - a telemedicine application in India that is web-based and allows mobile access
Their implementation is a bolt-on solution, that intercepts calls and verifies their authorization
Provide detailed architecture and implementation details

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
Include section that discusses differences between their model and STRBAC and STARBAC
Incorporate algorithms for determining whether a particular authorization is granted and use time-complexity arguments to discuss the effectiveness of each algorithm and to attempt to display that the added complexity of their model over traditional RBAC does not impace performance significantly

\item[{Does this model extend the core model?}] \leavevmode
Extends at all 4 levels
Adds ``extents'' which are design to extend the base entities such as a role, multiple RoleExtents can belong to a single role based off constraint or context

\end{description}

Notes.
\end{quote}

\chapter{Task}
\label{categories/task:task}\label{categories/task::doc}\begin{description}
\item[{Definition.}] \leavevmode
a usually assigned piece of work often to be finished within a certain time

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4623993}{A Task-Role Based Access Control Model with Multi-Constraints}

\item {} 
\href{http://en.cnki.com.cn/Article\_en/CJFDTOTAL-TXJS200911047.htm}{Team and Task Based RBAC Access Control Model}

\item {} 
\href{http://www.sciencedirect.com/science/article/pii/S0306437902000297}{Task–role-based access control model}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{A Task-Role Based Access Control Model with Multi-Constraints}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Not much reasoning or evidence as to the `why'

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
RBAC can't handle large enterprises with multiple applications and a single user base

\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
Discussion of the applications of the model, mentions of real companies using the implementation, and a mention of a system that exists

\end{itemize}

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
Unclear, as their examples and applications didn't contain much data

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
rigorous constraints definitions, mathematical specification of additional model elements

\item {} 
Attempts to combine TBAC and RBAC by placing a task between a role and a permission

\item {} 
not really extending the base RBAC model

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Team and Task Based RBAC Access Control Model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
\item {} 
Real world scenarios - hospital

\item {} 
Applications of TT-RBAC in an imaginary hospital scenario

\end{itemize}

\end{description}

What reasons and evidence do researchers claim RBAC needs extension?
\begin{description}
\item[{Did they implement the model?}] \leavevmode\begin{itemize}
\item {} 
Prototype system

\end{itemize}

\end{description}

Is there an evaluation? If yes, how did they do one? If no, why?
\begin{description}
\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Covers all 4 levels of RBAC, provides mathematical definitions of entities

\item {} 
Extension to RBAC at all four levels

\item {} 
Introduces teams that map to users, roles and tasks

\item {} 
Introduces tasks that map to teams and permissions

\end{itemize}

\end{description}

Notes.
\end{quote}

\bigskip\hrule{}\bigskip

\textbf{Task–role-based access control model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
Define a set of security requirements for access control in an enterprise environment

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
8 security requirements for enterprise environments

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
No implementation, not enough envidence other than the security requirements, the requirements are based off a few thought scenarios that seem based on a `typical' enterprise organizational setup

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
extension by adding Task notion to connect with roles and users

\end{itemize}

\item[{Notes.}] \leavevmode
Task has constraints, task sits between permission and roles, tasks used to compose workflows, and are more mutable than roles - task can be used for defining set of actions and help when the specific data objects are less known at creation time

\end{description}
\end{quote}

\chapter{Temporal}
\label{categories/temporal:temporal}\label{categories/temporal::doc}\begin{description}
\item[{Definition.}] \leavevmode
of or relating to the sequence of time or to a particular time

\item[{Papers}] \leavevmode\begin{itemize}
\item {} 
\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1363762}{A generalized temporal role-based access control model}

\end{itemize}

\end{description}

\bigskip\hrule{}\bigskip

\textbf{A generalized temporal role-based access control model}
\begin{quote}
\begin{description}
\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
In many organizations, processes and functions may have limited time spans or have periodic temporal durations

\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
\item {} 
Workflow based organizations need a time based constraint for disabling/enabling roles

\end{itemize}

\item[{Did they implement the model?}] \leavevmode
No.

\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
\item {} 
Model is limited to enabling/disabling of roles based on time constraints only

\item {} 
model is fully described from entities, to constraints, to conflict resolution to execution in pseudo-code

\item {} 
Evaluation lies in the construction of mathematical modeling and graphs to detail the usage and interaction of the model pieces

\item {} 
A few example scenarios are explained

\end{itemize}

\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
\item {} 
Extension of the base RBAC model

\item {} 
adds periodicity and duration constraints for role enabling

\item {} 
provides for conflict resolution of constraints

\end{itemize}

\end{description}

Notes.
\end{quote}

\begin{thebibliography}{1}
\bibitem[1]{1}{\phantomsection\label{intro:1} \DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{David}\DUspan{}{ }\DUspan{}{F.}}\DUspan{}{ }}\DUspan{}{Ferraiolo}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Ravi}\DUspan{}{ }\DUspan{}{S.}}\DUspan{}{ }}\DUspan{}{Sandhu}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Serban}\DUspan{}{ }\DUspan{}{I.}}\DUspan{}{ }}\DUspan{}{Gavrila}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{D.}\DUspan{}{ }\DUspan{}{Richard}}\DUspan{}{ }}\DUspan{}{Kuhn}}}\DUspan{}{, and }\DUspan{}{\DUspan{}{\DUspan{}{Ramaswamy}\DUspan{}{ }}\DUspan{}{Chandramouli}}}\DUspan{}{.}}\DUspan{}{ }\DUspan{}{\DUspan{}{Proposed NIST standard for role-based access control}\DUspan{}{.}}}\DUspan{}{ }\DUspan{}{\DUspan{}{\emph{\DUspan{}{ACM Transactions on Information and System Security}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{4}\DUspan{}{\DUspan{}{(}\DUspan{}{3}\DUspan{}{)}}\DUspan{}{:}}\DUspan{}{\DUspan{}{224}\DUspan{}{–}\DUspan{}{274}}}}\DUspan{}{, }\DUspan{}{2001}\DUspan{}{.}}}}
\end{thebibliography}

\renewcommand{\indexname}{Index}
\printindex
\end{document}

### 2013-02-03 old/intro.tex
Moves original latex documents into old/ directory and adds a latex build directory for current documentation state.

--- 2013-02-03 latex/RBAC-SLR.tex
@@
-% Generated by Sphinx.
-\def\sphinxdocclass{report}
-\documentclass[letterpaper,10pt,english]{sphinxmanual}
-\usepackage[utf8]{inputenc}
-\DeclareUnicodeCharacter{00A0}{\nobreakspace}
-\usepackage[T1]{fontenc}
-\usepackage{babel}
-\usepackage{times}
-\usepackage[Bjarne]{fncychap}
-\usepackage{longtable}
-\usepackage{sphinx}
-\usepackage{multirow}
+\section{Introduction} \label{sec:introduction}
 
-\title{RBAC-SLR Documentation}
-\date{February 02, 2013}
-\release{0.1}
-\author{Eric D. Helms}
-\newcommand{\sphinxlogo}{}
-\renewcommand{\releasename}{Release}
-\makeindex
-
-\makeatletter
-\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
-    \let\PYG@ul=\relax \let\PYG@tc=\relax%
-    \let\PYG@bc=\relax \let\PYG@ff=\relax}
-\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
-\def\PYG@toks#1+{\ifx\relax#1\empty\else%
-    \PYG@tok{#1}\expandafter\PYG@toks\fi}
-\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
-    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
-\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}
-
-\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.25,0.82}{##1}}}
-\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
-\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@cs\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{1.00,0.94,0.94}{\strut ##1}}}
-\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
-\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.19,0.19,0.19}{##1}}}
-\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.84,0.33,0.22}{##1}}}
-\expandafter\def\csname PYG@tok@nl\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.13,0.44}{##1}}}
-\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
-\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.38,0.68,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
-\expandafter\def\csname PYG@tok@nd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
-\expandafter\def\csname PYG@tok@ne\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.49}{##1}}}
-\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.44,0.63,0.82}{##1}}}
-\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.45}{##1}}}
-\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
-\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
-\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.32,0.47,0.09}{##1}}}
-\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.14,0.33,0.53}{##1}}}
-\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.56,0.13,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-
-\def\PYGZbs{\char`\\}
-\def\PYGZus{\char`\_}
-\def\PYGZob{\char`\{}
-\def\PYGZcb{\char`\}}
-\def\PYGZca{\char`\^}
-\def\PYGZam{\char`\&}
-\def\PYGZlt{\char`\<}
-\def\PYGZgt{\char`\>}
-\def\PYGZsh{\char`\#}
-\def\PYGZpc{\char`\%}
-\def\PYGZdl{\char`\$}
-\def\PYGZti{\char`\~}
-% for compatibility with earlier versions
-\def\PYGZat{@}
-\def\PYGZlb{[}
-\def\PYGZrb{]}
-\makeatother
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-\phantomsection\label{index::doc}
-
-Contents:
-
-\chapter{1. Introduction}
-\label{intro:introduction}\label{intro:systematic-literature-review-of-rbac-analysis-notes}\label{intro::doc}
-Why is the base model of RBAC extended by newer models?
-
-Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.
-In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
-RBAC assigns permissions of specific actions on resources to roles instead of individual users.
+Role-based access control (RBAC) models~\cite{ferraiolo:rbac} become popularly used
+to govern access to critical resources.
+In an RBAC model, roles represent a group of users who are
+involved in a specific job function in an organization. 
+RBAC assigns permissions of specific actions on resources to roles
+instead of individual users.
 Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.
 
-RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.
+RBAC is a generalized access control approach used for various applications including
+web services, database applications, and healthcare applications.
 RBAC has advantages in maintaining and managing organization's security policies.
-For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.
+For example, if a user is to access manager role's resources within a given
+organization, security policy administrators simply add the user to be associated with the manager role.
 
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
+RBAC is first introduced in 1990s, NIST proposed standard RBAC model~\cite{ferraiolo:rbac}.
 Standard RBAC model considers only role-user association and role hierarchy.
@@
 As a result, researchers could take time on reinventing the wheel.    
-But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
+%But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
 
@@
 We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
-\begin{description}
-\item[{Our research give benefits to a community as follows:}] \leavevmode\begin{itemize}
-\item {} 
-Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
 
-\item {} 
-Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
-
-\item {} 
-Our work shows a criteria in comparison among research results.
-
-\item {} 
-Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
-
+Our research give benefits to a community as follows:
+\begin{itemize}
+	\item Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.	
+	\item Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing
+	research results helps decide a standard of extended RBAC.
+	\item Our work shows a criteria in comparison among research results.
+	\item Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.			
 \end{itemize}
 
-\end{description}
+%Agent-Based\cite{yamazaki04:designing}
+%
+%
+%
+%Task-based
+%Temporal-role based
+%Inheritance
+%Location
+%Delegation
 
-\chapter{2. Methodology and Process}
-\label{process:methodology-and-process}\label{process::doc}
-For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
-\begin{itemize}
-\item {} 
-Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}
+%\cite[ferraiolorbac]
 
-\item {} 
-IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}
+%Introduction come here.
+%Various techniques have been proposed on regression testing of software programs~\cite{Rothermel:1996:ART:235681.235682, Graves:2001:ESR:367008.367020}.
 
-\item {} 
-ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}
-
-\item {} 
-CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}
-
-\end{itemize}
-
-For each of the criteria below, a search was performed on each of the search engines for a total of 12 data sets.  The criteria used were:
-\begin{itemize}
-\item {} 
-role based access control
-
-\item {} 
-RBAC
-
-\item {} 
-role-based access control
-
-\end{itemize}
-
-The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.
-For each criteria for each search engine, the results were captured until a stopping criteria was met.
-Each run was performed as follows:
-\begin{enumerate}
-\item {} 
-Remote call to search engine with current search start position and the current search criteria.
-
-\item {} 
-Parse results and extract paper title, authors and year of publication.
-
-\item {} 
-Compare results against stopping criteria.
-
-\end{enumerate}
-\begin{itemize}
-\item {} 
-If stopping criteria met, stop search.
-
-\item {} 
-If stopping criteria not met, increase search position by number of results and return to step 1.
-
-\end{itemize}
-
-The stopping criteria used was either after the first 1000 results, a limitation imposed by some of the search engines, or if ten consecutive results did not contain the search criteria phrase within the title.
-After gathering all 12 data sets, the data was combined into a master list by systematically comparing the bibliographic information for each.
-After producing a master list, a series of assessment rounds were performed to narrow the paper list and identify primary sources.
-
-There were a total of three elimination rounds.
-The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.
-Each round was performed as follows:
-\begin{itemize}
-\item {} 
-Each reviewer independently classified papers as relevant, irrelevant or uncertain.
-
-\item {} 
-Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
-
-\item {} 
-Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
-From this discussion, papers were either thrown out or kept until the next round of the review.
-Ties were broken by an indepedent party.
-
-\end{itemize}
-
-The results of the searches is summarised below:
-
-\begin{tabulary}{\linewidth}{|L|L|L|L|L|}
-\hline
-\textbf{
-Search Engines
-} & \textbf{
-RBAC
-} & \textbf{
-role based access control
-} & \textbf{
-role-based access control
-} & \textbf{
-Total
-}\\\hline
-
-Google Scholar
- & 
-651
- & 
-213
- & 
-435
- & 
-1299
-\\\hline
-
-ACM Portal
- & 
-500
- & 
-20
- & 
-720
- & 
-1240
-\\\hline
-
-IEEExplore
- & 
-200
- & 
-40
- & 
-230
- & 
-470
-\\\hline
-
-CiteSeerX
- & 
-100
- & 
-100
- & 
-150
- & 
-350
-\\\hline
- &  &  &  & \\\hline
-
-Totals
- & 
-1451
- & 
-373
- & 
-1535
- & 
-3359
-\\\hline
-
-Combined
- &  &  &  & 
-\textbf{1716}
-\\\hline
-\end{tabulary}
-
-\section{Research Questions}
-\label{process:research-questions}
-Based on our intial intent for the systematic literature review, and notes taken during the third phase of elimination, we have identified the following research questions:
-\begin{itemize}
-\item {} 
-RQ1. What are the deficiencies in current RBAC model to propose extended RBAC models?
-
-\item {} 
-RQ2. What are the motivations behind RBAC extensions?
-
-\item {} 
-RQ3. What are the categorizations of RBAC model proposed?
-
-\item {} 
-RQ4. For each categorization, what are the extended features of RBAC model proposed?
-
-\item {} 
-RQ5. Do these models have corresponding implementations in practice?
-
-\item {} 
-RQ6. How are extended RBAC models evaluated in theory and in practice?
-
-\item {} 
-RQ7. Are there any commonalties or generalizations across all categorizations?
-
-\item {} 
-RQ8. What domains or scenarios serve as inspiration for these extensions?
-
-\end{itemize}
-
-\chapter{RQ1: What are the deficiencies in current RBAC model?}
-\label{research_questions/RQ_1:rq1-what-are-the-deficiencies-in-current-rbac-model}\label{research_questions/RQ_1::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_1:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_1:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-no support for partial inheritance
-
-\item {} 
-inability to support permission assignment constraints
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Missing ability to apply constraints between role and permissions and enforce
-
-\item {} 
-Missing context that defines if a role can be activated at execution time
-
-\item {} 
-Role activation on a per object or process execution rather than session based
-
-\item {} 
-No allowance for context to check permission at execution moment based on context
-
-\item {} 
-activation of roles based on current context rather than session intiation
-
-\item {} 
-conflict resolution
-
-\item {} 
-RBAC cannot handle dynamic authorization and entities
-
-\item {} 
-Role cannot be resource dependent
-
-\item {} 
-RBAC is too simple
-
-\item {} 
-Missing granularity
-
-\item {} 
-Inability to handle dynamic nature of multiple projects with multiple resources
-
-\item {} 
-Too simple
-
-\item {} 
-Lacks context
-
-\item {} 
-Handling authorization between applications and within can be different
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-RBAC cannot handle cross-organizational roles and access in an administratively easy way
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Privacy regulations are complex, with many entities, RBAC lacks ability to handle all of these
-
-\item {} 
-Context, conflict detection and resolution
-
-\item {} 
-RBAC lacks ability to define purpose
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-No support for multiple security domains
-
-\item {} 
-No support for same security, different applications
-
-\item {} 
-Lack of flexibility
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Lack of contextual information
-
-\item {} 
-Role needs a defined schema
-
-\item {} 
-Role activation based on context
-
-\item {} 
-Giving a permission an operation
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Needs an entity to represent a task and connect a role and a permission through it
-
-\item {} 
-Need for team and tasks
-
-\item {} 
-Need for entity between permission and roles to hold constraint and allow for dynamic situations and entities
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-RBAC can't handle time based constraints for role enablement
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ2: What are the motivations behind RBAC extensions?}
-\label{research_questions/RQ_2:rq2-what-are-the-motivations-behind-rbac-extensions}\label{research_questions/RQ_2::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_2:common-themes}\begin{itemize}
-\item {} 
-Roles, users and permissions need context associations
-
-\item {} 
-Location and time constraints as well as contexts needed
-
-\item {} 
-RBAC lacks ability to link in objects
-
-\item {} 
-RBAC does not handle single user base and multiple organizations/applications
-
-\item {} 
-Role enablement based on context
-
-\item {} 
-User granting/revoking of roles based on context
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_2:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No support for partial inheritance
-
-\item {} 
-Permission assignment constraints
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Relationship between object and viewer can't be represented
-
-\item {} 
-Object content matters
-
-\item {} 
-Location and time constraints on the user
-
-\item {} 
-Departmental context around access
-
-\item {} 
-RBAC does not handle many small services to wrap around them
-
-\item {} 
-RBAC lacks security principle enforcement
-
-\item {} 
-Lack of specifying context, and permission checks of dynamic evaluation time rules
-
-\item {} 
-RBAC is static and lacks ability to handle dynamic authorization and definitions
-
-\item {} 
-RBAC ignores context
-
-\item {} 
-Role granulatiry not well controlled
-
-\item {} 
-Roles should depend on the resources
-
-\item {} 
-RBAC cannot handle multiple projects with multiple tasks and differing roles in each system but the same user
-
-\item {} 
-Lack of context support
-
-\item {} 
-Too simple for dynamic environments
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-Support cross-domain delegation
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Reduce administrative complexity of RBAC for multi-organizational
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Lack of components, purpose binding, conditions, obligations to handle privacy needs
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-Need to support multiple securtiy domains
-
-\item {} 
-Need to support same security with different applications
-
-\item {} 
-Role heirarchy adds complexity and reduces flexibility
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Lack of contextual information with roles, users and permissions
-
-\item {} 
-Role needs a schema to allow specification of attributes
-
-\item {} 
-Location based security scenarios
-
-\item {} 
-Role activation based on location
-
-\item {} 
-Support physical and logical locations
-
-\item {} 
-Attaching operations to permissions
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-RBAC doesn't inherently support location and time constraints
-
-\item {} 
-Spatial and temporal awareness of roles, inheritance and permissions
-
-\item {} 
-Context mechanisms needed
-
-\item {} 
-Physical and logical locations, hybrids
-
-\item {} 
-Role assignment based off time and space, as well as permissions
-
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-RBAC can't handle large enterprises with multiple applications and single user base
-
-\item {} 
-Security requirements for enterprises (look up these)
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Workflow based organizations need time base constraint for role enablement
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ3: What are the categorizations of proposed RBAC extensions?}
-\label{research_questions/RQ_3:rq3-what-are-the-categorizations-of-proposed-rbac-extensions}\label{research_questions/RQ_3::doc}
-Across the papers found, a number of higher level categories emerged that each proposed model fell in to.
-Some of these categories had singular primary papers, while others had multiple extension models proposed that were deemed independent of one another.
-
-For each category, the breakdown of the papers for each category is:
-
-\begin{tabulary}{\linewidth}{|L|L|}
-\hline
-\textbf{
-Paper
-} & \textbf{
-Category
-}\\\hline
-
-Alam, M. and Hafner, M. and Breu, R. 2006
- & 
-Constraint
-\\\hline
-
-Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001
- & 
-Context
-\\\hline
-
-Haibo, SHEN and Fan, HONG 2005
- & 
-Context
-\\\hline
-
-Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000
- & 
-Context
-\\\hline
-
-Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006
- & 
-Context
-\\\hline
-
-Motta, G.H.M.B. and Furuie, S.S. 2003
- & 
-Context
-\\\hline
-
-Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008
- & 
-Context
-\\\hline
-
-Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004
- & 
-Context
-\\\hline
-
-Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008
- & 
-Context
-\\\hline
-
-Thein, N. and others 2011
- & 
-Context
-\\\hline
-
-Zou, D. and He, L. and Jin, H. and Chen, X. 2009
- & 
-Context
-\\\hline
-
-Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010
- & 
-Delegation
-\\\hline
-
-Zhang, Z. and Zhang, X. and Sandhu, R. 2006
- & 
-Organizational
-\\\hline
-
-Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007
- & 
-Privacy
-\\\hline
-
-Masoumzadeh, A. and Joshi, J. 2008
- & 
-Privacy
-\\\hline
-
-Zhao, Y. and Zhao, Y. and Lu, H. 2008
- & 
-Resource
-\\\hline
-
-Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005
- & 
-Spatial
-\\\hline
-
-Ray, I. and Kumar, M. and Yu, L. 2006
- & 
-Spatial
-\\\hline
-
-Hansen, F. and Oleshchuk, V. 2003
- & 
-Spatial
-\\\hline
-
-Aich, S. and Sural, S. and Majumdar, A. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Chen, L. and Crampton, J. 2008
- & 
-Spatio-Temporal
-\\\hline
-
-Samuel, A. and Ghafoor, A. and Bertino, E. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Chandran, S. and Joshi, J. 2005
- & 
-Spatio-Temporal
-\\\hline
-
-Ray, I. and Toahchoodee, M. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009
- & 
-Spatio-Temporal
-\\\hline
-
-Yao, L. and Kong, X. and Xu, Z. 2008
- & 
-Task
-\\\hline
-
-ZHANG, S. and CHEN, X. and HOU, G. 2009
- & 
-Task
-\\\hline
-
-Oh, S. and Park, S. 2003
- & 
-Task
-\\\hline
-
-Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005
- & 
-Temporal
-\\\hline
-\end{tabulary}
-
-\section{By Category}
-\label{research_questions/RQ_3:by-category}\begin{description}
-\item[{Constraint (1)}] \leavevmode
-a limitation or restriction
-
-\item[{Context (8)}] \leavevmode
-circumstances in which an event occurs
-
-\item[{Delegation (1)}] \leavevmode
-the act of empowering to act for another
-
-\item[{Organizational (1)}] \leavevmode
-of or relating to an organization
-
-\item[{Privacy (1)}] \leavevmode
-freedom from unauthorized intrusion
-
-\item[{Resource (1)}] \leavevmode
-an available supply than be drawn on when needed
-
-\item[{Spatial (3)}] \leavevmode
-of or relating to facility in percieving relations in space
-
-\item[{Spatio/Temporal (6)}] \leavevmode
-combining space and time
-
-\item[{Task (3)}] \leavevmode
-a usually assigned piece of work often to be finished within a certain time
-
-\item[{Temporal (1)}] \leavevmode
-of or relating to the sequence of time or to a particular time
-
-\end{description}
-
-{\hfill\scalebox{1.000000}{\includegraphics{categories_map.png}}\hfill}
-
-\chapter{RQ4: For each categorization, what are the extended features of RBAC proposed?}
-\label{research_questions/RQ_4::doc}\label{research_questions/RQ_4:rq4-for-each-categorization-what-are-the-extended-features-of-rbac-proposed}
-Is this the same as RQ\_1 ?
-
-\section{Common Themes}
-\label{research_questions/RQ_4:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_4:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ5. Do these models have corresponding implementations in practice?}
-\label{research_questions/RQ_5::doc}\label{research_questions/RQ_5:rq5-do-these-models-have-corresponding-implementations-in-practice}
-When designing and proposing a model targeted at a feature that is rooted in practical
-usage by real software systems, bringing the model to life is strong evidence that the
-proposed model can work in practice.  The concept of authorization, and access control
-is rooted in a business need.  Thus, any access control model needs to be feasible
-in the real world not just on paper.  We analyzed the primary sources to see how many
-proposed models actually had implmenetations associated with them.  And quantified the
-type of implementation.  Whether the implementation was for a real system, for a prototype
-and/or used in a production environment.
-
-\section{Common Themes}
-\label{research_questions/RQ_5:common-themes}\begin{itemize}
-\item {} 
-21 No's
-
-\item {} 
-8 Yes's
-
-\item {} 
-Half the implementation's were prototypes
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_5:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-No - system architecture provided
-
-\item {} 
-No
-
-\item {} 
-Yes - prototype with VB6 and MS Access97
-
-\item {} 
-Yes - prototype system, system architecture, algorithms implemented, simulation scenarios
-
-\item {} 
-Yes - on an EHR system
-
-\item {} 
-Yes - prototype using XML container
-
-\item {} 
-Yes - web based API agent
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-Yes - integrated with telemedicine application, web based
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Yes - mentions of systems that exist using the model
-
-\item {} 
-Yes - prototype system
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ6. How are extended RBAC models evaluated in theory and in practice?}
-\label{research_questions/RQ_6:rq6-how-are-extended-rbac-models-evaluated-in-theory-and-in-practice}\label{research_questions/RQ_6::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_6:common-themes}\begin{itemize}
-\item {} 
-12 had no evaluation whatsoever
-
-\item {} 
-8 present example scenarios and how their model would apply
-
-\item {} 
-6 presented performance and time complexity analysis
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_6:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No evaluation
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Present hospital scenarios and explanation of how new model would handle
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-Scenarios with how model affects, performance test of the model vs. traditional RBAC
-
-\item {} 
-Description of a fake project management system that the model is designed for
-
-\item {} 
-Applied to conceptual situation
-
-\item {} 
-No
-
-\item {} 
-Example scenario comparing their model vs other models
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Comparison caluclation for number of roles vs standard model, numerical size comparisons against other proposed models
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Evaluate the performance of permission assignment and authorization checks for various sizes and context variables
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-Model comparison from a qualitative standpoint
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Algorithm and Big-O performance evaluation
-
-\item {} 
-No except minor scenarios and how the model handles
-
-\item {} 
-No other than mathematical specification
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Mathematical description
-
-\item {} 
-Mathematical analysis of role graph, complexity comparisons across different models
-
-\item {} 
-Pseudo example scenario
-
-\item {} 
-No
-
-\item {} 
-Detailed scenario with examination of how model handles situation
-
-\item {} 
-Discuess difference btween their model and other models, time-complexity analysis to show no performance impact with standard model
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Mathematical modeling and graphs to detail usage and interactions, explanation of example scenarios
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ7. Are there any commonalties or generalizations across all categorizations?}
-\label{research_questions/RQ_7:rq7-are-there-any-commonalties-or-generalizations-across-all-categorizations}\label{research_questions/RQ_7::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_7:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_7:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ8. What domains or scenarios serve as inspiration for these extensions?}
-\label{research_questions/RQ_8:rq8-what-domains-or-scenarios-serve-as-inspiration-for-these-extensions}\label{research_questions/RQ_8::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_8:common-themes}\begin{itemize}
-\item {} 
-Medical domain
-
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Mobile devices
-
-\item {} 
-Large-scale organizations with many sub-departments
-
-\item {} 
-Enterprise, organization workflows
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_8:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-Medical domain
-
-\item {} 
-Service Oriented Architecture
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Medical imaging database
-
-\item {} 
-Handling an insurance claim workflow
-
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Medical emergency overrides
-
-\item {} 
-Electronic health records
-
-\item {} 
-Supermarket
-
-\item {} 
-Project management system with large project, task and user base
-
-\item {} 
-Power load forecasting system (multi-level user environment)
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-Large-scale electronic health record systems
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Online tutoring system
-
-\item {} 
-State education system with schools, districts, etc.
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Privacy within the medical domain, HIPAA
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-N/A
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Hospital with multi-floors, departments
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Hospital using mobile devices with multiple user types on a single floor, departments
-
-\item {} 
-Mobile, peer-to-peer
-
-\item {} 
-Doctor out in the world needing access during a random medical emergency
-
-\item {} 
-Students at a university
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Large enterprises with multiple applications and a single user base
-
-\item {} 
-Hospital needs
-
-\item {} 
-Complex enterprise environments
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-General organizational needs for time constraints
-
-\end{itemize}
-
-\end{description}
-
-\chapter{Constraint}
-\label{categories/constraint::doc}\label{categories/constraint:constraint}\begin{description}
-\item[{Definition}] \leavevmode
-a limitation or restriction
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1501451}{A constraint based role based access control in the SECTET a model-driven approach}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A constraint based role based access control in the SECTET a model-driven approach}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example?}] \leavevmode
-SECTECT is a framework that provides workflow for Service oriented Architectures and is the primary target.
-
-\item[{Scenarios or examples.}] \leavevmode
-Presented generic scenarios from the medical domain as examples to describe how mechanisms worked.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-The claim is that with respect to Service Oriented Architecture, and the execution of web services, RBAC lacks needed features.
-
-\item {} 
-RBAC does not support partial inheritance
-
-\item {} 
-``total inheritance of permissions in RBAC is against the principle of least privilege''
-
-\item {} 
-ability for some sub-role permissions not be available to super role
-
-\item {} 
-``permission assignment constraints''
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No evidence of use in a real system.
-
-\item[{Is there an evaluation? If yes, how did they do one? If not, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Presents a modification to a core aspect of RBAC
-
-\item {} 
-Allow partial inheritance at the permission level by a superrole instead of treating the role as an atomic entity
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode
-Apply a restricted status to a permission which prevents a superrole from inheriting that permission
-
-\end{description}
-\end{quote}
-
-\chapter{Context}
-\label{categories/context::doc}\label{categories/context:context}\begin{description}
-\item[{Definition.}] \leavevmode
-circumstances in which an event occurs
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1232473}{A flexible content and context-based access control model for multimedia medical image database systems}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpl/login.jsp?tp=\&arnumber=1552897\&url=http\%3A\%2F\%2Fieeexplore.ieee.org\%2Fxpls\%2Fabs\_all.jsp\%3Farnumber\%3D1552897}{A Context-Aware Role-Based Access Control Model for Web Services}
-
-\item {} 
-\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.102.9126\&rep=rep1\&type=pdf}{A context-sensitive access control model and prototype implementation}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4079196}{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1229859}{A contextual role-based access control authorization model for electronic patient record}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4709140}{A Role and Context Based Access Control Model with UML}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1376833}{Designing an agent-based RBAC system for dynamic security policy}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4664701}{An extended RBAC model based on granular logic}
-
-\item {} 
-\href{http://ieeexplore.ieee.org.prox.lib.ncsu.edu/xpl/articleDetails.jsp?tp=\&arnumber=5967517\&contentType=Conference+Publications\&searchField\%3DSearch\_All\%26queryText\%3Dleveraging+access+control+mechanism+of+android+smartphone+using+context-related+role-based+access+control+model}{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}
-
-\item {} 
-\href{http://www.sciencedirect.com.prox.lib.ncsu.edu/science/article/pii/S1084804508000520}{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A flexible content and context-based access control model for multimedia medical image database systems}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Real world inspiration based upon medical image database.  One simple scenario shown where the model is applicable.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-What the image is of is important to who can view
-
-\item {} 
-Domain governs who can see a particular image (e.g. departments in a hospital)
-
-\item {} 
-Location for determining if user is accessing from secure location or location of importance (e.g. emergency room)
-
-\item {} 
-Time constraints for validity periods
-
-\item {} 
-Relationship between images and viewer
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Provided system architecture for entire medical imaging database with access control system. No implied implementation of model.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No evaluation.  Present proposed situation and hospital structure with scenario where new model would be applicable. No reason for lack of evaluation given.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Proposes adding of constraints between the Role-Permission relationship which must be satisfied in order for the holder of said role to be able to use those permissions
-
-\end{itemize}
-
-\item[{This appears to be an active mechanism during enforcement not assignment}] \leavevmode\begin{itemize}
-\item {} 
-Adds user attributes (e.g. company position, location, user name)
-
-\item {} 
-Adds a decision rule to role-permission relationship
-
-\item {} 
-Role-perm relationship contains 5 attributes: identifier, subject/role, action, target objects, constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Context-Aware Role-Based Access Control Model for Web Services}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-No.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-claim is that RBAC can not handle many small services that need roles/perms wrapped around them
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-services replace objects/actions
-
-\item {} 
-adds contexts that must be satisfied
-
-\item {} 
-splits roles into local and global, where local is reflective of actions within a service/domain and global roles are reflective of the permissions to access the service
-
-\item {} 
-roles are activated based on context at execution time
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-context oriented, provides what appears to be a different spin or implementation of heirarchies
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Context-sensitive access control model and prototype implementation}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Initiates the paper with case scenario ``handle an insurance claim'' workflow.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-points out contrasts between base RBAC and improvements clearly
-
-\item {} 
-claims that RBAC supports security principles but does not enforce the use of
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Prototype created using VB6 and MS Access97.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-largely using mechanisms prescribed by RBAC specifically SoD and LP see note about how RBAC provides for but not the use of the previous security practices
-
-\item {} 
-roles are activated whenever a tasks is chosen to be performed and checked if the user can have said role
-
-\item {} 
-provides more implementation of SoD and LP than additions to model
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-context oriented, provides what appears to be a different spin or implementation of heirarchies
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Inspired by the needs of access control in pervasive computing environments
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-``static method of management user-role-permissions''
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Created prototype system to verify model correctness and system architecture
-
-\item {} 
-Implemented the algorithms proposed by their model
-
-\item {} 
-Simulated scenarios but the details of the simulations are not in the paper, nor are the results in a consumable manner
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Claim to be an extension and follow the standard RBAC execution model except by adding two different checks where permissions are re-evaluated by the context of the current user according to the already specific rules
-
-\item {} 
-context aware rules that dynamically determine permissions
-
-\item {} 
-constraining the permissions based on a users location
-
-\item {} 
-assignment/removal of a role based on context information
-
-\item {} 
-connection of roles, granting a user certain role permissions when another user/role is present
-
-\item {} 
-Provide mechanism to deal with conflicts
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A contextual role-based access control authorization model for electronic patient record}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Traditional RBAC does not support emergency overrides based on the situation
-
-\item {} 
-Inspired by the needs of an EHR application in practice
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Claim RBAC is static and cannot handle the dynamic authorization and definitions needed by EHR application complexity
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Implemented using Java and LDAP on InCor EHR system
-
-\item {} 
-They quote number of users and rules, but provide no evidence of comparison or any observation data of the system in use
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Showed some discussion evidence for the requirement for these extensions to RBAC, and they implemented it, however, they provided little evidence as to the effectiveness
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Claim to be a context based extension
-
-\item {} 
-Focused on the authorization aspect of permissions for a given role
-
-\item {} 
-Contains both positive and negative permission authorizations
-
-\item {} 
-Provides for overriding of authorizations based on inheritance
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Role and Context Based Access Control Model with UML}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Describes a scenario whereby a supermarket could not effectively use RBAC to encapsulate all the various access scenarios they need
-
-\item {} 
-Propose a model that adds attributes and system context
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-RBAC ignores context
-
-\item {} 
-RBAC is too simple and unilateral and does not align with real world
-
-\item {} 
-Role granularity is not well controlled
-
-\item {} 
-Roles should depend on the resources not vice-versa
-
-\item {} 
-Roles should e assigned to one resource instead of resources being assigned to one role
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Prototype system created using XML container to contain roles and conditionals
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No evidence as to the comparison of the model in terms of the need over traditional RBAC other than conjecture around a scenario
-
-\item {} 
-The performance test is really only valid for their implementation and shows no grounded re-worldness
-
-\item {} 
-performed a performance analysis of traditional RBAC and C-RBAC based off their prototype
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode
-Extension of the traditional model adding context and maps roles to context and contex to resources with authorization and verification mixed in
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Designing an agent-based RBAC system for dynamic security policy}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-users and application sbecoming larger
-
-\item {} 
-many applications have dynamic attributes and defining scenarios for all of them is difficult
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-For their project manage system, traditional RBAC cannot handle the dynamic nature of multiple projects with multiple tasks that different users may be managing in one and not in another
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Talks about some weird web based API for the agent
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Discuss a fake project management system claiming this type of system is the target for this type of access control
-
-\item[{Does this model extend the core model?}] \leavevmode
-Defines model clearly
-Extension that defines an Abstract Role, Context Rules and Context Information and the actual role is decided upon based off the context inputs and the rules
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{An extended RBAC model based on granular logic}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Inspired by a muli-level user environment with a complicated authorization management - power load forecasting system
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-traditional RBAC does not provide context
-
-\item {} 
-RBAC is too simple for large dyanmic environments
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No evidence, applied in theory to a conceptual situation, no comparison
-
-\item[{Does this model extend the core model?}] \leavevmode
-Additions to specification, and rules around whether a user is authorized to perform an action based on their role and the context within the role
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Access control needs of smart phones
-
-\item {} 
-Installation of third party applications that a user needs to trust
-
-\item {} 
-User must grant device privileges to the application
-
-\item {} 
-Parents want to limit the amount of time kids use phone
-
-\item {} 
-User might can to limit accessbility to friends and admins
-
-\item {} 
-Companies want to limit data access by employee phones
-
-\item {} 
-User loses company phone in admin mode, context to prevent leakage
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-smartphone is centralized, user-centric system where identities are known in advance
-
-\item {} 
-smartphone has lots of contextual info
-
-\item {} 
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Adds objects, environmental context, policies and decisions to model
-
-\item {} 
-Object is an accessible entity
-
-\item {} 
-a property of the system at the moment of interaction
-
-\item {} 
-policy is the formal specification of the access control
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-precedential order of access privileges to prvent policy bugs
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-proliferation of distributed applications
-
-\item {} 
-example of limited disk space when a database read and update are performed to resolve which operation to give preference to
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-need flexible and dynamic authorization constraints
-
-\item {} 
-authorization differs between and within applications
-
-\item {} 
-Users may be granted access to the application and entities or may be granted access to objects with the application depending on context
-
-\item {} 
-role and permission constraints are also possible
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-analysis of example scenario, comparison of constraint vs other models,
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-adds object sets, entities, status and set of authorization attributes
-
-\item {} 
-permissions on objects with constraints
-
-\item {} 
-user and user constraints mapped to roles with permission attribute constraints
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-classification of constraints as either users eligibility to to use a resource/service or constraints on the users actual use of a resource (limiting what they can do)
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\chapter{Historical}
-\label{categories/historical::doc}\label{categories/historical:historical}\begin{description}
-\item[{Definition.}] \leavevmode
-based on or reconstructed from an event, custom, style, etc., in the past
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.182.8833\&rep=rep1\&type=pdf}{HRBAC: Historical Role-Based Access Control}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{HRBAC: Historical Role-Based Access Control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-a need to incorporate information of the past to make access control decisions
-
-\item {} 
-for example, ensuring a user had a set of previous roles before being granted a higher responsibility role (e.g. ``chief'' role needs to have once had a manager role)
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-in order to store and express historical entities of each base entity in order to apply constraints of current assignments and roles and permissions based off the past
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No, only examples of mechanisms
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Yes, extends each of the 4 levels of RBAC
-
-\item {} 
-incorporates new entities to represent historical version of each core entity
-
-\item {} 
-provides various historical constraints
-
-\item {} 
-provides algorithms for performing constraint resolution
-
-\item {} 
-highly defined model and entities
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Organizational}
-\label{categories/organizational:organizational}\label{categories/organizational::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to an organization
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4207551}{ROBAC: Scalable Role and Organization Based Access Control Models}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{ROBAC: Scalable Role and Organization Based Access Control Models}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-inspired by context from business-to-business and business-to-consumer context
-
-\item {} 
-a theoritical web based report delivery system e.g. a set of educational professionals from schools, districts, and states
-
-\item {} 
-also provide example of an online tutoring system
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-claim that their model can significantly reduce administrative complexity for multi-organization
-
-\item {} 
-more succint and intuitive than RBAC
-
-\item {} 
-same expressive power
-
-\end{itemize}
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-present comparison calculations for the number of roles that must be generated between their model and standard
-
-\item {} 
-provide numerical size comparisons with other proposed models (Role Templates, Team-based Access Control, Organizational Units
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension of RBAC to consider organization and role (a.k.a. context)
-
-\item {} 
-an organization must own the object, and the role must give permission on the type for a user to access
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Privacy}
-\label{categories/privacy::doc}\label{categories/privacy:privacy}\begin{description}
-\item[{Definition.}] \leavevmode
-freedom from unauthorized intrusion
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1266848}{Privacy-aware role-based access control}
-
-\item {} 
-\href{http://www.sis.pitt.edu/~amirreza/papers/masoumzadeh2008\_purbac.pdf}{PuRBAC: Purpose-Aware Role-Based Access Control}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Privacy-aware role-based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-inspired by conventional RBAC's inability to handle privacy
-
-\item {} 
-extra attributes are needed to support the intricacies of privacy
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}]
-due to the lack of basic components required by privacy regulations, especially purpose binding (i.e., data collected for one purpose should not used for another purpose without user consent), conditions, and obligations.
-Despite its limitations, existing access control technology can be used as a starting point for managing personal identifiable information
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Evaluated the performance of assigning permissions and checking authorization for various permission size bases as well as variety in number of context variables
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Defines additions to each level of the base RBAC model
-
-\item {} 
-Adds objects, purposes, actions, obligations and conditions to core
-
-\item {} 
-Includs context variables that can be very general, from space to time to other
-
-\item {} 
-Conflict detection and resolution mechanisms
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{PuRBAC: Purpose-Aware Role-Based Access Control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Web and companies with privacy policies make it harder to follow said policies and protect user data
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-privacy policies include purposes under which collected data can be used
-
-\item {} 
-current RBAC model does not include a entity or relationships to incorporate purpose
-
-\item {} 
-authorization requests should include a purpose for performing the intended action
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-No.
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No evaluation, conclusion includes note that further analysis needs to be done
-
-\item {} 
-Discusses entities added by this model, the way they were added and why this method is the best way to add purpose and support privacy
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extends the base and heirarchy portion of the model, adds in a hybrid heirarchy
-
-\item {} 
-connects purposes to permissions
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-claim to reduce policy complexity by associating a purpose with only a permission and not with roles and users as well
-
-\item {} 
-locks down the users ability to declare purpose, by putting authorization controls around the purposes
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\chapter{Resource}
-\label{categories/resource:resource}\label{categories/resource::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-an available supply that can be drawn on when needed
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4609645}{A Flexible Role- and Resource-Based Access Control Model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Flexible Role- and Resource-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-No.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-RBAC cannot support multiple security domains
-
-\item {} 
-cannot support same security with different applications
-
-\item {} 
-lack of flexibility in authorization
-
-\item {} 
-discard role heirarchy for directory structure to reduce complexity and increase flexibility
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Comparison of the two models from a qualitative standpoint
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Not an extension
-
-\item {} 
-Re-imagining of the hierarchy aspect and adding a resources entity
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Spatial}
-\label{categories/spatial::doc}\label{categories/spatial:spatial}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to facility in perceiving relations in space
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1063985}{GEO-RBAC: a spatially aware RBAC}
-
-\item {} 
-\href{http://www.springerlink.com/index/4N6XK46321M574UH.pdf}{LRBAC: A location-aware role-based access control model}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1285394}{Spatial role-based access control model for wireless networks}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{GEO-RBAC: a spatially aware RBAC}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Hospital based examples of activating roles, allowing doctors to only see patient records while in the department of the patient
-Extended example section that discusses how a hospital setting with some typical actors would perform simple patient access
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Lack of contextual information with roles, users and permissions
-Role needs a schema which allows specification of more attributes of the role
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Shows that this model can be be represented by first-order logic and that they can be evaluated in PTIME ?
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Introduces the concept of a spatial role
-
-\item {} 
-Extends the base model at each of the 4 levels
-
-\item {} 
-Specify mutual exclusion constraint for separation of duty
-
-\item {} 
-Split model into a model without constraints and an extensive model with constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{LRBAC: A location-aware role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Some brief real world concepts with regards to intricate security scenarios based on location of user or object
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-Allowing for different security scenarios based on location
-
-\item {} 
-Role activation based on location of object or user
-
-\item {} 
-Add objects to define physical or logical with location data
-
-\item {} 
-Attaches an operation to a permission
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Describes in full definition each object and it's role in the graph
-Describes a number of specific operations that would be performed and how the model handles them with respect to locations
-Description of how the core RBAC is related to each piece of their model
-No evaluation
-Minor scenarios
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Extension adding operation, object and role enabling based on a constraint
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Spatial role-based access control model for wireless networks}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Provide an application scenario of where this would be needed and used
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-standard RBAC cannot handle location based constraints and attributes
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Mathematical specification of each aspect
-Definitions of each element in their model
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension at every level of RBAC
-
-\item {} 
-includes separation of duty and heiarchry concerns specifically
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Spatio-Temporal}
-\label{categories/spatio_temporal:spatio-temporal}\label{categories/spatio_temporal::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-combining space and time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://www.springerlink.com/index/tq15n88122v63754.pdf}{STARBAC: Spatiotemporal role based access control}
-
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1368341}{On spatio-temporal constraints and inheritance in role-based access control}
-
-\item {} 
-\href{https://www.cerias.purdue.edu/tools\_and\_resources/bibtex\_archive/archive/2007-08.pdf}{A framework for specification and verification of generalized spatio-temporal role based access control model}
-
-\item {} 
-\href{http://www.springerlink.com/index/77j0j540004585wv.pdf}{LoT-RBAC: A Location and Time-Based RBAC Model}
-
-\item {} 
-\href{http://www.springerlink.com/index/C3T50338535HX115.pdf}{A spatio-temporal role-based access control model}
-
-\item {} 
-\href{http://www.springerlink.com/index/r1n6j114g5431174.pdf}{Role Based Access Control with Spatiotemporal Context for Mobile Applications}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{STARBAC: Spatiotemporal role based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Similar reasoning as other location based models, No formal evaluation, scenarios or real world examples.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Extension of RBAC and of SRBAC by combining spatial and temporal
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Mathematical explanations, definitions
-Includes descriptions and math behind the semantics
-
-\end{description}
-
-Does this model extend the core model?
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{On spatio-temporal constraints and inheritance in role-based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Believe that previous models focus on syntax and not semantics
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-pervasive computing environments requirement for spatial and temporal awareness
-
-\item {} 
-previous spatio-temporal do nt address interaction between these constraints and inheritance
-
-\item {} 
-makes 4 claims about how existing models do not handle the above
-
-\item {} 
-trusted entities to override restrictions
-
-\item {} 
-attempt to show compatibility with standard model and claim others do not
-
-\end{itemize}
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Examine previous models and point out deficiencies, use of graphs to show off relationships
-
-\item {} 
-Evaluation includes a mathematical examination of the role graphs, and ordering
-
-\item {} 
-authors note that there is added complexity in their method, but that is the price to pay to get user  and user-role application constraints
-
-\item {} 
-trade-off between complexity of policies and complexity of constraints
-
-\item {} 
-authors do comparisons of complexity for the various models
-
-\item {} 
-authors note that SoD should be considered in the future
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode
-Additions to the specification for constraints with a specific look at how to do spatial and temporal
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A framework for specification and verification of generalized spatio-temporal role based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Based off hospital example with mobile devices inspiring combining location and time based models
-Provides an example hospital access control policy that requires space and time - based on the layout of the floor of a hospital and the various users and entitie sthat might exist such as nurse, doctor and surgeons
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Need to be able to handle add attributes that location and spatial specification requires
-Add context mechanisms
-
-\item[{Did they implement the model?}] \leavevmode
-No
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-To some degree, an example, and some analysis of developing policy and how to deal with conflicts
-
-\item[{Does this model extend the core model?}] \leavevmode
-Builds on GTRBAC, extends at all 4 levels
-Adds role activation/deactivation
-Provides conflict resolution
-Includes a GSTRBAC policy specification model using Alloy
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{LoT-RBAC: A Location and Time-Based RBAC Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Inspired by growing needs of mobile, peer-to-peer devices with regards to activating and specifying permissions and roles based on time and location
-Detailed scenario of a person having a medical emergency, and a doctor on site needing to request patient data, as well as the progression through the ambulance, to hospital and citing such location based clues to help know if the doctor should be granted privileges such as being in or near his own car
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Previous location and time based RBAC models did not incorporate location for both the user and role
-In-depth, detailed location heirarchies and specifications are needed
-Physical, logical, and hybrid location
-
-\item[{Did they implement the model?}] \leavevmode
-No
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No
-
-\end{description}
-
-Does this model extend the core model?
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Spatio-temporal Role-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Growth of wireless networks, mobile devices, the move towards pervasive computing
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Claim that previously examined models of spatial, temporal, or spatio-temporal lack needed aspects
-Role assignment should be dependent on space and time
-Permissions need to be dependent on space and time not just roles
-
-\item[{Did they implement the model?}] \leavevmode
-No. Authors acknowledge need to implement.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.  Best they do is provide a detailed scenario with an examination of how such a scenario would be represented in their model. Author acknowledge detailed analysis is needed.
-
-\item[{Does this model extend the core model?}] \leavevmode
-Defines for Space - physical location, logical location and mapping functions and relationship definitions
-Defines for Time - time instant, time interval
-Adds objects to Core model, object can be physical or logical (computer or files e.g.)
-Defines a number of potential role hierarchy strategies
-Examines impact on both static and dynamic separation of duties for new attributes and mechanisms
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Role Based Access Control with Spatiotemporal Context for Mobile Applications}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Need and desire to put spatio-temporal context and constraint on roles and permissions - such as students at a university or hospital facilities
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Needs enhancements to be able to handle space and time attributes on roles and permissions
-
-\item[{Did they implement the model?}] \leavevmode
-Yes.  Integrated an access control system into iMedik - a telemedicine application in India that is web-based and allows mobile access
-Their implementation is a bolt-on solution, that intercepts calls and verifies their authorization
-Provide detailed architecture and implementation details
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Include section that discusses differences between their model and STRBAC and STARBAC
-Incorporate algorithms for determining whether a particular authorization is granted and use time-complexity arguments to discuss the effectiveness of each algorithm and to attempt to display that the added complexity of their model over traditional RBAC does not impace performance significantly
-
-\item[{Does this model extend the core model?}] \leavevmode
-Extends at all 4 levels
-Adds ``extents'' which are design to extend the base entities such as a role, multiple RoleExtents can belong to a single role based off constraint or context
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Task}
-\label{categories/task:task}\label{categories/task::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-a usually assigned piece of work often to be finished within a certain time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4623993}{A Task-Role Based Access Control Model with Multi-Constraints}
-
-\item {} 
-\href{http://en.cnki.com.cn/Article\_en/CJFDTOTAL-TXJS200911047.htm}{Team and Task Based RBAC Access Control Model}
-
-\item {} 
-\href{http://www.sciencedirect.com/science/article/pii/S0306437902000297}{Task–role-based access control model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Task-Role Based Access Control Model with Multi-Constraints}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Not much reasoning or evidence as to the `why'
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-RBAC can't handle large enterprises with multiple applications and a single user base
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Discussion of the applications of the model, mentions of real companies using the implementation, and a mention of a system that exists
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Unclear, as their examples and applications didn't contain much data
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-rigorous constraints definitions, mathematical specification of additional model elements
-
-\item {} 
-Attempts to combine TBAC and RBAC by placing a task between a role and a permission
-
-\item {} 
-not really extending the base RBAC model
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Team and Task Based RBAC Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Real world scenarios - hospital
-
-\item {} 
-Applications of TT-RBAC in an imaginary hospital scenario
-
-\end{itemize}
-
-\end{description}
-
-What reasons and evidence do researchers claim RBAC needs extension?
-\begin{description}
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Prototype system
-
-\end{itemize}
-
-\end{description}
-
-Is there an evaluation? If yes, how did they do one? If no, why?
-\begin{description}
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Covers all 4 levels of RBAC, provides mathematical definitions of entities
-
-\item {} 
-Extension to RBAC at all four levels
-
-\item {} 
-Introduces teams that map to users, roles and tasks
-
-\item {} 
-Introduces tasks that map to teams and permissions
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Task–role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Define a set of security requirements for access control in an enterprise environment
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-8 security requirements for enterprise environments
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No implementation, not enough envidence other than the security requirements, the requirements are based off a few thought scenarios that seem based on a `typical' enterprise organizational setup
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension by adding Task notion to connect with roles and users
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode
-Task has constraints, task sits between permission and roles, tasks used to compose workflows, and are more mutable than roles - task can be used for defining set of actions and help when the specific data objects are less known at creation time
-
-\end{description}
-\end{quote}
-
-\chapter{Temporal}
-\label{categories/temporal:temporal}\label{categories/temporal::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to the sequence of time or to a particular time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1363762}{A generalized temporal role-based access control model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A generalized temporal role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-In many organizations, processes and functions may have limited time spans or have periodic temporal durations
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-Workflow based organizations need a time based constraint for disabling/enabling roles
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Model is limited to enabling/disabling of roles based on time constraints only
-
-\item {} 
-model is fully described from entities, to constraints, to conflict resolution to execution in pseudo-code
-
-\item {} 
-Evaluation lies in the construction of mathematical modeling and graphs to detail the usage and interaction of the model pieces
-
-\item {} 
-A few example scenarios are explained
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Extension of the base RBAC model
-
-\item {} 
-adds periodicity and duration constraints for role enabling
-
-\item {} 
-provides for conflict resolution of constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\begin{thebibliography}{1}
-\bibitem[1]{1}{\phantomsection\label{intro:1} \DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{David}\DUspan{}{ }\DUspan{}{F.}}\DUspan{}{ }}\DUspan{}{Ferraiolo}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Ravi}\DUspan{}{ }\DUspan{}{S.}}\DUspan{}{ }}\DUspan{}{Sandhu}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Serban}\DUspan{}{ }\DUspan{}{I.}}\DUspan{}{ }}\DUspan{}{Gavrila}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{D.}\DUspan{}{ }\DUspan{}{Richard}}\DUspan{}{ }}\DUspan{}{Kuhn}}}\DUspan{}{, and }\DUspan{}{\DUspan{}{\DUspan{}{Ramaswamy}\DUspan{}{ }}\DUspan{}{Chandramouli}}}\DUspan{}{.}}\DUspan{}{ }\DUspan{}{\DUspan{}{Proposed NIST standard for role-based access control}\DUspan{}{.}}}\DUspan{}{ }\DUspan{}{\DUspan{}{\emph{\DUspan{}{ACM Transactions on Information and System Security}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{4}\DUspan{}{\DUspan{}{(}\DUspan{}{3}\DUspan{}{)}}\DUspan{}{:}}\DUspan{}{\DUspan{}{224}\DUspan{}{–}\DUspan{}{274}}}}\DUspan{}{, }\DUspan{}{2001}\DUspan{}{.}}}}
-\end{thebibliography}
-
-\renewcommand{\indexname}{Index}
-\printindex
-\end{document}
+%
+%Papers~\cite{ferraiolo:rbac,alam06:constraint,tzelepi01:flexible,haibo05:context,cholewka00:acontext-sensitive,huang06:pervasive, motta03:contextual}
+%
+%Papers~\cite{bao08:role, yamazaki04:designing, han08:extended, hasebe10:capability, mossakowski03:temporal, ren10:machine, zhang06:collaborative, aich07:STARBAC}
+%
+%Papers~\cite{kumar06:strbac,samuel07:spatio-temporal,ray07:spatio,chandran05:llt,aich09:role,yao08:task,oh03:task,zhou07:network} Papers~\cite{oh00:task,bertino01:trbac,li08:fine}
+%%Papers~\cite{oh00:task}
+%%Papers~\cite{bertino01:trbac}
+%%Papers~\cite{li08:fine}
+%Papers~\cite{joshi05:generalized, chen08:spatio-temporal}

### 2013-02-03 old/process.tex
Moves original latex documents into old/ directory and adds a latex build directory for current documentation state.

--- 2012-09-27 process.tex
@@
 
+For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines is:
+
+\begin{itemize}
+  \item Google Scholar - ttp://scholar.google.com
+  \item IEEExplore - http://ieeexplore.ieee.org
+  \item ACM Portal - http://dl.acm.org
+\end{itemize}
+
+For each of the criteria below, a search was performed on each of the search engines for a total of 9 data sets.  The criteria used were:
+
+\begin{itemize}
+  \item role based access control
+  \item RBAC
+  \item role-based access control
+\end{itemize}
+
+The search performed was done in an automated way using a set of scripts to query and collect data from each search engine.
+For each criteria for each search engine, the results were captured until a stopping criteria was met.
+For a given run was done as follows:
+
+\begin{enumerate}
+  \item Remote call to search engine with current search start position and the current search criteria.
+  \item Parse results and extract paper title, authors and year of publication.
+  \item Compare results against stopping criteria.
+    \begin{itemize}
+      \item If stopping criteria met, stop search.
+      \item If stopping criteria not met, increase search position by number of results and return to step 1.
+    \end{itemize}
+\end{enumerate}
+
+The stopping criteria used was either after the first 1000 results, a limitation imposed by some of the search engines, or if ten consecutive results did not contain any of the search criteria words within the title.
+After gathering all 9 data sets, the data was combined into a master list by systematically comparing the bibliographic information for each.
+After producing a master list a series of assessment rounds were performed to narrow the paper list and identify primary sources.
+
+There were a total of three elimination rounds.
+The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.
+Each round was performed as follows:
+
+\begin{itemize}
+  \item Each reviewer independently classified papers as relevant, irrelevant or uncertain.  
+  \item Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.  
+  \item Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
+From this discussion, papers were either thrown out or kept until the next round of the review.
+\end{itemize}
+
+The results of the searches was as follows:
+
+\begin{center}
+  \begin{tabular}{ l | c | c | c }
+    \hline
+     & RBAC & role based access control & role-based access control \\ \hline \hline
+    Google Scholar & 261 & 581 & 391 \\ \hline
+    ACM Portal & 321 & 281 & 261 \\ \hline
+    IEEExplore & 171 & 221 & 201 \\ \hline
+  \end{tabular}
+\end{center}
+
 We collect papers on extended role-based access control~\cite{ferraiolo:rbac}.
-We select 27 papers out of XXX papers.
 

### 2013-02-03 old/related.tex
Moves original latex documents into old/ directory and adds a latex build directory for current documentation state.

\section{Related Work}
\label{sec:related}

Various techniques have been proposed on regression testing of software programs~\cite{Rothermel:1996:ART:235681.235682, Graves:2001:ESR:367008.367020}.
These techniques aim to select test cases that could reveal different behaviors after modification
in programs.
%These techniques are related to regression-test selection \cite{Rothermel:1996:ART:235681.235682}, 
%\cite{Graves:2001:ESR:367008.367020}, and test-suite prioritization \cite{Elbaum:2000:PTC:347324.348910}.
Note that these techniques focus on changes at code level.
None of these techniques consider potential changes that can arise from code-related components (such as security policies specified separately).
Polices and general programs are fundamentally different in terms of structures, semantics, and
functionalities, etc. Therefore, techniques for regression
testing of programs are not suitable for addressing 
the test-selection problem for policy evolution.
%Our work is the first automatic test-selection approach for policy
%evolution.
\Comment{
Prior work that is closest to ours is Mouelhi et al.'s
work~\cite{mouelhi09:tranforming}.
They proposed a technique to transform functional test cases into security test cases. 
These test cases then be used to verify correctness of access control mechanisms in policy-based systems.
They defined various mutation operators at the policy level.
Given a policy and its mutated policy, the technique selects only impacted test cases to be transformed as security test cases
(with additional manual task to add code for checking security).
While this work focuses on security test case selection from initial test cases,
our work focuses on test selection problem impacted by policy changes
to reveal faults induced by policy changes.
}

Another approach closest to ours is Fisler et al.'s
approach~\cite{fisler05:verification}.
They developed a tool called Margrave that enables conducting change-impact analysis between two XACML
policies. We could use Margrave to identify semantic policy changes
between two policies.
However, Margrave does not support test selection as our work does.
% However, Margrave supports for only limited functionality of XACML. Moreover,

%where system behaviors are changed by policy evolution as
%our work does.

%does not supports for functionality of XACML sufficiently
%and  
%
%limited functionality of XACML
%policies
%
%Margrave does not handle with situations where
%program behaviors are changed by policy changes as our work focuses on.
%Moreover, Margrave supports limited functions of XACML, and

%in terms of regression testing. As application
%code can often interact with policies, our work is useful

\Comment{
To facilitate verifying and validating the correctness of policies, prior
research work has been done in the area of policy testing, which generates and executing test requests against a policy.
Martin et al.~\cite{martin06:defining} proposed a framework to detect policies faults by analyzing
policy test suites that involve 
requests-responses pairs. The framework is based on mutation operators~\cite{martin07:fault} that enable measuring fault detection capability. 
It uses a tool~\cite {martin07:automated} that aims at minimizing the test cases to be generated by analyzing the structural coverage of the policies.
Hu et al.~\cite{hu07:conformance} proposed a generic model-based conformance checking technique for access control policies written in
XACML.
These approaches focus on test request generation at a policy level (e.g., policies written in XACML).
Instead of generating new requests, our work focuses on regression testing at the implementation level (the system level) by
reusing existing hardcoded test cases.
}
%, which includes
%application code when a policy evolves.

%Margrave detects the presence of requests that may violate the specified properties.

%used these mutated policies to select impacted tests. The impacted tests are then tranformed to security tests (by added specific code to check the security mechanism). 
%Although this work focused on applications interacting with policies, it was not presented from a regression testing perspective.
%
%
%The most relevant paper that is related to our work was presented by Mouelhi et al. in~\cite{mouelhi09:tranforming}. The authors proposed transforming functional tests into security tests. 
%These tests can then be used to validate the access control mechanisms in policy-based systems. They defined mutation operators at the policy level and used these mutated policies to select impacted tests. The impacted tests are then tranformed to security tests (by added specific code to check the security mechanism). 
%Although this work focused on applications interacting with policies, it was not presented from a regression testing perspective.

%fault
%localization problem of ?rewall policies.
%
%This paper focuses on regression testing for policy-based software systems.

%Zhang et al \cite{DBLP:conf/isw/ZhangRG05} used model checking technique which is supported by a tool to evaluate access
%control policies specified in a modelling formalism called RW. Hughes et al. \cite{hughes04:automated} proposed using First-Order Logic 
%to automatically verify access control policies. Their model is based on specifying access to resources, they show that the access
%control policies in XACML can be translated into a normlized form that includes four classes: permit,
%deny, error, and not-applicable. A partial ordering among access control polices enable the authors to specify properties that are checked later
%with Alloy analyszer \cite{Alloy}. Nyanchama et. al. \cite{Nyanchama99therole} focused on policy analysis and verification related to 
%RBAC models. They used graphs to model hierarchies of users, roles and permissions and to solve role-role conflicts issues.
%While these different approaches focus on test request generation for validating the policy implementation (the XACML code), 
%the technique presented in this paper focuses on regression testing at the implementation level (the system level) which occurs when the access control policy evolves.

\Comment{
Our previous work~\cite{hu10:model,hu08:property} presented
model checking to verify various policies such as role-based
access control policies~\cite{ferraiolo92:role}.
Our previous work translates policies into specifications
in NuSMV~\cite{cimatti02:nusmv2}. NuSMV verifies the specifications against properties. 
Based on our previous work, we developed a tool, called ACPT\footnote{http://csrc.nist.gov/groups/SNS/acpt/index.html},
to verify and test access control policies.
However, our previous work does not detect conflicts
or redundancies between two policies, one major
focus of our current work. Moreover, our previous
work does not recommend any combinations (to the policy
authors) that can be recommended by our current work.

There exist tools to verify XACML policies. These tools require XACML policies in formal specification languages such as 
Alloy~\cite{jackson01:micromodularity}. Hughes et al.~\cite{hughes04:automated} proposed an approach to translate XACML policies to a model 
specified in the Alloy language~\cite{jackson01:micromodularity}. Then, the Alloy Analyzer verifies whether the model satisfies the given properties.
Kolovski et al.~\cite{kolovski07:analyzing} converted XACML policies into specifications in  description logics (DL) and adopted existing DL 
verification tools to verify the specifications. Fisler et al.~\cite{fisler05:verification} developed a Binding Decision Diagram (BDD) based 
verification tool, called Margrave, to verify XACML policies. Margrave parses XACML policies into the BDD structure. Moreover, Margrave supports 
change-impact analysis to report which decisions are changed after policy modification.
Similar to these previous approaches, our approach focuses on modeling and verifying policies via a symbolic model checking. Moreover, our approach
 verifies whether two policies may include conflicts and redundancies via a symbolic model checking. 
While these previous approaches do not help combining two policies, our approach recommends combinations of two policies that could satisfy 
user-specified properties.

%\textbf{Policy Combinations/Integration.}
%Bonatti et al.~\cite{Bonatti2} proposed combining sub-policies based on 2-valued algebra. 
Backes et al.~\cite{backes04algebra} proposed a 3-valued algebra to combine sub-policies considering privacy policies.
 They defined relations of combination with three operators: conjunction, disjunction, and scoping.
 Li at el.~\cite{Ninghui2009} proposed an algebra for combining sub-policies. 
They introduced new combining algorithms such as weak majority or strong majority.
For example, in the case of the strong majority combining algorithms, at least 2/3 
rules evaluated against a request should be evaluated to the same decision. Since the decision is the majority, 
the decision is returned as a final decision. They proposed algebras to generalize their combining algorithms.
%Mazzoleni et al.~\cite{Mazzoleni2006} proposed policy integration that can be applied to XACML policies. 
In their approach, policy authors can specify conditions to represent how their policies can be integrated with other policies. 
Their integration approach is based on analysis of two policies to be integrated. Policy authors measure similarity of two policies 
and integrate the policies based on given condition.

}
\Comment{
%\textbf{Policy Change Impact Analysis}
%\textbf{Policy Conflict and Redundancy}
%\textbf{Verification in NuSMV for other projects}

%interaction, but runtime ...more general stuff to do it.

\textbf{Policy Verification.}
There are verification tools available for XACML
policies. Hughes et
al.~\cite{hughes04:automated}
proposed an approach
to verify XACML policies by
translating the policies to the Alloy
language~\cite{jackson01:micromodularity}.
Then, they use Alloy Analyze to verify properties
against the translated policies. 
Zhang et
al.~\cite{zhang05:evaluating} developed a model-checking tool
to verify policies specified in
\Intro{RW} languages~\cite{zhang04:synthesis}, which can be converted to
XACML.
Kolovski et
al.~\cite{kolovski07:analyzing} formalize XACML policies with
description logics (DL), which are a decidable fragment of
first-order logic, and exploit existing DL verifiers to conduct
policy verification.
Schaad and Moffett~\cite{schaad02:lightweight} leverage
Alloy~\cite{jackson01:micromodularity} to check that role-based
access-control policies do not allow roles to be assigned to users
in ways that violate SoD constraints.
Similar to previous research work, our work also model and verify policies
via model checking. However, we extend
our work to detect conflict and redundancy via a symbolic checker.

\textbf{Policy Combinations/Integration.}
Bonatti et al.~\cite{Bonatti2} proposed combining sub-policies based on 2-valued algebra. Backes et al.~\cite{backes04algebra} proposed a 3-valued algebra to combine sub-policies considering privacy policies. They defined relations of combination with three operators: conjunction, disjunction, and scoping. Li at el.~\cite{Ninghui2009} proposed an algebra for combining sub-policies. They also introduce new combining algorithms such as weak majority or strong majority. For example, in case of the strong majority combining algorithms, at least 2/3 rules evaluated against a request should return the same decision. As the decision is majority, the decision is returned as a final decision. They propose algebras to generalize their combining algorithms. Moreover, they propose combinations of sub-policies, which may return policy evaluation errors and obligations.
Mazzoleni et al.~\cite{Mazzoleni2006} proposed policy integration that can be applied to XACML policies. In their approach, policy authors can specify conditions to represent how their policies can be integrated with other policies. Their integration approach is based on analysis of two policies to be integrated. Policy authors measure similarity of two policies and integrate the policies based on given condition.
We also propose that a methodology to combine two policies using property verification.
We next describe a methodology to recommend the combinations (of two policies) that could satisfy user-specified properties.

\textbf{Policy Testing.}

Our previous work~\cite{hu10:model} demonstrated an approach
to represent a policy and its properties as
corresponding finite state machine (FSM) model and temporal
logics (e.g., computation tree logic), respectively, using SMV specification language~\cite{cimatti02:nusmv2}.

Our previous work developed policy testing approaches for policy
structural coverage~\cite{martin06:defining}, request
generation~\cite{martin07:automated}, and mutation
testing~\cite{martin07:fault}.
% Our previous
%work~\cite{hu07:conformance} also proposed a generic model-based
%conformance checking approach for access control policies written in
%XACML.
These pieces of work do not rely on properties for generating
test requests to detect a fault in a policy. 
Our previous work~\cite{martin08:assessing} developed an approach
for measuring the quality of policy properties in policy
verification. Given user user-specified
properties, the quality of properties are measured
based on fault-detection capability.
}

### 2013-02-03 paper/rbac-slr.tex
Re-factor of location of latex elements.  Now main paper is stored under paper and within that each major section of the paper is stored under sections directory.

\documentclass[preprint,12pt]{elsarticle}

\usepackage{amssymb}

\journal{TBD}

%% DOCUMENT %%
\begin{document}

%% 
\begin{frontmatter}

\title{A Systematic Literature Review on Role-Based Access Control}

\author{}

\address{}

\input{sections/abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
\end{keyword}

\end{frontmatter}

%% SECTIONS %%
\input{sections/intro}
\input{sections/process}

%% REFERENCES %%
\bibliographystyle{elsarticle-num}
\bibliography{sections/rbacslr}

\end{document}

### 2013-02-03 paper/sections/abstract.tex
Re-factor of location of latex elements.  Now main paper is stored under paper and within that each major section of the paper is stored under sections directory.

\begin{abstract}

Role-based access control (RBAC) has become an increasingly popular access
control for various applications such as web applications and
database applications. RBAC restricts access to resources based on identity of subjects and/or groups
called roles. 
Since RBAC is introduced in 1998, researchers have proposed various extended models of RBAC.
For example, they define additional constraints among roles (e.g., temporal constraints or location constraints) or hierarchy relationship of roles.
Our goal of this work is to study extended models of RBAC and analyze their extended features and claimed research contributions
to find limitations of current RBAC models and what extent of extended features that can be used for future RBAC.
We conduct a systematic literature review by collecting and synthesizing relevant research
papers. We initially collect XXXX papers from various sources such as IEEE and ACM websites and selected 26 papers systemically.
We perform a comparative analysis to find relationships among extended models and RBAC.

\end{abstract}

### 2013-02-03 paper/sections/intro.tex
Re-factor of location of latex elements.  Now main paper is stored under paper and within that each major section of the paper is stored under sections directory.
Updates to latest intro and process documentation.

--- 2013-02-03 old/intro.tex
@@
 
-Role-based access control (RBAC) models~\cite{ferraiolo:rbac} become popularly used
-to govern access to critical resources.
-In an RBAC model, roles represent a group of users who are
-involved in a specific job function in an organization. 
-RBAC assigns permissions of specific actions on resources to roles
-instead of individual users.
+Why is the base model of RBAC extended by newer models?
+
+Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.
+In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
+RBAC assigns permissions of specific actions on resources to roles instead of individual users.
 Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.
 
-RBAC is a generalized access control approach used for various applications including
-web services, database applications, and healthcare applications.
+RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.
 RBAC has advantages in maintaining and managing organization's security policies.
-For example, if a user is to access manager role's resources within a given
-organization, security policy administrators simply add the user to be associated with the manager role.
+For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.
 
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model~\cite{ferraiolo:rbac}.
+RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
 Standard RBAC model considers only role-user association and role hierarchy.
@@
 As a result, researchers could take time on reinventing the wheel.
-%But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
+But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
 
@@
 We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
+\begin{description}
+\item[{Our research give benefits to a community as follows:}]
 
-Our research give benefits to a community as follows:
 \begin{itemize}
-	\item Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.	
-	\item Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing
-	research results helps decide a standard of extended RBAC.
-	\item Our work shows a criteria in comparison among research results.
-	\item Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.			
+\item {} 
+Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
+\item {} 
+Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
+\item {} 
+Our work shows a criteria in comparison among research results.
+\item {} 
+Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
 \end{itemize}
 
-%Agent-Based\cite{yamazaki04:designing}
-%
-%
-%
-%Task-based
-%Temporal-role based
-%Inheritance
-%Location
-%Delegation
-
-%\cite[ferraiolorbac]
-
-%Introduction come here.
-%Various techniques have been proposed on regression testing of software programs~\cite{Rothermel:1996:ART:235681.235682, Graves:2001:ESR:367008.367020}.
-
-%
-%Papers~\cite{ferraiolo:rbac,alam06:constraint,tzelepi01:flexible,haibo05:context,cholewka00:acontext-sensitive,huang06:pervasive, motta03:contextual}
-%
-%Papers~\cite{bao08:role, yamazaki04:designing, han08:extended, hasebe10:capability, mossakowski03:temporal, ren10:machine, zhang06:collaborative, aich07:STARBAC}
-%
-%Papers~\cite{kumar06:strbac,samuel07:spatio-temporal,ray07:spatio,chandran05:llt,aich09:role,yao08:task,oh03:task,zhou07:network} Papers~\cite{oh00:task,bertino01:trbac,li08:fine}
-%%Papers~\cite{oh00:task}
-%%Papers~\cite{bertino01:trbac}
-%%Papers~\cite{li08:fine}
-%Papers~\cite{joshi05:generalized, chen08:spatio-temporal}
+\end{description}

### 2013-02-03 paper/sections/process.tex
Updates to latest intro and process documentation.

--- 2013-02-03 latex/RBAC-SLR.tex
@@
-% Generated by Sphinx.
-\def\sphinxdocclass{report}
-\documentclass[letterpaper,10pt,english]{sphinxmanual}
-\usepackage[utf8]{inputenc}
-\DeclareUnicodeCharacter{00A0}{\nobreakspace}
-\usepackage[T1]{fontenc}
-\usepackage{babel}
-\usepackage{times}
-\usepackage[Bjarne]{fncychap}
-\usepackage{longtable}
-\usepackage{sphinx}
-\usepackage{multirow}
+\section{Methodology and Process} \label{sec:process}
 
-\title{RBAC-SLR Documentation}
-\date{February 02, 2013}
-\release{0.1}
-\author{Eric D. Helms}
-\newcommand{\sphinxlogo}{}
-\renewcommand{\releasename}{Release}
-\makeindex
-
-\makeatletter
-\def\PYG@reset{\let\PYG@it=\relax \let\PYG@bf=\relax%
-    \let\PYG@ul=\relax \let\PYG@tc=\relax%
-    \let\PYG@bc=\relax \let\PYG@ff=\relax}
-\def\PYG@tok#1{\csname PYG@tok@#1\endcsname}
-\def\PYG@toks#1+{\ifx\relax#1\empty\else%
-    \PYG@tok{#1}\expandafter\PYG@toks\fi}
-\def\PYG@do#1{\PYG@bc{\PYG@tc{\PYG@ul{%
-    \PYG@it{\PYG@bf{\PYG@ff{#1}}}}}}}
-\def\PYG#1#2{\PYG@reset\PYG@toks#1+\relax+\PYG@do{#2}}
-
-\expandafter\def\csname PYG@tok@gd\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@gu\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
-\expandafter\def\csname PYG@tok@gt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.25,0.82}{##1}}}
-\expandafter\def\csname PYG@tok@gs\endcsname{\let\PYG@bf=\textbf}
-\expandafter\def\csname PYG@tok@gr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@cm\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@vg\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@m\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@mh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@cs\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\colorbox[rgb]{1.00,0.94,0.94}{\strut ##1}}}
-\expandafter\def\csname PYG@tok@ge\endcsname{\let\PYG@it=\textit}
-\expandafter\def\csname PYG@tok@vc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@il\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@go\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.19,0.19,0.19}{##1}}}
-\expandafter\def\csname PYG@tok@cp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@gi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@gh\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
-\expandafter\def\csname PYG@tok@ni\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.84,0.33,0.22}{##1}}}
-\expandafter\def\csname PYG@tok@nl\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.13,0.44}{##1}}}
-\expandafter\def\csname PYG@tok@nn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
-\expandafter\def\csname PYG@tok@no\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.38,0.68,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@na\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@nb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@nc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.05,0.52,0.71}{##1}}}
-\expandafter\def\csname PYG@tok@nd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.33,0.33,0.33}{##1}}}
-\expandafter\def\csname PYG@tok@ne\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@nf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.49}{##1}}}
-\expandafter\def\csname PYG@tok@si\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.44,0.63,0.82}{##1}}}
-\expandafter\def\csname PYG@tok@s2\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@vi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@nt\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.02,0.16,0.45}{##1}}}
-\expandafter\def\csname PYG@tok@nv\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.38,0.84}{##1}}}
-\expandafter\def\csname PYG@tok@s1\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@gp\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
-\expandafter\def\csname PYG@tok@sh\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@ow\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@sx\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.78,0.36,0.04}{##1}}}
-\expandafter\def\csname PYG@tok@bp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@c1\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@kc\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@c\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.50,0.56}{##1}}}
-\expandafter\def\csname PYG@tok@mf\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@err\endcsname{\def\PYG@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
-\expandafter\def\csname PYG@tok@kd\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@ss\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.32,0.47,0.09}{##1}}}
-\expandafter\def\csname PYG@tok@sr\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.14,0.33,0.53}{##1}}}
-\expandafter\def\csname PYG@tok@mo\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@mi\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.13,0.50,0.31}{##1}}}
-\expandafter\def\csname PYG@tok@kn\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@o\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
-\expandafter\def\csname PYG@tok@kr\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@s\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@kp\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@w\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
-\expandafter\def\csname PYG@tok@kt\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.56,0.13,0.00}{##1}}}
-\expandafter\def\csname PYG@tok@sc\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@sb\endcsname{\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@k\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.00,0.44,0.13}{##1}}}
-\expandafter\def\csname PYG@tok@se\endcsname{\let\PYG@bf=\textbf\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-\expandafter\def\csname PYG@tok@sd\endcsname{\let\PYG@it=\textit\def\PYG@tc##1{\textcolor[rgb]{0.25,0.44,0.63}{##1}}}
-
-\def\PYGZbs{\char`\\}
-\def\PYGZus{\char`\_}
-\def\PYGZob{\char`\{}
-\def\PYGZcb{\char`\}}
-\def\PYGZca{\char`\^}
-\def\PYGZam{\char`\&}
-\def\PYGZlt{\char`\<}
-\def\PYGZgt{\char`\>}
-\def\PYGZsh{\char`\#}
-\def\PYGZpc{\char`\%}
-\def\PYGZdl{\char`\$}
-\def\PYGZti{\char`\~}
-% for compatibility with earlier versions
-\def\PYGZat{@}
-\def\PYGZlb{[}
-\def\PYGZrb{]}
-\makeatother
-
-\begin{document}
-
-\maketitle
-\tableofcontents
-\phantomsection\label{index::doc}
-
-Contents:
-
-\chapter{1. Introduction}
-\label{intro:introduction}\label{intro:systematic-literature-review-of-rbac-analysis-notes}\label{intro::doc}
-Why is the base model of RBAC extended by newer models?
-
-Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.
-In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
-RBAC assigns permissions of specific actions on resources to roles instead of individual users.
-Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.
-
-RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.
-RBAC has advantages in maintaining and managing organization's security policies.
-For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.
-
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
-Standard RBAC model considers only role-user association and role hierarchy.
-Since standard RBAC model has limitations such as specifying environmental constraints or context information
-Researchers developed extended models of RBAC to overcome the limitations.
-However, as researchers often develop their own specialized extended models of RBAC,
-their research cannot be generalized or compared with other research work appropriately.
-As a result, researchers could take time on reinventing the wheel.
-But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
-
-The goal of this work is to synthesize available research results on extended models of RBAC.
-We analyze their extended features and claimed research contributions to find limitations of current RBAC models and what extent of extended features by
-comparing with similar research work.
-We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
-\begin{description}
-\item[{Our research give benefits to a community as follows:}] \leavevmode\begin{itemize}
-\item {} 
-Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
-
-\item {} 
-Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
-
-\item {} 
-Our work shows a criteria in comparison among research results.
-
-\item {} 
-Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
-
-\end{itemize}
-
-\end{description}
-
-\chapter{2. Methodology and Process}
-\label{process:methodology-and-process}\label{process::doc}
 For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
@@
 Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}
-
 \item {} 
 IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}
-
 \item {} 
 ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}
-
 \item {} 
 CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}
-
 \end{itemize}
@@
 role based access control
-
 \item {} 
 RBAC
-
 \item {} 
 role-based access control
-
 \end{itemize}
@@
 Remote call to search engine with current search start position and the current search criteria.
-
 \item {} 
 Parse results and extract paper title, authors and year of publication.
-
 \item {} 
 Compare results against stopping criteria.
+\end{enumerate}
 
-\end{enumerate}
 \begin{itemize}
@@
 If stopping criteria met, stop search.
-
 \item {} 
 If stopping criteria not met, increase search position by number of results and return to step 1.
-
 \end{itemize}
@@
 Each round was performed as follows:
+
 \begin{itemize}
@@
 Each reviewer independently classified papers as relevant, irrelevant or uncertain.
-
 \item {} 
 Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
-
 \item {} 
@@
 Ties were broken by an indepedent party.
-
 \end{itemize}
@@
 
-Google Scholar
- & 
-651
- & 
-213
- & 
-435
- & 
-1299
+Google Scholar & 651 & 213 & 435 & 1299
 \\\hline
-
-ACM Portal
- & 
-500
- & 
-20
- & 
-720
- & 
-1240
+ACM Portal & 500 & 20 & 720 & 1240
 \\\hline
-
-IEEExplore
- & 
-200
- & 
-40
- & 
-230
- & 
-470
+IEEExplore & 200 & 40 & 230 & 470
 \\\hline
-
-CiteSeerX
- & 
-100
- & 
-100
- & 
-150
- & 
-350
+CiteSeerX & 100 & 100 & 150 & 350
 \\\hline
- &  &  &  & \\\hline
-
-Totals
- & 
-1451
- & 
-373
- & 
-1535
- & 
-3359
+ &  &  &  & 
+\\\hline
+Totals & 1451 & 373 & 1535 & 3359
 \\\hline
@@
 \end{tabulary}
-
-\section{Research Questions}
-\label{process:research-questions}
-Based on our intial intent for the systematic literature review, and notes taken during the third phase of elimination, we have identified the following research questions:
-\begin{itemize}
-\item {} 
-RQ1. What are the deficiencies in current RBAC model to propose extended RBAC models?
-
-\item {} 
-RQ2. What are the motivations behind RBAC extensions?
-
-\item {} 
-RQ3. What are the categorizations of RBAC model proposed?
-
-\item {} 
-RQ4. For each categorization, what are the extended features of RBAC model proposed?
-
-\item {} 
-RQ5. Do these models have corresponding implementations in practice?
-
-\item {} 
-RQ6. How are extended RBAC models evaluated in theory and in practice?
-
-\item {} 
-RQ7. Are there any commonalties or generalizations across all categorizations?
-
-\item {} 
-RQ8. What domains or scenarios serve as inspiration for these extensions?
-
-\end{itemize}
-
-\chapter{RQ1: What are the deficiencies in current RBAC model?}
-\label{research_questions/RQ_1:rq1-what-are-the-deficiencies-in-current-rbac-model}\label{research_questions/RQ_1::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_1:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_1:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-no support for partial inheritance
-
-\item {} 
-inability to support permission assignment constraints
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Missing ability to apply constraints between role and permissions and enforce
-
-\item {} 
-Missing context that defines if a role can be activated at execution time
-
-\item {} 
-Role activation on a per object or process execution rather than session based
-
-\item {} 
-No allowance for context to check permission at execution moment based on context
-
-\item {} 
-activation of roles based on current context rather than session intiation
-
-\item {} 
-conflict resolution
-
-\item {} 
-RBAC cannot handle dynamic authorization and entities
-
-\item {} 
-Role cannot be resource dependent
-
-\item {} 
-RBAC is too simple
-
-\item {} 
-Missing granularity
-
-\item {} 
-Inability to handle dynamic nature of multiple projects with multiple resources
-
-\item {} 
-Too simple
-
-\item {} 
-Lacks context
-
-\item {} 
-Handling authorization between applications and within can be different
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-RBAC cannot handle cross-organizational roles and access in an administratively easy way
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Privacy regulations are complex, with many entities, RBAC lacks ability to handle all of these
-
-\item {} 
-Context, conflict detection and resolution
-
-\item {} 
-RBAC lacks ability to define purpose
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-No support for multiple security domains
-
-\item {} 
-No support for same security, different applications
-
-\item {} 
-Lack of flexibility
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Lack of contextual information
-
-\item {} 
-Role needs a defined schema
-
-\item {} 
-Role activation based on context
-
-\item {} 
-Giving a permission an operation
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Needs an entity to represent a task and connect a role and a permission through it
-
-\item {} 
-Need for team and tasks
-
-\item {} 
-Need for entity between permission and roles to hold constraint and allow for dynamic situations and entities
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-RBAC can't handle time based constraints for role enablement
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ2: What are the motivations behind RBAC extensions?}
-\label{research_questions/RQ_2:rq2-what-are-the-motivations-behind-rbac-extensions}\label{research_questions/RQ_2::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_2:common-themes}\begin{itemize}
-\item {} 
-Roles, users and permissions need context associations
-
-\item {} 
-Location and time constraints as well as contexts needed
-
-\item {} 
-RBAC lacks ability to link in objects
-
-\item {} 
-RBAC does not handle single user base and multiple organizations/applications
-
-\item {} 
-Role enablement based on context
-
-\item {} 
-User granting/revoking of roles based on context
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_2:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No support for partial inheritance
-
-\item {} 
-Permission assignment constraints
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Relationship between object and viewer can't be represented
-
-\item {} 
-Object content matters
-
-\item {} 
-Location and time constraints on the user
-
-\item {} 
-Departmental context around access
-
-\item {} 
-RBAC does not handle many small services to wrap around them
-
-\item {} 
-RBAC lacks security principle enforcement
-
-\item {} 
-Lack of specifying context, and permission checks of dynamic evaluation time rules
-
-\item {} 
-RBAC is static and lacks ability to handle dynamic authorization and definitions
-
-\item {} 
-RBAC ignores context
-
-\item {} 
-Role granulatiry not well controlled
-
-\item {} 
-Roles should depend on the resources
-
-\item {} 
-RBAC cannot handle multiple projects with multiple tasks and differing roles in each system but the same user
-
-\item {} 
-Lack of context support
-
-\item {} 
-Too simple for dynamic environments
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-Support cross-domain delegation
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Reduce administrative complexity of RBAC for multi-organizational
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Lack of components, purpose binding, conditions, obligations to handle privacy needs
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-Need to support multiple securtiy domains
-
-\item {} 
-Need to support same security with different applications
-
-\item {} 
-Role heirarchy adds complexity and reduces flexibility
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Lack of contextual information with roles, users and permissions
-
-\item {} 
-Role needs a schema to allow specification of attributes
-
-\item {} 
-Location based security scenarios
-
-\item {} 
-Role activation based on location
-
-\item {} 
-Support physical and logical locations
-
-\item {} 
-Attaching operations to permissions
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-RBAC doesn't inherently support location and time constraints
-
-\item {} 
-Spatial and temporal awareness of roles, inheritance and permissions
-
-\item {} 
-Context mechanisms needed
-
-\item {} 
-Physical and logical locations, hybrids
-
-\item {} 
-Role assignment based off time and space, as well as permissions
-
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-RBAC can't handle large enterprises with multiple applications and single user base
-
-\item {} 
-Security requirements for enterprises (look up these)
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Workflow based organizations need time base constraint for role enablement
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ3: What are the categorizations of proposed RBAC extensions?}
-\label{research_questions/RQ_3:rq3-what-are-the-categorizations-of-proposed-rbac-extensions}\label{research_questions/RQ_3::doc}
-Across the papers found, a number of higher level categories emerged that each proposed model fell in to.
-Some of these categories had singular primary papers, while others had multiple extension models proposed that were deemed independent of one another.
-
-For each category, the breakdown of the papers for each category is:
-
-\begin{tabulary}{\linewidth}{|L|L|}
-\hline
-\textbf{
-Paper
-} & \textbf{
-Category
-}\\\hline
-
-Alam, M. and Hafner, M. and Breu, R. 2006
- & 
-Constraint
-\\\hline
-
-Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001
- & 
-Context
-\\\hline
-
-Haibo, SHEN and Fan, HONG 2005
- & 
-Context
-\\\hline
-
-Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000
- & 
-Context
-\\\hline
-
-Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006
- & 
-Context
-\\\hline
-
-Motta, G.H.M.B. and Furuie, S.S. 2003
- & 
-Context
-\\\hline
-
-Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008
- & 
-Context
-\\\hline
-
-Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004
- & 
-Context
-\\\hline
-
-Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008
- & 
-Context
-\\\hline
-
-Thein, N. and others 2011
- & 
-Context
-\\\hline
-
-Zou, D. and He, L. and Jin, H. and Chen, X. 2009
- & 
-Context
-\\\hline
-
-Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010
- & 
-Delegation
-\\\hline
-
-Zhang, Z. and Zhang, X. and Sandhu, R. 2006
- & 
-Organizational
-\\\hline
-
-Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007
- & 
-Privacy
-\\\hline
-
-Masoumzadeh, A. and Joshi, J. 2008
- & 
-Privacy
-\\\hline
-
-Zhao, Y. and Zhao, Y. and Lu, H. 2008
- & 
-Resource
-\\\hline
-
-Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005
- & 
-Spatial
-\\\hline
-
-Ray, I. and Kumar, M. and Yu, L. 2006
- & 
-Spatial
-\\\hline
-
-Hansen, F. and Oleshchuk, V. 2003
- & 
-Spatial
-\\\hline
-
-Aich, S. and Sural, S. and Majumdar, A. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Chen, L. and Crampton, J. 2008
- & 
-Spatio-Temporal
-\\\hline
-
-Samuel, A. and Ghafoor, A. and Bertino, E. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Chandran, S. and Joshi, J. 2005
- & 
-Spatio-Temporal
-\\\hline
-
-Ray, I. and Toahchoodee, M. 2007
- & 
-Spatio-Temporal
-\\\hline
-
-Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009
- & 
-Spatio-Temporal
-\\\hline
-
-Yao, L. and Kong, X. and Xu, Z. 2008
- & 
-Task
-\\\hline
-
-ZHANG, S. and CHEN, X. and HOU, G. 2009
- & 
-Task
-\\\hline
-
-Oh, S. and Park, S. 2003
- & 
-Task
-\\\hline
-
-Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005
- & 
-Temporal
-\\\hline
-\end{tabulary}
-
-\section{By Category}
-\label{research_questions/RQ_3:by-category}\begin{description}
-\item[{Constraint (1)}] \leavevmode
-a limitation or restriction
-
-\item[{Context (8)}] \leavevmode
-circumstances in which an event occurs
-
-\item[{Delegation (1)}] \leavevmode
-the act of empowering to act for another
-
-\item[{Organizational (1)}] \leavevmode
-of or relating to an organization
-
-\item[{Privacy (1)}] \leavevmode
-freedom from unauthorized intrusion
-
-\item[{Resource (1)}] \leavevmode
-an available supply than be drawn on when needed
-
-\item[{Spatial (3)}] \leavevmode
-of or relating to facility in percieving relations in space
-
-\item[{Spatio/Temporal (6)}] \leavevmode
-combining space and time
-
-\item[{Task (3)}] \leavevmode
-a usually assigned piece of work often to be finished within a certain time
-
-\item[{Temporal (1)}] \leavevmode
-of or relating to the sequence of time or to a particular time
-
-\end{description}
-
-{\hfill\scalebox{1.000000}{\includegraphics{categories_map.png}}\hfill}
-
-\chapter{RQ4: For each categorization, what are the extended features of RBAC proposed?}
-\label{research_questions/RQ_4::doc}\label{research_questions/RQ_4:rq4-for-each-categorization-what-are-the-extended-features-of-rbac-proposed}
-Is this the same as RQ\_1 ?
-
-\section{Common Themes}
-\label{research_questions/RQ_4:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_4:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ5. Do these models have corresponding implementations in practice?}
-\label{research_questions/RQ_5::doc}\label{research_questions/RQ_5:rq5-do-these-models-have-corresponding-implementations-in-practice}
-When designing and proposing a model targeted at a feature that is rooted in practical
-usage by real software systems, bringing the model to life is strong evidence that the
-proposed model can work in practice.  The concept of authorization, and access control
-is rooted in a business need.  Thus, any access control model needs to be feasible
-in the real world not just on paper.  We analyzed the primary sources to see how many
-proposed models actually had implmenetations associated with them.  And quantified the
-type of implementation.  Whether the implementation was for a real system, for a prototype
-and/or used in a production environment.
-
-\section{Common Themes}
-\label{research_questions/RQ_5:common-themes}\begin{itemize}
-\item {} 
-21 No's
-
-\item {} 
-8 Yes's
-
-\item {} 
-Half the implementation's were prototypes
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_5:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-No - system architecture provided
-
-\item {} 
-No
-
-\item {} 
-Yes - prototype with VB6 and MS Access97
-
-\item {} 
-Yes - prototype system, system architecture, algorithms implemented, simulation scenarios
-
-\item {} 
-Yes - on an EHR system
-
-\item {} 
-Yes - prototype using XML container
-
-\item {} 
-Yes - web based API agent
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-Yes - integrated with telemedicine application, web based
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Yes - mentions of systems that exist using the model
-
-\item {} 
-Yes - prototype system
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ6. How are extended RBAC models evaluated in theory and in practice?}
-\label{research_questions/RQ_6:rq6-how-are-extended-rbac-models-evaluated-in-theory-and-in-practice}\label{research_questions/RQ_6::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_6:common-themes}\begin{itemize}
-\item {} 
-12 had no evaluation whatsoever
-
-\item {} 
-8 present example scenarios and how their model would apply
-
-\item {} 
-6 presented performance and time complexity analysis
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_6:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-No evaluation
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Present hospital scenarios and explanation of how new model would handle
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-Scenarios with how model affects, performance test of the model vs. traditional RBAC
-
-\item {} 
-Description of a fake project management system that the model is designed for
-
-\item {} 
-Applied to conceptual situation
-
-\item {} 
-No
-
-\item {} 
-Example scenario comparing their model vs other models
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Comparison caluclation for number of roles vs standard model, numerical size comparisons against other proposed models
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Evaluate the performance of permission assignment and authorization checks for various sizes and context variables
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-Model comparison from a qualitative standpoint
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Algorithm and Big-O performance evaluation
-
-\item {} 
-No except minor scenarios and how the model handles
-
-\item {} 
-No other than mathematical specification
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Mathematical description
-
-\item {} 
-Mathematical analysis of role graph, complexity comparisons across different models
-
-\item {} 
-Pseudo example scenario
-
-\item {} 
-No
-
-\item {} 
-Detailed scenario with examination of how model handles situation
-
-\item {} 
-Discuess difference btween their model and other models, time-complexity analysis to show no performance impact with standard model
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\item {} 
-No
-
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Mathematical modeling and graphs to detail usage and interactions, explanation of example scenarios
-
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ7. Are there any commonalties or generalizations across all categorizations?}
-\label{research_questions/RQ_7:rq7-are-there-any-commonalties-or-generalizations-across-all-categorizations}\label{research_questions/RQ_7::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_7:common-themes}\begin{itemize}
-\item {} 
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_7:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-\end{itemize}
-
-\end{description}
-
-\chapter{RQ8. What domains or scenarios serve as inspiration for these extensions?}
-\label{research_questions/RQ_8:rq8-what-domains-or-scenarios-serve-as-inspiration-for-these-extensions}\label{research_questions/RQ_8::doc}
-
-\section{Common Themes}
-\label{research_questions/RQ_8:common-themes}\begin{itemize}
-\item {} 
-Medical domain
-
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Mobile devices
-
-\item {} 
-Large-scale organizations with many sub-departments
-
-\item {} 
-Enterprise, organization workflows
-
-\end{itemize}
-
-\section{By Category}
-\label{research_questions/RQ_8:by-category}\begin{description}
-\item[{Constraint}] \leavevmode\begin{itemize}
-\item {} 
-Medical domain
-
-\item {} 
-Service Oriented Architecture
-
-\end{itemize}
-
-\item[{Context}] \leavevmode\begin{itemize}
-\item {} 
-Medical imaging database
-
-\item {} 
-Handling an insurance claim workflow
-
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Medical emergency overrides
-
-\item {} 
-Electronic health records
-
-\item {} 
-Supermarket
-
-\item {} 
-Project management system with large project, task and user base
-
-\item {} 
-Power load forecasting system (multi-level user environment)
-
-\end{itemize}
-
-\item[{Delegation}] \leavevmode\begin{itemize}
-\item {} 
-Large-scale electronic health record systems
-
-\end{itemize}
-
-\item[{Organizational}] \leavevmode\begin{itemize}
-\item {} 
-Online tutoring system
-
-\item {} 
-State education system with schools, districts, etc.
-
-\end{itemize}
-
-\item[{Privacy}] \leavevmode\begin{itemize}
-\item {} 
-Privacy within the medical domain, HIPAA
-
-\end{itemize}
-
-\item[{Resource}] \leavevmode\begin{itemize}
-\item {} 
-N/A
-
-\end{itemize}
-
-\item[{Spatial}] \leavevmode\begin{itemize}
-\item {} 
-Hospital with multi-floors, departments
-
-\end{itemize}
-
-\item[{Spatio/Temporal}] \leavevmode\begin{itemize}
-\item {} 
-Pervasive computing environments
-
-\item {} 
-Hospital using mobile devices with multiple user types on a single floor, departments
-
-\item {} 
-Mobile, peer-to-peer
-
-\item {} 
-Doctor out in the world needing access during a random medical emergency
-
-\item {} 
-Students at a university
-
-\end{itemize}
-
-\item[{Task}] \leavevmode\begin{itemize}
-\item {} 
-Large enterprises with multiple applications and a single user base
-
-\item {} 
-Hospital needs
-
-\item {} 
-Complex enterprise environments
-
-\end{itemize}
-
-\item[{Temporal}] \leavevmode\begin{itemize}
-\item {} 
-General organizational needs for time constraints
-
-\end{itemize}
-
-\end{description}
-
-\chapter{Constraint}
-\label{categories/constraint::doc}\label{categories/constraint:constraint}\begin{description}
-\item[{Definition}] \leavevmode
-a limitation or restriction
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1501451}{A constraint based role based access control in the SECTET a model-driven approach}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A constraint based role based access control in the SECTET a model-driven approach}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example?}] \leavevmode
-SECTECT is a framework that provides workflow for Service oriented Architectures and is the primary target.
-
-\item[{Scenarios or examples.}] \leavevmode
-Presented generic scenarios from the medical domain as examples to describe how mechanisms worked.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-The claim is that with respect to Service Oriented Architecture, and the execution of web services, RBAC lacks needed features.
-
-\item {} 
-RBAC does not support partial inheritance
-
-\item {} 
-``total inheritance of permissions in RBAC is against the principle of least privilege''
-
-\item {} 
-ability for some sub-role permissions not be available to super role
-
-\item {} 
-``permission assignment constraints''
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No evidence of use in a real system.
-
-\item[{Is there an evaluation? If yes, how did they do one? If not, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Presents a modification to a core aspect of RBAC
-
-\item {} 
-Allow partial inheritance at the permission level by a superrole instead of treating the role as an atomic entity
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode
-Apply a restricted status to a permission which prevents a superrole from inheriting that permission
-
-\end{description}
-\end{quote}
-
-\chapter{Context}
-\label{categories/context::doc}\label{categories/context:context}\begin{description}
-\item[{Definition.}] \leavevmode
-circumstances in which an event occurs
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1232473}{A flexible content and context-based access control model for multimedia medical image database systems}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpl/login.jsp?tp=\&arnumber=1552897\&url=http\%3A\%2F\%2Fieeexplore.ieee.org\%2Fxpls\%2Fabs\_all.jsp\%3Farnumber\%3D1552897}{A Context-Aware Role-Based Access Control Model for Web Services}
-
-\item {} 
-\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.102.9126\&rep=rep1\&type=pdf}{A context-sensitive access control model and prototype implementation}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4079196}{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1229859}{A contextual role-based access control authorization model for electronic patient record}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4709140}{A Role and Context Based Access Control Model with UML}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1376833}{Designing an agent-based RBAC system for dynamic security policy}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4664701}{An extended RBAC model based on granular logic}
-
-\item {} 
-\href{http://ieeexplore.ieee.org.prox.lib.ncsu.edu/xpl/articleDetails.jsp?tp=\&arnumber=5967517\&contentType=Conference+Publications\&searchField\%3DSearch\_All\%26queryText\%3Dleveraging+access+control+mechanism+of+android+smartphone+using+context-related+role-based+access+control+model}{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}
-
-\item {} 
-\href{http://www.sciencedirect.com.prox.lib.ncsu.edu/science/article/pii/S1084804508000520}{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A flexible content and context-based access control model for multimedia medical image database systems}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Real world inspiration based upon medical image database.  One simple scenario shown where the model is applicable.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-What the image is of is important to who can view
-
-\item {} 
-Domain governs who can see a particular image (e.g. departments in a hospital)
-
-\item {} 
-Location for determining if user is accessing from secure location or location of importance (e.g. emergency room)
-
-\item {} 
-Time constraints for validity periods
-
-\item {} 
-Relationship between images and viewer
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Provided system architecture for entire medical imaging database with access control system. No implied implementation of model.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No evaluation.  Present proposed situation and hospital structure with scenario where new model would be applicable. No reason for lack of evaluation given.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Proposes adding of constraints between the Role-Permission relationship which must be satisfied in order for the holder of said role to be able to use those permissions
-
-\end{itemize}
-
-\item[{This appears to be an active mechanism during enforcement not assignment}] \leavevmode\begin{itemize}
-\item {} 
-Adds user attributes (e.g. company position, location, user name)
-
-\item {} 
-Adds a decision rule to role-permission relationship
-
-\item {} 
-Role-perm relationship contains 5 attributes: identifier, subject/role, action, target objects, constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Context-Aware Role-Based Access Control Model for Web Services}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-No.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-claim is that RBAC can not handle many small services that need roles/perms wrapped around them
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-services replace objects/actions
-
-\item {} 
-adds contexts that must be satisfied
-
-\item {} 
-splits roles into local and global, where local is reflective of actions within a service/domain and global roles are reflective of the permissions to access the service
-
-\item {} 
-roles are activated based on context at execution time
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-context oriented, provides what appears to be a different spin or implementation of heirarchies
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Context-sensitive access control model and prototype implementation}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Initiates the paper with case scenario ``handle an insurance claim'' workflow.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-points out contrasts between base RBAC and improvements clearly
-
-\item {} 
-claims that RBAC supports security principles but does not enforce the use of
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Prototype created using VB6 and MS Access97.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-largely using mechanisms prescribed by RBAC specifically SoD and LP see note about how RBAC provides for but not the use of the previous security practices
-
-\item {} 
-roles are activated whenever a tasks is chosen to be performed and checked if the user can have said role
-
-\item {} 
-provides more implementation of SoD and LP than additions to model
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-context oriented, provides what appears to be a different spin or implementation of heirarchies
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Inspired by the needs of access control in pervasive computing environments
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-``static method of management user-role-permissions''
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Created prototype system to verify model correctness and system architecture
-
-\item {} 
-Implemented the algorithms proposed by their model
-
-\item {} 
-Simulated scenarios but the details of the simulations are not in the paper, nor are the results in a consumable manner
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Claim to be an extension and follow the standard RBAC execution model except by adding two different checks where permissions are re-evaluated by the context of the current user according to the already specific rules
-
-\item {} 
-context aware rules that dynamically determine permissions
-
-\item {} 
-constraining the permissions based on a users location
-
-\item {} 
-assignment/removal of a role based on context information
-
-\item {} 
-connection of roles, granting a user certain role permissions when another user/role is present
-
-\item {} 
-Provide mechanism to deal with conflicts
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A contextual role-based access control authorization model for electronic patient record}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Traditional RBAC does not support emergency overrides based on the situation
-
-\item {} 
-Inspired by the needs of an EHR application in practice
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Claim RBAC is static and cannot handle the dynamic authorization and definitions needed by EHR application complexity
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Implemented using Java and LDAP on InCor EHR system
-
-\item {} 
-They quote number of users and rules, but provide no evidence of comparison or any observation data of the system in use
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Showed some discussion evidence for the requirement for these extensions to RBAC, and they implemented it, however, they provided little evidence as to the effectiveness
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Claim to be a context based extension
-
-\item {} 
-Focused on the authorization aspect of permissions for a given role
-
-\item {} 
-Contains both positive and negative permission authorizations
-
-\item {} 
-Provides for overriding of authorizations based on inheritance
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Role and Context Based Access Control Model with UML}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Describes a scenario whereby a supermarket could not effectively use RBAC to encapsulate all the various access scenarios they need
-
-\item {} 
-Propose a model that adds attributes and system context
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-RBAC ignores context
-
-\item {} 
-RBAC is too simple and unilateral and does not align with real world
-
-\item {} 
-Role granularity is not well controlled
-
-\item {} 
-Roles should depend on the resources not vice-versa
-
-\item {} 
-Roles should e assigned to one resource instead of resources being assigned to one role
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Prototype system created using XML container to contain roles and conditionals
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No evidence as to the comparison of the model in terms of the need over traditional RBAC other than conjecture around a scenario
-
-\item {} 
-The performance test is really only valid for their implementation and shows no grounded re-worldness
-
-\item {} 
-performed a performance analysis of traditional RBAC and C-RBAC based off their prototype
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode
-Extension of the traditional model adding context and maps roles to context and contex to resources with authorization and verification mixed in
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Designing an agent-based RBAC system for dynamic security policy}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-users and application sbecoming larger
-
-\item {} 
-many applications have dynamic attributes and defining scenarios for all of them is difficult
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-For their project manage system, traditional RBAC cannot handle the dynamic nature of multiple projects with multiple tasks that different users may be managing in one and not in another
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-Talks about some weird web based API for the agent
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Discuss a fake project management system claiming this type of system is the target for this type of access control
-
-\item[{Does this model extend the core model?}] \leavevmode
-Defines model clearly
-Extension that defines an Abstract Role, Context Rules and Context Information and the actual role is decided upon based off the context inputs and the rules
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{An extended RBAC model based on granular logic}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Inspired by a muli-level user environment with a complicated authorization management - power load forecasting system
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-traditional RBAC does not provide context
-
-\item {} 
-RBAC is too simple for large dyanmic environments
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No evidence, applied in theory to a conceptual situation, no comparison
-
-\item[{Does this model extend the core model?}] \leavevmode
-Additions to specification, and rules around whether a user is authorized to perform an action based on their role and the context within the role
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Access control needs of smart phones
-
-\item {} 
-Installation of third party applications that a user needs to trust
-
-\item {} 
-User must grant device privileges to the application
-
-\item {} 
-Parents want to limit the amount of time kids use phone
-
-\item {} 
-User might can to limit accessbility to friends and admins
-
-\item {} 
-Companies want to limit data access by employee phones
-
-\item {} 
-User loses company phone in admin mode, context to prevent leakage
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-smartphone is centralized, user-centric system where identities are known in advance
-
-\item {} 
-smartphone has lots of contextual info
-
-\item {} 
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Adds objects, environmental context, policies and decisions to model
-
-\item {} 
-Object is an accessible entity
-
-\item {} 
-a property of the system at the moment of interaction
-
-\item {} 
-policy is the formal specification of the access control
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-precedential order of access privileges to prvent policy bugs
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-proliferation of distributed applications
-
-\item {} 
-example of limited disk space when a database read and update are performed to resolve which operation to give preference to
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-need flexible and dynamic authorization constraints
-
-\item {} 
-authorization differs between and within applications
-
-\item {} 
-Users may be granted access to the application and entities or may be granted access to objects with the application depending on context
-
-\item {} 
-role and permission constraints are also possible
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-analysis of example scenario, comparison of constraint vs other models,
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-adds object sets, entities, status and set of authorization attributes
-
-\item {} 
-permissions on objects with constraints
-
-\item {} 
-user and user constraints mapped to roles with permission attribute constraints
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-classification of constraints as either users eligibility to to use a resource/service or constraints on the users actual use of a resource (limiting what they can do)
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\chapter{Historical}
-\label{categories/historical::doc}\label{categories/historical:historical}\begin{description}
-\item[{Definition.}] \leavevmode
-based on or reconstructed from an event, custom, style, etc., in the past
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.182.8833\&rep=rep1\&type=pdf}{HRBAC: Historical Role-Based Access Control}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{HRBAC: Historical Role-Based Access Control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-a need to incorporate information of the past to make access control decisions
-
-\item {} 
-for example, ensuring a user had a set of previous roles before being granted a higher responsibility role (e.g. ``chief'' role needs to have once had a manager role)
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-in order to store and express historical entities of each base entity in order to apply constraints of current assignments and roles and permissions based off the past
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-No
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No, only examples of mechanisms
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Yes, extends each of the 4 levels of RBAC
-
-\item {} 
-incorporates new entities to represent historical version of each core entity
-
-\item {} 
-provides various historical constraints
-
-\item {} 
-provides algorithms for performing constraint resolution
-
-\item {} 
-highly defined model and entities
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Organizational}
-\label{categories/organizational:organizational}\label{categories/organizational::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to an organization
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4207551}{ROBAC: Scalable Role and Organization Based Access Control Models}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{ROBAC: Scalable Role and Organization Based Access Control Models}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-inspired by context from business-to-business and business-to-consumer context
-
-\item {} 
-a theoritical web based report delivery system e.g. a set of educational professionals from schools, districts, and states
-
-\item {} 
-also provide example of an online tutoring system
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-claim that their model can significantly reduce administrative complexity for multi-organization
-
-\item {} 
-more succint and intuitive than RBAC
-
-\item {} 
-same expressive power
-
-\end{itemize}
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-present comparison calculations for the number of roles that must be generated between their model and standard
-
-\item {} 
-provide numerical size comparisons with other proposed models (Role Templates, Team-based Access Control, Organizational Units
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension of RBAC to consider organization and role (a.k.a. context)
-
-\item {} 
-an organization must own the object, and the role must give permission on the type for a user to access
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Privacy}
-\label{categories/privacy::doc}\label{categories/privacy:privacy}\begin{description}
-\item[{Definition.}] \leavevmode
-freedom from unauthorized intrusion
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1266848}{Privacy-aware role-based access control}
-
-\item {} 
-\href{http://www.sis.pitt.edu/~amirreza/papers/masoumzadeh2008\_purbac.pdf}{PuRBAC: Purpose-Aware Role-Based Access Control}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Privacy-aware role-based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-inspired by conventional RBAC's inability to handle privacy
-
-\item {} 
-extra attributes are needed to support the intricacies of privacy
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}]
-due to the lack of basic components required by privacy regulations, especially purpose binding (i.e., data collected for one purpose should not used for another purpose without user consent), conditions, and obligations.
-Despite its limitations, existing access control technology can be used as a starting point for managing personal identifiable information
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Evaluated the performance of assigning permissions and checking authorization for various permission size bases as well as variety in number of context variables
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Defines additions to each level of the base RBAC model
-
-\item {} 
-Adds objects, purposes, actions, obligations and conditions to core
-
-\item {} 
-Includs context variables that can be very general, from space to time to other
-
-\item {} 
-Conflict detection and resolution mechanisms
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{PuRBAC: Purpose-Aware Role-Based Access Control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Web and companies with privacy policies make it harder to follow said policies and protect user data
-
-\end{itemize}
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-privacy policies include purposes under which collected data can be used
-
-\item {} 
-current RBAC model does not include a entity or relationships to incorporate purpose
-
-\item {} 
-authorization requests should include a purpose for performing the intended action
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-No.
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No evaluation, conclusion includes note that further analysis needs to be done
-
-\item {} 
-Discusses entities added by this model, the way they were added and why this method is the best way to add purpose and support privacy
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extends the base and heirarchy portion of the model, adds in a hybrid heirarchy
-
-\item {} 
-connects purposes to permissions
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode\begin{itemize}
-\item {} 
-claim to reduce policy complexity by associating a purpose with only a permission and not with roles and users as well
-
-\item {} 
-locks down the users ability to declare purpose, by putting authorization controls around the purposes
-
-\end{itemize}
-
-\end{description}
-\end{quote}
-
-\chapter{Resource}
-\label{categories/resource:resource}\label{categories/resource::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-an available supply that can be drawn on when needed
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4609645}{A Flexible Role- and Resource-Based Access Control Model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Flexible Role- and Resource-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-No.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-RBAC cannot support multiple security domains
-
-\item {} 
-cannot support same security with different applications
-
-\item {} 
-lack of flexibility in authorization
-
-\item {} 
-discard role heirarchy for directory structure to reduce complexity and increase flexibility
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Comparison of the two models from a qualitative standpoint
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Not an extension
-
-\item {} 
-Re-imagining of the hierarchy aspect and adding a resources entity
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Spatial}
-\label{categories/spatial::doc}\label{categories/spatial:spatial}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to facility in perceiving relations in space
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1063985}{GEO-RBAC: a spatially aware RBAC}
-
-\item {} 
-\href{http://www.springerlink.com/index/4N6XK46321M574UH.pdf}{LRBAC: A location-aware role-based access control model}
-
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1285394}{Spatial role-based access control model for wireless networks}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{GEO-RBAC: a spatially aware RBAC}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Hospital based examples of activating roles, allowing doctors to only see patient records while in the department of the patient
-Extended example section that discusses how a hospital setting with some typical actors would perform simple patient access
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Lack of contextual information with roles, users and permissions
-Role needs a schema which allows specification of more attributes of the role
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Shows that this model can be be represented by first-order logic and that they can be evaluated in PTIME ?
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Introduces the concept of a spatial role
-
-\item {} 
-Extends the base model at each of the 4 levels
-
-\item {} 
-Specify mutual exclusion constraint for separation of duty
-
-\item {} 
-Split model into a model without constraints and an extensive model with constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{LRBAC: A location-aware role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Some brief real world concepts with regards to intricate security scenarios based on location of user or object
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-Allowing for different security scenarios based on location
-
-\item {} 
-Role activation based on location of object or user
-
-\item {} 
-Add objects to define physical or logical with location data
-
-\item {} 
-Attaches an operation to a permission
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Describes in full definition each object and it's role in the graph
-Describes a number of specific operations that would be performed and how the model handles them with respect to locations
-Description of how the core RBAC is related to each piece of their model
-No evaluation
-Minor scenarios
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Extension adding operation, object and role enabling based on a constraint
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Spatial role-based access control model for wireless networks}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Provide an application scenario of where this would be needed and used
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-standard RBAC cannot handle location based constraints and attributes
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Mathematical specification of each aspect
-Definitions of each element in their model
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension at every level of RBAC
-
-\item {} 
-includes separation of duty and heiarchry concerns specifically
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Spatio-Temporal}
-\label{categories/spatio_temporal:spatio-temporal}\label{categories/spatio_temporal::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-combining space and time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://www.springerlink.com/index/tq15n88122v63754.pdf}{STARBAC: Spatiotemporal role based access control}
-
-\item {} 
-\href{http://dl.acm.org/citation.cfm?id=1368341}{On spatio-temporal constraints and inheritance in role-based access control}
-
-\item {} 
-\href{https://www.cerias.purdue.edu/tools\_and\_resources/bibtex\_archive/archive/2007-08.pdf}{A framework for specification and verification of generalized spatio-temporal role based access control model}
-
-\item {} 
-\href{http://www.springerlink.com/index/77j0j540004585wv.pdf}{LoT-RBAC: A Location and Time-Based RBAC Model}
-
-\item {} 
-\href{http://www.springerlink.com/index/C3T50338535HX115.pdf}{A spatio-temporal role-based access control model}
-
-\item {} 
-\href{http://www.springerlink.com/index/r1n6j114g5431174.pdf}{Role Based Access Control with Spatiotemporal Context for Mobile Applications}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{STARBAC: Spatiotemporal role based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Similar reasoning as other location based models, No formal evaluation, scenarios or real world examples.
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Extension of RBAC and of SRBAC by combining spatial and temporal
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Mathematical explanations, definitions
-Includes descriptions and math behind the semantics
-
-\end{description}
-
-Does this model extend the core model?
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{On spatio-temporal constraints and inheritance in role-based access control}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Believe that previous models focus on syntax and not semantics
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-pervasive computing environments requirement for spatial and temporal awareness
-
-\item {} 
-previous spatio-temporal do nt address interaction between these constraints and inheritance
-
-\item {} 
-makes 4 claims about how existing models do not handle the above
-
-\item {} 
-trusted entities to override restrictions
-
-\item {} 
-attempt to show compatibility with standard model and claim others do not
-
-\end{itemize}
-
-\end{description}
-
-Did they implement the model?
-\begin{description}
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Examine previous models and point out deficiencies, use of graphs to show off relationships
-
-\item {} 
-Evaluation includes a mathematical examination of the role graphs, and ordering
-
-\item {} 
-authors note that there is added complexity in their method, but that is the price to pay to get user  and user-role application constraints
-
-\item {} 
-trade-off between complexity of policies and complexity of constraints
-
-\item {} 
-authors do comparisons of complexity for the various models
-
-\item {} 
-authors note that SoD should be considered in the future
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode
-Additions to the specification for constraints with a specific look at how to do spatial and temporal
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A framework for specification and verification of generalized spatio-temporal role based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Based off hospital example with mobile devices inspiring combining location and time based models
-Provides an example hospital access control policy that requires space and time - based on the layout of the floor of a hospital and the various users and entitie sthat might exist such as nurse, doctor and surgeons
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Need to be able to handle add attributes that location and spatial specification requires
-Add context mechanisms
-
-\item[{Did they implement the model?}] \leavevmode
-No
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-To some degree, an example, and some analysis of developing policy and how to deal with conflicts
-
-\item[{Does this model extend the core model?}] \leavevmode
-Builds on GTRBAC, extends at all 4 levels
-Adds role activation/deactivation
-Provides conflict resolution
-Includes a GSTRBAC policy specification model using Alloy
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{LoT-RBAC: A Location and Time-Based RBAC Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Inspired by growing needs of mobile, peer-to-peer devices with regards to activating and specifying permissions and roles based on time and location
-Detailed scenario of a person having a medical emergency, and a doctor on site needing to request patient data, as well as the progression through the ambulance, to hospital and citing such location based clues to help know if the doctor should be granted privileges such as being in or near his own car
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Previous location and time based RBAC models did not incorporate location for both the user and role
-In-depth, detailed location heirarchies and specifications are needed
-Physical, logical, and hybrid location
-
-\item[{Did they implement the model?}] \leavevmode
-No
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No
-
-\end{description}
-
-Does this model extend the core model?
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Spatio-temporal Role-Based Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Growth of wireless networks, mobile devices, the move towards pervasive computing
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Claim that previously examined models of spatial, temporal, or spatio-temporal lack needed aspects
-Role assignment should be dependent on space and time
-Permissions need to be dependent on space and time not just roles
-
-\item[{Did they implement the model?}] \leavevmode
-No. Authors acknowledge need to implement.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-No.  Best they do is provide a detailed scenario with an examination of how such a scenario would be represented in their model. Author acknowledge detailed analysis is needed.
-
-\item[{Does this model extend the core model?}] \leavevmode
-Defines for Space - physical location, logical location and mapping functions and relationship definitions
-Defines for Time - time instant, time interval
-Adds objects to Core model, object can be physical or logical (computer or files e.g.)
-Defines a number of potential role hierarchy strategies
-Examines impact on both static and dynamic separation of duties for new attributes and mechanisms
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Role Based Access Control with Spatiotemporal Context for Mobile Applications}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Need and desire to put spatio-temporal context and constraint on roles and permissions - such as students at a university or hospital facilities
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-Needs enhancements to be able to handle space and time attributes on roles and permissions
-
-\item[{Did they implement the model?}] \leavevmode
-Yes.  Integrated an access control system into iMedik - a telemedicine application in India that is web-based and allows mobile access
-Their implementation is a bolt-on solution, that intercepts calls and verifies their authorization
-Provide detailed architecture and implementation details
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode
-Include section that discusses differences between their model and STRBAC and STARBAC
-Incorporate algorithms for determining whether a particular authorization is granted and use time-complexity arguments to discuss the effectiveness of each algorithm and to attempt to display that the added complexity of their model over traditional RBAC does not impace performance significantly
-
-\item[{Does this model extend the core model?}] \leavevmode
-Extends at all 4 levels
-Adds ``extents'' which are design to extend the base entities such as a role, multiple RoleExtents can belong to a single role based off constraint or context
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\chapter{Task}
-\label{categories/task:task}\label{categories/task::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-a usually assigned piece of work often to be finished within a certain time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=4623993}{A Task-Role Based Access Control Model with Multi-Constraints}
-
-\item {} 
-\href{http://en.cnki.com.cn/Article\_en/CJFDTOTAL-TXJS200911047.htm}{Team and Task Based RBAC Access Control Model}
-
-\item {} 
-\href{http://www.sciencedirect.com/science/article/pii/S0306437902000297}{Task–role-based access control model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A Task-Role Based Access Control Model with Multi-Constraints}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Not much reasoning or evidence as to the `why'
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-RBAC can't handle large enterprises with multiple applications and a single user base
-
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Discussion of the applications of the model, mentions of real companies using the implementation, and a mention of a system that exists
-
-\end{itemize}
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Unclear, as their examples and applications didn't contain much data
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-rigorous constraints definitions, mathematical specification of additional model elements
-
-\item {} 
-Attempts to combine TBAC and RBAC by placing a task between a role and a permission
-
-\item {} 
-not really extending the base RBAC model
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Team and Task Based RBAC Access Control Model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode\begin{itemize}
-\item {} 
-Real world scenarios - hospital
-
-\item {} 
-Applications of TT-RBAC in an imaginary hospital scenario
-
-\end{itemize}
-
-\end{description}
-
-What reasons and evidence do researchers claim RBAC needs extension?
-\begin{description}
-\item[{Did they implement the model?}] \leavevmode\begin{itemize}
-\item {} 
-Prototype system
-
-\end{itemize}
-
-\end{description}
-
-Is there an evaluation? If yes, how did they do one? If no, why?
-\begin{description}
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Covers all 4 levels of RBAC, provides mathematical definitions of entities
-
-\item {} 
-Extension to RBAC at all four levels
-
-\item {} 
-Introduces teams that map to users, roles and tasks
-
-\item {} 
-Introduces tasks that map to teams and permissions
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{Task–role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-Define a set of security requirements for access control in an enterprise environment
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode
-8 security requirements for enterprise environments
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-No implementation, not enough envidence other than the security requirements, the requirements are based off a few thought scenarios that seem based on a `typical' enterprise organizational setup
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-extension by adding Task notion to connect with roles and users
-
-\end{itemize}
-
-\item[{Notes.}] \leavevmode
-Task has constraints, task sits between permission and roles, tasks used to compose workflows, and are more mutable than roles - task can be used for defining set of actions and help when the specific data objects are less known at creation time
-
-\end{description}
-\end{quote}
-
-\chapter{Temporal}
-\label{categories/temporal:temporal}\label{categories/temporal::doc}\begin{description}
-\item[{Definition.}] \leavevmode
-of or relating to the sequence of time or to a particular time
-
-\item[{Papers}] \leavevmode\begin{itemize}
-\item {} 
-\href{http://ieeexplore.ieee.org/xpls/abs\_all.jsp?arnumber=1363762}{A generalized temporal role-based access control model}
-
-\end{itemize}
-
-\end{description}
-
-\bigskip\hrule{}\bigskip
-
-\textbf{A generalized temporal role-based access control model}
-\begin{quote}
-\begin{description}
-\item[{Inspired by a real example? Scenarios or examples.}] \leavevmode
-In many organizations, processes and functions may have limited time spans or have periodic temporal durations
-
-\item[{What reasons and evidence do researchers claim RBAC needs extension?}] \leavevmode\begin{itemize}
-\item {} 
-Workflow based organizations need a time based constraint for disabling/enabling roles
-
-\end{itemize}
-
-\item[{Did they implement the model?}] \leavevmode
-No.
-
-\item[{Is there an evaluation? If yes, how did they do one? If no, why?}] \leavevmode\begin{itemize}
-\item {} 
-Model is limited to enabling/disabling of roles based on time constraints only
-
-\item {} 
-model is fully described from entities, to constraints, to conflict resolution to execution in pseudo-code
-
-\item {} 
-Evaluation lies in the construction of mathematical modeling and graphs to detail the usage and interaction of the model pieces
-
-\item {} 
-A few example scenarios are explained
-
-\end{itemize}
-
-\item[{Does this model extend the core model?}] \leavevmode\begin{itemize}
-\item {} 
-Extension of the base RBAC model
-
-\item {} 
-adds periodicity and duration constraints for role enabling
-
-\item {} 
-provides for conflict resolution of constraints
-
-\end{itemize}
-
-\end{description}
-
-Notes.
-\end{quote}
-
-\begin{thebibliography}{1}
-\bibitem[1]{1}{\phantomsection\label{intro:1} \DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{David}\DUspan{}{ }\DUspan{}{F.}}\DUspan{}{ }}\DUspan{}{Ferraiolo}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Ravi}\DUspan{}{ }\DUspan{}{S.}}\DUspan{}{ }}\DUspan{}{Sandhu}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{Serban}\DUspan{}{ }\DUspan{}{I.}}\DUspan{}{ }}\DUspan{}{Gavrila}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{\DUspan{}{D.}\DUspan{}{ }\DUspan{}{Richard}}\DUspan{}{ }}\DUspan{}{Kuhn}}}\DUspan{}{, and }\DUspan{}{\DUspan{}{\DUspan{}{Ramaswamy}\DUspan{}{ }}\DUspan{}{Chandramouli}}}\DUspan{}{.}}\DUspan{}{ }\DUspan{}{\DUspan{}{Proposed NIST standard for role-based access control}\DUspan{}{.}}}\DUspan{}{ }\DUspan{}{\DUspan{}{\emph{\DUspan{}{ACM Transactions on Information and System Security}}\DUspan{}{, }\DUspan{}{\DUspan{}{\DUspan{}{4}\DUspan{}{\DUspan{}{(}\DUspan{}{3}\DUspan{}{)}}\DUspan{}{:}}\DUspan{}{\DUspan{}{224}\DUspan{}{–}\DUspan{}{274}}}}\DUspan{}{, }\DUspan{}{2001}\DUspan{}{.}}}}
-\end{thebibliography}
-
-\renewcommand{\indexname}{Index}
-\printindex
-\end{document}

### 2013-02-05 paper/sections/extraction.tex
Some minor updates and the start of a data extraction section.

\section{Data Extraction} \label{sec:extraction}

After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends.  The first step was to take the individual data generated from the final elimination round and organize this information around the research questions.  During the paper reading round, and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.  Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.  The result of this paper categorization is displayed below:

\begin{tabular}{\linewidth}{|l|l|}
\hline
\textbf{Paper} & \textbf{Category}
\\\hline
Alam, M. and Hafner, M. and Breu, R. 2006 & Constraint \\\hline
Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 & Context \\\hline
Haibo, SHEN and Fan, HONG 2005 & Context \\\hline
Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 & Context \\\hline
Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 & Context \\\hline
Motta, G.H.M.B. and Furuie, S.S. 2003 & Context \\\hline
Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 & Context \\\hlin
Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 & Context \\\hline
Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 & Context \\\hline
Thein, N. and others 2011 & Context \\\hline
Zou, D. and He, L. and Jin, H. and Chen, X. 2009 & Context \\\hline
Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 & Delegation \\\hline
Zhang, Z. and Zhang, X. and Sandhu, R. 2006 & Organizational \\\hline
Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 & Privacy \\\hline
Masoumzadeh, A. and Joshi, J. 2008 & Privacy \\\hline
Zhao, Y. and Zhao, Y. and Lu, H. 2008 & Resource \\\hline
Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 & Spatial \\\hline
Ray, I. and Kumar, M. and Yu, L. 2006 & Spatial \\\hline
Hansen, F. and Oleshchuk, V. 2003 & Spatial \\\hline
Aich, S. and Sural, S. and Majumdar, A. 2007 & Spatio-Temporal \\\hline
Chen, L. and Crampton, J. 2008 & Spatio-Temporal \\\hline
Samuel, A. and Ghafoor, A. and Bertino, E. 2007 & Spatio-Temporal \\\hline
Chandran, S. and Joshi, J. 2005 & Spatio-Temporal \\\hline
Ray, I. and Toahchoodee, M. 2007 & Spatio-Temporal \\\hline
Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 & Spatio-Temporal \\\hline
Yao, L. and Kong, X. and Xu, Z. 2008 & Task \\\hline
ZHANG, S. and CHEN, X. and HOU, G. 2009 & Task \\\hline
Oh, S. and Park, S. 2003 & Task \\\hline
Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 & Temporal \\\hline
\end{tabular}

Given that there were multiple papers for some categories, the researchers decided to tackle all further research questions by first analyzing the research question on a per category basis and then looking across all categories for generalization and trends.

### 2013-02-05 paper/sections/intro.tex
Some minor updates and the start of a data extraction section.

@@
 
-Role-based access control (RBAC) models {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}} become popularly used to govern access to critical resources.
+Role-based access control (RBAC) models \cite{ferraiolo-rbac} become popularly used to govern access to critical resources.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model {\hyperref[intro:ferraiolo-rbac]{{[}1{]}}}
+RBAC is first introduced in 1990s, NIST proposed standard RBAC model \cite{ferraiolo-rbac}
 Standard RBAC model considers only role-user association and role hierarchy.

### 2013-02-05 paper/sections/process.tex
Some minor updates and the start of a data extraction section.

@@
 
+The systematic literature review process was developed ahead of time and agreed upon by the researchers following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
+The systematic literature review was performed in four stages:
+
+\begin{itemize}
+\item Development of a search strategy
+\item Elimination of papers based on title
+\item Elimination of papers based on abstract
+\item Elimination of papers based on content and matching to elimination criteria
+\end{itemize}
+
+\subsection{Search Strategy}
+
 For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
+
 \begin{itemize}
-\item {} 
-Google Scholar - \href{http://scholar.google.com}{http://scholar.google.com}
-\item {} 
-IEEExplore     - \href{http://ieeexplore.ieee.org}{http://ieeexplore.ieee.org}
-\item {} 
-ACM Portal     - \href{http://dl.acm.org}{http://dl.acm.org}
-\item {} 
-CiteSeerX      - \href{http://http://citeseerx.ist.psu.edu/index}{http://http://citeseerx.ist.psu.edu/index}
+\item Google Scholar - \url{http://scholar.google.com}
+\item IEEExplore - \url{http://ieeexplore.ieee.org}
+\item ACM Portal - \url{http://dl.acm.org}
+\item CiteSeerX - \url{http://http://citeseerx.ist.psu.edu/index}
 \end{itemize}
@@
 \begin{itemize}
-\item {} 
-role based access control
-\item {} 
-RBAC
-\item {} 
-role-based access control
+\item role based access control
+\item RBAC
+\item role-based access control
 \end{itemize}
@@
 Each run was performed as follows:
+
 \begin{enumerate}
-\item {} 
-Remote call to search engine with current search start position and the current search criteria.
-\item {} 
-Parse results and extract paper title, authors and year of publication.
-\item {} 
-Compare results against stopping criteria.
+\item Remote call to search engine with current search start position and the current search criteria.
+\item Parse results and extract paper title, authors and year of publication.
+\item Compare results against stopping criteria.
 \end{enumerate}
@@
 \begin{itemize}
-\item {} 
-If stopping criteria met, stop search.
-\item {} 
-If stopping criteria not met, increase search position by number of results and return to step 1.
+\item If stopping criteria met, stop search.
+\item If stopping criteria not met, increase search position by number of results and return to step 1.
 \end{itemize}
@@
 
-There were a total of three elimination rounds.
-The first round was based solely on title, the second on reading of the abstract and the last round was based on a full read of the paper and comparison to the research questions outlined.
-Each round was performed as follows:
+\subsection{Elimination Rounds}
+
+The elimination rounds were performed based on reading of the title, abstract and finally the paper itself.
+While each elimination stage had a unique set of criteria for elimination the general procedure for elimination for the researchers was as follows.
 
 \begin{itemize}
-\item {} 
-Each reviewer independently classified papers as relevant, irrelevant or uncertain.
-\item {} 
-Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
-\item {} 
-Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
+\item Each reviewer independently classified papers as relevant, irrelevant or uncertain.
+\item Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
+\item Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
 From this discussion, papers were either thrown out or kept until the next round of the review.
@@
 
+The title elimination round was based off of whether role based access control and model were mentioned directly.
+
+The second round of elimination was based off on reading the abstracts of the remaining papers.  The inclusion criteria for the abstract reading tried to answer the following questions:
+
+\begin{itemize}
+\item Does the abstract mention a proposed model?
+\item Does the abstract mention extension of role-based access control?
+\item Does the abstract mention either an implementation, evaluation or domain for their model?
+\end{itemize}
+
 The results of the searches is summarised below:
 
-\begin{tabulary}{\linewidth}{|L|L|L|L|L|}
+\begin{tabular}{|l|l|l|l|l|}
 \hline
-\textbf{
-Search Engines
-} & \textbf{
-RBAC
-} & \textbf{
-role based access control
-} & \textbf{
-role-based access control
-} & \textbf{
-Total
-}\\\hline
+\textbf{Search Engines} & 
+\textbf{RBAC} & 
+\textbf{role based access control} & 
+\textbf{role-based access control} & 
+\textbf{Total}
+\\\hline
 
@@
 
-Combined
- &  &  &  & 
-\textbf{1716}
+Combined &  &  &  & \textbf{1716}
 \\\hline
-\end{tabulary}
+\end{tabular}

### 2013-02-06 paper/sections/extraction.tex
Adds updated bibliography.

@@
 
-\begin{tabular}{\linewidth}{|l|l|}
+\begin{center}
+\begin{tabular}{|p{10cm}|p{3cm}|}
 \hline
@@
 \\\hline
-Alam, M. and Hafner, M. and Breu, R. 2006 & Constraint \\\hline
-Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 & Context \\\hline
-Haibo, SHEN and Fan, HONG 2005 & Context \\\hline
-Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 & Context \\\hline
-Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 & Context \\\hline
-Motta, G.H.M.B. and Furuie, S.S. 2003 & Context \\\hline
-Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 & Context \\\hlin
-Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 & Context \\\hline
-Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 & Context \\\hline
-Thein, N. and others 2011 & Context \\\hline
-Zou, D. and He, L. and Jin, H. and Chen, X. 2009 & Context \\\hline
-Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 & Delegation \\\hline
-Zhang, Z. and Zhang, X. and Sandhu, R. 2006 & Organizational \\\hline
-Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 & Privacy \\\hline
-Masoumzadeh, A. and Joshi, J. 2008 & Privacy \\\hline
-Zhao, Y. and Zhao, Y. and Lu, H. 2008 & Resource \\\hline
-Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 & Spatial \\\hline
-Ray, I. and Kumar, M. and Yu, L. 2006 & Spatial \\\hline
-Hansen, F. and Oleshchuk, V. 2003 & Spatial \\\hline
-Aich, S. and Sural, S. and Majumdar, A. 2007 & Spatio-Temporal \\\hline
-Chen, L. and Crampton, J. 2008 & Spatio-Temporal \\\hline
-Samuel, A. and Ghafoor, A. and Bertino, E. 2007 & Spatio-Temporal \\\hline
-Chandran, S. and Joshi, J. 2005 & Spatio-Temporal \\\hline
-Ray, I. and Toahchoodee, M. 2007 & Spatio-Temporal \\\hline
-Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 & Spatio-Temporal \\\hline
-Yao, L. and Kong, X. and Xu, Z. 2008 & Task \\\hline
-ZHANG, S. and CHEN, X. and HOU, G. 2009 & Task \\\hline
-Oh, S. and Park, S. 2003 & Task \\\hline
-Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 & Temporal \\\hline
+Alam, M. and Hafner, M. and Breu, R. 2006 \cite{alam06:constraint} & Constraint \\\hline
+Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 \cite{tzelepi01:flexible} & Context \\\hline
+Haibo, SHEN and Fan, HONG 2005 \cite{haibo05:context} & Context \\\hline
+Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
+Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 \cite{huang06:pervasive} & Context \\\hline
+Motta, G.H.M.B. and Furuie, S.S. 2003 \cite{motta03:contextual} & Context \\\hline
+Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 \cite{bao08:role} & Context \\\hline
+Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 \cite{yamazaki104:designing} & Context \\\hline
+Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 \cite{han08:extended} & Context \\\hline
+Thein, N. and others 2011 \cite{thein2011leveraging} & Context \\\hline
+Zou, D. and He, L. and Jin, H. and Chen, X. 2009 \cite{zou2009crbac} & Context \\\hline
+Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 \cite{hasebe10:capability} & Delegation \\\hline
+Zhang, Z. and Zhang, X. and Sandhu, R. 2006 \cite{zhang06:collaborative} & Organizational \\\hline
+Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 \cite{ni2010privacy} & Privacy \\\hline
+Masoumzadeh, A. and Joshi, J. 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
+Zhao, Y. and Zhao, Y. and Lu, H. 2008 \cite{zhao2008flexible} & Resource \\\hline
+Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 \cite{damian2007geo} & Spatial \\\hline
+Ray, I. and Kumar, M. and Yu, L. 2006 \cite{ray07:spatio} & Spatial \\\hline
+Hansen, F. and Oleshchuk, V. 2003 \cite{hansen2003spatial} & Spatial \\\hline
+Aich, S. and Sural, S. and Majumdar, A. 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
+Chen, L. and Crampton, J. 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
+Samuel, A. and Ghafoor, A. and Bertino, E. 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
+Chandran, S. and Joshi, J. 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
+Ray, I. and Toahchoodee, M. 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
+Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
+Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
+ZHANG, S. and CHEN, X. and HOU, G. 2009 \cite{zhou2007team} & Task \\\hline
+Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
+Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline
 \end{tabular}
+\end{center}
 

### 2013-02-06 paper/sections/intro.tex
Adds updated bibliography.

@@
 
-Role-based access control (RBAC) models \cite{ferraiolo-rbac} become popularly used to govern access to critical resources.
+Role-based access control (RBAC) models \cite{ferraiolo} become popularly used to govern access to critical resources.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model \cite{ferraiolo-rbac}
+RBAC is first introduced in 1990s, NIST proposed standard RBAC model \cite{ferraiolo}
 Standard RBAC model considers only role-user association and role hierarchy.

### 2013-02-06 paper/sections/terms.tex
Adding terms section.

\section{Core Role Based Access Control} \label{sec:core-rbac}

The NIST RBAC model is proposed by Ferraiolo et al.. The model
consists of four basic components:

\begin{itemize}
	\item a set of users \emph{Users}: A user can be a person or an agent.
	\item  a set of roles \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
	\item a set of permissions \emph{Permissions}: A permission refers to an access mode that
can be exercised on an object in the system and a session
relates a user to possibly many roles.
	\item a set of sessions \emph{Sessions}: In each session, a
user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
\end{itemize}

In the RBAC, a user can exercise a permission only if the user are assigned to a role.
In addition to the four basic components, two functions are defined:
the user role assignment (UA) and the role
permission assignment (PA) functions.
UA models assignment of users to roles.
PA models assignment of permissions to roles.

%Each user incorporates a session    
%The user function maps each session to a
%single user, whereas the role function establishes a
%mapping between a session and a set of roles activated
%by the corresponding user in the session.

\section{Terms and Definitions} \label{sec:terms}

We found that different definitions for the same terms. Therefore, we next describe
terms and definitions.

\begin{itemize}
%	\item Access Control Model:
%	\item Access Control Policy:
%	\item Access Control Requirements:
%	\item Rules:
%	\item Attributes-Based Access Control:
	\item Task-Based Access Control:
%	\item Purpose-Based Access Control:
	\item Agent-Based Access Control:						
\end{itemize}

\begin{itemize}

%	\item Role Hierarchies:

	\item Obligations: obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
	her/his office duty before accessing the resources.

	\item Inheritance:
	Inheritance defines an inheritance relationship among attributes such as roles. For example, the role structure for a company use
	employee role for employees. Department manager may inherit all permissions of the employee role. (Role-Based Access Control by F. Ferrailolo et al.)

	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$. SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$. (Role-Based Access Control by F. Ferrailolo et al.)

	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user. (Role-Based Access Control by F. Ferrailolo et al.)
	\end{itemize}

We define temporal and spatial constraints as follows:

\begin{itemize}
	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.   
	(Role-Based Access Control by F. Ferrailolo et al.)

	\item Spatial Constraints: Spatial constraints are location-based constraints in specifying access
	control policies. For example, in organizations, 	locations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
	In such cases, the part-time employee role should access required resources only when the user is in the location. 
	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.

\end{itemize}

### 2013-02-09 paper/sections/conclusion.tex
Adding some updated writing and notes on each section and what to put.

\section{Conclusion} \label{sec:conclusion}

P1: Mention core RBAC, note specific results for each research question

P2: Mention bigger results that cut across and any other peritentn information

### 2013-02-09 paper/sections/extraction.tex
Adding some updated writing and notes on each section and what to put.

@@
 \section{Data Extraction} \label{sec:extraction}
+
+P1: Describe data extraction process, including questions each reviewer based note taking off of
+
+P2: Describe how from the data extraction process cateogories of extension became evident
+
+P3: Describe how those categories were deemed relevant and a way to organize going forward
+
+P4: Describe the categories, show how each paper and their category
+
+P5: Discuss how the categories are inter-related and show diagram of relationships
 

### 2013-02-09 paper/sections/intro.tex
Adding some updated writing and notes on each section and what to put.

@@
 
-Role-based access control (RBAC) models \cite{ferraiolo} become popularly used to govern access to critical resources.
+P1: Note history of the RBAC model, statistics, and official status
+
+P2: Mention extensions exist in response to developments and new domains over time
+
+P3: Mention audience, why they should care, and why exploring extensions space is important in relation to the reason the standard was originally concieved
+
+P4: Talk about what this work is going to do, goal, brief process
+
+P5: List goals or contributions
+
+P6: Outline paper and sections
+
+Role based access control (RBAC) was first introduced in the 1990 when the National Instittute for Standards and Technology (NIST) requested that a unified standard be created.
+proposed standard RBAC model \cite{ferraiolo}
+
+Role-based access control (RBAC) models \cite{ferraiolo} became popular used to govern access to critical resources.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 
-RBAC is first introduced in 1990s, NIST proposed standard RBAC model \cite{ferraiolo}
 Standard RBAC model considers only role-user association and role hierarchy.
@@
 We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
-\begin{description}
-\item[{Our research give benefits to a community as follows:}]
+
+Our research give benefits to a community as follows:
 
 \begin{itemize}
-\item {} 
-Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
-\item {} 
-Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
-\item {} 
-Our work shows a criteria in comparison among research results.
-\item {} 
-Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
+\item Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
+\item Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
+\item Our work shows a criteria in comparison among research results.
+\item Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
 \end{itemize}
-
-\end{description}

### 2013-02-09 paper/sections/results.tex
Adding some updated writing and notes on each section and what to put.

\section{Results} \label{sec:results}

P1: Discuss focus on going through each research question individually, and combining results across at the end of the process

P2: List the research questions

S1: RQ1 and discuss by category and general results

S2: RQ2 and discuss by category and general results

S3: RQ3 and discuss by category and general results

S4: RQ4 and discuss by category and general results

S5: RQ5 and discuss by category and general results

### 2013-02-09 paper/sections/terms.tex
Adding some updated writing and notes on each section and what to put.

@@
 
-The NIST RBAC model is proposed by Ferraiolo et al.. The model
-consists of four basic components:
+P1: Describe RBAC entities
+
+P2: Note the 4 levels of RBAC
+
+P3: Concisely describe level 1
+
+P4: Concisely describe level 2
+
+P5: COncisely describe level 3
+
+P6: Concisely describe level 4
+
+Since the basis for our review is extensions to the core model, we will describe the core model, associated entities and other terminology encountered across the space of our review.
+The NIST RBAC model proposed by Ferraiolo et al.
+and later adopted as the official standard for RBAC by the International Committee for Information Technology Standards (INCITS) consists of four basic entities:
 
@@
 \item  a set of roles \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
-	\item a set of permissions \emph{Permissions}: A permission refers to an access mode that
-can be exercised on an object in the system and a session
-relates a user to possibly many roles.
-	\item a set of sessions \emph{Sessions}: In each session, a
-user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
+\item a set of permissions \emph{Permissions}: A permission refers to an access mode that can be exercised on an object in the system and a session relates a user to possibly many roles.
+\item a set of sessions \emph{Sessions}: In each session, a user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
 \end{itemize}
@@
 \section{Terms and Definitions} \label{sec:terms}
+
+NOTE: Not sure about this section since some of this is derived from our results
 

### 2013-02-17 paper/sections/intro.tex
Updates to results, process.

@@
 
-P5: List goals or contributions
+P5: List contributions
 
@@
 
-Role based access control (RBAC) was first introduced in the 1990 when the National Instittute for Standards and Technology (NIST) requested that a unified standard be created.
-proposed standard RBAC model \cite{ferraiolo}
-
-Role-based access control (RBAC) models \cite{ferraiolo} became popular used to govern access to critical resources.
+Role based access control (RBAC) was first introduced in the 1990s when the National Instittute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.
-
 RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.
@@
 
-Standard RBAC model considers only role-user association and role hierarchy.
+Since the introduction of the standardized RBAC model, innovation and the spread of software engineering into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.
+For example, RBAC is ill-equipped to handle the additional entities that need to be taken into consideration during authorization with the introduction of privacy concerns in domains such as healthcare  Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.
+
 Since standard RBAC model has limitations such as specifying environmental constraints or context information
@@
 
-The goal of this work is to synthesize available research results on extended models of RBAC.
-We analyze their extended features and claimed research contributions to find limitations of current RBAC models and what extent of extended features by
-comparing with similar research work.
+The goal of this work is to provide practioners an assessment of the state of extended models of RBAC..
+
 We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
 
-Our research give benefits to a community as follows:
+Our research provides contributions to the community by:
 
 \begin{itemize}
-\item Our work summarizes current extended RBAC research work and its contributions. By synthesizing the current results, our work shows a roadmap of current extended RBAC research.
-\item Our work guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
+\item Summarizes current extended RBAC research work and its contributions
+\item Guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
 \item Our work shows a criteria in comparison among research results.
-\item Our work helps identify the research challenges in the ares of security policies and suggest a future extension of RBAC.
+\item Identify the research challenges in the areas of security policies and suggest a future extension of RBAC
+\item Identify the deficiencies in RBAC extension evaluation
 \end{itemize}
+
+The rest of the paper is organized as follows.  Section 2. Section 3. Section 4. Section 5.

### 2013-02-17 paper/sections/process.tex
Adds updated bibliography.
Adding some updated writing and notes on each section and what to put.
Updates to results, process.

@@
-\section{Methodology and Process} \label{sec:process}
+\section{Process} \label{sec:process}
 
-The systematic literature review process was developed ahead of time and agreed upon by the researchers following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
-The systematic literature review was performed in four stages:
+The systematic literature review process was developed in whole prior to application and agreed upon by the researchers following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
+The systematic literature review process was broken down in to four stages and the rest of this section is broken down by each stage.
+The stages were:
 
@@
 
-For the first phase of our systematic literature review, an automated comprehensive search of multiple academic search engines was performed. The list of search engines were:
-
-\begin{itemize}
-\item Google Scholar - \url{http://scholar.google.com}
-\item IEEExplore - \url{http://ieeexplore.ieee.org}
-\item ACM Portal - \url{http://dl.acm.org}
-\item CiteSeerX - \url{http://http://citeseerx.ist.psu.edu/index}
-\end{itemize}
-
-For each of the criteria below, a search was performed on each of the search engines for a total of 12 data sets.  The criteria used were:
-\begin{itemize}
-\item role based access control
-\item RBAC
-\item role-based access control
-\end{itemize}
-
+For the first phase of our systematic literature review, a search strategy for finding papers was developed.
+The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search criteria.
 The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.
 For each criteria for each search engine, the results were captured until a stopping criteria was met.
-Each run was performed as follows:
+The search algorithm was performed as follows:
 
@@
 \item Compare results against stopping criteria.
-\end{enumerate}
-
 	\begin{itemize}
@@
 	\end{itemize}
+\end{enumerate}
 
@@
 After gathering all 12 data sets, the data was combined into a master list by systematically comparing the bibliographic information for each.
-After producing a master list, a series of assessment rounds were performed to narrow the paper list and identify primary sources.
+After producing a master list, a series of elimination rounds were performed to narrow the list of papers and identify primary sources.
 
@@
 \begin{itemize}
-\item Each reviewer independently classified papers as relevant, irrelevant or uncertain.
+\item Each reviewer independently classified papers as relevant, irrelevant or uncertain based off elimination criteria
 \item Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
@@
 
-The title elimination round was based off of whether role based access control and model were mentioned directly.
+\subsubsection{Title Elimination}
 
-The second round of elimination was based off on reading the abstracts of the remaining papers.  The inclusion criteria for the abstract reading tried to answer the following questions:
+The first round of elimination was performed by strict examination of the title.  Each researcher was tasked with deciding on elimination by answering the following questions:
+
+\begin{itemize}
+\item Did the title contain a reference to 'role based access control' or 'RBAC'? 
+\item Did the title contain a reference to 'model'?
+\end{itemize}
+
+\subsubsection{Abstract Elimination}
+
+The second round of elimination was based on strict reading of the abstracts of papers that survived title elimination.  Researchers read each abstract and evaluated relevancy based off:
 
@@
 
-The results of the searches is summarised below:
+\subsubsection{Content Elimination}
 
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\textbf{Search Engines} & 
-\textbf{RBAC} & 
-\textbf{role based access control} & 
-\textbf{role-based access control} & 
-\textbf{Total}
-\\\hline
+The final elimination round involved taking the entire paper into consideration and answering five questions that would serve as the basis for elimination.
+The data collected by answering these questions served as the basis towards answering the research questions.
+The questions each reviewer attempted to answer based on the content of the paper was:
 
-Google Scholar & 651 & 213 & 435 & 1299
-\\\hline
-ACM Portal & 500 & 20 & 720 & 1240
-\\\hline
-IEEExplore & 200 & 40 & 230 & 470
-\\\hline
-CiteSeerX & 100 & 100 & 150 & 350
-\\\hline
- &  &  &  & 
-\\\hline
-Totals & 1451 & 373 & 1535 & 3359
-\\\hline
-
-Combined &  &  &  & \textbf{1716}
-\\\hline
-\end{tabular}
+\begin{itemize}
+\item Does this model extend the core model?
+\item What reasons and evidence do researchers give that RBAC needs extension?
+\item Was the paper and subsequent model inspired by a real world example?
+\item Is there any evaluation of the proposed model? If yes, how did they do one? If no, why?
+\item Did the authors implement their model?
+\end{itemize}

### 2013-02-17 paper/sections/results.tex
Updates to results, process.

@@
 \section{Results} \label{sec:results}
+
+For our systematic literature review, the search strategy was applied to the following search engines based on popularity and database size:
+
+\begin{itemize}
+\item Google Scholar - \url{http://scholar.google.com}
+\item IEEExplore - \url{http://ieeexplore.ieee.org}
+\item ACM Portal - \url{http://dl.acm.org}
+\item CiteSeerX - \url{http://http://citeseerx.ist.psu.edu/index}
+\end{itemize}
+
+For each of the criteria below, a search was performed using the exact query for each of the search engines for a total of 12 data sets.  The criteria used were:
+
+\begin{itemize}
+\item role based access control
+\item RBAC
+\item role-based access control
+\end{itemize}
+
+The results of the search are summarised below:
+
+\begin{center}
+\begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
+
+\hline
+\textbf{Search Engines} & 
+\textbf{RBAC} & 
+\textbf{role based access control} & 
+\textbf{role-based access control} & 
+\textbf{Total}
+\\\hline
+
+Google Scholar & 651 & 213 & 435 & 1299 \\\hline
+ACM Portal & 500 & 20 & 720 & 1240 \\\hline
+IEEExplore & 200 & 40 & 230 & 470 \\\hline
+CiteSeerX & 100 & 100 & 150 & 350 \\\hline
+ &  &  &  & \\\hline
+Totals & 1451 & 373 & 1535 & 3359 \\\hline
+Combined &  &  &  & \textbf{1716} \\\hline
+
+\end{tabular}
+\end{center}
 
@@
 S5: RQ5 and discuss by category and general results
+
+\subsection{Extraction}
+
+P1: Describe data extraction process, including questions each reviewer based note taking off of
+
+P2: Describe how from the data extraction process cateogories of extension became evident
+
+P3: Describe how those categories were deemed relevant and a way to organize going forward
+
+P4: Describe the categories, show how each paper and their category
+
+P5: Discuss how the categories are inter-related and show diagram of relationships
+
+After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends.
+The first step was to take the individual data generated from the final elimination round and organize this information around the research questions.
+During the paper reading round, and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
+Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.
+The result of this paper categorization is displayed below:
+
+\begin{center}
+\begin{tabular}{|p{12.5cm}|p{3cm}|}
+\hline
+\textbf{Paper} & \textbf{Category}
+\\\hline
+Alam, M. and Hafner, M. and Breu, R. 2006 \cite{alam06:constraint} & Constraint \\\hline
+Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 \cite{tzelepi01:flexible} & Context \\\hline
+Haibo, SHEN and Fan, HONG 2005 \cite{haibo05:context} & Context \\\hline
+Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
+Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 \cite{huang06:pervasive} & Context \\\hline
+Motta, G.H.M.B. and Furuie, S.S. 2003 \cite{motta03:contextual} & Context \\\hline
+Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 \cite{bao08:role} & Context \\\hline
+Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 \cite{yamazaki104:designing} & Context \\\hline
+Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 \cite{han08:extended} & Context \\\hline
+Thein, N. and others 2011 \cite{thein2011leveraging} & Context \\\hline
+Zou, D. and He, L. and Jin, H. and Chen, X. 2009 \cite{zou2009crbac} & Context \\\hline
+Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 \cite{hasebe10:capability} & Delegation \\\hline
+Zhang, Z. and Zhang, X. and Sandhu, R. 2006 \cite{zhang06:collaborative} & Organizational \\\hline
+Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 \cite{ni2010privacy} & Privacy \\\hline
+Masoumzadeh, A. and Joshi, J. 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
+Zhao, Y. and Zhao, Y. and Lu, H. 2008 \cite{zhao2008flexible} & Resource \\\hline
+Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 \cite{damian2007geo} & Spatial \\\hline
+Ray, I. and Kumar, M. and Yu, L. 2006 \cite{ray07:spatio} & Spatial \\\hline
+Hansen, F. and Oleshchuk, V. 2003 \cite{hansen2003spatial} & Spatial \\\hline
+Aich, S. and Sural, S. and Majumdar, A. 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
+Chen, L. and Crampton, J. 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
+Samuel, A. and Ghafoor, A. and Bertino, E. 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
+Chandran, S. and Joshi, J. 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
+Ray, I. and Toahchoodee, M. 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
+Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
+Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
+ZHANG, S. and CHEN, X. and HOU, G. 2009 \cite{zhou2007team} & Task \\\hline
+Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
+Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline
+\end{tabular}
+\end{center}
+
+Given that there were multiple papers for some categories, the researchers decided to tackle all further research questions by first analyzing the research question on a per category basis and then looking across all categories for generalization and trends.

### 2013-02-17 paper/sections/results.tex
Adds analysis section for research questions.

@@
 \end{center}
-
-P1: Discuss focus on going through each research question individually, and combining results across at the end of the process
-
-P2: List the research questions
-
-S1: RQ1 and discuss by category and general results
-
-S2: RQ2 and discuss by category and general results
-
-S3: RQ3 and discuss by category and general results
-
-S4: RQ4 and discuss by category and general results
-
-S5: RQ5 and discuss by category and general results
 

### 2013-02-19 paper/rbac-slr.tex
Some minor updates and the start of a data extraction section.
Adding terms section.
Updates to results, process.
Adds analysis section for research questions.
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

@@
-\documentclass[preprint,12pt]{elsarticle}
+\documentclass[3p,12pt]{elsarticle}
 
 \usepackage{amssymb}
+\usepackage{hyperref}
 
@@
 
+\section{Outline}
+
+\begin{enumerate}
+\item Introduction
+    \begin{itemize}
+  	\item RBAC history, RBAC as a standard
+  	\item Audience that should care about our review
+  	\item Brief explanation of what has led to extensions cropping up
+  	\item Paper organization
+    \end{itemize}
+\item Background
+    \begin{itemize}
+   	\item What is RBAC
+   	\item Core RBAC and it's entities, and what it tries to solve
+    \end{itemize}
+\item Process
+\item Results
+    \begin{itemize}
+  	\item Data used in process, results of process for paper count
+  	\item Categorization that came out applying process (based on word usage)
+  	\item Results for each research question (raw data)
+    \end{itemize}
+\item Analysis
+    \begin{itemize}
+  	\item Explanation and definitions of categories based on results
+  	\item Examination of results in relation to each research question and larger trends drawn from data
+  	\item Examination of cross research question concerns an
+    \end{itemize}
+\item Conclusion
+\end{enumerate}
+
 %% SECTIONS %%
 \input{sections/intro}
+\input{sections/terms}
 \input{sections/process}
+\input{sections/results}
+\input{sections/analysis}
 

### 2013-02-19 paper/sections/analysis.tex
Adds analysis section for research questions.
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

--- 2013-02-09 paper/sections/results.tex
@@
-\section{Results} \label{sec:results}
+\section{Analysis} \label{sec:analysis}
 
-P1: Discuss focus on going through each research question individually, and combining results across at the end of the process
+\subsection{Research Question 5}
 
-P2: List the research questions
+Do the extension models have corresponding implementations in practice? \\
 
-S1: RQ1 and discuss by category and general results
+When designing and proposing a model targeted at a feature that is rooted in practical
+usage by real software systems, bringing the model to life is strong evidence that the
+proposed model can work in practice.  The concept of authorization, and access control
+is rooted in a business need.  Thus, any access control model needs to be feasible
+in the real world not just on paper.  We analyzed the primary sources to see how many
+proposed models actually had implmenetations associated with them.  And quantified the
+type of implementation.  Whether the implementation was for a real system, for a prototype
+and/or used in a production environment.
 
-S2: RQ2 and discuss by category and general results
+Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
+mention of a implementation or prototype.  Of the remianing 8 papers that did mention an implementation, half 
+were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
+system.
 
-S3: RQ3 and discuss by category and general results
+\subsection{Research Question 6}
 
-S4: RQ4 and discuss by category and general results
+How are extended RBAC models evaluated in theory and in practice? \\
 
-S5: RQ5 and discuss by category and general results
+Providing evaluation of a proposed model is a key component in establishing the models validity.  The papers were examined
+for evidence of evaluations ranging from performance to mathematical accuracy to application to real world scenarios.  Further,
+for each proposed model, the reviewers looked for evidence that the authors made comparisons between their own model and the base
+model as they pertained to claims made by the authors of why their model is needed.  The quantifiable evaluations looked for were:
+
+\begin{itemize}
+\item Time-based Performance
+\item Complexity analysis
+\item Comparison to standard RBAC
+\item Mathematical accuracy
+\item Example scenarios of the model in action
+\item Experimental analysis of the model
+\item Case study of the model in practice
+\end{itemize}
+
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
+and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
+or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
+grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
+and analsysis as a way to provide evaluation in the form of completeness.
+
+The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
+would tackle those scenarios.  Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
+or show how the standard model is deficient in tackling said scenarios.
+
+\subsection{Research Question X}
+
+What domains or scenarios serve as inspiration for these extensions? \\
+
+Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
+model extensions has been the presentation of business scenarios in various domains and how the model
+uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
+scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
+RBAC extensions.
+
+Upon examination of the primary sources, the most prevalent domains were:
+
+\begin{itemize}
+\item What domains or scenarios serve as inspiration for these extensions?
+\item Medical domain
+\item Pervasive computing environments
+\item Mobile devices
+\item Large-scale organizations with many sub-departments
+\item Enterprise, organization workflows
+\end{itemize}

### 2013-02-19 paper/sections/intro.tex
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

@@
 For example, RBAC is ill-equipped to handle the additional entities that need to be taken into consideration during authorization with the introduction of privacy concerns in domains such as healthcare  Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.
+However, as researchers often develop their own specialized extended models of RBAC, their research cannot be generalized or compared with other research work appropriately.
 
-Since standard RBAC model has limitations such as specifying environmental constraints or context information
-Researchers developed extended models of RBAC to overcome the limitations.
-However, as researchers often develop their own specialized extended models of RBAC,
-their research cannot be generalized or compared with other research work appropriately.
-As a result, researchers could take time on reinventing the wheel.
-But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?
+A systematic literature review is designed to yield quanitfiable results across a body of research in order to draw out similarities, trends and defiencies in an area as a whole.
+The review we have performed should serve as a future starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
+For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.
 
-The goal of this work is to provide practioners an assessment of the state of extended models of RBAC..
+The goal of this work is to provide practioners an assessment of the state of extended models of RBAC\ldots
 
-We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.
-
-Our research provides contributions to the community by:
+Our research provides contributions to the community in the following ways:
 

### 2013-02-20 paper/sections/abstract.tex
Cleans up outlines, adds numerous updates to various sections.

@@
 
-Role-based access control (RBAC) has become an increasingly popular access
-control for various applications such as web applications and
-database applications. RBAC restricts access to resources based on identity of subjects and/or groups
-called roles. 
-Since RBAC is introduced in 1998, researchers have proposed various extended models of RBAC.
-For example, they define additional constraints among roles (e.g., temporal constraints or location constraints) or hierarchy relationship of roles.
-Our goal of this work is to study extended models of RBAC and analyze their extended features and claimed research contributions
-to find limitations of current RBAC models and what extent of extended features that can be used for future RBAC.
-We conduct a systematic literature review by collecting and synthesizing relevant research
-papers. We initially collect XXXX papers from various sources such as IEEE and ACM websites and selected 26 papers systemically.
-We perform a comparative analysis to find relationships among extended models and RBAC.
+Since the introduction of the Role-based access control (RBAC) standard in the late 1990's, 
+RBAC has become an increasingly popular access control mechanism for enterprises such as 
+web applications, mobile computing and databases. RBAC restricts access to resources 
+based on the identity of subjects connected to logical groupings of permissions called roles.  
+The introduction of computing into new domains has led to proposals for extensions to the standard
+RBAC model; for example, defining additional constraints among roles, adding context 
+or defining new hierarchy relationships.
+\textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+We conducted a systematic literature review by collecting and synthesizing relevant research
+papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
+websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
+find relationships among extended models and analyze the state of RBAC extension evaluations.
 

### 2013-02-20 paper/sections/core_rbac.tex
Cleans up outlines, adds numerous updates to various sections.

--- 2013-02-09 paper/sections/terms.tex
@@
 \section{Core Role Based Access Control} \label{sec:core-rbac}
-
-P1: Describe RBAC entities
-
-P2: Note the 4 levels of RBAC
-
-P3: Concisely describe level 1
-
-P4: Concisely describe level 2
-
-P5: COncisely describe level 3
-
-P6: Concisely describe level 4
 
@@
 
-\section{Terms and Definitions} \label{sec:terms}
+P1: Describe RBAC entities
 
-NOTE: Not sure about this section since some of this is derived from our results
+P2: Note the 4 levels of RBAC
 
-We found that different definitions for the same terms. Therefore, we next describe
-terms and definitions.
+P3: Concisely describe level 1
 
-\begin{itemize}
-%	\item Access Control Model:
-%	\item Access Control Policy:
-%	\item Access Control Requirements:
-%	\item Rules:
-%	\item Attributes-Based Access Control:
-	\item Task-Based Access Control:
-%	\item Purpose-Based Access Control:
-	\item Agent-Based Access Control:						
-\end{itemize}
+P4: Concisely describe level 2
 
-\begin{itemize}
+P5: COncisely describe level 3
 
-%	\item Role Hierarchies:
-
-	\item Obligations: obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
-	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
-	her/his office duty before accessing the resources.
-
-	\item Inheritance:
-	Inheritance defines an inheritance relationship among attributes such as roles. For example, the role structure for a company use
-	employee role for employees. Department manager may inherit all permissions of the employee role. (Role-Based Access Control by F. Ferrailolo et al.)
-
-	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
-	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$.
-SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.
-(Role-Based Access Control by F.
-Ferrailolo et al.)
-
-	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
-more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
-	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.
-(Role-Based Access Control by F.
-Ferrailolo et al.)
-	\end{itemize}
-
-We define temporal and spatial constraints as follows:
-
-\begin{itemize}
-	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
-	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
-	In such cases, the part-time employee role should access required resources during the interval. 
-	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.   
-	(Role-Based Access Control by F. Ferrailolo et al.)
-
-	\item Spatial Constraints: Spatial constraints are location-based constraints in specifying access
-	control policies. For example, in organizations, 	locations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
-	In such cases, the part-time employee role should access required resources only when the user is in the location. 
-	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
-
-\end{itemize}
+P6: Concisely describe level 4

### 2013-02-20 paper/sections/intro.tex
Cleans up outlines, adds numerous updates to various sections.

@@
 \section{Introduction} \label{sec:introduction}
-
-Why is the base model of RBAC extended by newer models?
-
-P1: Note history of the RBAC model, statistics, and official status
-
-P2: Mention extensions exist in response to developments and new domains over time
-
-P3: Mention audience, why they should care, and why exploring extensions space is important in relation to the reason the standard was originally concieved
-
-P4: Talk about what this work is going to do, goal, brief process
-
-P5: List contributions
-
-P6: Outline paper and sections
 
@@
 
+\textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+
 A systematic literature review is designed to yield quanitfiable results across a body of research in order to draw out similarities, trends and defiencies in an area as a whole.
-The review we have performed should serve as a future starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
+The review we performed yielded 1716, of which XX were deemed primary sources for inclusion as model extensions to the RBAC standard model.
+This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
 For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.
-
-The goal of this work is to provide practioners an assessment of the state of extended models of RBAC\ldots
+Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.
 
@@
 
-The rest of the paper is organized as follows.  Section 2. Section 3. Section 4. Section 5.
+The rest of the paper is organized as follows.
+Section 2 covers background and the core RBAC model.
+Section 3 lays out the process used in conducting the review.
+Section 4 provides the results of the search.
+Section 5 tackles the research questions and analyzes the results.
+Section 6 contains the final conclusions.
+
+\begin{itemize}
+\item TODO: Add some references to the prevalence of RBAC in industry
+\item TODO: Add link to RBAC standard documents
+\item TODO: List research questions here or in process
+\end{itemize}

### 2013-02-20 paper/sections/results.tex
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.
Cleans up outlines, adds numerous updates to various sections.

@@
 
-For our systematic literature review, the search strategy was applied to the following search engines based on popularity and database size:
+\subsection{Search}
 
-\begin{itemize}
-\item Google Scholar - \url{http://scholar.google.com}
-\item IEEExplore - \url{http://ieeexplore.ieee.org}
-\item ACM Portal - \url{http://dl.acm.org}
-\item CiteSeerX - \url{http://http://citeseerx.ist.psu.edu/index}
-\end{itemize}
+For our systematic literature review, the search strategy was applied to four search engines and three search strings based on popularity and database size.
+This produced 12 different data sets that are combining for overlapping papers resulted in 1716 candidates.
+The paper count for each search engine and criteria are shown in Table \ref{tab:search_results}.
 
-For each of the criteria below, a search was performed using the exact query for each of the search engines for a total of 12 data sets.  The criteria used were:
-
-\begin{itemize}
-\item role based access control
-\item RBAC
-\item role-based access control
-\end{itemize}
-
-The results of the search are summarised below:
-
-\begin{center}
+\begin{table}
+\centering
 \begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
@@
 \end{tabular}
-\end{center}
+\caption{Paper counts after applying search strategy}
+\label{tab:search_results}
+\end{table}
+
+TODO: Add metrics for each portion of the elimination
 
 \subsection{Extraction}
-
-P1: Describe data extraction process, including questions each reviewer based note taking off of
-
-P2: Describe how from the data extraction process cateogories of extension became evident
-
-P3: Describe how those categories were deemed relevant and a way to organize going forward
-
-P4: Describe the categories, show how each paper and their category
-
-P5: Discuss how the categories are inter-related and show diagram of relationships
 
@@
 Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.
-The result of this paper categorization is displayed below:
+Each paper deemed a primary source and the papers associated categorization are shown in Table \ref{tab:categorization}.
 
-\begin{center}
+\begin{table}
+\centering
 \begin{tabular}{|p{12.5cm}|p{3cm}|}
@@
 Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline
+
 \end{tabular}
-\end{center}
+\caption{Primary sources grouped by categorization}
+\label{tab:categorization}
+\end{table}
 
 Given that there were multiple papers for some categories, the researchers decided to tackle all further research questions by first analyzing the research question on a per category basis and then looking across all categories for generalization and trends.
+
+\section{Category Definitions} \label{sec:terms}
+
+We found that different definitions for the same terms. Therefore, we next describe
+terms and definitions.
+
+\begin{itemize}
+%	\item Access Control Model:
+%	\item Access Control Policy:
+%	\item Access Control Requirements:
+%	\item Rules:
+%	\item Attributes-Based Access Control:
+	\item Task-Based Access Control:
+%	\item Purpose-Based Access Control:
+	\item Agent-Based Access Control:						
+\end{itemize}
+
+\begin{itemize}
+
+%	\item Role Hierarchies:
+
+	\item Obligations: obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
+	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
+	her/his office duty before accessing the resources.
+
+	\item Inheritance:
+	Inheritance defines an inheritance relationship among attributes such as roles. For example, the role structure for a company use
+	employee role for employees. Department manager may inherit all permissions of the employee role. (Role-Based Access Control by F. Ferrailolo et al.)
+
+	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
+	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$.
+SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.
+(Role-Based Access Control by F.
+Ferrailolo et al.)
+
+	\item Dynamic Separation of Duty: Dynamic SoD (DSD) is known to be
+more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
+	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.
+(Role-Based Access Control by F.
+Ferrailolo et al.)
+	\end{itemize}
+
+We define temporal and spatial constraints as follows:
+
+\begin{itemize}
+	\item Temporal Constraints: Temporal constraints are time-based constraints in specifying access
+	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
+	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
+	In such cases, the part-time employee role should access required resources during the interval. 
+	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.   
+	(Role-Based Access Control by F. Ferrailolo et al.)
+
+	\item Spatial Constraints: Spatial constraints are location-based constraints in specifying access
+	control policies. For example, in organizations, 	locations are enforced while a
+	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
+	In such cases, the part-time employee role should access required resources only when the user is in the location. 
+	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
+
+\end{itemize}

### 2013-02-20 paper/sections/analysis.tex
Cleans up outlines, adds numerous updates to various sections.
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

@@
 
-\subsection{Research Question 5}
+\subsection{Research Question 1}
+
+What are the deficiencies in the current RBAC model? \\
+
+Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
+permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
+security mechanism, a variety of extended RBAC models have been proposed over the years to meet
+security requirements. Our surveyed 29 papers address more than one deficiencies described as follows.
+
+\begin{itemize}
+%	\item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
+	\item Core RBAC does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
+	\item Core RBAC does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
+	\item Core RBAC does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
+	\item Core RBAC has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
+\end{itemize}
+
+\subsection{Research Question 2}
+
+What are the motivations behind RBAC extensions? \\
+
+When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
+Administrators improve existing core RBAC model due to following reasons:
+
+\begin{itemize}
+	\item Model needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
+	\item Model needs to incorporate additional attributes such as
+	task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new association between attributes
+	are introduced. 
+		\item Model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
+\end{itemize}
+
+\subsection{Research Question 3}
 
@@
 in the real world not just on paper.  We analyzed the primary sources to see how many
-proposed models actually had implmenetations associated with them.  And quantified the
+proposed models actually had implementations associated with them.  And quantified the
 type of implementation.  Whether the implementation was for a real system, for a prototype
@@
 Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
-mention of a implementation or prototype.  Of the remianing 8 papers that did mention an implementation, half 
+mention of a implementation or prototype.  Of the remaining 8 papers that did mention an implementation, half 
 were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
@@
 
-\subsection{Research Question 6}
+\subsection{Research Question 4}
 
@@
 grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
-and analsysis as a way to provide evaluation in the form of completeness.
+and analysis as a way to provide evaluation in the form of completeness.
 
@@
 
-\subsection{Research Question X}
+\subsection{Research Question 5}
 
@@
 \end{itemize}
+
+\subsection{Research Question 6}
+
+What commonalities or generalizations exist across all categorizations? \\

### 2013-02-20 paper/sections/intro.tex
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

@@
 
-Role based access control (RBAC) was first introduced in the 1990s when the National Instittute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.
+Role based access control (RBAC) was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 
-\textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+\textit{The goal of this work is to provide practitioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
 
-A systematic literature review is designed to yield quanitfiable results across a body of research in order to draw out similarities, trends and defiencies in an area as a whole.
+A systematic literature review is designed to yield quantifiable results across a body of research in order to draw out similarities, trends and deficiencies in an area as a whole.
 The review we performed yielded 1716, of which XX were deemed primary sources for inclusion as model extensions to the RBAC standard model.

### 2013-02-20 paper/sections/results.tex
adds RQ 1 and 2, new terms, and correct typos
add research questions 1 and 2.
add some of new terms
change/correct typos (in intro and other sections)

@@
 Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
-ZHANG, S. and CHEN, X. and HOU, G. 2009 \cite{zhou2007team} & Task \\\hline
+Zhang, S. and Chen, X. and Hou, G. 2009 \cite{zhou2007team} & Task \\\hline
 Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
@@
 %	\item Attributes-Based Access Control:
-	\item Task-Based Access Control:
+
+%http://www.hpi.uni-potsdam.de/fileadmin/hpi/Forschung/Publikationen/Dissertationen/Diss_Zhou.pdf
+% A Task-Role Based Access Control Model with Multi-Constraints
+	\item Task-role-based access control: a task is a fundamental unit of a business activity.
+Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
+Roles are first associated with tasks, which are associated permissions.
+For example, the employee role is associated with a task, which is to write a report.
+Then, this task is associated with a permission.
+
+% Team and Task Based RBAC Access Control Model	
+	\item Team-based access control: a team encapsulates a group of users, who have various roles.
+A team member can activate only roles that can be assigned to a team.
+Moreover, team-based access control may decide maximum number of permissions such as roles (that are assigned to a team) having a maximum number of permissions to the team.
 %	\item Purpose-Based Access Control:
-	\item Agent-Based Access Control:						
+%	\item Agent-based access control:						
 \end{itemize}
@@
 %	\item Role Hierarchies:
+
+	\item Delegation: delegation is to assign permissions (to access specific resources) available to a user to another user.
+	A employee $E_1$ may delegate her/his permissions to another employee $E_2$ while $E_1$ is on leave.
+
+%	 obligations are requirements, which should be fulfilled before or after authorization decision is enforced.
+%	Consider that a user has permission to access specific resources. For example, obligation is that the user should complete
+%	her/his office duty before accessing the resources.
 
@@
 SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.
-(Role-Based Access Control by F.
-Ferrailolo et al.)
 
@@
 	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.
-(Role-Based Access Control by F.
-Ferrailolo et al.)
 	\end{itemize}
@@
 	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
-	(Role-Based Access Control by F. Ferrailolo et al.)
 

### 2013-02-27 paper/sections/abstract.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

@@
 
-Since the introduction of the Role-based access control (RBAC) standard in the late 1990's, 
-RBAC has become an increasingly popular access control mechanism for enterprises such as 
+\noindent
+{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's, 
+RBAC has become an increasingly popular access control mechanism for applications such as 
 web applications, mobile computing and databases. RBAC restricts access to resources 
@@
 The introduction of computing into new domains has led to proposals for extensions to the standard
-RBAC model; for example, defining additional constraints among roles, adding context 
-or defining new hierarchy relationships.
-\textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
-We conducted a systematic literature review by collecting and synthesizing relevant research
+RBAC model; for example, defining additional constraints among roles, adding context or defining new heirarchy relationships.
+\\
+\\
+\noindent
+{\bf Aim}: \textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+\\
+\\
+\noindent
+{\bf Method}: We conducted a systematic literature review by collecting and synthesizing relevant research
+papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
+websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
+find relationships among extended models and analyze the state of RBAC extension evaluations.
+\\
+\\
+\noindent
+{\bf Results}:
+\\
+\\
+\noindent
+{\bf Conclusions}: We conducted a systematic literature review by collecting and synthesizing relevant research
 papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 

### 2013-02-27 paper/sections/analysis.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

@@
 
-What are the deficiencies in the current RBAC model? \\
-
-Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
-permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
-security mechanism, a variety of extended RBAC models have been proposed over the years to meet
-security requirements. Our surveyed 29 papers address more than one deficiencies described as follows.
+What categorizations exist within RBAC extensions? \\
 
 \begin{itemize}
-%	\item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
-	\item Core RBAC does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
-	\item Core RBAC does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
-	\item Core RBAC does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
-	\item Core RBAC has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
+
+  \item Constraint:
+
+  \item Context:
+
+  \item Organizational:
+
+  \item Privacy:
+
+  \item Resource:
+
+  \item Spatial:  Spatial constraints are location-based constraints in specifying access
+	control policies. For example, in organizations, 	locations are enforced while a
+	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
+	In such cases, the part-time employee role should access required resources only when the user is in the location. 
+	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
+
+  \item Spatio-Temporal:
+
+  \item Task: A task is a fundamental unit of a business activity.
+Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
+Roles are first associated with tasks, which are associated permissions.
+For example, the employee role is associated with a task, which is to write a report.
+Then, this task is associated with a permission.
+
+  \item Temporal:  Temporal constraints are time-based constraints in specifying access
+	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
+	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
+	In such cases, the part-time employee role should access required resources during the interval. 
+	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
+
 \end{itemize}
@@
 
-When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
-Administrators improve existing core RBAC model due to following reasons:
+Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
+permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
+security mechanism, a variety of extended RBAC models have been proposed over the years to meet
+security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
 
 \begin{itemize}
+\item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
+\item Core RBAC does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
+\item Core RBAC does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
+\item Core RBAC does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
+\item Core RBAC has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
 \item Model needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
@@
 
-Do the extension models have corresponding implementations in practice? \\
+Do the extension models have corresponding implementations? \\
 
@@
 
-How are extended RBAC models evaluated in theory and in practice? \\
+How are extension to RBAC evaluated theoritically and in practice? \\
 
@@
 \begin{itemize}
-\item Time-based Performance
-\item Complexity analysis
-\item Comparison to standard RBAC
-\item Mathematical accuracy
-\item Example scenarios of the model in action
+\item Time-based Performance \cite{ni2010privacy}, \cite{aich09:role}
+\item Complexity analysis \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{chen08:spatio-temporal}, \cite{aich09:role}
+\item Comparison to standard RBAC \cite{bao08:role}, \cite{zou2009crbac}, \cite{zhang06:collaborative}, \cite{zhao2008flexible}, \cite{ray07:spatio}
+\item Mathematical modeling \cite{damian2007geo}, \cite{hansen2003spatial}, \cite{aich07:STARBAC}, \cite{chen08:spatio-temporal}, \cite{joshi05:generalized}
+\item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki104:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{ray07:spatio}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
 \item Experimental analysis of the model
-\item Case study of the model in practice
+\item Case study of the model in practice \cite{motta03:contextual}
 \end{itemize}
@@
 
+TODO: Add references to list above.
+
 \subsection{Research Question 5}
 
-What domains or scenarios serve as inspiration for these extensions? \\
+What domains have RBAC extensions been created for? \\
 

### 2013-02-27 paper/sections/appendix.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

\section{Appendix} \label{sec:appendix}

\begin{table}
\centering
\begin{tabular}{|p{12.5cm}|p{3cm}|}
\hline
\textbf{Paper} & \textbf{Category}
\\\hline
A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint} & Constraint \\\hline

A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible} & Context \\\hline

A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline

A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive} & Context \\\hline

A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual} & Context \\\hline

A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role} & Context \\\hline

An extended RBAC model based on granular logic, 2008 \cite{jian2008extended} & Context \\\hline

Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing} & Context \\\hline

An extended RBAC model based on granular logic, 2008 \cite{han08:extended} & Context \\\hline

Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging} & Context \\\hline

CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac} & Context \\\hline

ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative} & Organizational \\\hline

Privacy-aware role-based access control, 2007 \cite{ni2010privacy} & Privacy \\\hline

PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline

A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible} & Resource \\\hline

GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo} & Spatial \\\hline

LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio} & Spatial \\\hline

Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial} & Spatial \\\hline

STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline

On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline

A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline

LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline

A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline

Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role} & Spatio-Temporal \\\hline

A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task} & Task \\\hline

Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team} & Task \\\hline

Task-role-based access control model, 2003 \cite{oh2003task} & Task \\\hline

A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized} & Temporal \\\hline

\end{tabular}
\caption{Primary sources grouped by categorization}
\label{tab:categorization}
\end{table}

### 2013-02-27 paper/sections/process.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

@@
 
-The systematic literature review process was developed in whole prior to application and agreed upon by the researchers following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
+The authors adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
 The systematic literature review process was broken down in to four stages and the rest of this section is broken down by each stage.
@@
 \begin{itemize}
-\item Development of a search strategy
-\item Elimination of papers based on title
-\item Elimination of papers based on abstract
-\item Elimination of papers based on content and matching to elimination criteria
+\item Step 1: Development of a search strategy
+\item Step 2: Elimination of papers based on title criteria
+\item Step 3: Elimination of papers based on abstract criteria
+\item Step 4: Elimination of papers based on content and matching to elimination criteria
 \end{itemize}
 
-\subsection{Search Strategy}
+\subsection{Step 1: Search Strategy}
 
-For the first phase of our systematic literature review, a search strategy for finding papers was developed.
-The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search criteria.
-The search performed was done in an automated way using a set of scripts to query and collect data from each search engine with the criteria string as input.
-For each criteria for each search engine, the results were captured until a stopping criteria was met.
+For the first phase of our systematic literature review, the authors developed a search strategy for finding papers.
+The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search terms.
+The search was performed by applying each search term to each engine incrementally until a stopping criteria was met.
+Table \ref{tab:search_results} lists the four search engines along the left most column with the three search term across the top row along with the papers for each criteria and engine combination.
 The search algorithm was performed as follows:
@@
 \begin{enumerate}
-\item Remote call to search engine with current search start position and the current search criteria.
+\item Call to search engine with current search position and current search term.
 \item Parse results and extract paper title, authors and year of publication.
-\item Compare results against stopping criteria.
+\item Compare results against stopping criteria:
 	\begin{itemize}
-	\item If stopping criteria met, stop search.
-	\item If stopping criteria not met, increase search position by number of results and return to step 1.
+	\item If the size of the result set is greater than or equal to 1000 then stop.
+	\item If the last ten results did not contain the search term phrase within the tite then stop.
 	\end{itemize}
+\item If stopping criteria not met, increment search position and go back to step one.
 \end{enumerate}
 
-The stopping criteria used was either after the first 1000 results, a limitation imposed by some of the search engines, or if ten consecutive results did not contain the search criteria phrase within the title.
-After gathering all 12 data sets, the data was combined into a master list by systematically comparing the bibliographic information for each.
-After producing a master list, a series of elimination rounds were performed to narrow the list of papers and identify primary sources.
+The result set size stopping criteria was chosen due to a technical limitation of some search engines.
+The stopping criteria related to the last ten titles is meant to stop after relevant results are no longer being returned by the search engine.
+After gathering all 12 data sets, the authors combined the papers into a master list by systematically comparing the bibliographic information for each.
+The master list of 1716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
 
-\subsection{Elimination Rounds}
+\begin{table}
+\centering
+\begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
+
+\hline
+ & 
+\textbf{RBAC} & 
+\textbf{role based access control} & 
+\textbf{role-based access control} & 
+\textbf{Total}
+\\\hline
+
+Google Scholar & 651 & 213 & 435 & 1299 \\\hline
+ACM Portal & 500 & 20 & 720 & 1240 \\\hline
+IEEExplore & 200 & 40 & 230 & 470 \\\hline
+CiteSeerX & 100 & 100 & 150 & 350 \\\hline
+ &  &  &  & \\\hline
+Totals & 1451 & 373 & 1535 & 3359 \\\hline
+Combined &  &  &  & \textbf{1716} \\\hline
+
+\end{tabular}
+\caption{Paper counts after applying search strategy}
+\label{tab:search_results}
+\end{table}
+
+TODO: Add elimination stats.
+
+\subsection{Steps 2-4: Elimination Rounds}
 
@@
 \begin{itemize}
-\item Each reviewer independently classified papers as relevant, irrelevant or uncertain based off elimination criteria
+\item The two first authors independently classified papers as relevant, irrelevant or uncertain based off elimination criteria
 \item Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
@@
 From this discussion, papers were either thrown out or kept until the next round of the review.
-Ties were broken by an indepedent party.
 \end{itemize}
 
-\subsubsection{Title Elimination}
+\subsubsection{Step 2: Title Elimination}
 
@@
 
-\subsubsection{Abstract Elimination}
+\subsubsection{Step 3: Abstract Elimination}
 
@@
 
-\subsubsection{Content Elimination}
+\subsubsection{Step 4: Content Elimination}
 
@@
 
-\begin{itemize}
-\item Does this model extend the core model?
-\item What reasons and evidence do researchers give that RBAC needs extension?
-\item Was the paper and subsequent model inspired by a real world example?
-\item Is there any evaluation of the proposed model? If yes, how did they do one? If no, why?
-\item Did the authors implement their model?
-\end{itemize}
+\begin{enumerate}
+\item Does this model extend the core model? (Exclusion)
+\item Do the researchers give evidence that RBAC needs extension? (Inclusion)
+\item Was the paper and subsequent model inspired by a real world example?  (Conditional Inclusion)
+\item Did the researchers offer any evaluation of the proposed model? If yes, how did they do one? If no, why? (Conditional Inclusion)
+\item Did the authors implement their model? (Inclusion)
+\end{enumerate}
+
+Question 1 was a definitive exclusion criteria as any paper that failed in the affirmative was rejected.
+Questions 3 and 4 were marked as conditional includes given that they were connected in making a decision.
+A paper that met Question 3 but not 4, or met 4 but not 3 would be included since in some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.

### 2013-02-27 paper/sections/results.tex
Updates to paper list into appendix. Adds structured abstract. Moves paper counts into process.

--- 2013-02-17 paper/sections/results.tex
@@
 
-For our systematic literature review, the search strategy was applied to the following search engines based on popularity and database size:
-
-\begin{itemize}
-\item Google Scholar - \url{http://scholar.google.com}
-\item IEEExplore - \url{http://ieeexplore.ieee.org}
-\item ACM Portal - \url{http://dl.acm.org}
-\item CiteSeerX - \url{http://http://citeseerx.ist.psu.edu/index}
-\end{itemize}
-
-For each of the criteria below, a search was performed using the exact query for each of the search engines for a total of 12 data sets.  The criteria used were:
-
-\begin{itemize}
-\item role based access control
-\item RBAC
-\item role-based access control
-\end{itemize}
-
-The results of the search are summarised below:
-
-\begin{center}
-\begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
-
-\hline
-\textbf{Search Engines} & 
-\textbf{RBAC} & 
-\textbf{role based access control} & 
-\textbf{role-based access control} & 
-\textbf{Total}
-\\\hline
-
-Google Scholar & 651 & 213 & 435 & 1299 \\\hline
-ACM Portal & 500 & 20 & 720 & 1240 \\\hline
-IEEExplore & 200 & 40 & 230 & 470 \\\hline
-CiteSeerX & 100 & 100 & 150 & 350 \\\hline
- &  &  &  & \\\hline
-Totals & 1451 & 373 & 1535 & 3359 \\\hline
-Combined &  &  &  & \textbf{1716} \\\hline
-
-\end{tabular}
-\end{center}
-
 \subsection{Extraction}
-
-P1: Describe data extraction process, including questions each reviewer based note taking off of
-
-P2: Describe how from the data extraction process cateogories of extension became evident
-
-P3: Describe how those categories were deemed relevant and a way to organize going forward
-
-P4: Describe the categories, show how each paper and their category
-
-P5: Discuss how the categories are inter-related and show diagram of relationships
 
@@
 Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.
-The result of this paper categorization is displayed below:
-
-\begin{center}
-\begin{tabular}{|p{12.5cm}|p{3cm}|}
-\hline
-\textbf{Paper} & \textbf{Category}
-\\\hline
-Alam, M. and Hafner, M. and Breu, R. 2006 \cite{alam06:constraint} & Constraint \\\hline
-Tzelepi, Sofia K. and Koukopoulos, Dimitrios K. and Pangalos, George 2001 \cite{tzelepi01:flexible} & Context \\\hline
-Haibo, SHEN and Fan, HONG 2005 \cite{haibo05:context} & Context \\\hline
-Damian G. Cholewka and Reinhardt A. Botha and Jan H. P. Eloff 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
-Huang, X. and Wang, H. and Chen, Z. and Lin, J. 2006 \cite{huang06:pervasive} & Context \\\hline
-Motta, G.H.M.B. and Furuie, S.S. 2003 \cite{motta03:contextual} & Context \\\hline
-Bao, Y. and Song, J. and Wang, D. and Shen, D. and Yu, G. 2008 \cite{bao08:role} & Context \\\hline
-Yamazaki, W. and Hiraishi, H. and Mizoguchi, F. 2004 \cite{yamazaki104:designing} & Context \\\hline
-Jian-min, H. and Xi-yu, L. and Hui-qun, Y. and Jun, T. 2008 \cite{han08:extended} & Context \\\hline
-Thein, N. and others 2011 \cite{thein2011leveraging} & Context \\\hline
-Zou, D. and He, L. and Jin, H. and Chen, X. 2009 \cite{zou2009crbac} & Context \\\hline
-Hasebe, K. and Mabuchi, M. and Matsushita, A. 2010 \cite{hasebe10:capability} & Delegation \\\hline
-Zhang, Z. and Zhang, X. and Sandhu, R. 2006 \cite{zhang06:collaborative} & Organizational \\\hline
-Ni, Q. and Trombetta, A. and Bertino, E. and Lobo, J. 2007 \cite{ni2010privacy} & Privacy \\\hline
-Masoumzadeh, A. and Joshi, J. 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
-Zhao, Y. and Zhao, Y. and Lu, H. 2008 \cite{zhao2008flexible} & Resource \\\hline
-Bertino, E. and Catania, B. and Damiani, M.L. and Perlasca, P. 2005 \cite{damian2007geo} & Spatial \\\hline
-Ray, I. and Kumar, M. and Yu, L. 2006 \cite{ray07:spatio} & Spatial \\\hline
-Hansen, F. and Oleshchuk, V. 2003 \cite{hansen2003spatial} & Spatial \\\hline
-Aich, S. and Sural, S. and Majumdar, A. 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
-Chen, L. and Crampton, J. 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
-Samuel, A. and Ghafoor, A. and Bertino, E. 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
-Chandran, S. and Joshi, J. 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
-Ray, I. and Toahchoodee, M. 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
-Aich, S. and Mondal, S. and Sural, S. and Majumdar, A. 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
-Yao, L. and Kong, X. and Xu, Z. 2008 \cite{yao2008task} & Task \\\hline
-ZHANG, S. and CHEN, X. and HOU, G. 2009 \cite{zhou2007team} & Task \\\hline
-Oh, S. and Park, S. 2003 \cite{oh2003task} & Task \\\hline
-Joshi, J.B.D. and Bertino, E. and Latif, U. and Ghafoor, A. 2005 \cite{joshi05:generalized} & Temporal \\\hline
-\end{tabular}
-\end{center}
+Each paper deemed a primary source and the papers associated categorization are shown in Table \ref{tab:categorization}.
 

### 2013-02-27 paper/sections/intro.tex
Adds research questions to introductionwq

@@
 
-\textit{The goal of this work is to provide practitioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+\textit{The goal of this work is to provide practitioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.} With respect to our goal, the authors addressed the following research questions:
+
+\begin{itemize}
+\item What categorizations exist within extensions to RBAC?
+\item What are the motivations behind extensions to RBAC?
+\item Do the extensions to RBAC have corresponding implementations?
+\item How are extensions to RBAC evaluated theoritically and in practice?
+\item What domains have extensions to RBAC been created for?
+\item What commonalities or generalizations exist across all categorizations?
+\end{itemize}
 

### 2013-03-03 paper/sections/analysis.tex
Adds heirarchy table in appendix. Adds structured abstract.

@@
 \section{Analysis} \label{sec:analysis}
-
-\subsection{Research Question 1}
-
-What categorizations exist within RBAC extensions? \\
-
-\begin{itemize}
-
-  \item Constraint:
-
-  \item Context:
-
-  \item Organizational:
-
-  \item Privacy:
-
-  \item Resource:
-
-  \item Spatial:  Spatial constraints are location-based constraints in specifying access
-	control policies. For example, in organizations, 	locations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
-	In such cases, the part-time employee role should access required resources only when the user is in the location. 
-	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
-
-  \item Spatio-Temporal:
-
-  \item Task: A task is a fundamental unit of a business activity.
-Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
-Roles are first associated with tasks, which are associated permissions.
-For example, the employee role is associated with a task, which is to write a report.
-Then, this task is associated with a permission.
-
-  \item Temporal:  Temporal constraints are time-based constraints in specifying access
-	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
-	In such cases, the part-time employee role should access required resources during the interval. 
-	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
-
-\end{itemize}
 

### 2013-03-03 paper/sections/appendix.tex
Adds heirarchy table in appendix. Adds structured abstract.

@@
 
-\begin{table}
+\begin{sidewaystable}
 \centering
-\begin{tabular}{|p{12.5cm}|p{3cm}|}
-\hline
-\textbf{Paper} & \textbf{Category}
-\\\hline
-A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint} & Constraint \\\hline
+\vline
+\begin{tabularx}{\linewidth}{ | X | X | X | X | }
+\cline{1-4}
 
-A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible} & Context \\\hline
+\multicolumn{1}{ c }{\multirow{2}{*} {\rotatebox{90}{ Context } }} & 
+\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}
 
-A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\ \cline{2-4}
 
-A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}
 
-A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\
 
-A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive}} \\ \cline{2-4}
 
-An extended RBAC model based on granular logic, 2008 \cite{jian2008extended} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive}} \\ \cline{2-4}
 
-Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual}} \\ \cline{2-4}
 
-An extended RBAC model based on granular logic, 2008 \cite{han08:extended} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role}} \\ \cline{2-4}
 
-Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{jian2008extended}} \\ \cline{2-4}
 
-CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac} & Context \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing}} \\ \cline{2-4}
 
-ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative} & Organizational \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{han08:extended}} \\ \cline{2-4}
 
-Privacy-aware role-based access control, 2007 \cite{ni2010privacy} & Privacy \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging}} \\ \cline{2-4}
 
-PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac} & Privacy \\\hline
+\multicolumn{1}{ c }{}  &
+\multicolumn{3}{ |l| }{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac}} \\ \cline{2-4}
 
-A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible} & Resource \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org. }}}} &
+\multicolumn{2}{ |l| }{ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative}} \\[2.75ex]  \cline{2-4}
 
-GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo} & Spatial \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{ Priv. }}}} &
+\multicolumn{2}{ |l| }{Privacy-aware role-based access control, 2007 \cite{ni2010privacy}} \\[1ex] \cline{3-4}
 
-LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio} & Spatial \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac}} \\[1ex] \cline{2-4}
 
-Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial} & Spatial \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Res. }}}} &
+\multicolumn{2}{ |l| }{A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible}} \\[2.5ex]  \cline{2-4}
 
-STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Task }}}} &
+\multicolumn{2}{ |l| }{A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task}} \\ \cline{3-4}
 
-On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team}} \\ \cline{3-4}
 
-A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{Task-role-based access control model, 2003 \cite{oh2003task}} \\ \cline{2-4}
 
-LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{10}{*}{\rotatebox{90}{ Spatio-Temporal }}}} &
+\multicolumn{2}{ |l| }{STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC}} \\[0.25ex] \cline{3-4}
 
-A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal}} \\[0.25ex] \cline{3-4}
 
-Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role} & Spatio-Temporal \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal}} \\[0.25ex] \cline{3-4}
 
-A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task} & Task \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt}} \\[0.25ex] \cline{3-4}
 
-Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team} & Task \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio}} \\[0.25ex] \cline{3-4}
 
-Task-role-based access control model, 2003 \cite{oh2003task} & Task \\\hline
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
+\multicolumn{2}{ |l| }{Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role}} \\[0.25ex] \cline{3-4}
 
-A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized} & Temporal \\\hline
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
+\multicolumn{1}{ |l| }{GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo}} \\[0.25ex] \cline{4-4}
 
-\end{tabular}
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
+\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio}} \\[0.25ex] \cline{4-4}
+
+\multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
+\multicolumn{1}{ |l| }{Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial}} \\[0.25ex] \cline{3-4}
+
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp. }}}} &
+\multicolumn{1}{ |l| }{A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized}} \\[5ex]
+
+\cline{1-4}
+\end{tabularx}
+\vline
 \caption{Primary sources grouped by categorization}
 \label{tab:categorization}
-\end{table}
+\end{sidewaystable}

### 2013-03-03 paper/sections/categorization.tex
Adds heirarchy table in appendix. Adds structured abstract.

\subsection{Research Question 1}

After deciding upon a set of primary sources, the authors noticed that each paper fell into one or more categorization's of extensions to RBAC.
The categorizations were extracted by looking at how they are defined, what are their characteristics and how do individual categories relate
to other categories, and encapsulated by the following research question:

What categorizations exist within RBAC extensions? \\

\subsection{Definitions}

The primary sources contained either within the title, or by multiple references within the body of the text a reference to a quantifier for the
type of extension their model was.  For example, the paper "Privacy-aware role-based access control" \cite{ni2010privacy} contained "privacy" in
the title leading to the papers categorization of Privacy.  By comparison, the paper "An extended RBAC model based on granular logic" \cite{jian2008extended}
does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was 
context based.  The authors offer definitions for each categorization and their characteristics based on the primary sources and the english definitions
for each.

\begin{itemize}

  \item Constraint:

  \item Context:

  \item Organizational:

  \item Privacy:

  \item Resource:

  \item Spatial:  Spatial constraints are location-based constraints in specifying access
	control policies. For example, in organizations, 	locations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
	In such cases, the part-time employee role should access required resources only when the user is in the location. 
	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.

  \item Spatio-Temporal:

  \item Task: A task is a fundamental unit of a business activity. Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions. Roles are first associated with tasks, which are associated permissions. For example, the employee role is associated with a task, which is to write a report. Then, this task is associated with a permission.

  \item Temporal:  Temporal constraints are time-based constraints in specifying access
	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
	In such cases, the part-time employee role should access required resources during the interval. 
	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.

\end{itemize}

\subsection{Relationships}

### 2013-03-04 paper/sections/process.tex
Adds elimination totals.

@@
 The master list of 1716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
+Table \ref{tab:eliminations} shows the total number of papers selected by each author for each round and how many papers from the disjoint set for each round survived to the next round.
 
@@
 
-The first round of elimination was performed by strict examination of the title.  Each researcher was tasked with deciding on elimination by answering the following questions:
+The first round of elimination was performed by strict examination based on the title.  Each author was tasked with deciding on elimination by answering the following questions:
 
@@
 A paper that met Question 3 but not 4, or met 4 but not 3 would be included since in some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.
+
+\begin{table}
+\centering
+\begin{tabular}{ c c | c | c | c | }
+\cline{3-5}
+
+ & \multicolumn{1}{ c| }{} & \textbf{Title} & \textbf{Abstract} & \textbf{Paper Body} \\ \hline
+
+\multicolumn{1}{ |c| }{Eric}  &  & 305 & 86 & 46 \\ \hline
+\multicolumn{1}{ |c| }{Hwang} &  & 176 & 102 & 42 \\ \hline
+
+\multicolumn{1}{ c| }{} & Overlap & 249 & 51 & 24 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Disjoint & 232 & 137 & 64 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Rejected & 208 & 116 & 59 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Retained & 41 & 21 & 5 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & \textbf{Total} & 290 & 72 & 29 \\ \cline{2-5}
+
+\end{tabular}
+\caption{Elimination Rounds}
+\label{tab:search_results}
+\end{table}

### 2013-03-12 paper/sections/abstract.tex
Commiting updated sections from JeeHyun and a png version of the core model.

@@
 web applications, mobile computing and databases. RBAC restricts access to resources 
-based on the identity of subjects connected to logical groupings of permissions called roles.  
+based on the identity of subjects connected to logical groupings of permissions.  
 The introduction of computing into new domains has led to proposals for extensions to the standard
@@
 \noindent
-{\bf Aim}: \textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+{\bf Aim}: \textit{Researchers looking to expand on the body of knowledge surrounding extensions to RBAC need a starting place to prevent re-invention of the wheel.  
+Developers need to find and select an access control mode based on their project requirements. 
+The goal of this work is to provide practioners an assessment of the current extensions to RBAC, and researchers with insights into the state of RBAC extension models and how they are evaluated.}
 \\
@@
 \noindent
-{\bf Method}: We conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
-websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
-find relationships among extended models and analyze the state of RBAC extension evaluations.
+{\bf Method}: The authors conducted a systematic literature review by collecting and synthesizing relevant research
+papers in the area of RBAC extensions. We initially collected 1716 papers from electronic digital libraries
+IEEE, Google Scholar, CiteSeerX and ACM and selected 29 primary sources.  We performed a comparative 
+analysis of the primary sources to find relationships among extended models and analyze the state of RBAC extension evaluations.
 \\
@@
 \noindent
-{\bf Results}:
+{\bf Results}: Our results showed that extensions to RBAC fall into a number of categorizations that in turn all
+fall under the single category of context based extensions.  We look into the motivations behind RBAC extensions and
+the domains that are leading factors in developing these newer models. We also quantify the current evaluation methods
+used when RBAC extension models are presented to the research community.
 \\
@@
 \noindent
-{\bf Conclusions}: We conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
-websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
-find relationships among extended models and analyze the state of RBAC extension evaluations.
+{\bf Conclusions}:
 

### 2013-03-12 paper/sections/analysis.tex
Commiting updated sections from JeeHyun and a png version of the core model.

@@
 What commonalities or generalizations exist across all categorizations? \\
+
+Core or any extended role-based access control is used in various aspects of computer systems.
+In order to reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
+We found that propositional logic is used to describe access control model across all categorizations.
+Propositional logic is concerned with propositions and their logical relationships.
+In propositional logic, simple (i.e., atomic) or compound condition at given context is evaluated to true or false based on specified rules and access control logic.
+Researchers are concerned to extend limited set of propositions specific to core RBAC to meet real-world scenarios such as dynamic constraints, temporal, or spatial constraints.
+However, semantic meanings of such propositions are various based on researchers' intention.

### 2013-03-12 paper/sections/categorization.tex
Adds elimination totals.
Commiting updated sections from JeeHyun and a png version of the core model.

@@
 does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was 
-context based.  The authors offer definitions for each categorization and their characteristics based on the primary sources and the english definitions
+context based.  The authors offer definitions for each categorization and their characteristics based on the primary sources and the English definitions
 for each.
@@
 
-  \item Constraint:
+  \item Constraint: Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles.
+This constraint is either static and dynamic.
+For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician.
+This example describes doctor's permission with conditional restriction, "only when the doctor is assigned as the designated primary care physician".
 
-  \item Context:
+  \item Context: Access control model integrates context information.
+In access control, context often refers to user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
 
-  \item Organizational:
+  \item Organizational: Organizational access control is concerned with access control associated with multiple organizations.
+Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
 
-  \item Privacy:
+  \item Privacy: Access control model can be extended to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
 
-  \item Resource:
+  \item Resource: Access control model can be extended to handle any system resources (e.g., a file, printer, terminal, database record, etc) in a flexible way.
 
@@
 
-  \item Spatio-Temporal:
+  \item Spatio-Temporal: Spatio-temporal constraints are combination of spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
 

### 2013-03-12 paper/sections/core_rbac.tex
Commiting updated sections from JeeHyun and a png version of the core model.

@@
 
-%Each user incorporates a session    
-%The user function maps each session to a
-%single user, whereas the role function establishes a
-%mapping between a session and a set of roles activated
-%by the corresponding user in the session.
+\begin{figure}[ht]
+    \centering
+        \includegraphics[width=4.0in]{sections/core-model.png}
+    \caption{\label{fig:overview}RBAC Core Diagram in NIST RBAC model \cite{ferraiolokuhn}.}
+\end{figure}
 
-P1: Describe RBAC entities
-
-P2: Note the 4 levels of RBAC
-
-P3: Concisely describe level 1
-
-P4: Concisely describe level 2
-
-P5: COncisely describe level 3
-
-P6: Concisely describe level 4
+Figure~\ref{fig:overview} presents an overview of core RBAC model diagram where elements and their relations are described.
+Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
+Permissions are associated with possible users' pre-defined operation on an object (e.g., execute a file).
+Note that, at user or role activation, a session associated with user or role is established.

### 2013-03-24 paper/sections/abstract.tex
Breaks each research quesiton into it's own major section.

--- 2013-02-27 paper/sections/abstract.tex
@@
 \noindent
-{\bf Aim}: \textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+{\bf Aim}: Our work aims to aid researchers looking to expand on the body of knowledge surrounding extensions to RBAC need a starting place to prevent re-invention of the wheel and
+developers needing to find and select an access control mode based on their project requirements. 
+\textit{The goal of this work is to provide practioners an assessment of the current extensions to RBAC to aid with choosing a model that meets their requirements, and researchers with insights into the state of RBAC extension models and how they are evaluated.}
 \\
@@
 {\bf Method}: We conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
-websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
-find relationships among extended models and analyze the state of RBAC extension evaluations.
+papers in the area of RBAC extensions. The review we performed yielded 1716 papers, of which 29 were deemed primary sources 
+for inclusion as model extensions to the RBAC standard model. The papers we collected were from electronic digital libraries
+IEEE, Google Scholar, CiteSeerX and ACM based on a set of exclusion and inclusion criteria.  
+We performed a comparative analysis of the primary sources to find relationships among extended models and to analyze the state of RBAC extension evaluations.
 \\
@@
 \noindent
-{\bf Results}:
+{\bf Results}: Our results showed that extensions to RBAC fall into a number of categories that in turn all
+fall under the single category of context based extensions.  We look into the motivations behind RBAC extensions and
+the domains that are leading factors in developing these newer models. We also quantify the current evaluation methods
+used when RBAC extension models are presented to the research community.
 \\
@@
 \noindent
-{\bf Conclusions}: We conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
-websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
-find relationships among extended models and analyze the state of RBAC extension evaluations.
+{\bf Conclusions}: The extensions to RBAC fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial and Temporal falling under the general category of context.
+Domains, such as healthcare and mobile computing, were idenitfied as motivations behind the development of extensions to the RBAC model.
+Our literature review showed that the state of RBAC model evaluation could benefit from the research community given most model evaluations seen within the papers were based on hypothetical situations with little to no case studies or implementations in practice.
 

### 2013-03-24 paper/sections/analysis.tex
Breaks each research quesiton into it's own major section.

@@
-\section{Analysis} \label{sec:analysis}
+\section{Motivations} \label{sec:motivations}
 
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
@@
 \end{itemize}
+
+\section{Implementations} \label{sec:implementations}
 
@@
 
-Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
+Of the 29 papers surveyed, there was a lack of implementation with 21 of the paper providing no
 mention of a implementation or prototype.  Of the remaining 8 papers that did mention an implementation, half 
@@
 system.
+
+\section{Evaluations} \label{sec:evaluations}
 
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item Time-based Performance \cite{ni2010privacy}, \cite{aich09:role}
@@
 
-TODO: Add references to list above.
+\section{Domains} \label{sec:domains}
 
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item What domains or scenarios serve as inspiration for these extensions?
@@
 \end{itemize}
+
+\section{Generaliations} \label{sec:generalizations}
 

### 2013-03-24 paper/sections/appendix.tex
Breaks each research quesiton into it's own major section.

@@
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}
-
-\multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\
+\multicolumn{3}{ |l| }{ A Context-Aware Role-Based Access Control Model for Web Services, 2005 \cite{haibo2005context}} \\ \cline{2-4}
 
@@
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki104:designing}} \\ \cline{2-4}
+\multicolumn{3}{ |l| }{Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki04:designing}} \\ \cline{2-4}
 
@@
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
-\multicolumn{1}{ |l| }{GEO-RBAC: a spatially aware RBAC, 2005 \cite{damian2007geo}} \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ |l| }{GEO-RBAC: a spatially aware RBAC, 2005 \cite{damiani2007geo}} \\[0.25ex] \cline{4-4}
 

### 2013-03-24 paper/sections/categorization.tex
Breaks each research quesiton into it's own major section.

@@
+\section{Categorization} \label{sec:categorization}
+
+After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends.
+The first step was to take the individual data generated from the final elimination round and organize this information around the research questions.
+During the paper reading round, and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
+Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.
+Each paper deemed a primary source and the papers associated categorization are shown in Table \ref{tab:categorization}.
+
 \subsection{Research Question 1}
 
-After deciding upon a set of primary sources, the authors noticed that each paper fell into one or more categorization's of extensions to RBAC.
-The categorizations were extracted by looking at how they are defined, what are their characteristics and how do individual categories relate
-to other categories, and encapsulated by the following research question:
-
-What categorizations exist within RBAC extensions? \\
-
-\subsection{Definitions}
+What categories exist within RBAC extensions? \\
 
@@
 \end{itemize}
-
-\subsection{Relationships}

### 2013-03-24 paper/sections/conclusion.tex
Breaks each research quesiton into it's own major section.

@@
 
-P1: Mention core RBAC, note specific results for each research question
-
-P2: Mention bigger results that cut across and any other peritentn information
+The extensions to RBAC were revealed to fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial and Temporal falling under the general category of context.
+The categories each had properties specific to their implementation, but were seen to generalize to being specialized instances of context tailored to the entities or actions the categories covered.
+A number of domains were idenitfied as being the motivations behind needing extensions to the core RBAC model.  The domains, such as healthcare, presenting new challenges the previous models were not
+required to design for.  Our literature review showed that the state of RBAC model evaluation needs focused from the research community given most model evaluations seen within the papers were
+based on hypothetical situations with little to no case studies or implementations in practice.

### 2013-03-24 paper/sections/core_rbac.tex
Breaks each research quesiton into it's own major section.

@@
-\section{Core Role Based Access Control} \label{sec:core-rbac}
+\section{Role Based Access Control Standard} \label{sec:core-rbac}
 
@@
 The NIST RBAC model proposed by Ferraiolo et al.
-and later adopted as the official standard for RBAC by the International Committee for Information Technology Standards (INCITS) consists of four basic entities:
+\cite{ferraiolokuhn} and later adopted as the official standard for RBAC by the International Committee for Information Technology Standards (INCITS) consists of four basic entities:
 
 \begin{itemize}
-\item a set of users \emph{Users}: A user can be a person or an agent.
-\item  a set of roles \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
-\item a set of permissions \emph{Permissions}: A permission refers to an access mode that can be exercised on an object in the system and a session relates a user to possibly many roles.
-\item a set of sessions \emph{Sessions}: In each session, a user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
+\setlength{\itemsep}{0.25pt}
+\item a set of \emph{Users}: A user can be a person or an agent.
+\item a set of \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
+\item a set of \emph{Permissions}: A permission refers to an access mode that can be exercised on an object in the system and a session relates a user to possibly many roles.
+\item a set of \emph{Sessions}: In each session, a user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
 \end{itemize}
 
-In the RBAC, a user can exercise a permission only if the user are assigned to a role.
+In RBAC, a user can exercise a permission only if the user is assigned to a role.
 In addition to the four basic components, two functions are defined:

### 2013-03-24 paper/sections/intro.tex
Adds heirarchy table in appendix. Adds structured abstract.
Commiting updated sections from JeeHyun and a png version of the core model.
Breaks each research quesiton into it's own major section.

@@
 
+Innovation and the use of RBAC into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.
+For example, Ni, Q.
+and Bertino, E.
+et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare.
+Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.
+However, as researchers often develop their own specialized extended models of RBAC, their research cannot be generalized or compared with other research work appropriately.
+
 Role based access control (RBAC) was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}.
 In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
@@
 
-Since the introduction of the standardized RBAC model, innovation and the spread of software engineering into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.
-For example, RBAC is ill-equipped to handle the additional entities that need to be taken into consideration during authorization with the introduction of privacy concerns in domains such as healthcare  Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.
-However, as researchers often develop their own specialized extended models of RBAC, their research cannot be generalized or compared with other research work appropriately.
-
-\textit{The goal of this work is to provide practitioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.} With respect to our goal, the authors addressed the following research questions:
+\textit{The goal of this work is to provide practioners an assessment of the current extensions to RBAC to aid with choosing a model that meets their requirements, and researchers with insights into the state of RBAC extension models and how they are evaluated.} With respect to our goal, the authors addressed the following research questions:
 
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item What categorizations exist within extensions to RBAC?
@@
 
-A systematic literature review is designed to yield quantifiable results across a body of research in order to draw out similarities, trends and deficiencies in an area as a whole.
-The review we performed yielded 1716, of which XX were deemed primary sources for inclusion as model extensions to the RBAC standard model.
+The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the core RBAC model.
+The review we performed yielded 1716, of which 29 were deemed primary sources for inclusion as model extensions to the RBAC standard model.
 This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
@@
 \begin{itemize}
-\item Summarizes current extended RBAC research work and its contributions
+\setlength{\itemsep}{0.25pt}
+\item Summarizes current extended RBAC research work and its contributions.
 \item Guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
-\item Our work shows a criteria in comparison among research results.
-\item Identify the research challenges in the areas of security policies and suggest a future extension of RBAC
-\item Identify the deficiencies in RBAC extension evaluation
+\item Identifies the research challenges in the areas of security policies and suggest a future extension of RBAC.
+\item Identifies the deficiencies in RBAC extension evaluation.
 \end{itemize}
@@
 Section 3 lays out the process used in conducting the review.
-Section 4 provides the results of the search.
-Section 5 tackles the research questions and analyzes the results.
+Section 4 discusses the categorization break down of the papers.
+Section 5 analyzes the rest of the research questions.
 Section 6 contains the final conclusions.
-
-\begin{itemize}
-\item TODO: Add some references to the prevalence of RBAC in industry
-\item TODO: Add link to RBAC standard documents
-\item TODO: List research questions here or in process
-\end{itemize}

### 2013-04-02 paper/sections/abstract.tex
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

--- 2013-02-27 paper/sections/abstract.tex
@@
 \noindent
-{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's, 
+{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's by the National Institute of Standards and Technology (NIST),
 RBAC has become an increasingly popular access control mechanism for applications such as 
 web applications, mobile computing and databases. RBAC restricts access to resources 
-based on the identity of subjects connected to logical groupings of permissions called roles.  
+based on the identity of subjects connected to logical groupings of permissions.  
 The introduction of computing into new domains has led to proposals for extensions to the standard
@@
 \noindent
-{\bf Aim}: \textit{The goal of this work is to provide practioners an assessment of the state of extended models of RBAC and researchers with insights into the lack of robust evaluation of RBAC extension models.}
+{\bf Aim}: The goal of our work is to aid practioners and researchers in choosing an exentsion to the RBAC standard model, and in understanding
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establisment of a 
+set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
+that have led to extension of the RBAC standard.
 \\
@@
 \noindent
-{\bf Results}:
+{\bf Results}: Our results showed that extensions to RBAC can be classified under seven categories: Constraint, Context, 
+Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial and Temporal. We found a range of motivations behind RBAC extensions from 
+lack of fine-grained controls to a need for context placed on permissions, roles and assignment.  We found that 12 models presented no evidence
+of an evaluation, 8 models used example scenarios and 6 others provided some complexity and time analysis.  Of the 29 papers, we identified only
+8 that provided in an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
+that inspired were the model domain with 9 of the 29 models, enterprise workflows with 5 of 29, and mobile computing with 5 of 29.
 \\
@@
 \noindent
-{\bf Conclusions}: We conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collect XXXX papers from sources such as IEEE and ACM 
-websites and selected X primary sources.  We performed a comparative analysis of the primary sources to 
-find relationships among extended models and analyze the state of RBAC extension evaluations.
+{\bf Conclusions}: The extensions to RBAC found by our systematic literature review fall into seven categoriies that upon closer inspection can be 
+aligned under the single category of context. Two domains, healthcare and mobile computing, were idenitfied as predominant motivations behind the 
+development of extensions to the RBAC model. Our literature review showed that the state of evaluation of RBAC model evaluation is limited with 1 
+paper doing an evaluation comparing itself to the core RBAC model, 12 of the 29 models lacking any identifiable evaluation and 8 presenting example 
+scenarios of their model in action, the predominant evaluation method.
 

### 2013-04-02 paper/sections/analysis.tex
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

@@
 \item What domains or scenarios serve as inspiration for these extensions?
-\item Medical domain
-\item Pervasive computing environments
-\item Mobile devices
-\item Large-scale organizations with many sub-departments
-\item Enterprise, organization workflows
+\item Medical domain \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{motta03:contextual}, \cite{ni2010privacy}, \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{samuel07:spatio-temporal}, \cite{aich09:role}, \cite{zhou2007team}
+\item Pervasive computing environments \cite{huang06:pervasive}, \cite{chen08:spatio-temporal}, \cite{ray07:spatio}
+\item Web applications \cite{masoumzadeh2008purbac}
+\item Mobile computing \cite{thein2011leveraging}, \cite{zou2009crbac}, \cite{chandran05:llt}, \cite{ray07:spatio}, \cite{aich09:role}
+\item Large-scale organizations with many sub-departments \cite{yamazaki04:designing}, \cite{han08:extended}, \cite{yao2008task}
+\item Enterprise, organization workflows \cite{cholewka00:acontext-sensitive}, \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{oh2003task}, \cite{joshi05:generalized}
 \end{itemize}

### 2013-04-02 paper/sections/appendix.tex
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

@@
 \multicolumn{1}{ c }{\multirow{2}{*} {\rotatebox{90}{ Context } }} & 
-\multicolumn{3}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\ \cline{2-4}
-
-\multicolumn{1}{ c }{}  &
 \multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\ \cline{2-4}
@@
 \multicolumn{3}{ |l| }{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac}} \\ \cline{2-4}
+
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Const. }}}} &
+\multicolumn{2}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\[2.75ex] \cline{2-4}
 

### 2013-04-02 paper/sections/categorization.tex
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

--- 2013-03-12 paper/sections/categorization.tex
@@
-\subsection{Research Question 1}
+\section{Categorization} \label{sec:categorization}
 
-After deciding upon a set of primary sources, the authors noticed that each paper fell into one or more categorization's of extensions to RBAC.
-The categorizations were extracted by looking at how they are defined, what are their characteristics and how do individual categories relate
-to other categories, and encapsulated by the following research question:
+RQ1: What categories exist within RBAC extensions? \\
 
-What categorizations exist within RBAC extensions? \\
+During the paper reading phase, the authors noted that a variety of themes were present amongst the extensions. 
+For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly in the title of the paper and the name of their model. 
+Some papers presented this direct prouncement of their theme, while others were less obvious. 
+Thus, the authors developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. The guidelines were:
+
+\begin{itemize}
+\item Self Declaration - Does the title or name of the model classify itself?
+\item Repetition of Phrase - Does the body of the paper present the same phrase repeatedly when discussing their model?
+\item Self Assessment - Do the authors of the paper directly identify a quantifier for their paper within the body of the paper?
+\item Simple - Can the category be represented by a single quantifier?
+\end{itemize}
+
+If we take the previous example above, the paper "Privacy-aware role-based access control" \cite{ni2010privacy} contained "privacy" in
+the title and in the name of the model leading to the creation of the Privacy categorization and the subsequent placement of the paper under that category.
+By comparison, the paper "An extended RBAC model based on granular logic" \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.
+
+The authors offer definitions for each category based on the data extracted from the primary sources and the English definitions for each word.
 
@@
 
-The primary sources contained either within the title, or by multiple references within the body of the text a reference to a quantifier for the
-type of extension their model was.  For example, the paper "Privacy-aware role-based access control" \cite{ni2010privacy} contained "privacy" in
-the title leading to the papers categorization of Privacy.  By comparison, the paper "An extended RBAC model based on granular logic" \cite{jian2008extended}
-does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was 
-context based.  The authors offer definitions for each categorization and their characteristics based on the primary sources and the English definitions
-for each.
-
 \begin{itemize}
 
-  \item Constraint: Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles.
+  \item Context: Access control model integrates context information.
+In access control, context often refers to user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
+
+  \item Constraint (Const): Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles.
 This constraint is either static and dynamic.
@@
 
-  \item Context: Access control model integrates context information.
-In access control, context often refers to user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
-
-  \item Organizational: Organizational access control is concerned with access control associated with multiple organizations.
+  \item Organizational (Org): Organizational access control is concerned with access control associated with multiple organizations.
 Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
 
-  \item Privacy: Access control model can be extended to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
+  \item Privacy (Priv): Access control model can be extended to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
 
-  \item Resource: Access control model can be extended to handle any system resources (e.g., a file, printer, terminal, database record, etc) in a flexible way.
+  \item Resource (Res): Access control model can be extended to handle any system resources (e.g., a file, printer, terminal, database record, etc) in a flexible way.
+
+  \item Task: A task is a fundamental unit of a business activity.
+Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
+Roles are first associated with tasks, which are associated permissions.
+For example, the employee role is associated with a task, which is to write a report.
+Then, this task is associated with a permission.
+
+  \item Spatio-Temporal: Spatio-temporal constraints are combination of spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
 
@@
 
-  \item Spatio-Temporal: Spatio-temporal constraints are combination of spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
-For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
-
-  \item Task: A task is a fundamental unit of a business activity.
-Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
-Roles are first associated with tasks, which are associated permissions.
-For example, the employee role is associated with a task, which is to write a report.
-Then, this task is associated with a permission.
-
-  \item Temporal:  Temporal constraints are time-based constraints in specifying access
+  \item Temporal (Temp):  Temporal constraints are time-based constraints in specifying access
 	control policies. For example, in organizations, 	periodic temporal durations are enforced while a
@@
 
-\subsection{Relationships}
+\subsection{Analysis and Discussion}
+
+The 29 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
+category and further, displays the perceived heirarchy of the categories. 
+The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and
+the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.
+
+When looking across all categories, the authors noted that each category added some new features on top of the standard RBAC model that were domain specific.
+These domain specific features were under the surface adding contextual relationships between the core user, permission and role entities.  Thus, the authors
+concluded that all categories stemmed from the context category, of which some primary sources were already deemed direct members.
+
+For example, in the case of Privacy the models added entities such as purpose binding to represent within the model data collected for one purpose should not used for another purpose without user consent {\cite{ni2010privacy}.
+While the new entity provided by the Privacy based models is inspired by domains such as healthcare where privacy is of legal concern, the underlying mechanism that drives purpose binding is providing context around making an access control decision.
+The system must take into account not just a static set of permissions a user has through their roles, but also the context of the data being accessed as that data relates to privacy policy. 
+In the spatio-temporal models, a users location and the time of day are two factors that can be taken into account when activating a role or verifying a permission.  
+The concepts of location and time are properties of the user and place specific contexts around the role and permission entities.

### 2013-04-02 paper/sections/intro.tex
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

@@
 
-Innovation and the use of RBAC into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.
-For example, Ni, Q.
-and Bertino, E.
-et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare.
-Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.
-However, as researchers often develop their own specialized extended models of RBAC, their research cannot be generalized or compared with other research work appropriately.
+Innovation and the use of role based access control (RBAC) into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
+For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
+meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare. 
+Since the introduciton of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features onto the core.  
+A cursory examination of the extension model papers does not provide good comparisons. Further, no generalizations among models have been made which inherently defies the reasoning behind
+creating a standard.
 
-Role based access control (RBAC) was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu, et al \cite{sandhu1996role}.
-The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}.
-In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization.
-RBAC assigns permissions of specific actions on resources to roles instead of individual users.
-Therefore, in order to gain roles' permission on specific resources, users acquire appropriate roles first.
-RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.
-RBAC has advantages in maintaining and managing organization's security policies.
+The RBAC standard was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) 
+requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
+proposed by Sandhu, et al \cite{sandhu1996role}.  In 2004, this standard was adopted as ANSI/INCITS 359-2004.  The development of a standard was inspired by an economic impact study done 
+during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}
+that showed the cost savings of RBAC in implementation and maintenance.
+RBAC is a generalized access control approach used for various applications including web services, 
+database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  
 For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.
+In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions 
+of specific actions on resources to roles instead of individual users. Therefore, in order to gain a roles' permission on specific resources, 
+users acquire appropriate roles first.
 
@@
 Section 3 lays out the process used in conducting the review.
-Section 4 discusses the categorization break down of the papers.
-Section 5 analyzes the rest of the research questions.
-Section 6 contains the final conclusions.
+Sections 4-9 present the analysis and discussion of the research questions.
+Section 10 contains the final conclusions.

### 2013-04-02 paper/sections/process.tex
Breaks each research quesiton into it's own major section.
Adding adjustments to abstract and introduction with focus on standards. Re-ordeirng the categories and adding short names for categories as seen in the table. Adding analysis and discussion section for categories.

@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item Step 1: Development of a search strategy
@@
 \begin{enumerate}
+\setlength{\itemsep}{0.25pt}
 \item Call to search engine with current search position and current search term.
@@
 
-TODO: Add elimination stats.
-
 \subsection{Steps 2-4: Elimination Rounds}
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item The two first authors independently classified papers as relevant, irrelevant or uncertain based off elimination criteria
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item Did the title contain a reference to 'role based access control' or 'RBAC'? 
@@
 \begin{itemize}
+\setlength{\itemsep}{0.25pt}
 \item Does the abstract mention a proposed model?
@@
 \begin{enumerate}
+\setlength{\itemsep}{0.25pt}
 \item Does this model extend the core model? (Exclusion)
@@
 
- & \multicolumn{1}{ c| }{} & \textbf{Title} & \textbf{Abstract} & \textbf{Paper Body} \\ \hline
+ & \multicolumn{1}{ c| }{} & \textbf{Title} & \textbf{Abstract} & \textbf{Content} \\ \hline
 
@@
 \end{table}
+
+\subsection{Extraction}
+
+After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends.
+The first step was to take the individual data generated from the final elimination round and organize this information around the research questions.
+During the paper reading round, and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
+Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.

### 2013-04-02 paper/sections/intro.tex
Adding NIST RBAC standard in Intro
why RBAC stand is initiated.

@@
 requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
-proposed by Sandhu, et al \cite{sandhu1996role}.  In 2004, this standard was adopted as ANSI/INCITS 359-2004.  The development of a standard was inspired by an economic impact study done 
-during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}
+proposed by Sandhu, et al \cite{sandhu1996role}.
+In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS).
+INCITS is an international organization, which approves and maintains information technology standards.
+INCITS operates under rules approved by American National Standards Institute (ANSI).
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}
 that showed the cost savings of RBAC in implementation and maintenance.
+More specifically, NIST initiated the development of RBAC standard because NIST found that IT developers and users in government and industry need consistent and uniform definition of RBAC features.
+In prior to the development of RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on the unified definition of RBAC features.
+As a result, IT developers and users have uncertainty and confusion of RBAC features.
+Due to this uncertainty and confusion, not only development and maintenance cost of RBAC but also frequencies of security violations increase.
+RBAC standard address this issue by describing definition and functional specification of RBAC features using a reference model.
+
 RBAC is a generalized access control approach used for various applications including web services, 

### 2013-04-03 paper/sections/intro.tex
Adding an updated introduction.

--- 2013-04-02 paper/sections/intro.tex
@@
 
-Innovation and the use of role based access control (RBAC) into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
+Innovation and the use of role based access control (RBAC) in new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
 For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
 meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare. 
-Since the introduciton of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features onto the core.  
-A cursory examination of the extension model papers does not provide good comparisons. Further, no generalizations among models have been made which inherently defies the reasoning behind
-creating a standard.
+Since the introduciton of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features on top of the core RBAC model. 
+A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
+Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
 
@@
 requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
-proposed by Sandhu, et al \cite{sandhu1996role}.  In 2004, this standard was adopted as ANSI/INCITS 359-2004.  The development of a standard was inspired by an economic impact study done 
-during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}
-that showed the cost savings of RBAC in implementation and maintenance.
-RBAC is a generalized access control approach used for various applications including web services, 
-database applications, and healthcare applications.  RBAC has advantages in maintaining and managing organization's security policies.  
-For example, if a user is to access manager role's resources within a given organization, security policy administrators simply add the user to be associated with the manager role.
-In an RBAC model, roles represent a group of users who are involved in a specific job function in an organization. RBAC assigns permissions 
-of specific actions on resources to roles instead of individual users. Therefore, in order to gain a roles' permission on specific resources, 
-users acquire appropriate roles first.
+proposed by Sandhu, et al \cite{sandhu1996role}.  
+In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
+NIST found that IT developers and users in government and industry need consistent and uniform definition of RBAC features. 
+Prior to the development of RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on the unified definition of RBAC features. 
+As a result, IT developers and users have uncertainty, confusion and a lack of interoperability of RBAC features. 
+Due to this uncertainty and confusion, not only do the development and maintenance cost of RBAC increase but also frequencies of security violations.
+RBAC standard address this issue by describing definition and functional specification of RBAC features using a reference model.
 
-\textit{The goal of this work is to provide practioners an assessment of the current extensions to RBAC to aid with choosing a model that meets their requirements, and researchers with insights into the state of RBAC extension models and how they are evaluated.} With respect to our goal, the authors addressed the following research questions:
+RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  
+RBAC has advantages in maintaining and managing an organization's security policies.  
+For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
+In the standard RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
+Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning indvidual permissions to individual users.
+As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves.
+
+\textit{The goal of our work is to aid practioners and researchers in choosing an exentsion to the RBAC standard model, and in understanding
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establisment of a 
+set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
+that have led to extension of the RBAC standard.} With respect to our goal, the authors addressed the following research questions:
 

### 2013-04-03 paper/sections/categorization.tex
Minor updates to process and categories.

@@
 
-During the paper reading phase, the authors noted that a variety of themes were present amongst the extensions. 
-For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly in the title of the paper and the name of their model. 
-Some papers presented this direct prouncement of their theme, while others were less obvious. 
+During the paper reading phase, the authors noted that categorical themes were present among the extensions. 
+For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly within the title of the paper and the name of their model. 
+Some papers presented this direct prouncement of a theme, while others were less obvious. 
 Thus, the authors developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. The guidelines were:

### 2013-04-03 paper/sections/analysis.tex
Adding analysis and discussion of domains.

@@
 
-\subsection{Research Question 2}
-
-What are the motivations behind RBAC extensions? \\
+RQ2: What are the motivations behind RBAC extensions? \\
 
@@
 
-\subsection{Research Question 3}
-
-Do the extension models have corresponding implementations? \\
+RQ3: Do the extension models have corresponding implementations? \\
 
@@
 
-\subsection{Research Question 4}
-
-How are extension to RBAC evaluated theoritically and in practice? \\
+RQ4: How are extension to RBAC evaluated theoritically and in practice? \\
 
@@
 
+\subsection{Analysis and Discussion}
+
 \section{Domains} \label{sec:domains}
 
-\subsection{Research Question 5}
-
-What domains have RBAC extensions been created for? \\
+RQ5: What domains have RBAC extensions been created for? \\
 
@@
 uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
-scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
-RBAC extensions.
-
-Upon examination of the primary sources, the most prevalent domains were:
+scenarios might serve to illuminate a trend worth further examination into the reasaon for the explosion of
+RBAC extensions.  The authors identified domains presented within the primary sources by looking for example
+scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
+We found that the domains mentioned and their associated sources are:
 
@@
 \end{itemize}
+
+The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 29 mentioning scenarios or requirements of that industry.
+Mobile computing and enterprise workflows were each represented by 5 papers claiming to be influenced by the requirements for access control within these domains. The final set
+of domains was pervasive computing environments and large-scale organizations with 3 each and web applications with 1.  There were 4 papers without any direct mention of a domain
+since Aich et al. \cite{aich09:role} fall under both the medical domain and mobile computing.
+
+\subsection{Analysis and Discussion}
+
+The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of extension models.  
+Further, the authors noted that the categories associated with papers identifying the medical domain was not limited to one or two but cut across
+each of the nine categories except for Organization and Resource. 
+The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
+Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
+the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
+of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
+of both the RBAC standard and current regulations.
+
+The RBAC standard has been re-enforced by the economic impact that standardization has had on enterprises needing to apply access control.  The
+inclusion of extension models targeted at the enterprise workflow domain is indicative of the expansion of requirements for enterprises. Developers
+and researchers should take care when looking at extension models designed to addresses the newer requirements of enterprise workflows in order to
+achieve the same economic implementation and maintainability benefits the original standard model presents.
+
+Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
+was introduced. 
+The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
+For a domain that was roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.
 

### 2013-04-03 paper/sections/intro.tex
Adding analysis and discussion of domains.

@@
 \setlength{\itemsep}{0.25pt}
-\item What categorizations exist within extensions to RBAC?
-\item What are the motivations behind extensions to RBAC?
-\item Do the extensions to RBAC have corresponding implementations?
-\item How are extensions to RBAC evaluated theoritically and in practice?
-\item What domains have extensions to RBAC been created for?
-\item What commonalities or generalizations exist across all categorizations?
+\item RQ1: What categorizations exist within extensions to RBAC?
+\item RQ2: What are the motivations behind extensions to RBAC?
+\item RQ3: Do the extensions to RBAC have corresponding implementations?
+\item RQ4: How are extensions to RBAC evaluated theoritically and in practice?
+\item RQ5: What domains have extensions to RBAC been created for?
+\item RQ6: What commonalities or generalizations exist across all categorizations?
 \end{itemize}

### 2013-04-08 paper/sections/abstract.tex
Committing updates based on comments from reading group.

--- 2013-03-12 paper/sections/abstract.tex
@@
 \noindent
-{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's, 
-RBAC has become an increasingly popular access control mechanism for applications such as 
-web applications, mobile computing and databases. RBAC restricts access to resources 
-based on the identity of subjects connected to logical groupings of permissions.  
-The introduction of computing into new domains has led to proposals for extensions to the standard
-RBAC model; for example, defining additional constraints among roles, adding context or defining new heirarchy relationships.
+{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's by the National Institute of Standards and Technology (NIST),
+computing in new domains has led to proposals for extensions to the standard RBAC model; for example, defining additional constraints among roles, adding context or defining new hierarchy relationships.
 \\
@@
 \noindent
-{\bf Aim}: \textit{Researchers looking to expand on the body of knowledge surrounding extensions to RBAC need a starting place to prevent re-invention of the wheel.  
-Developers need to find and select an access control mode based on their project requirements. 
-The goal of this work is to provide practioners an assessment of the current extensions to RBAC, and researchers with insights into the state of RBAC extension models and how they are evaluated.}
+{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establishment of a 
+set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
+that have led to extension of the RBAC standard.
 \\
@@
 \noindent
-{\bf Method}: The authors conducted a systematic literature review by collecting and synthesizing relevant research
-papers in the area of RBAC extensions. We initially collected 1716 papers from electronic digital libraries
-IEEE, Google Scholar, CiteSeerX and ACM and selected 29 primary sources.  We performed a comparative 
-analysis of the primary sources to find relationships among extended models and analyze the state of RBAC extension evaluations.
+{\bf Method}: We performed a systematic literature review that yielded 1716 papers, of which 28 were deemed primary sources for inclusion as model extensions to the RBAC standard model. 
 \\
@@
 \noindent
-{\bf Results}: Our results showed that extensions to RBAC fall into a number of categorizations that in turn all
-fall under the single category of context based extensions.  We look into the motivations behind RBAC extensions and
-the domains that are leading factors in developing these newer models. We also quantify the current evaluation methods
-used when RBAC extension models are presented to the research community.
+{\bf Results}: Our results show that extensions to RBAC can be classified under eight categories: Constraint, Context, 
+Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We found that 12 models presented no evidence
+of an evaluation, eight models used example scenarios and 6 others provided complexity and time analysis.  We identified only
+eight of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
+that inspired extensions were the medical domain with nine of the 28 models, enterprise workflows with five of 28, and mobile computing with five of 28.
 \\
@@
 \noindent
-{\bf Conclusions}:
+{\bf Conclusions}: The extensions to RBAC found by our systematic literature review fall into eight categories that upon closer inspection can be 
+aligned under the single category of context. Our literature review showed that the state of evaluation of RBAC model evaluation is limited with one
+paper doing an evaluation comparing itself to the core RBAC model, eight presenting example 
+scenarios of their model in action, and 12 of the 28 models lacking any identifiable evaluation.
 

### 2013-04-08 paper/sections/analysis.tex
Committing updates based on comments from reading group.

@@
 
+\subsection{Analysis and Discussion}
+
 \section{Implementations} \label{sec:implementations}
@@
 
-Of the 29 papers surveyed, there was a lack of implementation with 21 of the paper providing no
+\begin{table}
+\centering
+\begin{tabular}{ | c | c | }
+\cline{1-2}
+
+\textbf{Implementation Type} & \textbf{Paper Count} \\ \cline{1-2}
+Enterprise Implementation & 4 \\ \cline{1-2}
+Prototype Implementation & 4 \\ \cline{1-2}
+No Implementation & 20 \\
+
+\cline{1-2}
+\end{tabular}
+\caption{Implementation types found and the count of primary sources}
+\label{tab:implementations}
+\end{table}
+
+Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
+Of the 28 papers surveyed, there was a lack of implementation with 20 of the paper providing no
 mention of a implementation or prototype.  Of the remaining 8 papers that did mention an implementation, half 
@@
 system.
+
+\subsection{Analysis and Discussion}
+
+The RBAC standard was designed with enterprises in mind such that when practioners implemented RBAC into their systems
+there would be a reasonable assurance being based off a well thought out model.  As extensions to the standard model
+come along, thought and time should be given to how features and nuances of their models may impact implementation
+in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
+with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as 4 papers did, should be
+a prototype implementation of the model for review by both pracitioners and researchers. Of the models that produced an 
+implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
 
@@
 \setlength{\itemsep}{0.25pt}
-\item What domains or scenarios serve as inspiration for these extensions?
 \item Medical domain \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{motta03:contextual}, \cite{ni2010privacy}, \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{samuel07:spatio-temporal}, \cite{aich09:role}, \cite{zhou2007team}
@@
 
-The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 29 mentioning scenarios or requirements of that industry.
+The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 28 mentioning scenarios or requirements of that industry.
 Mobile computing and enterprise workflows were each represented by 5 papers claiming to be influenced by the requirements for access control within these domains. The final set
@@
 
-\subsection{Research Question 6}
-
-What commonalities or generalizations exist across all categorizations? \\
+RQ6: What commonalities or generalizations exist across all categorizations? \\
 
@@
 However, semantic meanings of such propositions are various based on researchers' intention.
+
+\subsection{Analysis and Discussion}

### 2013-04-08 paper/sections/core_rbac.tex
Adding an updated introduction.
Committing updates based on comments from reading group.

@@
 
+RBAC provides effective and efficient permissions management for operations, especially when sharing resources in an organizations.
+Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
+of RBAC existed amongst practioners or within the research community. 
+Without a unified definition of RBAC, software developers describe similar concepts and features of RBAC models using different terminology, which was shown to slow implementation of RBAC.  
+Moreover, in cases where organizations are concerned with adopting RBAC,
+evaluation and comparison of RBAC technologies developed by different vendors was difficult.
+NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.
+
+NIST's work can directly benefit organizations by lowering cost of early phase R\&D and implementation costs of RBAC.
+Since RBAC is introduced, 2010 RTI report showed that the the rate of RBAC adoption is rapidly growing for recent years. 
+The analysts estimate that, by the end of 2010, at least a part of permission of systems use RBAC for more than 50\% of IT users at organizations, which
+hire more than 500 employees.
+For economic estimate analysis, the analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry, of which \$1.1 billion is attributable to NIST RBAC standard work.
+
 Since the basis for our review is extensions to the core model, we will describe the core model, associated entities and other terminology encountered across the space of our review.
+
+\subsection{Core RBAC}
+
 The NIST RBAC model proposed by Ferraiolo et al.
-\cite{ferraiolokuhn} and later adopted as the official standard for RBAC by the International Committee for Information Technology Standards (INCITS) consists of four basic entities:
+\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three types of RBAC models: core RBAC, hierarchal RBAC, and constrained RBAC.
+The four entities of the standard RBAC model are:
 
@@
 In addition to the four basic components, two functions are defined:
-the user role assignment (UA) and the role
-permission assignment (PA) functions.
-UA models assignment of users to roles.
-PA models assignment of permissions to roles.
+the user role assignment (UA) and the role and permission assignment (PA) functions.
+User role assignment models assignment of users to roles.
+Permission role assignment models assignment of permissions to roles.
 
@@
 Note that, at user or role activation, a session associated with user or role is established.
+
+\subsection{Hierarchal RBAC}
+
+Hierarchal RBAC model adds role hierarchies (RH) feature to core RBAC model.
+This model incorporates a structure of roles in an organization using inheritance relationship among attributes such as roles.
+The role structure in an organization may use
+a role $r_1$, who inherits all permissions of another role $r_2$.
+For example, a manager role may inherit all permissions of a employee role.
+Role hierarchies help simplify access control policy creation and maintenance by reducing the number of
+individual role assignments to a user.
+Formally, role inheritance relation is shown as \textit{RH} $\subseteq$ \textit{ROLES} $\times$ \textit{ROLES} describing many-to-many mapping role inheritance relation.
+General role hierarchies can be extended to use the concept of multiple inheritance where
+a role $r_1$, who inherits all permissions from more than one roles.
+
+\subsection{Constrained RBAC}
+
+Constrained RBAC incorporates static and dynamic separation of duty relations to the RBAC model. Separation of 
+duty relations enforce conflict of interest among roles. This model defines two types of separation of duty relations; static and dynamic.
+
+\begin{itemize}
+	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
+	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$.
+SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.
+	\item Dynamic Separation of Duty (DSD): Dynamic SoD is known to be
+more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
+	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.	
+\end{itemize}

### 2013-04-17 paper/sections/abstract.tex
Further trimming the abstract.

@@
 {\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's by the National Institute of Standards and Technology (NIST),
-computing in new domains has led to proposals for extensions to the standard RBAC model; for example, defining additional constraints among roles, adding context or defining new hierarchy relationships.
+computing in new domains has led to proposals for extensions to the RBAC standard model; for example adding context around permission granting and role activation.
 \\
@@
 {\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
-how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establishment of a 
-set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
-that have led to extension of the RBAC standard.
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model. We accomplished this through: 
+establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the 
+motivations that have led to extension of the RBAC standard model.
 \\
@@
 \noindent
-{\bf Method}: We performed a systematic literature review that yielded 1716 papers, of which 28 were deemed primary sources for inclusion as model extensions to the RBAC standard model. 
+{\bf Method}: We performed a systematic literature review that yielded 1716 papers, of which 28 were deemed primary sources for inclusion as extensions to the RBAC standard model. 
 \\

### 2013-04-17 paper/sections/analysis.tex
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

@@
 
-RQ2: What are the motivations behind RBAC extensions? \\
+RQ2: What are the motivations behind extensions to RBAC?
+
+\subsection{Results}
 
@@
 
-RQ3: Do the extension models have corresponding implementations? \\
+RQ3: Do the extension models have corresponding implementations?
+
+\subsection{Results}
 
@@
 \centering
+\caption{Implementation types found and the count of primary sources}
 \begin{tabular}{ | c | c | }
@@
 \end{tabular}
-\caption{Implementation types found and the count of primary sources}
 \label{tab:implementations}
@@
 
-RQ4: How are extension to RBAC evaluated theoritically and in practice? \\
+RQ4: How are extension to RBAC evaluated theoeetically and in practice?
+
+\subsection{Results}
 
@@
 
-RQ5: What domains have RBAC extensions been created for? \\
+RQ5: What domains have extensions to RBAC been created for?
+
+\subsection{Results}
 
@@
 Further, the authors noted that the categories associated with papers identifying the medical domain was not limited to one or two but cut across
-each of the nine categories except for Organization and Resource. 
+each of the eight categories except for Organization. 
 The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
@@
 
-RQ6: What commonalities or generalizations exist across all categorizations? \\
+RQ6: What commonalities or generalizations exist across all categories?
+
+\subsection{Results}
 

### 2013-04-17 paper/sections/categorization.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

@@
 
-RQ1: What categories exist within RBAC extensions? \\
+RQ1: What categories exist within RBAC extensions?
+
+\subsection{Results}
 
@@
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly within the title of the paper and the name of their model. 
-Some papers presented this direct prouncement of a theme, while others were less obvious. 
-Thus, the authors developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. The guidelines were:
+Some papers presented this direct pronouncement of a theme, while others were less obvious. 
+Thus, we developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. 
+In developing these guidelines, we attempted to define each category by a single noun-phrase descriptor.
+The guidelines were:
 
 \begin{itemize}
-\item Self Declaration - Does the title or name of the model classify itself?
+\item Model Name - Does the name of the model classify itself?
+\item Self Assessment - Do the authors of the paper directly identify a descriptor for their model within the body of the paper?
 \item Repetition of Phrase - Does the body of the paper present the same phrase repeatedly when discussing their model?
-\item Self Assessment - Do the authors of the paper directly identify a quantifier for their paper within the body of the paper?
-\item Simple - Can the category be represented by a single quantifier?
 \end{itemize}
 
-If we take the previous example above, the paper "Privacy-aware role-based access control" \cite{ni2010privacy} contained "privacy" in
-the title and in the name of the model leading to the creation of the Privacy categorization and the subsequent placement of the paper under that category.
-By comparison, the paper "An extended RBAC model based on granular logic" \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.
+If we take the previous example above, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in
+the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
+By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.
 
-The authors offer definitions for each category based on the data extracted from the primary sources and the English definitions for each word.
+We offer definitions for each of the 8 observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
+Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Table \ref{tab:categorization}.
 
@@
 For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician.
-This example describes doctor's permission with conditional restriction, "only when the doctor is assigned as the designated primary care physician".
+This example describes doctor's permission with conditional restriction, ``only when the doctor is assigned as the designated primary care physician''.
 
@@
   \item Privacy (Priv): Access control model can be extended to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
-
-  \item Resource (Res): Access control model can be extended to handle any system resources (e.g., a file, printer, terminal, database record, etc) in a flexible way.
 
@@
 
-The 29 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
+The 28 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
 category and further, displays the perceived heirarchy of the categories. 
@@
 The concepts of location and time are properties of the user and place specific contexts around the role and permission entities.
+
+\greybox{We found eight categories that exist within extensions to RBAC: Constraint, Context, 
+Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other
+seven categories are all fall under the category of Context.}

### 2013-04-17 paper/sections/intro.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.

@@
 \setlength{\itemsep}{0.25pt}
-\item RQ1: What categorizations exist within extensions to RBAC?
+\item RQ1: What categories exist within extensions to RBAC?
 \item RQ2: What are the motivations behind extensions to RBAC?
 \item RQ3: Do the extensions to RBAC have corresponding implementations?
-\item RQ4: How are extensions to RBAC evaluated theoritically and in practice?
+\item RQ4: How are extensions to RBAC evaluated theoretically and in practice?
 \item RQ5: What domains have extensions to RBAC been created for?
-\item RQ6: What commonalities or generalizations exist across all categorizations?
+\item RQ6: What commonalities or generalizations exist across all categories?
 \end{itemize}
@@
 The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the core RBAC model.
-The review we performed yielded 1716, of which 29 were deemed primary sources for inclusion as model extensions to the RBAC standard model.
+The review we performed yielded 1716, of which 28 were deemed primary sources for inclusion as model extensions to the RBAC standard model.
 This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
@@
 \setlength{\itemsep}{0.25pt}
-\item Summarizes current extended RBAC research work and its contributions.
-\item Guides a direction for a standard of extended RBAC. Understanding the categorization and the motivation of the existing research results helps decide a standard of extended RBAC.
-\item Identifies the research challenges in the areas of security policies and suggest a future extension of RBAC.
-\item Identifies the deficiencies in RBAC extension evaluation.
+\item Summarizes current research on extensions to RBAC.
+\item Guides a direction for a new standard of an extended RBAC model.
 \end{itemize}

### 2013-04-17 paper/sections/analysis.tex
commit analysis and background
backgroup updates
RQ2 and RQ6 analysis and discussion

@@
 \subsection{Analysis and Discussion}
+
+The core RBAC standard represents four elements such as roles, sessions, users, and permissions, and their relations.
+While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices.
+Such extended RBAC models are formally presented to describe its behaviors.
+In the selected papers, researchers illustrate how an extended model incorporates new element based on the correct RBAC and describe their relations with pre-defined elements in the core RBAC.
+
+We classify motivations behind RBAC extensions to these three categories:
+
+\begin{itemize}
+	\item Addition: an extended model incorporates additional elements with regards to context/constraints and their relations with existing elements of the core RBAC.
+Models in this category does not change the core model.
+We found OO papers in this category.
+	\item Modification:  an extended model incorporates modification of existing features in the core RBAC.
+For example, changes of attribute relations such as role hierarchy is to show modification.
+For models in this category, changes occur within the core model.
+We found OO papers in this category.
+	\item Combination: an extended model is combination of RBAC and an another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
+For this combination, new association between attributes of two models should be introduced.
+Models in this category requires a new association between attributes of different access control models to work together.
+We found OO papers in this category.
+\end{itemize}
 
@@
 \subsection{Analysis and Discussion}
+
+Since NIST proposed RBAC standard using propositional logic, researchers describe extended RBAC models using propositional logic.
+Given context, propositional logic is used to evaluate true or false based on specified rules and access control logic.
+As access control is typically evaluated to either true (Permit) or false (Deny) based on predicates, propositional logic is sufficient to describe key ideas and definitions.
+We found that OO papers of RBAC extensions use propositional logic to describe its extended model.
+However, propositional logic has limitations.
+While this logic is simple, this logic does not support for reasoning about RBAC, which helps reduce the administrative complexity of associations such as user- role associations.
+An alternative logic to describe RBAC is first-order logic.
+This logic is sufficient to describe RBAC and extended RBAC.
+Moreover, this logic supports for concise and elegant formulation of the RBAC model and its relation.
+First-Order logic is expressive enough to concisely represent access control systems.
+First-Order logic uses relations, variables, and quantifiers.

### 2013-04-17 paper/sections/core_rbac.tex
commit analysis and background
backgroup updates
RQ2 and RQ6 analysis and discussion

@@
 
-Since the basis for our review is extensions to the core model, we will describe the core model, associated entities and other terminology encountered across the space of our review.
+The NIST RBAC model proposed by Ferraiolo et al.
+\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three types of RBAC models: core RBAC, hierarchal RBAC, and constrained RBAC.
+Since the basis for our review is extensions to the core model, we describe the core model, associated entities and other terminology encountered across the space of our review.
+Then, we describe hierarchal RBAC and constrained RBAC, which are developed by incoporating new features into the core RBAC.
 
@@
 
-The NIST RBAC model proposed by Ferraiolo et al.
-\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three types of RBAC models: core RBAC, hierarchal RBAC, and constrained RBAC.
 The four entities of the standard RBAC model are:

### 2013-04-17 paper/sections/intro.tex
Commit Intro
Commit Intro

@@
 
-Innovation and the use of role based access control (RBAC) in new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
-For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
-meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare. 
-Since the introduciton of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features on top of the core RBAC model. 
-A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
-Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
+Access control models controls which principal can access which resources in what manner.
+Role based access control (RBAC) is one of access control models, which is used especially for maintaining and managing an organization's access control based on roles instead of individual users.
+The RBAC model has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in the 1990s.
+The RBAC model is widely used for protection of security and privacy of various applications including web services, database applications, and healthcare applications.
+In the RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
+Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning individual permissions to individual users.
+As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves.
+The RBAC model has advantage for managing access control.
+For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
 
-The RBAC standard was first introduced in the 1990s when the National Institute for Standards and Technology (NIST) 
+The RBAC standard was first proposed in the 1990s when NIST 
 requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
@@
 The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
-NIST found that IT developers and users in government and industry need consistent and uniform definition of RBAC features. 
 Prior to the development of RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on the unified definition of RBAC features.
-As a result, IT developers and users have uncertainty, confusion and a lack of interoperability of RBAC features. 
-Due to this uncertainty and confusion, not only do the development and maintenance cost of RBAC increase but also frequencies of security violations.
-RBAC standard address this issue by describing definition and functional specification of RBAC features using a reference model.
+RBAC standard addressed this issue by describing definition and functional specification of RBAC features using a reference model.
 
-RBAC is a generalized access control approach used for various applications including web services, database applications, and healthcare applications.  
-RBAC has advantages in maintaining and managing an organization's security policies.  
-For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
-In the standard RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
-Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning indvidual permissions to individual users.
-As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves.
+Innovation and the use of RBAC in new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
+For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
+meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare. 
+Since the introduction of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features on top of the core RBAC model. 
+A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
+Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
 
-\textit{The goal of our work is to aid practioners and researchers in choosing an exentsion to the RBAC standard model, and in understanding
-how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establisment of a 
+\textit{The goal of our work is to aid practioners and researchers in choosing an extension to the RBAC standard model, and in understanding
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establishment of a 
 set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations

### 2013-04-24 paper/sections/analysis.tex
Adding some edits and another analysis and discussion section.

@@
 
-Core RBAC model provides an abstraction of authorization model based on user role assignment (UA) and the role
+The RBAC standard model provides an abstraction of authorization based on user role assignment (UA) and role
 permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
@@
 
-The core RBAC standard represents four elements such as roles, sessions, users, and permissions, and their relations.
+The RBAC standard model presents four entities: roles, sessions, users, and permissions, and their relations.
 While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices.
@@
 
-RQ4: How are extension to RBAC evaluated theoeetically and in practice?
+RQ4: How are extension to RBAC evaluated theoretically and in practice?
 
@@
 
-Providing evaluation of a proposed model is a key component in establishing the models validity.  The papers were examined
-for evidence of evaluations ranging from performance to mathematical accuracy to application to real world scenarios.  Further,
-for each proposed model, the reviewers looked for evidence that the authors made comparisons between their own model and the base
-model as they pertained to claims made by the authors of why their model is needed.  The quantifiable evaluations looked for were:
+The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
+evaluation types and which primary sources provided which type.  In some cases a single primary source provided mutliple evaluation
+types.
 
@@
 
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
-and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Eight models presented example scenarios
+and how application of their model would apply and resolve the situation.  Six of the models provided some form of performance
 or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
-grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
+grew, and the size of the role space for the extension model compared to standard RBAC. Four models provided mathematical descriptions
 and analysis as a way to provide evaluation in the form of completeness. 
-
 The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
@@
 \subsection{Analysis and Discussion}
+
+When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
+Further, in the case of extensions to the RBAC standard model, the model should be accompanied by validation of the model as a stand-alone
+access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
+extenion models are lacking.
+
+The primary source of validation a developer or practioner may encounter is a qualitative discussion of real-world
+scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
+authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rarer cases, 
+one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
+of how an extension model handles a real-world scenario provides developers and practioners ancedotal evidence at best
+for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
+model, developers may be left implementing a more complex model to address their requirements when the standard model would
+have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
+case studies of a model in action provide developers with evidence that the model works as intended when applied.
+
+When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
+Two of the primary sources examined provided time-based performance analysis of their extension model compared
+to the RBAC standard model. This inclusion of time analysis provides some assurances to developers that any non-functional
+requirements surrounding time to compute authorizations compete or beat the standard model. Further, four of the models provided
+some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
+over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
+need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
+entities.
 
@@
 The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
-For a domain that was roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.
+For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.
 

### 2013-04-24 paper/sections/core_rbac.tex
Adding some edits and another analysis and discussion section.

@@
 
-RBAC provides effective and efficient permissions management for operations, especially when sharing resources in an organizations.
+RBAC provides effective and efficient permissions management for operations, especially when sharing resources withinin an organization.
 Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
 of RBAC existed amongst practioners or within the research community. 
-Without a unified definition of RBAC, software developers describe similar concepts and features of RBAC models using different terminology, which was shown to slow implementation of RBAC.  
-Moreover, in cases where organizations are concerned with adopting RBAC,
+Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
+This lack of consistent terminology was shown to slow the implementation of RBAC \cite{o20102010}.  
+Moreover, in cases where organizations were concerned with adopting RBAC,
 evaluation and comparison of RBAC technologies developed by different vendors was difficult.
-NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.
+NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for 
+RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.
 

### 2013-04-24 paper/sections/intro.tex
Adding some edits and another analysis and discussion section.

@@
 
-Access control models controls which principal can access which resources in what manner.
-Role based access control (RBAC) is one of access control models, which is used especially for maintaining and managing an organization's access control based on roles instead of individual users.
+Role based access control (RBAC) is used for maintaining and managing an organization's access control based on assigning permissions to roles instead of assigning the permissions directly to individual users.
 The RBAC model has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in the 1990s.
@@
 
-\textit{The goal of our work is to aid practioners and researchers in choosing an extension to the RBAC standard model, and in understanding
-how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model through: establishment of a 
-set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the reasoning and motivations
-that have led to extension of the RBAC standard.} With respect to our goal, the authors addressed the following research questions:
+\textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
+how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model. We accomplished this through: 
+establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the 
+motivations that have led to extension of the RBAC standard model.} With respect to our goal, the authors addressed the following research questions:
 

### 2013-04-24 paper/sections/abstract.tex
Abstract updates - updating results and conclusion.

@@
 {\bf Results}: Our results show that extensions to RBAC can be classified under eight categories: Constraint, Context, 
-Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We found that 12 models presented no evidence
-of an evaluation, eight models used example scenarios and 6 others provided complexity and time analysis.  We identified only
+Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We identified only
 eight of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
@@
 \noindent
-{\bf Conclusions}: The extensions to RBAC found by our systematic literature review fall into eight categories that upon closer inspection can be 
-aligned under the single category of context. Our literature review showed that the state of evaluation of RBAC model evaluation is limited with one
+{\bf Conclusions}: When examined as an entire group, the eight categories can be aligned under the single category of context. 
+Our literature review showed that the state of evaluation of RBAC model evaluation is limited with one
 paper doing an evaluation comparing itself to the core RBAC model, eight presenting example 
-scenarios of their model in action, and 12 of the 28 models lacking any identifiable evaluation.
+scenarios of their model in action, and 12 of the 28 models lacking any identifiable evaluation. The landscape of extensions
+to the RBAC standard model serves as a basis from which a new, and vetted, extended standard RBAC model should be derived.
 

### 2013-04-24 paper/sections/categorization.tex
Updates to the category section - minor edits, and definition re-structuring.

@@
 
-  \item Context: Access control model integrates context information.
-In access control, context often refers to user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
+  \item Context: The extension model integrates contextual information into the RBAC standard model.
+Context is defined as a user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
 
-  \item Constraint (Const): Access control model in this category extends constraints, which are conditional restrictions on permissions of given roles.
-This constraint is either static and dynamic.
+  \item Constraint (Const): The extension model provides conditional restrictions on permissions of given roles.
+The constraint is either static or dynamic.
 For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician.
-This example describes doctor's permission with conditional restriction, ``only when the doctor is assigned as the designated primary care physician''.
+This example describes a doctor's permission with conditional restriction - ``only when the doctor is assigned as the designated primary care physician can the doctor modify a particular medical record''.
 
-  \item Organizational (Org): Organizational access control is concerned with access control associated with multiple organizations.
+  \item Organizational (Org): The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
 Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
 
-  \item Privacy (Priv): Access control model can be extended to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
+  \item Privacy (Priv): The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
 
-  \item Task: A task is a fundamental unit of a business activity.
+  \item Task: The extension model provides a task entity which are associated with permissions and roles.
+A task is a fundamental unit of a business activity.
 Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
@@
 
-  \item Spatio-Temporal: Spatio-temporal constraints are combination of spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+  \item Spatio-Temporal: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
 For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
 
-  \item Spatial:  Spatial constraints are location-based constraints in specifying access
+  \item Spatial: The extension model provides spatial constraints, location-based constraints in specifying access
 	control policies. For example, in organizations, locations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only in specific location.
+	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
 	In such cases, the part-time employee role should access required resources only when the user is in the location. 
-	Spatial constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
+	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.
 
-  \item Temporal (Temp):  Temporal constraints are time-based constraints in specifying access
+  \item Temporal (Temp):  The extension model provides temporal constraints, time-based constraints in specifying access
 	control policies. For example, in organizations, periodic temporal durations are enforced while a
-	specific role is permitted to conduct an action. Consider that part-time employee works only from 9:00 a.m. to 3:00 p.m.
+	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
 	In such cases, the part-time employee role should access required resources during the interval. 
-	Temporal constraints can incorporate either on roles, user-role assignments, or role-permission assignments.
+	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.
 
@@
 
-The 28 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their assumed
+The 28 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their designated
 category and further, displays the perceived heirarchy of the categories. 
@@
 
-For example, in the case of Privacy the models added entities such as purpose binding to represent within the model data collected for one purpose should not used for another purpose without user consent {\cite{ni2010privacy}.
+For example, in the case of the Privacy category, the models added entities such as purpose binding to represent within the model data collected for one purpose should not used for another purpose without user consent {\cite{ni2010privacy}.
 While the new entity provided by the Privacy based models is inspired by domains such as healthcare where privacy is of legal concern, the underlying mechanism that drives purpose binding is providing context around making an access control decision.
@@
 Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other
-seven categories are all fall under the category of Context.}
+seven categories fall under the category of Context.}

### 2013-04-25 paper/sections/intro.tex
Introduction updates - minor edits to some sentences and phrases.

@@
 
-Role based access control (RBAC) is used for maintaining and managing an organization's access control based on assigning permissions to roles instead of assigning the permissions directly to individual users.
-The RBAC model has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in the 1990s.
-The RBAC model is widely used for protection of security and privacy of various applications including web services, database applications, and healthcare applications.
-In the RBAC model, roles represent a set of permissions needed to perform a particular job function within an organization.  
-Any number of users who are involved in that specific job function within the organization can then be assigned to the role to inherit the required access instead of assigning individual permissions to individual users.
-As the number of permissions and users scales upward, the ability to logically group both users and permissions into smaller numbers of entities becomes paramount for managing access control as an organization grows and evolves.
-The RBAC model has advantage for managing access control.
-For example, if a user requires access to resources associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
+Role based access control (RBAC) is used for maintaining and managing an organization's 
+access control based on assigning permissions to roles, and roles to users, instead of 
+assigning the permissions directly to individual users. 
+The use of an RBAC model has become popular since the National Institute for Standards and Technology (NIST) 
+first proposed the RBAC standard model in the 1990s \cite{o20102010}. The RBAC standard model is widely used for securing 
+various applications including web services, database applications, and healthcare applications. 
+In the RBAC standard model, roles represent a set of permissions needed to perform a particular job function within an organization.  
+One to many users who are involved in that specific job function within the organization can then be 
+assigned to the role to inherit the required access instead of assigning individual permissions to individual users. 
+The ability to logically group both users and permissions into smaller total entities becomes paramount for managing access control
+as an organization grows and the number of permissions and users scales upward, 
+The RBAC standard model has advantages for managing access control. For example, if a user requires access to resources 
+associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
 
@@
 In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
-The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2011 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
-Prior to the development of RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on the unified definition of RBAC features.
-RBAC standard addressed this issue by describing definition and functional specification of RBAC features using a reference model.
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
+Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on a unified definition of RBAC or feature set. 
+The RBAC standard addressed this issue by providing definitions and a functional specification of RBAC features using a reference model.
 
@@
 For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
-meet privacy protection requirements`` with the introduction of privacy concerns in domains such as healthcare. 
+meet privacy protection requirements'' with the introduction of privacy concerns in domains such as healthcare. 
 Since the introduction of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features on top of the core RBAC model. 
@@
 
-The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the core RBAC model.
-The review we performed yielded 1716, of which 28 were deemed primary sources for inclusion as model extensions to the RBAC standard model.
-This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel.
-For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.
-Further, the review provides insight into the state of evaluation, or lack thereof, within the RBAC extension model community.
+The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC standard model. 
+The review we performed yielded 1716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC standard model.  
+This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and a reference guide for discovering
+the current state of extensions to the RBAC standard model. 
+For developer's looking to find a model to fit their potentially unique access control needs, this review will provide a basis for comparison and easy look-up for what extension model to use.  
+Further, the review provides insight into the state of the art in evaluating RBAC extension models.
 
@@
 \item Summarizes current research on extensions to RBAC.
-\item Guides a direction for a new standard of an extended RBAC model.
+\item Guides a direction for a new standard based on an extension to the RBAC standard model.
 \end{itemize}
@@
 Sections 4-9 present the analysis and discussion of the research questions. 
-Section 10 contains the final conclusions.
+Section 10 is a general discussion section brings together all the research questions.
+Section 11 contains the final conclusions.

### 2013-04-25 paper/sections/core_rbac.tex
Background - Adds edits for the background section.

@@
 
-NIST's work can directly benefit organizations by lowering cost of early phase R\&D and implementation costs of RBAC.
-Since RBAC is introduced, 2010 RTI report showed that the the rate of RBAC adoption is rapidly growing for recent years. 
+NIST's work can directly benefit organizations by lowering cost of early phase R\&D and the implementation of RBAC.
+Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the the rate of RBAC adoption has rapidly grown over recent years \cite{o20102010}. 
 The analysts estimate that, by the end of 2010, at least a part of permission of systems use RBAC for more than 50\% of IT users at organizations, which
@@
 
-The NIST RBAC model proposed by Ferraiolo et al.
-\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three types of RBAC models: core RBAC, hierarchal RBAC, and constrained RBAC.
-Since the basis for our review is extensions to the core model, we describe the core model, associated entities and other terminology encountered across the space of our review.
+The NIST RBAC standard model proposed by Ferraiolo et al.
+\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three levels of RBAC: core RBAC, hierarchal RBAC, and constrained RBAC.
+Since the basis for our review is extensions to the RBAC standard model, we describe the core model, associated entities and other terminology encountered across the space of our review.
 Then, we describe hierarchal RBAC and constrained RBAC, which are developed by incoporating new features into the core RBAC.

### 2013-04-25 paper/sections/core_rbac.tex
Spelling corrections.

@@
 
-RBAC provides effective and efficient permissions management for operations, especially when sharing resources withinin an organization.
+RBAC provides effective and efficient permissions management for operations, especially when sharing resources within an organization.
 Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
-of RBAC existed amongst practioners or within the research community. 
+of RBAC existed among practitioners or within the research community. 
 Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
@@
 The NIST RBAC standard model proposed by Ferraiolo et al.
-\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three levels of RBAC: core RBAC, hierarchal RBAC, and constrained RBAC.
+\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three levels of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
 Since the basis for our review is extensions to the RBAC standard model, we describe the core model, associated entities and other terminology encountered across the space of our review.
-Then, we describe hierarchal RBAC and constrained RBAC, which are developed by incoporating new features into the core RBAC.
+Then, we describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into the core RBAC.
 
@@
 
-\subsection{Hierarchal RBAC}
+\subsection{Hierarchical RBAC}
 
-Hierarchal RBAC model adds role hierarchies (RH) feature to core RBAC model.
+Hierarchical RBAC model adds role hierarchies (RH) feature to core RBAC model.
 This model incorporates a structure of roles in an organization using inheritance relationship among attributes such as roles.

### 2013-04-25 paper/sections/analysis.tex
Minor formatting clean-up of analysis section.

@@
 \subsection{Results}
-
-The RBAC standard model provides an abstraction of authorization based on user role assignment (UA) and role
-permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
-security mechanism, a variety of extended RBAC models have been proposed over the years to meet
-security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
 
@@
 \subsection{Analysis and Discussion}
+
+The RBAC standard model provides an abstraction on top of authorization based on user role assignment (UA) and role
+permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
+security mechanism, a variety of extended RBAC models have been proposed over the years to meet
+security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
 
@@
 
-The RBAC standard was designed with enterprises in mind such that when practioners implemented RBAC into their systems
+The RBAC standard was designed with enterprises in mind such that when practitioners implemented RBAC into their systems
 there would be a reasonable assurance being based off a well thought out model.  As extensions to the standard model
@@
 with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as 4 papers did, should be
-a prototype implementation of the model for review by both pracitioners and researchers. Of the models that produced an 
+a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
 implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
@@
 Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
-evaluation types and which primary sources provided which type.  In some cases a single primary source provided mutliple evaluation
+evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
 types.
@@
 access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
-extenion models are lacking.
+extension models are lacking.
 
-The primary source of validation a developer or practioner may encounter is a qualitative discussion of real-world
+The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
 scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
@@
 one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
-of how an extension model handles a real-world scenario provides developers and practioners ancedotal evidence at best
+of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
 for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
@@
 uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
-scenarios might serve to illuminate a trend worth further examination into the reasaon for the explosion of
+scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
 RBAC extensions.  The authors identified domains presented within the primary sources by looking for example
@@
 was introduced. 
-The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
+The domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
 For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.

### 2013-04-25 paper/sections/discussion.tex
Add discussion section
add discussion section and update bib file as well

\section{Discussion} \label{sec:discussion}
%\subsection{Find appropriate extended Role-based access control models for a system}

This section discusses two issues: first one is about evaluation of extended role-based access control models and second one is to implement security system based on access control models.

\subsection{Metrics for Extended Access Control Models Evaluation }

Researchers provide metrics for access control system evaluation or security system in general~\cite{hu2012:NIST7874}. They collected metrics, which can be applicable for not only access control models, but also requirements and system implementation, and extended application such as verification and testing. From the metrics, we select metrics that can be applicable for evaluation of extended RBAC models as follows:

\begin{itemize}
	\item \textbf{Ease of Privilege Assignments}: this metric measures number of steps for assigning, changing, deleting a privilege for users and roles when an end-user or administrators manage RBAC. If one takes more steps for such a task, it may give more chance to make human or system errors due more steps. If you takes less step, it may reduce to make mistakes.  While extended RBAC may has similar features since from RBAC, we measure worst cases after extended, and gap between these two models.

	\item \textbf{Syntactic and semantic support for specifying AC rules}: this metric checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints. This support helps specify complex and flexible access controls. For example, given a spatial RBAC model, one may measure whether spatial constraints can be specified using various logic expressions.

	\item \textbf{Delegation of administrative capabilities}: this metric measure how easy and secure when an access control model supports for privilege delegation of administrators to another user.

	\item \textbf{Capable of combining different policy models}: this metric measures whether an access model has a feature to combine more than two policies.

	\item \textbf{Capable of conflict resolution}: These metric checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.

	\item \textbf{Least privilege principle support}: This metric checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning. Least privilege principle support can reduce damage when malicious users gain legitimate privileges.

	\item \textbf{Separation of Duty (SoD)}: This metric checks whether an access model supports for static SoD and/or dynamic SoD. Note that NIST Standard supports both static SoD and dynamic SoD described in Section~\ref{sec:core-rbac}.
\end{itemize}

One can measure these metrics for a given extended access control model. These metrics focus on comparing specific features support of access control models.

\subsection{Extended RBAC Adoption}

In order to adopt a model in practice, software developers should implement the model in real system environments based on intended use of the models. Typically, a model provides abstract formal representation of extended RBAC and its operation. To configure intended use (i.e., which subjects can access which resources), we define an access control policy, which specifies high-level rules such as which subjects can access to which resources. Given an access control model and policy, security mechanism provides low-level implementation in a system, which controls access of roles.

A formal representation of access control models is a critical step of designing high-level abstraction where which elements are used and how elements are operated.
One of the objectives of a formal representation is to help ensure the correct behaviors through formal verification. Given the formal representation of access control models, software developers may prove of properties (e.g., safety, consistency and completeness) and check whether these properties are satisfied.

While the concept of the RBAC model is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the RBAC model is can be interpreted in more than one interpretation due to its complexity. In order to bridge gap between RBAC standard and its implementation, NIST published RBAC Implementation and Interoperability standard (RIIS/ ANSI INCITS-459) in 2011. This standard specifies how to implement RBAC system, which is consistent with NIST RBAC standard. Moreover, this standard describes interoperability specification where one RBAC implementation can be translated to another one.

This standard does not provide a specific guideline for implementation of various extended RBAC models.
Moreover, when software developers implemented extended RBAC models, they require clear security requirements of extended role-based access control, which identifies security objectives, intended environments, and assumed threats. Moreover, software developers understand how this extended RBAC overlaps with other access control models when more than two access control models are integrated into a single system. A survey~\cite{o20102010} shows that over 50\% of users at organizations with more than 500 employees are given some of their permissions to access resources based on RBAC. To adopt RBAC into a system, organizations often use hybrid approaches, which combine RBAC and access control lists since specific user types, systems, and workflow may not be effective to manage access based on roles. Therefore, when RBAC is extended and combined with other access control models, software developers often require to provide a formal representation, which can support of proof of the model. Then, software developers implement extended models based on NIST standard of RBAC implementation. Especially, software developers should understand intended use of extended access control models to meet a new system requirement. For example, for a spatial RBAC model, software developers incorporate additional spatial constraints, which can be either static or dynamic, in practice into the RBAC model.

### 2013-04-25 paper/sections/categorization.tex
Adding some edits to the discussion section.

@@
 
-  \item Context: The extension model integrates contextual information into the RBAC standard model.
+  \item \textbf{Context}: The extension model integrates contextual information into the RBAC standard model.
 Context is defined as a user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
 
-  \item Constraint (Const): The extension model provides conditional restrictions on permissions of given roles.
+  \item \textbf{Constraint (Const)}: The extension model provides conditional restrictions on permissions of given roles.
 The constraint is either static or dynamic.
@@
 
-  \item Organizational (Org): The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
+  \item \textbf{Organizational (Org)}: The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
 Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
 
-  \item Privacy (Priv): The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
+  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
 
-  \item Task: The extension model provides a task entity which are associated with permissions and roles.
+  \item \textbf{Task}: The extension model provides a task entity which are associated with permissions and roles.
 A task is a fundamental unit of a business activity.
@@
 
-  \item Spatio-Temporal: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
 For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
 
-  \item Spatial: The extension model provides spatial constraints, location-based constraints in specifying access
+  \item \textbf{Spatial}: The extension model provides spatial constraints, location-based constraints in specifying access
 	control policies. For example, in organizations, locations are enforced while a
@@
 
-  \item Temporal (Temp):  The extension model provides temporal constraints, time-based constraints in specifying access
+  \item \textbf{Temporal (Temp)}:  The extension model provides temporal constraints, time-based constraints in specifying access
 	control policies. For example, in organizations, periodic temporal durations are enforced while a

### 2013-04-25 paper/sections/discussion.tex
Adding some edits to the discussion section.

@@
 \section{Discussion} \label{sec:discussion}
-%\subsection{Find appropriate extended Role-based access control models for a system}
 
-This section discusses two issues: first one is about evaluation of extended role-based access control models and second one is to implement security system based on access control models.
+The research questions we identified, presented results for and analyzed provide specific views of portions of the extension model landscape. 
+By stepping back, and looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
+We break the discussion into two issues: recommendations for the enhancement of the evaluation of extensions to role-based access control models and guidelines for developers needing to choose an extension model to meet their requirements.
 
-\subsection{Metrics for Extended Access Control Models Evaluation }
+\subsection{Metrics for the Evaluation of Extensions to RBAC}
 
-Researchers provide metrics for access control system evaluation or security system in general~\cite{hu2012:NIST7874}.
-They collected metrics, which can be applicable for not only access control models, but also requirements and system implementation, and extended application such as verification and testing.
-From the metrics, we select metrics that can be applicable for evaluation of extended RBAC models as follows:
+A small percentage of primary sources presented substantial evaluations or implementations of their model in practice.
+This in turn leaves researchers and practioners little data from which to compare one model to the next.
+And given the wide swath of popluar and critical domains being targeted by RBAC extension models, we propose a set of metrics by which RBAC extension models should be evaluated.
+Hu, Vincent C.
+and Scarfone, Karen provide metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
+Hu, Vincent C.
+and Scarfone, Karen collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing.
+From these metrics, and guided by the data surrounding how extension models are currently evaluated, we select metrics that can be applicable for evaluation of extended RBAC models as follows:
 
 \begin{itemize}
-	\item \textbf{Ease of Privilege Assignments}: this metric measures number of steps for assigning, changing, deleting a privilege for users and roles when an end-user or administrators manage RBAC.
-If one takes more steps for such a task, it may give more chance to make human or system errors due more steps.
-If you takes less step, it may reduce to make mistakes.
-While extended RBAC may has similar features since from RBAC, we measure worst cases after extended, and gap between these two models.
+	\item \textbf{Ease of Privilege Assignments}: measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manage RBAC.
+If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity.
+While extended RBAC may have similar features to that of the RBAC standard model, we measure worst cases after extended, and the gap between an extension model and the RBAC standard model.
 
-	\item \textbf{Syntactic and semantic support for specifying AC rules}: this metric checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints.
+	\item \textbf{Syntactic and semantic support for specifying AC rules}: checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints.
 This support helps specify complex and flexible access controls.
@@
 
-	\item \textbf{Delegation of administrative capabilities}: this metric measure how easy and secure when an access control model supports for privilege delegation of administrators to another user.
+	\item \textbf{Delegation of administrative capabilities}: measures how easy and secure an access control model supports for privilege delegation of administrators to another user.
 
-	\item \textbf{Capable of combining different policy models}: this metric measures whether an access model has a feature to combine more than two policies.
+	\item \textbf{Capable of combining different policy models}: measures whether an access model has a feature to combine more than two policies.
 
-	\item \textbf{Capable of conflict resolution}: These metric checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.
+	\item \textbf{Capable of conflict resolution}: checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.
 
-	\item \textbf{Least privilege principle support}: This metric checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning.
+	\item \textbf{Least privilege principle support}: checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning.
 Least privilege principle support can reduce damage when malicious users gain legitimate privileges.
 
-	\item \textbf{Separation of Duty (SoD)}: This metric checks whether an access model supports for static SoD and/or dynamic SoD.
+	\item \textbf{Separation of Duty (SoD)}: checks whether an access model supports for static SoD and/or dynamic SoD.
 Note that NIST Standard supports both static SoD and dynamic SoD described in Section~\ref{sec:core-rbac}.
@@
 
-One can measure these metrics for a given extended access control model. These metrics focus on comparing specific features support of access control models.
+We recommend these metrics be incorporated in any extension models presented hence forth. A future work could include providing metric measurements for the primary sources prsented here. 
+A developer could attemot to apply these metrics to an extension model not already evaluated when attempting to choose a model. 
+These metrics focus on comparing specific features support of access control models.
 
-\subsection{Extended RBAC Adoption}
+\subsection{Adoption of Extensions to RBAC}
 

### 2013-04-25 paper/sections/analysis.tex
update analysis
update analysis

@@
 
+We first describe motivations described in 28 papers surveyed.
+
 \begin{itemize}
 \setlength{\itemsep}{0.25pt}
-\item Core RBAC model needs additional contextual information and constraints to develop fine-grained policies in practice.
-\item Core RBAC does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
-\item Core RBAC does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
-\item Core RBAC does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
-\item Core RBAC has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
-\item Model needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
-\item Model needs to incorporate additional attributes such as
+\item The RBAC standard needs additional contextual information and constraints to develop fine-grained policies in practice.
+\item The RBAC standard does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
+\item The RBAC standard does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
+\item The RBAC standard does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
+\item The RBAC standard has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
+\item The RBAC standard needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
+\item The RBAC standard needs to incorporate additional attributes such as
 task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new association between attributes
 are introduced. 
-\item Model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
+\item RBAC model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
 \end{itemize}
+
+We classify motivations behind RBAC extensions to these three categories:
+
+\begin{itemize}
+	\item Addition: an extended model incorporates additional elements with regards to context/constraints and their relations with existing elements of the RBAC standard.
+Models in this category does not change the core model.
+We found 23 papers in this category.
+	\item Modification: an extended model incorporates modification of existing features in the RBAC standard.
+For example, changes of attribute relations such as role hierarchy is to show modification.
+For models in this category, changes occur within the core model.
+We found 1 papers in this category.
+	\item Combination: an extended model is combination of RBAC and an another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
+For this combination, new association between attributes of two models should be introduced.
+Models in this category requires a new association between attributes of different access control models to work together.
+We found 4 papers in this category.
+\end{itemize}
+
+%\begin{table}
+%\centering
+%\caption{Motivation categories found and the count of primary sources}
+%\begin{tabular}{ | c | c | }
+%\cline{1-2}
+%
+%\textbf{Motivation Category} & \textbf{Paper Count} \\ \cline{1-2}
+%Addition & 23 \\ \cline{1-2}
+%Modification & 1 \\ \cline{1-2}
+%Combination & 4 \\
+%
+%\cline{1-2}
+%\end{tabular}
+%\label{tab:motivations}
+%\end{table}
+%
+%Table \ref{tab:motivations} shows the breakdown of motivations found within the primary sources.
+%The result shows that 23 papers out of 28 papers are classified into ``Addition'' category. Only
+%one paper and four papers are classified into ``Modification'' category and ``Combination'' category, respectively.
 
@@
 
-The RBAC standard model provides an abstraction on top of authorization based on user role assignment (UA) and role
+The RBAC standard provides a formal access model, which is an abstraction on top of authorization based on user role assignment (UA) and role
 permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
@@
 
-The RBAC standard model presents four entities: roles, sessions, users, and permissions, and their relations.
+The RBAC standard presents four entities: roles, sessions, users, and permissions, and their relations.
 While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices.
-Such extended RBAC models are formally presented to describe its behaviors.
+In such environments, the RBAC standard needs additional constraints and dynamic context to support dynamic environments such as spatial and location.
+For this category, more than 80\% of papers are classified into ``Addition'' category.
+Besides, one paper~\cite{alam06:constraint} modifies relationship among elements to support partial inheritence.
+RBAC is shown to be combined with other access control models such as team-based access control and task-based access control~\cite{zhou2007team, oh2003task}.
+
+%Such extended RBAC models are formally presented to describe its behaviors.
 In the selected papers, researchers illustrate how an extended model incorporates new element based on the correct RBAC and describe their relations with pre-defined elements in the core RBAC.
-
-We classify motivations behind RBAC extensions to these three categories:
-
-\begin{itemize}
-	\item Addition: an extended model incorporates additional elements with regards to context/constraints and their relations with existing elements of the core RBAC.
-Models in this category does not change the core model.
-We found OO papers in this category.
-	\item Modification: an extended model incorporates modification of existing features in the core RBAC.
-For example, changes of attribute relations such as role hierarchy is to show modification.
-For models in this category, changes occur within the core model.
-We found OO papers in this category.
-	\item Combination: an extended model is combination of RBAC and an another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
-For this combination, new association between attributes of two models should be introduced.
-Models in this category requires a new association between attributes of different access control models to work together.
-We found OO papers in this category.
-\end{itemize}
 
@@
 
-Core or any extended role-based access control is used in various aspects of computer systems.
+The RBAC standard is used in various aspects of computer systems.
 In order to reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.

### 2013-04-25 paper/sections/core_rbac.tex
minor update (spell) for background

@@
 Constrained RBAC incorporates static and dynamic separation of duty relations to the RBAC model. Separation of 
-duty relations enforce conflict of interest among roles. This model defines two types of separation of duty relations; static and dynamic.
+duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations; static and dynamic.
 
@@
 	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$.
-SSoD is known to be too rigid for practical use in cases where a user should have permissions as either $Role_A$ and $Role_B$.
+SSoD is known to be too rigid for practical use in cases where a user should have permissions when a user is assigned to both $Role_A$ and $Role_B$.
 	\item Dynamic Separation of Duty (DSD): Dynamic SoD is known to be
 more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
-	where multiple roles can be associated with a single user and , given a context, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.	
+	where multiple roles can be associated with a single user and, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.	
 \end{itemize}

### 2013-04-25 paper/sections/analysis.tex
fixing after spell check and grammar check based on word
spell check and grammar check based on word

@@
 task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new association between attributes
-are introduced. 
+is introduced. 
 \item RBAC model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
@@
 \begin{itemize}
-	\item Addition: an extended model incorporates additional elements with regards to context/constraints and their relations with existing elements of the RBAC standard.
-Models in this category does not change the core model.
+	\item Addition: The extension model incorporates additional elements with regards to context/constraints and their relations with existing elements of the RBAC standard.
+Models in this category do not change the RBAC standard model.
 We found 23 papers in this category.
-	\item Modification: an extended model incorporates modification of existing features in the RBAC standard.
-For example, changes of attribute relations such as role hierarchy is to show modification.
-For models in this category, changes occur within the core model.
-We found 1 papers in this category.
-	\item Combination: an extended model is combination of RBAC and an another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
+	\item Modification: The extension model incorporates modification of existing features in the RBAC standard.
+For example, changes of attribute relations such as role hierarchy are to show modification.
+For models in this category, changes occur within the the RBAC standard model.
+We found one paper in this category.
+	\item Combination: The extension model is combination of RBAC and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
 For this combination, new association between attributes of two models should be introduced.
-Models in this category requires a new association between attributes of different access control models to work together.
-We found 4 papers in this category.
+Models in this category require a new association between attributes of different access control models to work together.
+We found four papers in this category.
 \end{itemize}
@@
 For this category, more than 80\% of papers are classified into ``Addition'' category.
-Besides, one paper~\cite{alam06:constraint} modifies relationship among elements to support partial inheritence.
-RBAC is shown to be combined with other access control models such as team-based access control and task-based access control~\cite{zhou2007team, oh2003task}.
+Besides, one paper~\cite{alam06:constraint} describes modification of existing relationship among elements to support partial inheritance.
+The RBAC standard is shown to be combined with other access control models such as team-based access control and task-based access control~\cite{zhou2007team, oh2003task}.
 
@@
 usage by real software systems, bringing the model to life is strong evidence that the
-proposed model can work in practice. The concept of authorization, and access control
+proposed model can work in practice. The concept of authorization and access control
 is rooted in a business need. Thus, any access control model needs to be feasible
 in the real world not just on paper. We analyzed the primary sources to see how many
-proposed models actually had implementations associated with them.  And quantified the
-type of implementation. Whether the implementation was for a real system, for a prototype
+proposed models actually had implementations associated with them.  Then, we quantified
+types of implementations. Whether the implementation was for a real system, for a prototype
 and/or used in a production environment.
@@
 Of the 28 papers surveyed, there was a lack of implementation with 20 of the paper providing no
-mention of a implementation or prototype.  Of the remaining 8 papers that did mention an implementation, half 
+mention of an implementation or prototype.  Of the remaining eight papers that did mention an implementation, half 
 were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
@@
 in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
-with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as 4 papers did, should be
+with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as four papers did, should be
 a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
@@
 
-RQ4: How are extension to RBAC evaluated theoretically and in practice?
+RQ4: How are extensions to RBAC evaluated theoretically and in practice?
 
@@
 The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
-Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
+Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
 evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
@@
 scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rarer cases, 
+authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rare cases, 
 one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
@@
 
-The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 28 mentioning scenarios or requirements of that industry.
-Mobile computing and enterprise workflows were each represented by 5 papers claiming to be influenced by the requirements for access control within these domains. The final set
-of domains was pervasive computing environments and large-scale organizations with 3 each and web applications with 1.  There were 4 papers without any direct mention of a domain
+The predominant domain for which extension models have been generated for is that of the medical domain with nine of 28 mentioning scenarios or requirements of that industry.
+Mobile computing and enterprise workflows were each represented by five papers claiming to be influenced by the requirements for access control within these domains. The final set
+of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.  There were four papers without any direct mention of a domain
 since Aich et al. \cite{aich09:role} fall under both the medical domain and mobile computing.
@@
 The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of extension models.  
-Further, the authors noted that the categories associated with papers identifying the medical domain was not limited to one or two but cut across
+Further, the authors noted that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
 each of the eight categories except for Organization. 
@@
 
-\section{Generaliations} \label{sec:generalizations}
+\section{Generalizations} \label{sec:generalizations}
 
@@
 As access control is typically evaluated to either true (Permit) or false (Deny) based on predicates, propositional logic is sufficient to describe key ideas and definitions.
-We found that OO papers of RBAC extensions use propositional logic to describe its extended model.
+We found that 28 papers of RBAC extensions use propositional logic to describe its extended model.
 However, propositional logic has limitations.

### 2013-04-25 paper/sections/categorization.tex
fixing after spell check and grammar check based on word
spell check and grammar check based on word

@@
 
-We offer definitions for each of the 8 observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
+We offer definitions for each of the eight observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
 Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Table \ref{tab:categorization}.
@@
 
-  \item \textbf{Task}: The extension model provides a task entity which are associated with permissions and roles.
+  \item \textbf{Task}: The extension model provides task elements which are associated with permissions and roles.
 A task is a fundamental unit of a business activity.
@@
 For example, the employee role is associated with a task, which is to write a report.
-Then, this task is associated with a permission.
+Then, this task is associated with a permission element.
 
@@
 
-The 28 primary sources produced a set of 8 heirarchical categories. Table \ref{tab:categorization} summarizes each primary source under their designated
-category and further, displays the perceived heirarchy of the categories. 
+The 28 primary sources produced a set of eight hierarchical categories. Table \ref{tab:categorization} summarizes each primary source under their designated
+category and further, displays the perceived hierarchy of the categories. 
 The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and

### 2013-04-25 paper/sections/discussion.tex
fixing after spell check and grammar check based on word
spell check and grammar check based on word

@@
 A small percentage of primary sources presented substantial evaluations or implementations of their model in practice.
-This in turn leaves researchers and practioners little data from which to compare one model to the next.
-And given the wide swath of popluar and critical domains being targeted by RBAC extension models, we propose a set of metrics by which RBAC extension models should be evaluated.
-Hu, Vincent C.
-and Scarfone, Karen provide metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
-Hu, Vincent C.
-and Scarfone, Karen collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing.
+This in turn leaves researchers and practitioners little data from which to compare one model to the next.
+And given the wide swath of popular and critical domains being targeted by RBAC extension models, we propose a set of metrics by which RBAC extension models should be evaluated.
+Hu and Scarfone provide metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
+They collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing.
 From these metrics, and guided by the data surrounding how extension models are currently evaluated, we select metrics that can be applicable for evaluation of extended RBAC models as follows:
@@
 \begin{itemize}
-	\item \textbf{Ease of Privilege Assignments}: measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manage RBAC.
+	\item \textbf{Ease of Privilege Assignments}: measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC.
 If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity.
@@
 
-We recommend these metrics be incorporated in any extension models presented hence forth. A future work could include providing metric measurements for the primary sources prsented here. 
-A developer could attemot to apply these metrics to an extension model not already evaluated when attempting to choose a model. 
+We recommend these metrics be incorporated in any extension models presented hence forth. A future work could include providing metric measurements for the primary sources presented here. 
+A developer could attempt to apply these metrics to an extension model not already evaluated when attempting to choose a model. 
 These metrics focus on comparing specific features support of access control models.

### 2013-04-25 paper/sections/process.tex
Minor updates to process and categories.
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.
fixing after spell check and grammar check based on word
spell check and grammar check based on word

@@
-\section{Process} \label{sec:process}
+\section{Methodology and Process} \label{sec:process}
 
@@
 The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search terms.
-The search was performed by applying each search term to each engine incrementally until a stopping criteria was met.
+The search was performed by applying each search term to each engine incrementally until stopping criteria were met.
 Table \ref{tab:search_results} lists the four search engines along the left most column with the three search term across the top row along with the papers for each criteria and engine combination.
@@
 	\item If the size of the result set is greater than or equal to 1000 then stop.
-	\item If the last ten results did not contain the search term phrase within the tite then stop.
+	\item If the last ten results did not contain the search term phrase within the title then stop.
 	\end{itemize}
@@
 \centering
+\caption{Paper counts after applying search strategy}
 \begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
@@
 \end{tabular}
-\caption{Paper counts after applying search strategy}
 \label{tab:search_results}
@@
 \item Does the abstract mention extension of role-based access control?
-\item Does the abstract mention either an implementation, evaluation or domain for their model?
+\item Does the abstract mention an implementation, evaluation or domain for their model?
 \end{itemize}
@@
 
-Question 1 was a definitive exclusion criteria as any paper that failed in the affirmative was rejected.
+Question 1 was a definitive exclusion criterion as any paper that failed in the affirmative was rejected.
 Questions 3 and 4 were marked as conditional includes given that they were connected in making a decision.
@@
 \centering
+\caption{Elimination Rounds}
 \begin{tabular}{ c c | c | c | c | }
@@
 \multicolumn{1}{ c| }{} & Retained & 41 & 21 & 5 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & \textbf{Total} & 290 & 72 & 29 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & \textbf{Total} & 290 & 72 & 28 \\ \cline{2-5}
 
 \end{tabular}
-\caption{Elimination Rounds}
 \label{tab:search_results}

### 2013-05-01 paper/sections/abstract.tex
address comments
John's comments

@@
 \noindent
-{\bf Context}: Since the introduction of the role based access control (RBAC) standard in the late 1990's by the National Institute of Standards and Technology (NIST),
-computing in new domains has led to proposals for extensions to the RBAC standard model; for example adding context around permission granting and role activation.
+{\bf Context}: Since the National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) standard using RBAC reference models (RBAC$_{m}$ in short) in the late 1990's, computing in new domains has led to proposals for extensions to the RBAC$_{m}$; for example adding context around permission granting and role activation.
 \\
@@
 \noindent
-{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
-how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model. We accomplished this through: 
-establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the 
-motivations that have led to extension of the RBAC standard model.
+{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
+how extensions to RBAC$_{m}$ are evaluated by providing an assessment of the state of extensions to the RBAC$_{m}$. We accomplished this through: 
+establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models, and a breakdown of the 
+motivations that have led to extension of the RBAC$_{m}$.
 \\
@@
 \noindent
-{\bf Method}: We performed a systematic literature review that yielded 1716 papers, of which 28 were deemed primary sources for inclusion as extensions to the RBAC standard model. 
+{\bf Method}: We performed a systematic literature review that yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extensions to the RBAC$_{m}$. 
 \\
@@
 \noindent
-{\bf Results}: Our results show that extensions to RBAC can be classified under eight categories: Constraint, Context, 
-Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We identified only
+{\bf Results}: Our results show that extensions to the RBAC$_{m}$ can be classified under eight categories: Constraint, Context, 
+Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal. We identified only
 eight of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
@@
 {\bf Conclusions}: When examined as an entire group, the eight categories can be aligned under the single category of context. 
-Our literature review showed that the state of evaluation of RBAC model evaluation is limited with one
-paper doing an evaluation comparing itself to the core RBAC model, eight presenting example 
-scenarios of their model in action, and 12 of the 28 models lacking any identifiable evaluation. The landscape of extensions
-to the RBAC standard model serves as a basis from which a new, and vetted, extended standard RBAC model should be derived.
+Our literature review shows that the state of evaluation of RBAC model evaluation is limited with one
+paper doing an evaluation comparing itself to the RBAC$_{m}$, eight presenting example 
+scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern. The landscape of extensions
+to the RBAC$_{m}$ serves as a basis from which a new, and vetted, extended RBAC reference models should be derived.
 

### 2013-05-01 paper/sections/analysis.tex
address comments
John's comments

@@
 
-RQ2: What are the motivations behind extensions to RBAC?
+RQ2: What are the motivations behind RBAC extensions?
 
@@
 
-We first describe motivations described in 28 papers surveyed.
+We first describe motivations described in the 28 papers surveyed.
 
@@
 \setlength{\itemsep}{0.25pt}
-\item The RBAC standard needs additional contextual information and constraints to develop fine-grained policies in practice.
-\item The RBAC standard does not incorporate context. Therefore, RBAC belongs to static access control model, which may not capture changes in environments.
-\item The RBAC standard does not support for various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
-\item The RBAC standard does not provide an abstraction to additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
-\item The RBAC standard has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
-\item The RBAC standard needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
-\item The RBAC standard needs to incorporate additional attributes such as
-task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new association between attributes
-is introduced. 
-\item RBAC model needs to improve existing features such delegation and role hierarchy to meet fine-grained access control. 
+\item The RBAC$_{m}$ needs additional contextual information and constraints to develop fine-grained policies.
+\item The RBAC$_{m}$ does not incorporate context. Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
+\item The RBAC$_{m}$ does not support various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
+\item The RBAC$_{m}$ does not provide an abstraction for additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
+\item The RBAC$_{m}$ has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
+%\item The RBAC standard needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
+\item The RBAC$_{m}$ needs to incorporate additional attributes such as
+task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new associations between attributes
+are introduced. 
+\item The RBAC$_{m}$ needs to improve existing features such as delegation and role hierarchy to provide fine-grained access control. 
 \end{itemize}
@@
 	\item Addition: The extension model incorporates additional elements with regards to context/constraints and their relations with existing elements of the RBAC standard.
-Models in this category do not change the RBAC standard model.
+Models in this category do not change the RBAC$_{m}$.
 We found 23 papers in this category.
-	\item Modification: The extension model incorporates modification of existing features in the RBAC standard.
-For example, changes of attribute relations such as role hierarchy are to show modification.
-For models in this category, changes occur within the the RBAC standard model.
+	\item Modification: The extension model incorporates modification of existing features in the RBAC$_{m}$.
+For example, changes of attribute relations such as role hierarchy fall into this modification category.
+For models in this category, changes occur within the RBAC$_{m}$.
 We found one paper in this category.
-	\item Combination: The extension model is combination of RBAC and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, organizational roles.
-For this combination, new association between attributes of two models should be introduced.
-Models in this category require a new association between attributes of different access control models to work together.
+	\item Combination: The extension model is a combination of RBAC$_{m}$ and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, and organizational roles.
+For this combination, new associations between attributes of two models should be introduced.
+Models in this category require new associations between attributes of different access control models to work together.
 We found four papers in this category.
@@
 
-The RBAC standard provides a formal access model, which is an abstraction on top of authorization based on user role assignment (UA) and role
-permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
+The RBAC standard provides RBAC$_{m}$, which is an abstraction on top of authorization based on $UA$ and $PA$. Since RBAC$_{m}$ may not provide a fine-grained access control for a sophisticated
 security mechanism, a variety of extended RBAC models have been proposed over the years to meet
@@
 
-The RBAC standard presents four entities: roles, sessions, users, and permissions, and their relations.
-While this model is considered fundamental in any RBAC systems, this model can be extended to be applicable for emerging applications such as healthcare and mobile devices.
+RBAC$_{m}$ presents four entities: roles, sessions, users, and permissions, and their relations.
+While RBAC$_{m}$ is considered fundamental in any RBAC systems, RBAC$_{m}$ can be extended to be applicable for emerging applications such as healthcare and mobile devices.
 In such environments, the RBAC standard needs additional constraints and dynamic context to support dynamic environments such as spatial and location.
@@
 The RBAC standard was designed with enterprises in mind such that when practitioners implemented RBAC into their systems
-there would be a reasonable assurance being based off a well thought out model.  As extensions to the standard model
+there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC$_{m}$
 come along, thought and time should be given to how features and nuances of their models may impact implementation
@@
 would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
-or show how the standard model is deficient in tackling said scenarios.
+or show how the RBAC$_{m}$ is deficient in tackling said scenarios.
 
@@
 When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
-Further, in the case of extensions to the RBAC standard model, the model should be accompanied by validation of the model as a stand-alone
+Further, in the case of extensions to the RBAC$_{m}$, the model should be accompanied by validation of the model as a stand-alone
 access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
@@
 scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rare cases, 
+authors provide some discussion of how the RBAC$_{m}$ is deficient in tackling the scenario. In rare cases, 
 one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
 of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
-for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
-model, developers may be left implementing a more complex model to address their requirements when the standard model would
+for what types of situations the proposed model could handle.
+Further, by not providing a comparison to the RBAC$_{m}$, developers may be left implementing a more complex model to address their requirements when the RBAC$_{m}$ would
 have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
@@
 Two of the primary sources examined provided time-based performance analysis of their extension model compared
-to the RBAC standard model. This inclusion of time analysis provides some assurances to developers that any non-functional
-requirements surrounding time to compute authorizations compete or beat the standard model. Further, four of the models provided
+to the RBAC$_{m}$. This inclusion of time analysis provides some assurances to developers that any non-functional
+requirements surrounding time to compute authorizations compete or beat the RBAC$_{m}$. Further, four of the models provided
 some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
@@
 
-RQ5: What domains have extensions to RBAC been created for?
+RQ5: What domains have RBAC extensions been created for?
 
@@
 and researchers should take care when looking at extension models designed to addresses the newer requirements of enterprise workflows in order to
-achieve the same economic implementation and maintainability benefits the original standard model presents.
+achieve the same economic implementation and maintainability benefits the RBAC$_{m}$ presents.
 
@@
 This logic is sufficient to describe RBAC and extended RBAC.
-Moreover, this logic supports for concise and elegant formulation of the RBAC model and its relation.
+Moreover, this logic supports for concise and elegant formulation of the RBAC$_{m}$ and its relation.
 First-Order logic is expressive enough to concisely represent access control systems.

### 2013-05-01 paper/sections/categorization.tex
address comments
John's comments

@@
 
-During the paper reading phase, the authors noted that categorical themes were present among the extensions. 
-For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy directly within the title of the paper and the name of their model. 
-Some papers presented this direct pronouncement of a theme, while others were less obvious. 
-Thus, we developed a set of guidelines to aide in determining a set of extension categories based upon observations during the paper reading phase. 
-In developing these guidelines, we attempted to define each category by a single noun-phrase descriptor.
+During the paper reading phase, we identify categorical themes, which are
+central topic or context of RBAC extensions. 
+For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
+Some papers presented this direct pronouncement of theme of RBAC extensions, while others were less obvious. 
+Thus, we developed a set of guidelines to aide in determining a set of eight categories based upon observations during the paper reading phase. 
+In developing these guidelines, we defined each category by a single noun-phrase descriptor.
 The guidelines were:
@@
 
-If we take the previous example above, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in
+The previous example paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in
 the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
-By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name, but in reading the body of the paper the authors assessed that this paper was context based.
-
+By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name.
+However, in reading the body of the paper, we determined that this paper discussed RBAC extension based on context.
 We offer definitions for each of the eight observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
@@
   \item \textbf{Context}: The extension model integrates contextual information into the RBAC standard model.
-Context is defined as a user's current state and environment information (e.g., location, time, system resources, network state, network security configuration, etc) which may affect user's access privileges.
+Context is defined as a user's current state and environment (e.g., location, time, system resources, network state, network security configuration, etc) which may affect the user's access privileges.
 
@@
 The constraint is either static or dynamic.
-For example, a doctor can modify any medical record for which the doctor is assigned as the designated primary care physician.
-This example describes a doctor's permission with conditional restriction - ``only when the doctor is assigned as the designated primary care physician can the doctor modify a particular medical record''.
+For example, a doctor may modify any medical record for which the doctor is assigned as the designated primary care physician.
+This example describes a doctor's permission with a conditional restriction - ``only when the doctor is assigned as the designated primary care physician may modify a particular medical record''.
 
@@
 
-  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclose and management of personally identifiable information such as name, address, data of birth, etc.
+  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, data of birth, etc.
 
@@
 Different from core RBAC, in task-role-based access control model, roles are not directly associated with permissions.
-Roles are first associated with tasks, which are associated permissions.
-For example, the employee role is associated with a task, which is to write a report.
-Then, this task is associated with a permission element.
+Roles are associated with tasks that are associated with permissions.
+For example, the employee role is associated with a task, for example, to write a report.
+Then, this task is associated with a permission.
 
   \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
-For example, specific locations are enforced while a role is permitted to conduct an action from 8 am to 5 pm.
+For example, specific locations permit roles to conduct actions from 8:00 a.m.
+to 5:00 p.m.
 
@@
 	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
-	In such cases, the part-time employee role should access required resources only when the user is in the location. 
+	In such cases, the part-time employee role should access required resources only when the user is in that location. 
 	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.
@@
 	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
-	In such cases, the part-time employee role should access required resources during the interval. 
+	In such cases, the part-time employee role should be allowed to access required resources during the interval. 
 	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.
@@
 The 28 primary sources produced a set of eight hierarchical categories. Table \ref{tab:categorization} summarizes each primary source under their designated
-category and further, displays the perceived hierarchy of the categories. 
+category and furthermore, displays the perceived hierarchy of the categories. 
 The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and
@@
 When looking across all categories, the authors noted that each category added some new features on top of the standard RBAC model that were domain specific.
-These domain specific features were under the surface adding contextual relationships between the core user, permission and role entities.  Thus, the authors
+These domain specific features were under the surface these features adding contextual relationships between the core user, permission and role entities.  Thus, the authors
 concluded that all categories stemmed from the context category, of which some primary sources were already deemed direct members.

### 2013-05-01 paper/sections/conclusion.tex
fixing after spell check and grammar check based on word
spell check and grammar check based on word
address comments
John's comments

@@
 
-The extensions to RBAC were revealed to fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial and Temporal falling under the general category of context.
+The RBAC extensions were revealed to fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context.
 The categories each had properties specific to their implementation, but were seen to generalize to being specialized instances of context tailored to the entities or actions the categories covered.
-A number of domains were idenitfied as being the motivations behind needing extensions to the core RBAC model.  The domains, such as healthcare, presenting new challenges the previous models were not
+A number of domains were identified as being the motivations behind needing extensions to the core RBAC model.  The domains, such as healthcare, presenting new challenges the previous models were not
 required to design for.  Our literature review showed that the state of RBAC model evaluation needs focused from the research community given most model evaluations seen within the papers were

### 2013-05-01 paper/sections/core_rbac.tex
address comments
John's comments

@@
-\section{Role Based Access Control Standard} \label{sec:core-rbac}
+\section{Role-Based Access Control Standard} \label{sec:core-rbac}
 
@@
 Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
-This lack of consistent terminology was shown to slow the implementation of RBAC \cite{o20102010}.  
+This lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
 Moreover, in cases where organizations were concerned with adopting RBAC,
@@
 
-NIST's work can directly benefit organizations by lowering cost of early phase R\&D and the implementation of RBAC.
+NIST's work can directly benefit organizations by lowering cost of early phase Research and Development, and implementation of RBAC.
 Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the the rate of RBAC adoption has rapidly grown over recent years \cite{o20102010}. 
-The analysts estimate that, by the end of 2010, at least a part of permission of systems use RBAC for more than 50\% of IT users at organizations, which
-hire more than 500 employees.
-For economic estimate analysis, the analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry, of which \$1.1 billion is attributable to NIST RBAC standard work.
+The analysts estimate that by the end of 2010 at least a portion of permission of systems will use RBAC for more than half of users at organizations with more than 500 employees.
+The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard work saved \$1.1 billion.
 
-The NIST RBAC standard model proposed by Ferraiolo et al.
-\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three levels of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
-Since the basis for our review is extensions to the RBAC standard model, we describe the core model, associated entities and other terminology encountered across the space of our review.
-Then, we describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into the core RBAC.
+The NIST RBAC standard proposed by Ferraiolo et al.
+\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
+Each component includes a RBAC reference model.
+RBAC$_{m}$ includes RBAC reference models of these three components.
+We first describe the core RBAC, associated entities and other terminology encountered across the space of our review.
+We next describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into the core RBAC.
 
@@
 
-The four entities of the standard RBAC model are:
+The four entities of the core RBAC reference model are:
 
@@
 In RBAC, a user can exercise a permission only if the user is assigned to a role.
-In addition to the four basic components, two functions are defined:
-the user role assignment (UA) and the role and permission assignment (PA) functions.
-User role assignment models assignment of users to roles.
-Permission role assignment models assignment of permissions to roles.
+In addition to the four basic entities, two functions are defined:
+user assignment ($UA$) and permission assignment ($PA$) functions.
+$UA$ assignment of users to roles.
+$PA$ assignment of permissions to roles.
 
@@
         \includegraphics[width=4.0in]{sections/core-model.png}
-    \caption{\label{fig:overview}RBAC Core Diagram in NIST RBAC model \cite{ferraiolokuhn}.}
+    \caption{\label{fig:overview}Diagram in Core RBAC reference model\cite{ferraiolokuhn}.}
 \end{figure}
 
-Figure~\ref{fig:overview} presents an overview of core RBAC model diagram where elements and their relations are described.
+Figure~\ref{fig:overview} presents an overview of core RBAC reference model diagram where elements and their relations are described.
 Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
@@
 
-Hierarchical RBAC model adds role hierarchies (RH) feature to core RBAC model.
-This model incorporates a structure of roles in an organization using inheritance relationship among attributes such as roles.
+The hierarchical RBAC reference model adds role hierarchies ($RH$) feature to the core RBAC reference model.
+$RH$ incorporates a structure of roles in an organization using inheritance relationships among attributes such as roles.
 The role structure in an organization may use
-a role $r_1$, who inherits all permissions of another role $r_2$.
+a role $r_1$, which inherits all permissions of another role $r_2$.
 For example, a manager role may inherit all permissions of a employee role.
 Role hierarchies help simplify access control policy creation and maintenance by reducing the number of
-individual role assignments to a user.
-Formally, role inheritance relation is shown as \textit{RH} $\subseteq$ \textit{ROLES} $\times$ \textit{ROLES} describing many-to-many mapping role inheritance relation.
-General role hierarchies can be extended to use the concept of multiple inheritance where
-a role $r_1$, who inherits all permissions from more than one roles.
+individual role assignments for a user.
+Formally, the role inheritance relation is shown as \textit{RH} $\subseteq$ \textit{ROLES} $\times$ \textit{ROLES} describing the many-to-many mapping role inheritance relation.
+General role hierarchies can be extended to use the concept of multiple inheritances where
+$r_1$ inherits all permissions from more than one roles.
 
@@
 
-Constrained RBAC incorporates static and dynamic separation of duty relations to the RBAC model. Separation of 
+The constrained RBAC reference model incorporates separation of duty relations to  the core or the hierarchical RBAC reference. Separation of 
 duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations; static and dynamic.
@@
 \begin{itemize}
-	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-role assignments statically that are associated with a user. On situations
-	where multiple roles can be associated with a single user and roles $Role_A$ and $Role_B$ are conflicting each other, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$.
+	\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-roles, which can be assigned to a single user statically.
+Consider that roles $Role_A$ and $Role_B$ are conflicting with each other.
+On situations
+	where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ statically.
 SSoD is known to be too rigid for practical use in cases where a user should have permissions when a user is assigned to both $Role_A$ and $Role_B$.
 	\item Dynamic Separation of Duty (DSD): Dynamic SoD is known to be
-more flexible than SSD. DSD restricts the conflicting-role assignments dynamically that are associated with a user. On situations
-	where multiple roles can be associated with a single user and, roles $Role_A$ and $Role_B$ are conflicting each other dynamically, no permission is given to a user.	
+more flexible than SSoD.
+DSD restricts the conflicting-role assignments dynamically that are associated with a user.
+Consider that roles $Role_A$ and $Role_B$ are conflicting with each other.
+For the situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ dynamically.
 \end{itemize}

### 2013-05-01 paper/sections/discussion.tex
address comments
John's comments

@@
 
-\subsection{Metrics for the Evaluation of Extensions to RBAC}
+\subsection{Metrics for the Evaluation of RBAC Extensions}
 
@@
 If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity.
-While extended RBAC may have similar features to that of the RBAC standard model, we measure worst cases after extended, and the gap between an extension model and the RBAC standard model.
+While extended RBAC may have similar features to that of the RBAC$_{m}$, we measure worst cases after extended, and the gap between an extension model and the RBAC$_{m}$.
 
@@
 
-While the concept of the RBAC model is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the RBAC model is can be interpreted in more than one interpretation due to its complexity.
+While the concept of the RBAC$_{m}$ is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the RBAC$_{m}$ is can be interpreted in more than one interpretation due to its complexity.
 In order to bridge gap between RBAC standard and its implementation, NIST published RBAC Implementation and Interoperability standard (RIIS/ ANSI INCITS-459) in 2011.

### 2013-05-01 paper/sections/intro.tex
fixing after spell check and grammar check based on word
spell check and grammar check based on word
address comments
John's comments

@@
 
-Role based access control (RBAC) is used for maintaining and managing an organization's 
+Role-Based Access Control (RBAC) is widely used for maintaining and managing an organization's 
 access control based on assigning permissions to roles, and roles to users, instead of 
 assigning the permissions directly to individual users.
-The use of an RBAC model has become popular since the National Institute for Standards and Technology (NIST) 
-first proposed the RBAC standard model in the 1990s \cite{o20102010}. The RBAC standard model is widely used for securing 
+RBAC is used for securing 
 various applications including web services, database applications, and healthcare applications. 
-In the RBAC standard model, roles represent a set of permissions needed to perform a particular job function within an organization.  
-One to many users who are involved in that specific job function within the organization can then be 
-assigned to the role to inherit the required access instead of assigning individual permissions to individual users. 
-The ability to logically group both users and permissions into smaller total entities becomes paramount for managing access control
-as an organization grows and the number of permissions and users scales upward, 
-The RBAC standard model has advantages for managing access control. For example, if a user requires access to resources 
-associated with the manager role within a given organization, security policy administrators need only associate the user with the manager role.
+In RBAC, roles represent a set of permissions needed to perform a particular job function within an organization.  
+Multiple users who are involved in that specific job function within the organization can then be 
+assigned to a single role to inherit the required access. 
+The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows
+and the number of permissions and users scales upward.
+As permissions can be managed by a role instead of a user, RBAC has an advantage to significantly reduce complexity of security
+administration.
+For example, if a user requires access to resources 
+associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead
+of assigning permissions of the manager role to the user.
 
-The RBAC standard was first proposed in the 1990s when NIST 
+The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) 
+first proposed RBAC standard in the 1990s \cite{o20102010}.
+The RBAC standard was first proposed when NIST 
 requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
@@
 Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on a unified definition of RBAC or feature set.
-The RBAC standard addressed this issue by providing definitions and a functional specification of RBAC features using a reference model.
+The RBAC standard includes RBAC reference models (RBAC$_{m}$ in short).
+RBAC$_{m}$ serve as a basis for defining scope of RBAC features and functional specifications of RBAC features.
 
-Innovation and the use of RBAC in new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle \cite{kuhn2010adding}.  
-For example, Ni, Q. et al \cite{ni2010privacy} state that RBAC is ``not designed to enforce privacy policies and barely 
+Innovation and the use of RBAC in new domains has led to scenarios that, by some accounts, the RBAC$_{m}$ cannot handle \cite{kuhn2010adding}.  
+For example, Ni et al. \cite{ni2010privacy} state that RBAC$_{m}$ are ``not designed to enforce privacy policies and barely 
 meet privacy protection requirements'' with the introduction of privacy concerns in domains such as healthcare. 
-Since the introduction of the NIST RBAC standard in 2000, extension models to the RBAC standard have appeared in the literature, each adding one or more features on top of the core RBAC model. 
-A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
+Since the introduction of the RBAC standard, extension models to the RBAC$_{m}$ have appeared in the literature, each adding one or more features on top of components in the RBAC$_{m}$.
+%A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
 Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
 
-\textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC standard model, and in understanding
-how extensions to RBAC are evaluated by providing an assessment of the state of extensions to the RBAC standard model. We accomplished this through: 
-establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models and a breakdown of the 
-motivations that have led to extension of the RBAC standard model.} With respect to our goal, the authors addressed the following research questions:
+\textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
+how extensions to RBAC$_{m}$ are evaluated by providing an assessment of RBAC extensions. We accomplished this through: 
+establishment of a set of RBAC extension categories, an examination of the state of the art in evaluations of RBAC extension, and a breakdown of the 
+motivations that have led to extensions to RBAC$_{m}$.} To accomplish this goal, we seek to answer following research questions:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item RQ1: What categories exist within extensions to RBAC?
-\item RQ2: What are the motivations behind extensions to RBAC?
-\item RQ3: Do the extensions to RBAC have corresponding implementations?
-\item RQ4: How are extensions to RBAC evaluated theoretically and in practice?
-\item RQ5: What domains have extensions to RBAC been created for?
+\item RQ1: What categories exist within extensions to RBAC$_{m}$?
+\item RQ2: What are the motivations behind extensions to RBAC$_{m}$?
+\item RQ3: Do the extensions to RBAC$_{m}$ have corresponding implementations?
+\item RQ4: How are extensions to RBAC$_{m}$ evaluated theoretically and in practice?
+\item RQ5: What domains have extensions to RBAC$_{m}$ been created for?
 \item RQ6: What commonalities or generalizations exist across all categories?
@@
 
-The authors performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC standard model. 
-The review we performed yielded 1716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC standard model.  
-This review is intended to serve as a starting place for researcher's looking to tackle new problems in the realm of authorization and a reference guide for discovering
-the current state of extensions to the RBAC standard model. 
-For developer's looking to find a model to fit their potentially unique access control needs, this review will provide a basis for comparison and easy look-up for what extension model to use.  
-Further, the review provides insight into the state of the art in evaluating RBAC extension models.
+We performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC$_{m}$. 
+The review we performed yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC$_{m}$.  
+This review is intended to serve as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
+the current state of extensions to the RBAC$_{m}$. 
+For developers looking to find a model to fit their potentially unique access control needs, this review provides a basis for comparison and easy look up for what extension model to use.  
+Furthermore, the review provides insight into the state of the art in evaluating RBAC extension models.
 
-Our research provides contributions to the community in the following ways:
+Our research provides contributions to the community in following ways:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item Summarizes current research on extensions to RBAC.
-\item Guides a direction for a new standard based on an extension to the RBAC standard model.
+\item Summarizes current research on extensions to RBAC$_{m}$.
+\item Guides a direction for a new standard based on an extension to the RBAC$_{m}$.
 \end{itemize}
@@
 The rest of the paper is organized as follows. 
-Section 2 covers background and the core RBAC model. 
-Section 3 lays out the process used in conducting the review. 
-Sections 4-9 present the analysis and discussion of the research questions. 
-Section 10 is a general discussion section brings together all the research questions.
-Section 11 contains the final conclusions.
+Section~\ref{sec:core-rbac} presents background and the RBAC standard. 
+Section~\ref{sec:process} presents methodology and process, which we used in conducting the systemic literature review. 
+Sections~\ref{sec:categorization}-\ref{sec:generalizations} present analysis and discussion of the research questions. 
+Section~\ref{sec:discussion} discusses issues about extensions to RBAC$_{m}$.
+Section~\ref{sec:conclusion} concludes the paper.

### 2013-05-01 paper/sections/process.tex
address comments
John's comments

@@
 
-The authors adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
-The systematic literature review process was broken down in to four stages and the rest of this section is broken down by each stage.
+We adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
+The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage.
 The stages were:
@@
 
-For the first phase of our systematic literature review, the authors developed a search strategy for finding papers.
+For the first phase of our systematic literature review, we developed a search strategy for finding papers.
 The search strategy was executed by an automated comprehensive search taking as input a set of academic search engines and a list of search terms.
-The search was performed by applying each search term to each engine incrementally until stopping criteria were met.
-Table \ref{tab:search_results} lists the four search engines along the left most column with the three search term across the top row along with the papers for each criteria and engine combination.
+The search was performed by applying each search term to each engine incrementally until the stopping criteria were met.
+Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination.
 The search algorithm was performed as follows:
@@
 The stopping criteria related to the last ten titles is meant to stop after relevant results are no longer being returned by the search engine.
-After gathering all 12 data sets, the authors combined the papers into a master list by systematically comparing the bibliographic information for each.
-The master list of 1716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
-Table \ref{tab:eliminations} shows the total number of papers selected by each author for each round and how many papers from the disjoint set for each round survived to the next round.
+After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
+%The master list of 1,716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
+Out of the master list of 1,716 papers, two reviewers, namely Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
+Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.
 
@@
 
-The elimination rounds were performed based on reading of the title, abstract and finally the paper itself.
-While each elimination stage had a unique set of criteria for elimination the general procedure for elimination for the researchers was as follows.
+The elimination rounds were conducted based on reading of the title, abstract, and finally the paper itself.
+While each elimination stage had a unique set of criteria for elimination, the general procedure for elimination for the researchers was as follows.
 
@@
 \setlength{\itemsep}{0.25pt}
-\item The two first authors independently classified papers as relevant, irrelevant or uncertain based off elimination criteria
+\item The two first authors independently classified papers as relevant, irrelevant or uncertain based on elimination criteria
 \item Those papers marked as relevant by both reviewers were kept and those marked irrelevant by both were thrown out.
-\item Papers marked as relevant, or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
+\item Papers marked as relevant or irrelevant by a single reviewer were combined with all papers marked as uncertain and discussed by both reviewers.
 From this discussion, papers were either thrown out or kept until the next round of the review.
@@
 
-The first round of elimination was performed by strict examination based on the title.  Each author was tasked with deciding on elimination by answering the following questions:
+The first round of elimination was performed by examination based on the title.  Each author was tasked with deciding on elimination by answering the following questions:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item Did the title contain a reference to 'role based access control' or 'RBAC'? 
+\item Did the title contain a reference to 'role-based access control' or 'RBAC'? 
 \item Did the title contain a reference to 'model'?
@@
 
-The second round of elimination was based on strict reading of the abstracts of papers that survived title elimination.  Researchers read each abstract and evaluated relevancy based off:
+The second round of elimination was based on reading of the abstracts of papers that survived title elimination.  Researchers read each abstract and evaluated relevancy based off:
 
@@
 \item Does the abstract mention extension of role-based access control?
-\item Does the abstract mention an implementation, evaluation or domain for their model?
+\item Does the abstract mention an implementation, evaluation, or domain for their model?
 \end{itemize}
@@
 
-The final elimination round involved taking the entire paper into consideration and answering five questions that would serve as the basis for elimination.
+The final elimination round involved reading the entire paper into consideration and answering five questions that would serve as the basis for elimination.
 The data collected by answering these questions served as the basis towards answering the research questions.
-The questions each reviewer attempted to answer based on the content of the paper was:
+Each reviewer seeks to answer following questions based on the content of the paper:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item Does this model extend the core model? (Exclusion)
-\item Do the researchers give evidence that RBAC needs extension? (Inclusion)
+\item Does this model extend the RBAC$_{m}$? (Exclusion)
+\item Do the researchers give evidence that RBAC$_{m}$ needs extension? (Inclusion)
 \item Was the paper and subsequent model inspired by a real world example?  (Conditional Inclusion)
@@
 Questions 3 and 4 were marked as conditional includes given that they were connected in making a decision.
-A paper that met Question 3 but not 4, or met 4 but not 3 would be included since in some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.
+A paper that met question 3 but not 4, or met 4 but not 3 would be included since in some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.
 
@@
 
-\multicolumn{1}{ |c| }{Eric}  &  & 305 & 86 & 46 \\ \hline
-\multicolumn{1}{ |c| }{Hwang} &  & 176 & 102 & 42 \\ \hline
+\multicolumn{1}{ |c| }{Reviewer 1}  &  & 305 & 86 & 46 \\ \hline
+\multicolumn{1}{ |c| }{Reviewer 2} &  & 176 & 102 & 42 \\ \hline
 
@@
 \end{tabular}
-\label{tab:search_results}
+\label{tab:eliminations}
 \end{table}
@@
 The first step was to take the individual data generated from the final elimination round and organize this information around the research questions.
-During the paper reading round, and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
-Thus, the first step undertaken was to answer the question of what categories exist for the RBAC model extensions and what papers feel into what categories.
+During the paper reading round and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
+Thus, the first step undertaken was to answer the question of what categories exist for extensions to RBAC$_{m}$ and what papers fell into what categories.

### 2013-05-01 paper/sections/analysis.tex
pdf upload and minior update

@@
 \begin{itemize}
-	\item Addition: The extension model incorporates additional elements with regards to context/constraints and their relations with existing elements of the RBAC standard.
+	\item Addition: The extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC$_{m}$.
 Models in this category do not change the RBAC$_{m}$.
@@
 For this category, more than 80\% of papers are classified into ``Addition'' category.
-Besides, one paper~\cite{alam06:constraint} describes modification of existing relationship among elements to support partial inheritance.
+Besides, one paper~\cite{alam06:constraint} describes modification of existing relationship among entities to support partial inheritance.
 The RBAC standard is shown to be combined with other access control models such as team-based access control and task-based access control~\cite{zhou2007team, oh2003task}.

### 2013-05-01 paper/sections/discussion.tex
pdf upload and minior update

@@
 
-A formal representation of access control models is a critical step of designing high-level abstraction where which elements are used and how elements are operated.
+A formal representation of access control models is a critical step of designing high-level abstraction where which entities are used and how these entities are operated.
 One of the objectives of a formal representation is to help ensure the correct behaviors through formal verification.

### 2013-05-01 paper/sections/intro.tex
pdf upload and minior update

@@
 In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
-The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010 \footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010\footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
 Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on a unified definition of RBAC or feature set.

### 2013-05-08 paper/sections/analysis.tex
Minor reference updates.

@@
 \item Comparison to standard RBAC \cite{bao08:role}, \cite{zou2009crbac}, \cite{zhang06:collaborative}, \cite{zhao2008flexible}, \cite{ray07:spatio}
-\item Mathematical modeling \cite{damian2007geo}, \cite{hansen2003spatial}, \cite{aich07:STARBAC}, \cite{chen08:spatio-temporal}, \cite{joshi05:generalized}
-\item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki104:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{ray07:spatio}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
+\item Mathematical modeling \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{aich07:STARBAC}, \cite{chen08:spatio-temporal}, \cite{joshi05:generalized}
+\item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki04:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
 \item Experimental analysis of the model

### 2013-05-08 paper/sections/abstract.tex
Updates to address grammatical issues.

@@
 {\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
-how extensions to RBAC$_{m}$ are evaluated by providing an assessment of the state of extensions to the RBAC$_{m}$. We accomplished this through: 
+how extensions to RBAC$_{m}$ are evaluated by providing an assessment of the state of extensions to the RBAC$_{m}$. We accomplished this through
 establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models, and a breakdown of the 
@@
 Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal. We identified only
-eight of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
-that inspired extensions were the medical domain with nine of the 28 models, enterprise workflows with five of 28, and mobile computing with five of 28.
+8 of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
+that inspired extensions were the medical domain with 9 of the 28 models, enterprise workflows with 5 of 28, and mobile computing with 5 of 28.
 \\
@@
 {\bf Conclusions}: When examined as an entire group, the eight categories can be aligned under the single category of context. 
-Our literature review shows that the state of evaluation of RBAC model evaluation is limited with one
-paper doing an evaluation comparing itself to the RBAC$_{m}$, eight presenting example 
+Our literature review shows that the state of evaluation of RBAC model evaluation is limited with 1
+paper doing an evaluation comparing itself to the RBAC$_{m}$, 8 presenting example 
 scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern. The landscape of extensions

### 2013-05-08 paper/sections/analysis.tex
Updates to address grammatical issues.

@@
 mention of an implementation or prototype.  Of the remaining eight papers that did mention an implementation, half 
-were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
+were simply prototypes developed by the authors whereas the other half were claimed to be implemented within a real
 system.
@@
 and how application of their model would apply and resolve the situation.  Six of the models provided some form of performance
-or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
-grew, and the size of the role space for the extension model compared to standard RBAC. Four models provided mathematical descriptions
-and analysis as a way to provide evaluation in the form of completeness. 
-The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
+or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
+as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
+Four models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
 would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
@@
 
-The predominant domain for which extension models have been generated for is that of the medical domain with nine of 28 mentioning scenarios or requirements of that industry.
+The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 28 mentioning scenarios or requirements of that industry.
 Mobile computing and enterprise workflows were each represented by five papers claiming to be influenced by the requirements for access control within these domains. The final set
@@
 inclusion of extension models targeted at the enterprise workflow domain is indicative of the expansion of requirements for enterprises. Developers
-and researchers should take care when looking at extension models designed to addresses the newer requirements of enterprise workflows in order to
+and researchers should take care when looking at extension models designed to address the newer requirements of enterprise workflows in order to
 achieve the same economic implementation and maintainability benefits the RBAC$_{m}$ presents.

### 2013-05-08 paper/sections/categorization.tex
pdf upload and minior update
Updates to address grammatical issues.

@@
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
-Some papers presented this direct pronouncement of theme of RBAC extensions, while others were less obvious. 
+Some papers presented this direct pronouncement of theme of RBAC extensions, whereas others were less obvious. 
 Thus, we developed a set of guidelines to aide in determining a set of eight categories based upon observations during the paper reading phase. 
 In developing these guidelines, we defined each category by a single noun-phrase descriptor.
-The guidelines were:
+The guidelines were
 
@@
 
-  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, data of birth, etc.
+  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or data of birth.
 
-  \item \textbf{Task}: The extension model provides task elements which are associated with permissions and roles.
+  \item \textbf{Task}: The extension model provides task entities which are associated with permissions and roles.
 A task is a fundamental unit of a business activity.
@@
   \item \textbf{Spatial}: The extension model provides spatial constraints, location-based constraints in specifying access
-	control policies. For example, in organizations, locations are enforced while a
+	control policies. For example, in organizations, locations are enforced whereas a
 	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
@@
   \item \textbf{Temporal (Temp)}:  The extension model provides temporal constraints, time-based constraints in specifying access
-	control policies. For example, in organizations, periodic temporal durations are enforced while a
+	control policies. For example, in organizations, periodic temporal durations are enforced whereas a
 	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
@@
 
-When looking across all categories, the authors noted that each category added some new features on top of the standard RBAC model that were domain specific.
-These domain specific features were under the surface these features adding contextual relationships between the core user, permission and role entities.  Thus, the authors
+When looking across all categories, we noted that each category added some new features on top of the standard RBAC model that were domain specific.
+These domain specific features were under the surface these features adding contextual relationships between the core user, permission and role entities.  Thus, we
 concluded that all categories stemmed from the context category, of which some primary sources were already deemed direct members.
 
-For example, in the case of the Privacy category, the models added entities such as purpose binding to represent within the model data collected for one purpose should not used for another purpose without user consent {\cite{ni2010privacy}.
+For example, in the case of the Privacy category, the models added entities such as purpose binding to represent within the model data collected for one purpose should not be used for another purpose without user consent {\cite{ni2010privacy}.
 While the new entity provided by the Privacy based models is inspired by domains such as healthcare where privacy is of legal concern, the underlying mechanism that drives purpose binding is providing context around making an access control decision.

### 2013-05-08 paper/sections/core_rbac.tex
pdf upload and minior update
Updates to address grammatical issues.

@@
 
-Figure~\ref{fig:overview} presents an overview of core RBAC reference model diagram where elements and their relations are described.
+Figure~\ref{fig:overview} presents an overview of core RBAC reference model diagram where entities and their relations are described.
 Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
@@
 a role $r_1$, which inherits all permissions of another role $r_2$.
-For example, a manager role may inherit all permissions of a employee role.
+For example, a manager role may inherit all permissions of an employee role.
 Role hierarchies help simplify access control policy creation and maintenance by reducing the number of
@@
 The constrained RBAC reference model incorporates separation of duty relations to  the core or the hierarchical RBAC reference. Separation of 
-duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations; static and dynamic.
+duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations: static and dynamic.
 

### 2013-05-08 paper/sections/discussion.tex
Updates to address grammatical issues.

@@
 A survey~\cite{o20102010} shows that over 50\% of users at organizations with more than 500 employees are given some of their permissions to access resources based on RBAC.
-To adopt RBAC into a system, organizations often use hybrid approaches, which combine RBAC and access control lists since specific user types, systems, and workflow may not be effective to manage access based on roles.
+To adopt RBAC into a system, organizations often use hybrid approaches, which combine RBAC and access control lists because specific user types, systems, and workflow may not be effective to manage access based on roles.
 Therefore, when RBAC is extended and combined with other access control models, software developers often require to provide a formal representation, which can support of proof of the model.

### 2013-05-08 paper/sections/intro.tex
Updates to address grammatical issues.

@@
 \textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
-how extensions to RBAC$_{m}$ are evaluated by providing an assessment of RBAC extensions. We accomplished this through: 
+how extensions to RBAC$_{m}$ are evaluated by providing an assessment of RBAC extensions. We accomplished this through 
 establishment of a set of RBAC extension categories, an examination of the state of the art in evaluations of RBAC extension, and a breakdown of the 

### 2013-05-08 paper/sections/process.tex
Updates to address grammatical issues.

@@
 The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage.
-The stages were:
+The stages were
 
@@
 Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination.
-The search algorithm was performed as follows:
+The search algorithm was performed as follows
 
@@
 Questions 3 and 4 were marked as conditional includes given that they were connected in making a decision.
-A paper that met question 3 but not 4, or met 4 but not 3 would be included since in some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.
+A paper that met question 3 but not 4, or met 4 but not 3 would be included because for some cases the real world examples served as research evaluations and without this conditional include the paper list size would be too small to be significant.
 

### 2013-05-10 paper/sections/abstract.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 \noindent
-{\bf Context}: Since the National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) standard using RBAC reference models (RBAC$_{m}$ in short) in the late 1990's, computing in new domains has led to proposals for extensions to the RBAC$_{m}$; for example adding context around permission granting and role activation.
+{\bf Context}: Since the National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) standard using RBAC Reference Model in the late 1990's, computing in new domains has led to proposals for extensions to the RBAC Reference Model; for example adding context around permission granting and role activation.
 \\
@@
 \noindent
-{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
-how extensions to RBAC$_{m}$ are evaluated by providing an assessment of the state of extensions to the RBAC$_{m}$. We accomplished this through
-establishment of a set of extension categories, an examination of the state of the art in evaluations of extension models, and a breakdown of the 
-motivations that have led to extension of the RBAC$_{m}$.
+{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
+how RBAC extension models are evaluated by providing an assessment of the state of the RBAC extension models. We accomplished this through
+establishment of a set of extension categories, an examination of the state of the art in evaluations of the RBAC extension models, and a breakdown of the motivations that have led to the RBAC extension models.
 \\
@@
 \noindent
-{\bf Method}: We performed a systematic literature review that yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extensions to the RBAC$_{m}$. 
+{\bf Method}: We performed a systematic literature review that yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as the RBAC extension models. 
 \\
@@
 \noindent
-{\bf Results}: Our results show that extensions to the RBAC$_{m}$ can be classified under eight categories: Constraint, Context, 
+{\bf Results}: Our results show that the RBAC extension models can be classified under eight categories: Constraint, Context, 
 Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal. We identified only
@@
 Our literature review shows that the state of evaluation of RBAC model evaluation is limited with 1
-paper doing an evaluation comparing itself to the RBAC$_{m}$, 8 presenting example 
-scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern. The landscape of extensions
-to the RBAC$_{m}$ serves as a basis from which a new, and vetted, extended RBAC reference models should be derived.
+paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example 
+scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern.
+The landscape of the RBAC extension models serves as a basis from which a new, and vetted, RBAC extension model should be derived.
 

### 2013-05-10 paper/sections/analysis.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 
-RQ2: What are the motivations behind RBAC extensions?
+RQ2: What are the motivations behind the RBAC extension models?
 
@@
 \setlength{\itemsep}{0.25pt}
-\item The RBAC$_{m}$ needs additional contextual information and constraints to develop fine-grained policies.
-\item The RBAC$_{m}$ does not incorporate context. Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
-\item The RBAC$_{m}$ does not support various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
-\item The RBAC$_{m}$ does not provide an abstraction for additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
-\item The RBAC$_{m}$ has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
+\item The RBAC Reference Model needs additional contextual information and constraints to develop fine-grained policies.
+\item The RBAC Reference Model does not incorporate context. Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
+\item The RBAC Reference Model does not support various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
+\item The RBAC Reference Model does not provide an abstraction for additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
+\item The RBAC Reference Model has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
 %\item The RBAC standard needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
-\item The RBAC$_{m}$ needs to incorporate additional attributes such as
+\item The RBAC Reference Model needs to incorporate additional attributes such as
 task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new associations between attributes
 are introduced. 
-\item The RBAC$_{m}$ needs to improve existing features such as delegation and role hierarchy to provide fine-grained access control. 
+\item The RBAC Reference Model needs to improve existing features such as delegation and role hierarchy to provide fine-grained access control. 
 \end{itemize}
 
-We classify motivations behind RBAC extensions to these three categories:
+We classify motivations behind RBAC extension models to these three categories:
 
 \begin{itemize}
-	\item Addition: The extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC$_{m}$.
-Models in this category do not change the RBAC$_{m}$.
-We found 23 papers in this category.
-	\item Modification: The extension model incorporates modification of existing features in the RBAC$_{m}$.
+	\item Addition: The RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
+Models in this category do not change the RBAC Reference Model.
+	\item Modification: The RBAC extension model incorporates modification of existing features in the RBAC Reference Model.
 For example, changes of attribute relations such as role hierarchy fall into this modification category.
-For models in this category, changes occur within the RBAC$_{m}$.
-We found one paper in this category.
-	\item Combination: The extension model is a combination of RBAC$_{m}$ and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, and organizational roles.
+For models in this category, changes occur within the RBAC Reference Model.
+	\item Combination: The RBAC extension model is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, and organizational roles.
 For this combination, new associations between attributes of two models should be introduced.
 Models in this category require new associations between attributes of different access control models to work together.
-We found four papers in this category.
 \end{itemize}
+
+1 paper~\cite{alam06:constraint} is classified into ``Modification'' category since the paper describes modification of existing relationship among entities to support partial inheritance.
+2 papers~\cite{zhou2007team, oh2003task} are classified into ``Combination'' category since the papers describe an RBAC extension model, which is combination of the RBAC Reference Model and other access control models such as team-based access control and task-based access control.
+The remaining 25 papers are classified into ``Addition'' category.
 
@@
 
-The RBAC standard provides RBAC$_{m}$, which is an abstraction on top of authorization based on $UA$ and $PA$. Since RBAC$_{m}$ may not provide a fine-grained access control for a sophisticated
-security mechanism, a variety of extended RBAC models have been proposed over the years to meet
-security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
-
-RBAC$_{m}$ presents four entities: roles, sessions, users, and permissions, and their relations.
-While RBAC$_{m}$ is considered fundamental in any RBAC systems, RBAC$_{m}$ can be extended to be applicable for emerging applications such as healthcare and mobile devices.
+The RBAC standard provides the RBAC Reference Model, which is an abstraction on top of authorization based on $UA$ and $PA$.
+Since the RBAC Reference Model may not provide a fine-grained access control for a sophisticated
+security mechanism, a variety of RBAC extension models have been proposed over the years to meet
+security requirements.
+When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
+RBAC Reference Model presents four entities: roles, sessions, users, and permissions, and their relations.
+While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitation on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
 In such environments, the RBAC standard needs additional constraints and dynamic context to support dynamic environments such as spatial and location.
-For this category, more than 80\% of papers are classified into ``Addition'' category.
-Besides, one paper~\cite{alam06:constraint} describes modification of existing relationship among entities to support partial inheritance.
-The RBAC standard is shown to be combined with other access control models such as team-based access control and task-based access control~\cite{zhou2007team, oh2003task}.
 
@@
 
-RQ3: Do the extension models have corresponding implementations?
+RQ3: Do the RBAC extension models have corresponding implementations?
 
@@
 Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
-Of the 28 papers surveyed, there was a lack of implementation with 20 of the paper providing no
-mention of an implementation or prototype.  Of the remaining eight papers that did mention an implementation, half 
-were simply prototypes developed by the authors whereas the other half were claimed to be implemented within a real
-system.
+Of the 28 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
+system. The remaining 20 papers provide no
+mention of an implementation or prototype.
+
+%Of the remaining eight papers that did mention an implementation, half 
+%were simply prototypes developed by the authors whereas the other half were claimed to be implemented within a real
+%system.
 
@@
 
-The RBAC standard was designed with enterprises in mind such that when practitioners implemented RBAC into their systems
-there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC$_{m}$
+The RBAC standard is designed with enterprises in mind such that when practitioners implemented RBAC into their systems,
+there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC Reference Mode
 come along, thought and time should be given to how features and nuances of their models may impact implementation
@@
 \item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki04:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
-\item Experimental analysis of the model
+%\item Experimental analysis of the model
 \item Case study of the model in practice \cite{motta03:contextual}
@@
 
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Eight models presented example scenarios
-and how application of their model would apply and resolve the situation.  Six of the models provided some form of performance
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
+and how application of their model would apply and resolve the situation.  6 models provided some form of performance
 or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
 as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-Four models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
 The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
 would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
-or show how the RBAC$_{m}$ is deficient in tackling said scenarios.
+or show how the Reference Model is deficient in tackling said scenarios.
 
@@
 When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
-Further, in the case of extensions to the RBAC$_{m}$, the model should be accompanied by validation of the model as a stand-alone
+Further, in the case of extensions to the RBAC Reference Model, the model should be accompanied by validation of the model as a stand-alone
 access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
@@
 scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC$_{m}$ is deficient in tackling the scenario. In rare cases, 
-one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
+authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario. In rare cases, 
+one paper~\cite{motta03:contextual} from our results, a case study is performed to examine how the proposed model works in practice. Discussions
 of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
 for what types of situations the proposed model could handle.
-Further, by not providing a comparison to the RBAC$_{m}$, developers may be left implementing a more complex model to address their requirements when the RBAC$_{m}$ would
+Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
 have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
@@
 Two of the primary sources examined provided time-based performance analysis of their extension model compared
-to the RBAC$_{m}$. This inclusion of time analysis provides some assurances to developers that any non-functional
-requirements surrounding time to compute authorizations compete or beat the RBAC$_{m}$. Further, four of the models provided
+to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
+requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
 some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
@@
 scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
-RBAC extensions.  The authors identified domains presented within the primary sources by looking for example
+RBAC extensions.  We identified domains presented within the primary sources by looking for example
 scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
@@
 Mobile computing and enterprise workflows were each represented by five papers claiming to be influenced by the requirements for access control within these domains. The final set
-of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.  There were four papers without any direct mention of a domain
-since Aich et al. \cite{aich09:role} fall under both the medical domain and mobile computing.
+of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.
+There were three papers~\cite{jian2008extended, aich07:STARBAC, haibo2005context} without any direct mention of a domain since Aich et al.
+\cite{aich09:role} fall under both the medical domain and mobile computing.
 
@@
 
-The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of extension models.  
-Further, the authors noted that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
-each of the eight categories except for Organization. 
+\begin{figure}[ht]
+    \centering
+        \includegraphics[width=4.0in]{sections/dist_domains_byYear.eps}
+\vspace{-0.2 in}
+    \caption{\label{fig:dist_domains}Domain distribution by 3-year period. Y-axis represents the number of corresponding domain papers.
+    X-axis represents 3-year period from 2000 to 2011.}
+\end{figure}
+
+The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of RBAC extension models.  
+Moreover, we observed that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
+each of the eight categories except for the Organizational category. 
 The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
@@
 and researchers should take care when looking at extension models designed to address the newer requirements of enterprise workflows in order to
-achieve the same economic implementation and maintainability benefits the RBAC$_{m}$ presents.
+achieve the same economic implementation and maintainability benefits the RBAC Reference Model presents.
 
-Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
-was introduced. 
-The domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
-For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.
+Figure~\ref{fig:dist_domains} shows domain distribution by 3-year period.
+We observed that medical and enterprise domain papers constantly appear for every period in Figure~\ref{fig:dist_domains}.
+For a domain that has roots in medical and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity. 
+Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone was introduced.
+Since 2004, the domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
 
@@
 In order to reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
-We found that propositional logic is used to describe access control model across all categorizations.
+The authors use propositional logic to describe access control model across all categorizations.
 Propositional logic is concerned with propositions and their logical relationships.
@@
 Given context, propositional logic is used to evaluate true or false based on specified rules and access control logic.
-As access control is typically evaluated to either true (Permit) or false (Deny) based on predicates, propositional logic is sufficient to describe key ideas and definitions.
+As access control is typically evaluated to either true or false based on predicates, propositional logic is sufficient to describe key ideas and definitions.
 We found that 28 papers of RBAC extensions use propositional logic to describe its extended model.
-However, propositional logic has limitations.
-While this logic is simple, this logic does not support for reasoning about RBAC, which helps reduce the administrative complexity of associations such as user- role associations.
-An alternative logic to describe RBAC is first-order logic.
-This logic is sufficient to describe RBAC and extended RBAC.
-Moreover, this logic supports for concise and elegant formulation of the RBAC$_{m}$ and its relation.
+However, propositional logic has limitations. While this logic is simple, this logic does not support for reasoning about RBAC, which helps
+reduce the administrative complexity of associations such as user- role associations.
+In order to support for reasoning of access control, one may describe the RBAC extended models in first-order logic. 
+Given an RBAC extension model, Samuel et al.~\cite{samuel07:spatio-temporal} proposed verification of the model using a specification language, which is based first-order logic.
+This logic is sufficient to model RBAC and extended RBAC for reasoning.
+Moreover, this logic supports for concise and elegant formulation of the Reference Model and its relation.
 First-Order logic is expressive enough to concisely represent access control systems.

### 2013-05-10 paper/sections/categorization.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 
-RQ1: What categories exist within RBAC extensions?
+RQ1: What categories exist within the RBAC extension models?
 
@@
 During the paper reading phase, we identify categorical themes, which are
-central topic or context of RBAC extensions. 
+central topic or context of the RBAC extension models. 
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
-Some papers presented this direct pronouncement of theme of RBAC extensions, whereas others were less obvious. 
+Some papers presented this direct pronouncement of theme of the RBAC extension models, whereas others were less obvious. 
 Thus, we developed a set of guidelines to aide in determining a set of eight categories based upon observations during the paper reading phase. 
 In developing these guidelines, we defined each category by a single noun-phrase descriptor.
-The guidelines were
+The guidelines were as follows:
 
@@
 
-  \item \textbf{Organizational (Org)}: The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
+  \item \textbf{Organizational (Org)}: The RBAC extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
 Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
 
-  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or data of birth.
+  \item \textbf{Privacy (Priv)}: The RBAC extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or data of birth.
 
-  \item \textbf{Task}: The extension model provides task entities which are associated with permissions and roles.
+  \item \textbf{Task}: The RBAC extension model provides task entities which are associated with permissions and roles.
 A task is a fundamental unit of a business activity.
@@
 
-  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+  \item \textbf{Spatio-Temporal}: The RBAC extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
 For example, specific locations permit roles to conduct actions from 8:00 a.m.
@@
 
-  \item \textbf{Spatial}: The extension model provides spatial constraints, location-based constraints in specifying access
+  \item \textbf{Spatial}: The RBAC extension model provides spatial constraints, location-based constraints in specifying access
 	control policies. For example, in organizations, locations are enforced whereas a
@@
 
-  \item \textbf{Temporal (Temp)}:  The extension model provides temporal constraints, time-based constraints in specifying access
+  \item \textbf{Temporal (Temp)}:  The RBAC extension model provides temporal constraints, time-based constraints in specifying access
 	control policies. For example, in organizations, periodic temporal durations are enforced whereas a
@@
 
+\begin{figure}[ht]
+    \centering
+        \includegraphics[width=6.0in]{sections/category_structrure.eps}
+\vspace{-0.7 in}
+    \caption{\label{fig:category_structrure}Structure of categories within the RBAC extension models.}
+\end{figure}
+
 \subsection{Analysis and Discussion}
 
-The 28 primary sources produced a set of eight hierarchical categories. Table \ref{tab:categorization} summarizes each primary source under their designated
-category and furthermore, displays the perceived hierarchy of the categories. 
-The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and
-the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.
+The 28 primary sources produced a set of eight hierarchical categories.
+Table \ref{tab:categorization} summarizes each primary source under their designated category and furthermore, displays the perceived hierarchy of the categories.
+Figure~\ref{fig:category_structrure} shows the hierarchy structure among categories.
+The Constraint, Organizational, Privacy, Task, and Spatial and Temporal categories can be special cases of the Context category.
+The Spatio-Temporal category subsumes Spatial and Temporal categories.
+The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.
 
-When looking across all categories, we noted that each category added some new features on top of the standard RBAC model that were domain specific.
+When looking across all categories, we noted that each category added domain specific features on top of the RBAC Reference Model.
 These domain specific features were under the surface these features adding contextual relationships between the core user, permission and role entities.  Thus, we
@@
 
-\greybox{We found eight categories that exist within extensions to RBAC: Constraint, Context, 
+\greybox{We found eight categories that exist within the RBAC extension models: Constraint, Context, 
 Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other

### 2013-05-10 paper/sections/conclusion.tex
Updates to address dr. xie's comments
address dr. xie's comments

--- 2013-02-09 paper/sections/conclusion.tex
@@
 
-P1: Mention core RBAC, note specific results for each research question
-
-P2: Mention bigger results that cut across and any other peritentn information
+The RBAC extension models were revealed to fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context.
+The categories each had properties specific to their implementation, but were seen to generalize to being specialized instances of context tailored to the entities or actions the categories covered.
+A number of domains were identified as being the motivations behind needing extensions to the RBAC Reference model.
+The domains, such as healthcare, presenting new challenges the previous models were not
+required to design for.
+Our literature review showed that the state of RBAC extension model evaluation needs focused from the research community given most model evaluations seen within the papers were
+based on hypothetical situations with little to no case studies or implementations in practice.

### 2013-05-10 paper/sections/core_rbac.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 NIST's work can directly benefit organizations by lowering cost of early phase Research and Development, and implementation of RBAC.
-Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the the rate of RBAC adoption has rapidly grown over recent years \cite{o20102010}. 
+Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown over recent years \cite{o20102010}. 
 The analysts estimate that by the end of 2010 at least a portion of permission of systems will use RBAC for more than half of users at organizations with more than 500 employees.
-The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard work saved \$1.1 billion.
+The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard saved \$1.1 billion.
 
@@
 \cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
-Each component includes a RBAC reference model.
-RBAC$_{m}$ includes RBAC reference models of these three components.
+Each component includes its corresponding RBAC Reference model.
 We first describe the core RBAC, associated entities and other terminology encountered across the space of our review.
@@
 
-The four entities of the core RBAC reference model are:
+The four entities of the core RBAC Reference Model are:
 
@@
 user assignment ($UA$) and permission assignment ($PA$) functions.
-$UA$ assignment of users to roles.
-$PA$ assignment of permissions to roles.
+$UA$ represents assignment of users to roles.
+$PA$ represents assignment of permissions to roles.
 
@@
 
-The hierarchical RBAC reference model adds role hierarchies ($RH$) feature to the core RBAC reference model.
+The hierarchical RBAC Reference Model adds role hierarchies ($RH$) feature to the core RBAC Reference Model.
 $RH$ incorporates a structure of roles in an organization using inheritance relationships among attributes such as roles.
@@
 General role hierarchies can be extended to use the concept of multiple inheritances where
-$r_1$ inherits all permissions from more than one roles.
+$r_1$ inherits all permissions from more than one role.
 
@@
 
-The constrained RBAC reference model incorporates separation of duty relations to  the core or the hierarchical RBAC reference. Separation of 
+The constrained RBAC Reference Model incorporates separation of duty relations to  the core or the hierarchical RBAC Reference Model. Separation of 
 duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations: static and dynamic.

### 2013-05-10 paper/sections/discussion.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 
-The research questions we identified, presented results for and analyzed provide specific views of portions of the extension model landscape. 
+The research questions we identified, presented results for and analyzed provide specific views of portions of the RBAC extension model landscape. 
 By stepping back, and looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
-We break the discussion into two issues: recommendations for the enhancement of the evaluation of extensions to role-based access control models and guidelines for developers needing to choose an extension model to meet their requirements.
+We break the discussion into two issues: recommendations for the enhancement of the evaluation of RBAC extension models to role-based access control models and guidelines for developers needing to choose an RBAC extension model to meet their requirements.
 
-\subsection{Metrics for the Evaluation of RBAC Extensions}
+\subsection{Metrics for the Evaluation of RBAC Extension Models}
 
@@
 This in turn leaves researchers and practitioners little data from which to compare one model to the next.
-And given the wide swath of popular and critical domains being targeted by RBAC extension models, we propose a set of metrics by which RBAC extension models should be evaluated.
-Hu and Scarfone provide metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
+And given the wide swath of popular and critical domains being targeted by RBAC extension models, Hu and Scarfone proposed metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
 They collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing.
-From these metrics, and guided by the data surrounding how extension models are currently evaluated, we select metrics that can be applicable for evaluation of extended RBAC models as follows:
-
-\begin{itemize}
-	\item \textbf{Ease of Privilege Assignments}: measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC.
+For example, ``Ease of Privilege Assignments'' metric measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC.
 If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity.
-While extended RBAC may have similar features to that of the RBAC$_{m}$, we measure worst cases after extended, and the gap between an extension model and the RBAC$_{m}$.
-
-	\item \textbf{Syntactic and semantic support for specifying AC rules}: checks whether privileges and constraints can be specified using complex logic expressions such as AND, OR, $\leq$ when specifying constraints.
-This support helps specify complex and flexible access controls.
-For example, given a spatial RBAC model, one may measure whether spatial constraints can be specified using various logic expressions.
-
-	\item \textbf{Delegation of administrative capabilities}: measures how easy and secure an access control model supports for privilege delegation of administrators to another user.
-
-	\item \textbf{Capable of combining different policy models}: measures whether an access model has a feature to combine more than two policies.
-
-	\item \textbf{Capable of conflict resolution}: checks whether an access model has a feature to resolve conflicts when two rules in a policy may be applicable for the same request, but evaluate different decisions: One grants permission and the other one denies permission.
-
-	\item \textbf{Least privilege principle support}: checks whether an access model supports for least privilege principle where a subject can access least set of information and resources that are necessary for its legitimate functioning.
-Least privilege principle support can reduce damage when malicious users gain legitimate privileges.
-
-	\item \textbf{Separation of Duty (SoD)}: checks whether an access model supports for static SoD and/or dynamic SoD.
-Note that NIST Standard supports both static SoD and dynamic SoD described in Section~\ref{sec:core-rbac}.
-\end{itemize}
-
-We recommend these metrics be incorporated in any extension models presented hence forth. A future work could include providing metric measurements for the primary sources presented here. 
-A developer could attempt to apply these metrics to an extension model not already evaluated when attempting to choose a model. 
+This metric can be applicable for not only the RBAC Reference Model but also RBAC extension models.
+%Future work could include providing metric measurements for the primary sources presented in this paper. 
+We plan to use these metrics proposed by Hu and Scarfone for any extension models presented in this paper. 
+A developer could attempt to apply these metrics to an RBAC extension model not already evaluated when attempting to choose a model. 
 These metrics focus on comparing specific features support of access control models.
 
-\subsection{Adoption of Extensions to RBAC}
+\subsection{Adoption of RBAC Extension Models}
 
@@
 
-While the concept of the RBAC$_{m}$ is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the RBAC$_{m}$ is can be interpreted in more than one interpretation due to its complexity.
+While the concept of the Reference Model is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the Reference Model is can be interpreted in more than one interpretation due to its complexity.
 In order to bridge gap between RBAC standard and its implementation, NIST published RBAC Implementation and Interoperability standard (RIIS/ ANSI INCITS-459) in 2011.

### 2013-05-10 paper/sections/intro.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
-proposed by Sandhu, et al \cite{sandhu1996role}.  
+proposed by Sandhu et al.~\cite{sandhu1996role}.  
 In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
@@
 Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on a unified definition of RBAC or feature set.
-The RBAC standard includes RBAC reference models (RBAC$_{m}$ in short).
-RBAC$_{m}$ serve as a basis for defining scope of RBAC features and functional specifications of RBAC features.
+The RBAC standard includes RBAC Reference Model.
+The RBAC Reference Model serves as a basis for defining scope of RBAC features and functional specifications of RBAC features.
 
-Innovation and the use of RBAC in new domains has led to scenarios that, by some accounts, the RBAC$_{m}$ cannot handle \cite{kuhn2010adding}.  
-For example, Ni et al. \cite{ni2010privacy} state that RBAC$_{m}$ are ``not designed to enforce privacy policies and barely 
+Innovation and the use of RBAC in new domains have led to scenarios that, by some accounts, the RBAC Reference Model cannot handle \cite{kuhn2010adding}.  
+For example, Ni et al. \cite{ni2010privacy} state that the RBAC Reference Model is ``not designed to enforce privacy policies and barely 
 meet privacy protection requirements'' with the introduction of privacy concerns in domains such as healthcare. 
-Since the introduction of the RBAC standard, extension models to the RBAC$_{m}$ have appeared in the literature, each adding one or more features on top of components in the RBAC$_{m}$.
+Since the introduction of the RBAC standard, RBAC extension models have appeared in the literature, each adding one or more features on top of components in the RBAC Reference Model.
 %A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
@@
 
-\textit{The goal of our work is to aid practitioners and researchers in choosing an extension to the RBAC$_{m}$, and in understanding
-how extensions to RBAC$_{m}$ are evaluated by providing an assessment of RBAC extensions. We accomplished this through 
-establishment of a set of RBAC extension categories, an examination of the state of the art in evaluations of RBAC extension, and a breakdown of the 
-motivations that have led to extensions to RBAC$_{m}$.} To accomplish this goal, we seek to answer following research questions:
+\textit{The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
+how the RBAC extension models are evaluated by providing an assessment of RBAC extensions. We accomplished this through 
+establishment of a set of RBAC extension categories, an examination of the state of the art in evaluations of the RBAC extension models, and a breakdown of the motivations that have led to the RBAC extension models.} To accomplish this goal, we seek to answer following research questions:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item RQ1: What categories exist within extensions to RBAC$_{m}$?
-\item RQ2: What are the motivations behind extensions to RBAC$_{m}$?
-\item RQ3: Do the extensions to RBAC$_{m}$ have corresponding implementations?
-\item RQ4: How are extensions to RBAC$_{m}$ evaluated theoretically and in practice?
-\item RQ5: What domains have extensions to RBAC$_{m}$ been created for?
+\item RQ1: What categories exist within the RBAC extension models?
+\item RQ2: What are the motivations behind the RBAC extension models?
+\item RQ3: Do the RBAC extension models have corresponding implementations?
+\item RQ4: How are the RBAC extension models evaluated theoretically and in practice?
+\item RQ5: What domains have the RBAC extension models been created for?
 \item RQ6: What commonalities or generalizations exist across all categories?
@@
 
-We performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC$_{m}$. 
-The review we performed yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC$_{m}$.  
+We performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC Reference Model. 
+The review we performed yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.  
 This review is intended to serve as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
-the current state of extensions to the RBAC$_{m}$. 
+the current state of the RBAC extension models. 
 For developers looking to find a model to fit their potentially unique access control needs, this review provides a basis for comparison and easy look up for what extension model to use.  
@@
 \setlength{\itemsep}{0.25pt}
-\item Summarizes current research on extensions to RBAC$_{m}$.
-\item Guides a direction for a new standard based on an extension to the RBAC$_{m}$.
+\item Summarizes current research on the RBAC extension models.
+\item Guides a direction for a new standard based on the RBAC extension models.
 \end{itemize}
@@
 Sections~\ref{sec:categorization}-\ref{sec:generalizations} present analysis and discussion of the research questions. 
-Section~\ref{sec:discussion} discusses issues about extensions to RBAC$_{m}$.
+Section~\ref{sec:discussion} discusses issues about the RBAC extension models.
 Section~\ref{sec:conclusion} concludes the paper.

### 2013-05-10 paper/sections/process.tex
Updates to address dr. xie's comments
address dr. xie's comments

@@
 
-We adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes \cite{kitchenham2007guidelines}.
+We adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes~\cite{kitchenham2007guidelines}.
 The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage.
-The stages were
+The stages were as follows:
 
@@
 Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination.
-The search algorithm was performed as follows
+The search algorithm was performed as follows:
 
@@
 The result set size stopping criteria was chosen due to a technical limitation of some search engines.
-The stopping criteria related to the last ten titles is meant to stop after relevant results are no longer being returned by the search engine.
+The stopping criteria related to the last ten titles are meant to stop after relevant results are no longer being returned by the search engine.
 After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
@@
 \setlength{\itemsep}{0.25pt}
-\item Does this model extend the RBAC$_{m}$? (Exclusion)
-\item Do the researchers give evidence that RBAC$_{m}$ needs extension? (Inclusion)
+\item Does this model extend the RBAC Reference Model (Exclusion)
+\item Do the researchers give evidence that the RBAC Reference Model needs extension? (Inclusion)
 \item Was the paper and subsequent model inspired by a real world example?  (Conditional Inclusion)
@@
 During the paper reading round and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
-Thus, the first step undertaken was to answer the question of what categories exist for extensions to RBAC$_{m}$ and what papers fell into what categories.
+Thus, the first step undertaken was to answer the question of what categories exist for the RBAC extension models and what papers fell into what categories.

### 2013-05-12 paper/sections/abstract.tex
Updates to address dr. williams' comments

@@
 \noindent
-{\bf Context}: Since the National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) standard using RBAC Reference Model in the late 1990's, computing in new domains has led to proposals for extensions to the RBAC Reference Model; for example adding context around permission granting and role activation.
+{\bf Context}: Since the US National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) as documented in the RBAC Reference Model in the late 1990's, domain-specific extensions have been proposed.
+For example, the mobile domain has surfaced a need for permission granting and role activation.
 \\
@@
 {\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
-how RBAC extension models are evaluated by providing an assessment of the state of the RBAC extension models. We accomplished this through
-establishment of a set of extension categories, an examination of the state of the art in evaluations of the RBAC extension models, and a breakdown of the motivations that have led to the RBAC extension models.
+how RBAC extension models are evaluated by providing an assessment of the state of the RBAC extension models.
 \\
@@
 \noindent
-{\bf Method}: We performed a systematic literature review that yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as the RBAC extension models. 
+{\bf Method}: We performed a systematic literature review that began with 1,716 paper of which 28 were deemed primary sources for inclusion as the RBAC extension models.
+We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
 \\
@@
 \noindent
-{\bf Results}: Our results show that the RBAC extension models can be classified under eight categories: Constraint, Context, 
-Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal. We identified only
-8 of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype. We found that the primary domains 
+{\bf Results}: We classified the RBAC extension models into eight extension categories: Constraint, Context, 
+Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal.
+only 8 of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype.
+The primary domains
 that inspired extensions were the medical domain with 9 of the 28 models, enterprise workflows with 5 of 28, and mobile computing with 5 of 28.
@@
 \noindent
-{\bf Conclusions}: When examined as an entire group, the eight categories can be aligned under the single category of context. 
+{\bf Conclusions}: All of the eight RBAC extension categories deal with context. 
 Our literature review shows that the state of evaluation of RBAC model evaluation is limited with 1
 paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example 
-scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern.
-The landscape of the RBAC extension models serves as a basis from which a new, and vetted, RBAC extension model should be derived.
+scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern. The magnitude of extensions to the RBAC standard suggests a revised standard may be beneficial.
 

### 2013-05-12 paper/sections/core_rbac.tex
Updates to address dr. williams' comments

@@
 NIST's work can directly benefit organizations by lowering cost of early phase Research and Development, and implementation of RBAC.
-Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown over recent years \cite{o20102010}. 
-The analysts estimate that by the end of 2010 at least a portion of permission of systems will use RBAC for more than half of users at organizations with more than 500 employees.
+Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
 The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard saved \$1.1 billion.
 
-The NIST RBAC standard proposed by Ferraiolo et al.
-\cite{ferraiolo} and later adopted as the official standard for RBAC by the INCITS includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
+The RBAC standard includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
 Each component includes its corresponding RBAC Reference model.
-We first describe the core RBAC, associated entities and other terminology encountered across the space of our review.
+We first describe the core RBAC, associated entities and other terminology.
 We next describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into the core RBAC.
@@
 \setlength{\itemsep}{0.25pt}
-\item a set of \emph{Users}: A user can be a person or an agent.
-\item a set of \emph{Roles}: A role is a collection of permissions to perform a specific job function in an organization.
-\item a set of \emph{Permissions}: A permission refers to an access mode that can be exercised on an object in the system and a session relates a user to possibly many roles.
-\item a set of \emph{Sessions}: In each session, a user can be assigned to some of the roles, only when the corresponding role is enabled for activation for that time.		
+\item \emph{Users}: A user is defined as a human being.
+Although the concept of a user can be extended to include machines, networks, or intelligent autonomous agents, the definition is limited to a person in the RBAC standard.
+\item \emph{Roles}: A role is a job function within the context of an organization with some associated semantics regarding the authority and responsibility conferred on the user assigned to the role.
+\item \emph{Permissions}: A permission is an approval to perform an operation on one or more RBAC protected objects in the system.
+\item \emph{Sessions}: A session is a mapping between a user and an activated subset of roles that are assigned to the user.
 \end{itemize}
 
-In RBAC, a user can exercise a permission only if the user is assigned to a role.
+In RBAC, a user can exercise a permission only if the user is assigned to their role.
 In addition to the four basic entities, two functions are defined:
@@
     \centering
-        \includegraphics[width=4.0in]{sections/core-model.png}
+        \includegraphics[width=4.0in]{sections/core-model.eps}
     \caption{\label{fig:overview}Diagram in Core RBAC reference model\cite{ferraiolokuhn}.}
@@
 individual role assignments for a user. 
-Formally, the role inheritance relation is shown as \textit{RH} $\subseteq$ \textit{ROLES} $\times$ \textit{ROLES} describing the many-to-many mapping role inheritance relation.
+Concept of the role inheritance describes the many-to-many mapping role inheritance relations among roles.
+Therefore, more than one role (e.g., two roles $r_1$ and $r_1'$) can inherit all permissions of $r_2$ 
 General role hierarchies can be extended to use the concept of multiple inheritances where
-$r_1$ inherits all permissions from more than one role.
+$r_1$ inherits all permissions from more than one role (e.g., two roles $r_2$ and $r_2'$) .
 
@@
 SSoD is known to be too rigid for practical use in cases where a user should have permissions when a user is assigned to both $Role_A$ and $Role_B$.
+	For example, accounting clerk role requests a check and accounting manager role approves the check. These two roles must be
+	mutually exclusive to avoid a situation where one approves the check that the one requested.
+
 	\item Dynamic Separation of Duty (DSD): Dynamic SoD is known to be
@@
 For the situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ dynamically.
+Consider that one can be assigned to the accounting clerk and the accounting manager role at the same time.
+In such a situation, DSD can enforce that the accounting manager role does not approve her/his requested checks but
+approve only checks that others requested. This feature requires handling dynamic environments.
 \end{itemize}

### 2013-05-12 paper/sections/intro.tex
Updates to address dr. williams' comments

@@
 
-Role-Based Access Control (RBAC) is widely used for maintaining and managing an organization's 
+Software systems use access control mechanisms to determine which subjects can access
+which resources.
+Role-Based Access Control (RBAC) is a widely used access control mechanism designed for maintaining and managing an organization's 
 access control based on assigning permissions to roles, and roles to users, instead of 
@@
 associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead
-of assigning permissions of the manager role to the user.
+of assigning a set of individual permissions of the manager role.
 
 The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) 
-first proposed RBAC standard in the 1990s \cite{o20102010}.
+first proposed the RBAC standard in the 1990s \cite{o20102010}.
 The RBAC standard was first proposed when NIST 
@@
 The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010\footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
-Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC features without general agreement on a unified definition of RBAC or feature set.
-The RBAC standard includes RBAC Reference Model.
+Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles).
+The RBAC standard includes the RBAC Reference Model.
 The RBAC Reference Model serves as a basis for defining scope of RBAC features and functional specifications of RBAC features.
 
-Innovation and the use of RBAC in new domains have led to scenarios that, by some accounts, the RBAC Reference Model cannot handle \cite{kuhn2010adding}.  
+Since the introduction of the RBAC standard, researchers proposed domain-specific extensions such as features of dynamic context or privacy, which
+can be critical in medical and mobile domains.
+RBAC extension models have appeared in the literature, each adding one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
 For example, Ni et al. \cite{ni2010privacy} state that the RBAC Reference Model is ``not designed to enforce privacy policies and barely 
-meet privacy protection requirements'' with the introduction of privacy concerns in domains such as healthcare. 
-Since the introduction of the RBAC standard, RBAC extension models have appeared in the literature, each adding one or more features on top of components in the RBAC Reference Model.
+meet privacy protection requirements'', with the introduction of privacy concerns in the medical domain.
+They proposed an RBAC extension model for privacy concerns incorporated in the RBAC Reference Model.
 %A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
@@
 \textit{The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
-how the RBAC extension models are evaluated by providing an assessment of RBAC extensions. We accomplished this through 
-establishment of a set of RBAC extension categories, an examination of the state of the art in evaluations of the RBAC extension models, and a breakdown of the motivations that have led to the RBAC extension models.} To accomplish this goal, we seek to answer following research questions:
+how the RBAC extension models are evaluated by providing an assessment of RBAC extensions.
+We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
+} To accomplish this goal, we seek to answer following research questions:
 
@@
 
-We performed a systematic literature review designed to explore the current body of research in the area of extensions to the RBAC Reference Model. 
-The review we performed yielded 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.  
-This review is intended to serve as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
-the current state of the RBAC extension models. 
-For developers looking to find a model to fit their potentially unique access control needs, this review provides a basis for comparison and easy look up for what extension model to use.  
-Furthermore, the review provides insight into the state of the art in evaluating RBAC extension models.
-
+We performed a systematic literature review to explore the current body of research in the area of extensions to the RBAC Reference Model. 
+The review began with 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.
+%This review is intended to serve as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
+%the current state of the RBAC extension models. 
+%For developers looking to find a model to fit their potentially unique access control needs, this review provides a basis for comparison and easy look up for what extension model to use.  
+%Furthermore, the review provides insight into the state of the art in evaluating RBAC extension models.
 Our research provides contributions to the community in following ways:
@@
 \setlength{\itemsep}{0.25pt}
-\item Summarizes current research on the RBAC extension models.
-\item Guides a direction for a new standard based on the RBAC extension models.
+\item Serves as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
+the current state of the RBAC extension models. 
+\item Provides a basis for comparison and easy look up for what extension model to use for developers looking to find a model to fit their potentially unique access control needs
+\item Provides insight in evaluating RBAC extension models.
+%\item Summarizes current research on the RBAC extension models.
+%\item Guides a direction for a new standard based on the RBAC extension models.
 \end{itemize}

### 2013-05-12 paper/sections/process.tex
Updates to address dr. williams' comments

@@
 %The master list of 1,716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
-Out of the master list of 1,716 papers, two reviewers, namely Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
+Out of the master list of 1,716 papers, two reviewers, namely Author 1 and Author 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
 Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.
@@
 
-\multicolumn{1}{ |c| }{Reviewer 1}  &  & 305 & 86 & 46 \\ \hline
-\multicolumn{1}{ |c| }{Reviewer 2} &  & 176 & 102 & 42 \\ \hline
+\multicolumn{1}{ |c| }{Author 1}  &  & 305 & 86 & 46 \\ \hline
+\multicolumn{1}{ |c| }{Author 2} &  & 176 & 102 & 42 \\ \hline
 

### 2013-05-13 paper/sections/process.tex
Minor updates to the process and numbers surrounding elimination.

--- 2013-05-10 paper/sections/process.tex
@@
 After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
-%The master list of 1,716 papers was used to perform a series of elimination rounds to narrow the list of papers and identify primary sources.
-Out of the master list of 1,716 papers, two reviewers, namely Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
+Out of the master list of 1,716 papers, two reviewers, denoted by Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
 Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.
@@
 
+The title elimination round resulted in Reviewer 1 selecting 305 papers, and Reviewer 2 selecting 176 papers with 149 papers of overlap between the two. 
+Between the two reviewers selections, there were 332 papers not in common, of which, 149 were rejected and 141 retained after a second review.
+
 \subsubsection{Step 3: Abstract Elimination}
@@
 \end{itemize}
+
+The abstract elimination round resulted in Reviewer 1 selecting 86 papers, and Reviewer 2 selecting 102 papers with 51 papers of overlap between the two. 
+Between the two reviewers selections, there were 137 papers not in common, of which, 116 were rejected and 21 retained after a second review.
 
@@
 
+The content elimination round resulted in Reviewer 1 selecting 46 papers, and Reviewer 2 selecting 42 papers with 24 papers of overlap between the two. 
+Between the two reviewers selections, there were 64 papers not in common, of which, 59 were rejected and 5 retained after a second review.
+
 \begin{table}
@@
 
-\multicolumn{1}{ c| }{} & Overlap & 249 & 51 & 24 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & Disjoint & 232 & 137 & 64 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & Rejected & 208 & 116 & 59 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & Retained & 41 & 21 & 5 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & \textbf{Total} & 290 & 72 & 28 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Overlap & 149 & 51 & 24 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Disjoint & 332 & 137 & 64 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Rejected & 191 & 116 & 60 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Retained & 141 & 21 & 5 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & \textbf{Num Left} & 290 & 72 & 28 \\ \cline{2-5}
 

### 2013-05-15 paper/sections/abstract.tex
Removing duplicate paper and updating total counts.

@@
 \noindent
-{\bf Method}: We performed a systematic literature review that began with 1,716 paper of which 28 were deemed primary sources for inclusion as the RBAC extension models.
+{\bf Method}: We performed a systematic literature review that began with 1,716 paper of which 27 were deemed primary sources for inclusion as the RBAC extension models.
 We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
@@
 Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal.
-only 8 of the 28 papers provided an implementation of their model in the form of an enterprise application or prototype.
+only 8 of the 27 papers provided an implementation of their model in the form of an enterprise application or prototype.
 The primary domains
-that inspired extensions were the medical domain with 9 of the 28 models, enterprise workflows with 5 of 28, and mobile computing with 5 of 28.
+that inspired extensions were the medical domain with 9 of the 27 models, enterprise workflows with 5 of 27, and mobile computing with 5 of 27.
 \\
@@
 paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example 
-scenarios of their model in action, and 12 of the 28 models lacking an evaluation that we could discern. The magnitude of extensions to the RBAC standard suggests a revised standard may be beneficial.
+scenarios of their model in action, and 12 of the 27 models lacking an evaluation that we could discern. The magnitude of extensions to the RBAC standard suggests a revised standard may be beneficial.
 

### 2013-05-15 paper/sections/analysis.tex
Updates to address dr. williams' comments
Removing duplicate paper and updating total counts.

@@
 
-We first describe motivations described in the 28 papers surveyed.
+We first describe motivations described in the 27 papers surveyed.
 
@@
 
-We classify motivations behind RBAC extension models to these three categories:
+We classify motivations behind the RBAC extension models to these three categories:
 
@@
 %Table \ref{tab:motivations} shows the breakdown of motivations found within the primary sources.
-%The result shows that 23 papers out of 28 papers are classified into ``Addition'' category. Only
+%The result shows that 23 papers out of 27 papers are classified into ``Addition'' category. Only
 %one paper and four papers are classified into ``Modification'' category and ``Combination'' category, respectively.
@@
 Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
-Of the 28 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
+Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
 system. The remaining 20 papers provide no
@@
 
-The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
 Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
@@
 
-The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 28 mentioning scenarios or requirements of that industry.
+The predominant domain for which extension models have been generated for is that of the medical domain with 9 of 27 mentioning scenarios or requirements of that industry.
 Mobile computing and enterprise workflows were each represented by five papers claiming to be influenced by the requirements for access control within these domains. The final set
@@
 As access control is typically evaluated to either true or false based on predicates, propositional logic is sufficient to describe key ideas and definitions.
-We found that 28 papers of RBAC extensions use propositional logic to describe its extended model.
+We found that 27 papers of RBAC extensions use propositional logic to describe its extended model.
 However, propositional logic has limitations. While this logic is simple, this logic does not support for reasoning about RBAC, which helps

### 2014-02-04 paper/sections/analysis.tex
Adding latest PDF and PNGs of images to be able to build with Latek.
Section 5.1 and 5.2 updates

@@
 
-We first describe motivations described in the 27 papers surveyed.
+During the paper reading phase, we identify motivations of the RBAC extension models by category.
+We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues.
+For example, the papers in the Spatial Category argue that the RBAC reference model does not support spatial constraints, which are used in determining users' access decisions based on spatial dimensions (i.e., location).
+To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model.
+We describe motivations behind the RBAC extension models by category as follows:
 
 \begin{itemize}
-\setlength{\itemsep}{0.25pt}
-\item The RBAC Reference Model needs additional contextual information and constraints to develop fine-grained policies.
-\item The RBAC Reference Model does not incorporate context. Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
-\item The RBAC Reference Model does not support various constraints such as temporal and spatial constraints to design sophisticated policies on demand.
-\item The RBAC Reference Model does not provide an abstraction for additional user-defined attributes	(e.g., task and team) and their association with existing attributes.
-\item The RBAC Reference Model has limitation on delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.  
-%\item The RBAC standard needs to incorporate additional contextual information and constraints to develop fine-grained policies in practice.
-\item The RBAC Reference Model needs to incorporate additional attributes such as
-task, team, purpose, organizational roles and collaborative activities over existing attributes. Moreover, new associations between attributes
-are introduced. 
-\item The RBAC Reference Model needs to improve existing features such as delegation and role hierarchy to provide fine-grained access control. 
+
+  \item \textbf{Context}: The RBAC Reference Model does not support the notion of context constraints  related to changes in environments.
+Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
+
+  \item \textbf{Const}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.
+
+  \item \textbf{Org}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations.
+The model needs to reduce the administrative complexity of RBAC across multiple organizations.
+
+  \item \textbf{Priv}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.
+
+  \item \textbf{Task}: The RBAC extension model does not support the notion of a task, team, purpose, and organizational roles, which help specify a business activity in enterprises.
+
+  \item \textbf{Spatio-Temporal}: The RBAC extension model does not support spatial (location-based) and temporal (time-based) constraints, which specify role-assignment, role-activation and permissions based on location and time.
+
+  \item \textbf{Spatial}: The RBAC extension model does not support the notion of spatial (location-based)  constraints, which specify role-assignment and permissions based on location.
+
+  \item \textbf{Temp}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.
+
 \end{itemize}
-
-We classify motivations behind the RBAC extension models to these three categories:
-
-\begin{itemize}
-	\item Addition: The RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
-Models in this category do not change the RBAC Reference Model.
-	\item Modification: The RBAC extension model incorporates modification of existing features in the RBAC Reference Model.
-For example, changes of attribute relations such as role hierarchy fall into this modification category.
-For models in this category, changes occur within the RBAC Reference Model.
-	\item Combination: The RBAC extension model is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations such as task, team, purpose, and organizational roles.
-For this combination, new associations between attributes of two models should be introduced.
-Models in this category require new associations between attributes of different access control models to work together.
-\end{itemize}
-
-1 paper~\cite{alam06:constraint} is classified into ``Modification'' category since the paper describes modification of existing relationship among entities to support partial inheritance.
-2 papers~\cite{zhou2007team, oh2003task} are classified into ``Combination'' category since the papers describe an RBAC extension model, which is combination of the RBAC Reference Model and other access control models such as team-based access control and task-based access control.
-The remaining 25 papers are classified into ``Addition'' category.
-
-%\begin{table}
-%\centering
-%\caption{Motivation categories found and the count of primary sources}
-%\begin{tabular}{ | c | c | }
-%\cline{1-2}
-%
-%\textbf{Motivation Category} & \textbf{Paper Count} \\ \cline{1-2}
-%Addition & 23 \\ \cline{1-2}
-%Modification & 1 \\ \cline{1-2}
-%Combination & 4 \\
-%
-%\cline{1-2}
-%\end{tabular}
-%\label{tab:motivations}
-%\end{table}
-%
-%Table \ref{tab:motivations} shows the breakdown of motivations found within the primary sources.
-%The result shows that 23 papers out of 27 papers are classified into ``Addition'' category. Only
-%one paper and four papers are classified into ``Modification'' category and ``Combination'' category, respectively.
 
@@
 The RBAC standard provides the RBAC Reference Model, which is an abstraction on top of authorization based on $UA$ and $PA$.
-Since the RBAC Reference Model may not provide a fine-grained access control for a sophisticated
-security mechanism, a variety of RBAC extension models have been proposed over the years to meet
-security requirements.
-When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.
 RBAC Reference Model presents four entities: roles, sessions, users, and permissions, and their relations.
-While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitation on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
-In such environments, the RBAC standard needs additional constraints and dynamic context to support dynamic environments such as spatial and location.
+While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitations on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
 
-%Such extended RBAC models are formally presented to describe its behaviors.
-In the selected papers, researchers illustrate how an extended model incorporates new element based on the correct RBAC and describe their relations with pre-defined elements in the core RBAC.
+The Context, Org, Priv, Spatio-Temporal, Spatial, and Temp categories claim that the RBAC extension model does not support the notion of specific constraints.
+The RBAC extension model in these categories incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
+The extension model does not change the RBAC Reference Model.
+
+The Const category claims that the RBAC extension model has limitations on its features.
+The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model.
+For models in this category, changes occur within the RBAC Reference Model.
+
+The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles.
+Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations.
+For this combination, new associations between attributes of two models should be introduced.
+Models in this category require new associations between attributes of different access control models to work together.
 
@@
     \centering
-        \includegraphics[width=4.0in]{sections/dist_domains_byYear.eps}
+        \includegraphics[width=4.0in]{sections/dist_domains_byYear.png}
 \vspace{-0.2 in}

### 2014-02-05 paper/sections/abstract.tex
Adding edits to various sections.

@@
 \noindent
-{\bf Context}: Since the US National Institute of Standards and Technology (NIST) proposed Role-Based Access Control (RBAC) as documented in the RBAC Reference Model in the late 1990's, domain-specific extensions have been proposed.
+{\bf Context}: Since the United States National Institute of Standards and Technology proposed Role-Based Access Control (RBAC) as documented in the RBAC Reference Model in the late 1990's, domain-targeted extensions have been proposed.
 For example, the mobile domain has surfaced a need for permission granting and role activation.
@@
 {\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
-how RBAC extension models are evaluated by providing an assessment of the state of the RBAC extension models.
+how RBAC extension models are evaluated by providing an assessment of the state of RBAC extension models.
 \\
@@
 \noindent
-{\bf Method}: We performed a systematic literature review that began with 1,716 paper of which 27 were deemed primary sources for inclusion as the RBAC extension models.
-We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
+{\bf Method}: We performed a systematic literature review of RBAC extension models that began with 1,716 papers of which 27 were deemed as primary sources for inclusion.
 \\
@@
 \noindent
-{\bf Results}: We classified the RBAC extension models into eight extension categories: Constraint, Context, 
+{\bf Results}: We identified and classified the RBAC extension models into eight extension categories: Constraint, Context, 
 Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal.
-only 8 of the 27 papers provided an implementation of their model in the form of an enterprise application or prototype.
+Only 8 of the 27 papers provided an implementation of their model in the form of an enterprise application or prototype.
 The primary domains
@@
 \noindent
-{\bf Conclusions}: All of the eight RBAC extension categories deal with context. 
-Our literature review shows that the state of evaluation of RBAC model evaluation is limited with 1
-paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example 
-scenarios of their model in action, and 12 of the 27 models lacking an evaluation that we could discern. The magnitude of extensions to the RBAC standard suggests a revised standard may be beneficial.
+{\bf Conclusions}: Our literature review shows that all eight of the RBAC extension categories we identified deal with context.
+The state of evaluation of RBAC model extensions is limited with 1 paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example
+scenarios of their model in action, and 12 of the 27 models lacking an evaluation that we could discern.
+The magnitude and scope of extensions to the RBAC standard suggests a revised standard may be beneficial.
 

### 2014-02-05 paper/sections/analysis.tex
Adding edits to various sections.

@@
 
+\greybox{Motivations for RBAC extension models vary across category but are rooted in adding new entities or relationships to allow authorization flexibility in meeting the demands of emerging requirements.}
+
 \section{Implementations} \label{sec:implementations}
@@
 
-%Of the remaining eight papers that did mention an implementation, half 
-%were simply prototypes developed by the authors whereas the other half were claimed to be implemented within a real
-%system.
-
 \subsection{Analysis and Discussion}
@@
 implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
+
+\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for.}
 
@@
 entities.
+
+\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
 
@@
 
+\greybox{RBAC extension models were found targeting domains such as medical, pervasive computing, web applications, enterprise workflows and mobile computing.}
+
 \section{Generalizations} \label{sec:generalizations}

### 2014-02-05 paper/sections/core_rbac.tex
Adding latest PDF and PNGs of images to be able to build with Latek.
Adding edits to various sections.

@@
 
-NIST's work can directly benefit organizations by lowering cost of early phase Research and Development, and implementation of RBAC.
+NIST's work can directly benefit organizations by lowering the cost of early phase research and development, and implementation of RBAC.
 Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
@@
 The RBAC standard includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
-Each component includes its corresponding RBAC Reference model.
+Each component includes a corresponding RBAC Reference model.
 We first describe the core RBAC, associated entities and other terminology.
-We next describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into the core RBAC.
+We next describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into core RBAC.
 
@@
 
-In RBAC, a user can exercise a permission only if the user is assigned to their role.
+In RBAC, a user can exercise a permission only if the user is assigned to a role that contains the permission.
 In addition to the four basic entities, two functions are defined:
@@
     \centering
-        \includegraphics[width=4.0in]{sections/core-model.eps}
+        \includegraphics[width=4.0in]{sections/core-model.png}
     \caption{\label{fig:overview}Diagram in Core RBAC reference model\cite{ferraiolokuhn}.}
@@
 
-Figure~\ref{fig:overview} presents an overview of core RBAC reference model diagram where entities and their relations are described.
+Figure~\ref{fig:overview} presents an overview of the core RBAC reference model where entities and their relations are described.
 Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
@@
 
-The hierarchical RBAC Reference Model adds role hierarchies ($RH$) feature to the core RBAC Reference Model.
+The hierarchical RBAC Reference Model adds role hierarchies ($RH$) as a feature to the core RBAC Reference Model.
 $RH$ incorporates a structure of roles in an organization using inheritance relationships among attributes such as roles.
@@
 
-The constrained RBAC Reference Model incorporates separation of duty relations to  the core or the hierarchical RBAC Reference Model. Separation of 
-duty relations enforces conflict of interest among roles. This model defines two types of separation of duty relations: static and dynamic.
+The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. 
+Separation of duty relations enforce conflicts of interest among roles. This model defines two types of separation of duty relations: static and dynamic.
 
@@
 \item Static Separation of Duty (SSoD): SSoD restricts the conflicting-roles, which can be assigned to a single user statically. 
-Consider that roles $Role_A$ and $Role_B$ are conflicting with each other.
-On situations
-	where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ statically.
+Suppose roles $Role_A$ and $Role_B$ conflict with each other.
+On situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ statically.
 SSoD is known to be too rigid for practical use in cases where a user should have permissions when a user is assigned to both $Role_A$ and $Role_B$. 
-	For example, accounting clerk role requests a check and accounting manager role approves the check. These two roles must be
-	mutually exclusive to avoid a situation where one approves the check that the one requested.
+For example, an accounting clerk role requests a check and an accounting manager role approves the check. 
+These two roles must be mutually exclusive to avoid a situation where one approves the check that the other requested.
 
-	\item Dynamic Separation of Duty (DSD): Dynamic SoD is known to be
-more flexible than SSoD.
+\item Dynamic Separation of Duty (DSD): DSD is known to be more flexible than SSoD. 
 DSD restricts the conflicting-role assignments dynamically that are associated with a user. 
-Consider that roles $Role_A$ and $Role_B$ are conflicting with each other.
-For the situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ dynamically.
-Consider that one can be assigned to the accounting clerk and the accounting manager role at the same time.
-In such a situation, DSD can enforce that the accounting manager role does not approve her/his requested checks but
-approve only checks that others requested. This feature requires handling dynamic environments.
+Suppose roles $Role_A$ and $Role_B$ conflict with each other. 
+For situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ dynamically.	
+Consider that a user can be assigned to the accounting clerk and accounting manager role at the same time.
+In such a situation, DSD can enforce that the accounting manager role does not approve her/his requested checks but can only approve checks that others have requested.
 \end{itemize}

### 2014-02-05 paper/sections/intro.tex
Removing duplicate paper and updating total counts.
Adding edits to various sections.

@@
 and the number of permissions and users scales upward.
-As permissions can be managed by a role instead of a user, RBAC has an advantage to significantly reduce complexity of security
+As permissions can be managed by a role instead of a user, RBAC has been shown to significantly reduce complexity of security
 administration.
@@
 associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead
-of assigning a set of individual permissions of the manager role.
+of assigning a set of individual permissions to the user that represent the manager role.
 
@@
 first proposed the RBAC standard in the 1990s \cite{o20102010}.
-The RBAC standard was first proposed when NIST 
-requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
+NIST requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
 proposed by Sandhu et al.~\cite{sandhu1996role}.  
 In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
-The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010\footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}} that showed the cost savings of RBAC implementation and maintenance.
-Prior to the development of the RBAC standard, vendors proposed and implemented their own defined RBAC without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles).
-The RBAC standard includes the RBAC Reference Model.
-The RBAC Reference Model serves as a basis for defining scope of RBAC features and functional specifications of RBAC features.
+The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010\footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}.
+The study showed the cost savings of RBAC implementation and maintenance.
+Prior to the development of the RBAC standard, vendors proposed and implemented their own RBAC definition without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles).
+The RBAC standard includes the RBAC Reference Model which serves as a basis for defining the scope of RBAC features and functional specifications of RBAC features.
 
-Since the introduction of the RBAC standard, researchers proposed domain-specific extensions such as features of dynamic context or privacy, which
-can be critical in medical and mobile domains.
-RBAC extension models have appeared in the literature, each adding one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
-For example, Ni et al. \cite{ni2010privacy} state that the RBAC Reference Model is ``not designed to enforce privacy policies and barely 
-meet privacy protection requirements'', with the introduction of privacy concerns in the medical domain.
-They proposed an RBAC extension model for privacy concerns incorporated in the RBAC Reference Model.
-%A cursory examination of the extension model papers does not provide good comparisons for developers and researchers to build upon. 
+Since the introduction of the RBAC standard, researchers have proposed domain-targeted extensions that add one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
+For example, there are extensions that provide dynamic context or privacy, which can be critical in medical and mobile domains.
+Further, Ni et al.
+\cite{ni2010privacy} proposed an RBAC extension model to incorporate privacy concerns in to the RBAC Reference Model noting that the RBAC Reference Model is ``not designed to enforce privacy policies and barely meet privacy protection requirements'' with the introduction of privacy concerns in to the medical domain.
 Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
@@
 We performed a systematic literature review to explore the current body of research in the area of extensions to the RBAC Reference Model. 
-The review began with 1,716 papers, of which 28 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.
-%This review is intended to serve as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
-%the current state of the RBAC extension models. 
-%For developers looking to find a model to fit their potentially unique access control needs, this review provides a basis for comparison and easy look up for what extension model to use.  
-%Furthermore, the review provides insight into the state of the art in evaluating RBAC extension models.
+The review began with 1,716 papers, of which 27 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.
 Our research provides contributions to the community in following ways:
@@
 the current state of the RBAC extension models. 
-\item Provides a basis for comparison and easy look up for what extension model to use for developers looking to find a model to fit their potentially unique access control needs
-\item Provides insight in evaluating RBAC extension models.
-%\item Summarizes current research on the RBAC extension models.
+\item Provides a basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs
+\item Summarizes current evaluation methods used in research on RBAC extension models.
 %\item Guides a direction for a new standard based on the RBAC extension models.

### 2014-02-05 paper/sections/process.tex
Removing duplicate paper and updating total counts.
Adding edits to various sections.

@@
 
-We adopted and applied a systematic literature review following recommendations from Kitchenham's suggested processes~\cite{kitchenham2007guidelines}.
+We adopted and applied a systematic literature review process following recommendations from Kitchenham's suggested processes~\cite{kitchenham2007guidelines}.
 The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage.
@@
 \item Step 3: Elimination of papers based on abstract criteria
-\item Step 4: Elimination of papers based on content and matching to elimination criteria
+\item Step 4: Elimination of papers based on content and elimination criteria
 \end{itemize}
@@
 
-The final elimination round involved reading the entire paper into consideration and answering five questions that would serve as the basis for elimination.
+The final elimination round involved reading the entire paper and answering five questions that would serve as the basis for elimination.
 The data collected by answering these questions served as the basis towards answering the research questions.
@@
 \multicolumn{1}{ c| }{} & Disjoint & 332 & 137 & 64 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & Rejected & 191 & 116 & 60 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & Rejected & 191 & 116 & 61 \\ \cline{2-5}
 \multicolumn{1}{ c| }{} & Retained & 141 & 21 & 5 \\ \cline{2-5}
-\multicolumn{1}{ c| }{} & \textbf{Num Left} & 290 & 72 & 28 \\ \cline{2-5}
+\multicolumn{1}{ c| }{} & \textbf{Num Left} & 290 & 72 & 27 \\ \cline{2-5}
 
@@
 The first step was to take the individual data generated from the final elimination round and organize this information around the research questions. 
-During the paper reading round and resulting data, the fact that the papers were logically falling into a number of categorizations became evident.
+During the paper reading round and resulting data, the fact that the papers were falling into a number of categorizations became evident. 
 Thus, the first step undertaken was to answer the question of what categories exist for the RBAC extension models and what papers fell into what categories.

### 2014-02-08 paper/sections/categorization.tex
Removing duplicate paper and updating total counts.
Adding latest PDF and PNGs of images to be able to build with Latek.
Adding edits to various sections.
updates - 1. Reorganization results section 2. update Figure 2

@@
 RQ1: What categories exist within the RBAC extension models?
-
-\subsection{Results}
 
@@
 
-\subsection{Definitions}
+\subsection{Results}
 
@@
 
-  \item \textbf{Privacy (Priv)}: The RBAC extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or data of birth.
+  \item \textbf{Privacy (Priv)}: The RBAC extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or date of birth.
 
@@
     \centering
-        \includegraphics[width=6.0in]{sections/category_structrure.eps}
-\vspace{-0.7 in}
+        \includegraphics[width=6.0in]{sections/category_structrure.png}
+%\vspace{-0.2 in}
     \caption{\label{fig:category_structrure}Structure of categories within the RBAC extension models.}
@@
 
-The 28 primary sources produced a set of eight hierarchical categories.
+The 27 primary sources produced a set of eight hierarchical categories.
 Table \ref{tab:categorization} summarizes each primary source under their designated category and furthermore, displays the perceived hierarchy of the categories.

### 2014-02-08 paper/sections/intro.tex
Updates for introduction references and research question refining.

@@
 
-Software systems use access control mechanisms to determine which subjects can access
-which resources.
-Role-Based Access Control (RBAC) is a widely used access control mechanism designed for maintaining and managing an organization's 
-access control based on assigning permissions to roles, and roles to users, instead of 
-assigning the permissions directly to individual users.
-RBAC is used for securing 
-various applications including web services, database applications, and healthcare applications. 
+Software systems use access control mechanisms to determine which subjects can access which resources.
+Role-Based Access Control (RBAC) is a widely used access control mechanism designed for maintaining and managing an organization's  access control based on assigning permissions to roles, and roles to users, instead of assigning the permissions directly to individual users.
+RBAC is used for securing various applications including web services, database applications, and healthcare applications. 
 In RBAC, roles represent a set of permissions needed to perform a particular job function within an organization.  
-Multiple users who are involved in that specific job function within the organization can then be 
-assigned to a single role to inherit the required access. 
-The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows
-and the number of permissions and users scales upward.
-As permissions can be managed by a role instead of a user, RBAC has been shown to significantly reduce complexity of security
-administration.
-For example, if a user requires access to resources 
-associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead
-of assigning a set of individual permissions to the user that represent the manager role.
+Multiple users who are involved in that specific job function within the organization can then be assigned to a single role to inherit the required access. 
+The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows and the number of permissions and users scales upward.
+As permissions can be managed by role instead of by user, RBAC has been shown to significantly reduce complexity of security administration\cite{impact2002}.
+For example, if a user requires access to resources associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead of assigning a set of individual permissions.
 
-The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) 
-first proposed the RBAC standard in the 1990s \cite{o20102010}.
-NIST requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework 
-proposed by Sandhu et al.~\cite{sandhu1996role}.  
-In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute (ANSI) and the InterNational Committee for Information Technology Standards (INCITS). 
-The development of a standard was inspired by an economic impact study done during the 1990s and later confirmed in 2010\footnote{\url{http://csrc.nist.gov/groups/SNS/rbac/documents/20101219_RBAC2_Final_Report.pdf}}.
+The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in 2000 \cite{sandhu2000nist}.
+NIST requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu et al.~\cite{sandhu1996role}.  
+In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute\footnote{http://www.ansi.org/} (ANSI) and the InterNational Committee for Information Technology Standards\footnote{http://www.incits.org/} (INCITS).
+The development of a standard was inspired by an economic impact study done during the 1990s\cite{impact1996}, again in 2002\cite{impact2002} and later confirmed in 2010\cite{o20102010}. 
 The study showed the cost savings of RBAC implementation and maintenance. 
 Prior to the development of the RBAC standard, vendors proposed and implemented their own RBAC definition without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles).
-The RBAC standard includes the RBAC Reference Model which serves as a basis for defining the scope of RBAC features and functional specifications of RBAC features.
+The RBAC standard includes the RBAC Reference Model which serves as a basis for defining the scope and functional specifications of RBAC features.
 
 Since the introduction of the RBAC standard, researchers have proposed domain-targeted extensions that add one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
-For example, there are extensions that provide dynamic context or privacy, which can be critical in medical and mobile domains.
+For example, extensions that target the medical and mobile domains provide dynamic context or privacy around the access control policies.
 Further, Ni et al.
 \cite{ni2010privacy} proposed an RBAC extension model to incorporate privacy concerns in to the RBAC Reference Model noting that the RBAC Reference Model is ``not designed to enforce privacy policies and barely meet privacy protection requirements'' with the introduction of privacy concerns in to the medical domain.
-Further, these extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and increase interoperability \cite{o20102010}.
+These extension models are each building upon and adding features to a standard that was designed to reduce the economic impact experienced by enterprises and to increase interoperability \cite{o20102010}.
 
 \textit{The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
-how the RBAC extension models are evaluated by providing an assessment of RBAC extensions.
-We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
-} To accomplish this goal, we seek to answer following research questions:
+how the RBAC extension models are evaluated by providing an assessment of the state of RBAC extension model1s.} We established a set of extension categories, examined the state of the art in evaluations of the RBAC extension models, and categorized the motivations that have led to the RBAC extension models.
+To accomplish this goal, we seek to answer following research questions:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item RQ1: What categories exist within the RBAC extension models?
-\item RQ2: What are the motivations behind the RBAC extension models?
-\item RQ3: Do the RBAC extension models have corresponding implementations?
-\item RQ4: How are the RBAC extension models evaluated theoretically and in practice?
-\item RQ5: What domains have the RBAC extension models been created for?
-\item RQ6: What commonalities or generalizations exist across all categories?
+\item RQ1: How can RBAC extension models be classified?
+\item RQ2: What are motivations behind the RBAC extension models?
+\item RQ3: How are RBAC extension models implemented and evaluated by their authors?
+\item RQ4: What domains have been targeted by RBAC extension models?
+\item RQ5: What commonalities exist across RBAC extension models?
 \end{itemize}
@@
 The review began with 1,716 papers, of which 27 were deemed primary sources for inclusion as extension models to the RBAC Reference Model.
-Our research provides contributions to the community in following ways:
+Our research provides the following:
 
@@
 \setlength{\itemsep}{0.25pt}
-\item Serves as a starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
+\item A starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
 the current state of the RBAC extension models. 
-\item Provides a basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs
-\item Summarizes current evaluation methods used in research on RBAC extension models.
-%\item Guides a direction for a new standard based on the RBAC extension models.
+\item A basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs.
+\item A summary of current evaluation methods used in research on RBAC extension models.
+\item A starting direction for a new standard based on conclusions drawn from the RBAC extension models.
 \end{itemize}

### 2014-02-08 paper/sections/analysis.tex
updates - 1. Reorganization results section 2. update Figure 2
Updating data display tables.

@@
 
-\subsection{Results}
+%\subsection{Results}
 
@@
 To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model.
+
+\subsection{Results}
+
 We describe motivations behind the RBAC extension models by category as follows:
@@
 
-\subsection{Results}
-
 When designing and proposing a model targeted at a feature that is rooted in practical
@@
 
-\begin{table}
-\centering
-\caption{Implementation types found and the count of primary sources}
-\begin{tabular}{ | c | c | }
-\cline{1-2}
-
-\textbf{Implementation Type} & \textbf{Paper Count} \\ \cline{1-2}
-Enterprise Implementation & 4 \\ \cline{1-2}
-Prototype Implementation & 4 \\ \cline{1-2}
-No Implementation & 20 \\
-
-\cline{1-2}
-\end{tabular}
-\label{tab:implementations}
-\end{table}
+\subsection{Results}
 
@@
 
+\begin{table}
+\centering
+\caption{Implementation types found and the count of primary sources}
+\vspace{0.1 in}
+\begin{tabular*}{.9\linewidth}{| l | p{.5\linewidth} | r | }
+\cline{1-3}
+
+\textbf{Implementation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
+Enterprise
+&
+\cite{motta03:contextual}
+\cite{aich09:role}
+\cite{yao2008task}
+&
+3 \\ \cline{1-3}
+Prototype
+& 
+\cite{cholewka00:acontext-sensitive}
+\cite{huang06:pervasive}
+\cite{bao08:role}
+\cite{zhou07:network}
+&
+4 \\ \cline{1-3}
+None
+&
+\cite{alam06:constraint}
+\cite{tzelepi01:flexible}
+\cite{yamazaki04:designing}
+\cite{jian2008extended}
+\cite{thein2011leveraging}
+\cite{zou2009crbac}
+\cite{haibo05:context}
+\cite{zhang06:collaborative}
+\cite{ni2010privacy}
+\cite{masoumzadeh2008purbac}
+\cite{damiani2007geo}
+\cite{ray2006lrbac}
+\cite{hansen2003spatial}
+\cite{aich07:STARBAC}
+\cite{chen08:spatio-temporal}
+\cite{samuel07:spatio-temporal}
+\cite{chandran05:llt}
+\cite{ray07:spatio}
+\cite{oh2003task}
+\cite{joshi05:generalized}
+&
+20 \\
+
+\cline{1-3}
+\end{tabular*}
+\label{tab:implementations}
+\end{table}
+
 \section{Evaluations} \label{sec:evaluations}
@@
 RQ4: How are extensions to RBAC evaluated theoretically and in practice?
-
-\subsection{Results}
 
@@
 
-\begin{itemize}
-\setlength{\itemsep}{0.25pt}
-\item Time-based Performance \cite{ni2010privacy}, \cite{aich09:role}
-\item Complexity analysis \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{chen08:spatio-temporal}, \cite{aich09:role}
-\item Comparison to standard RBAC \cite{bao08:role}, \cite{zou2009crbac}, \cite{zhang06:collaborative}, \cite{zhao2008flexible}, \cite{ray07:spatio}
-\item Mathematical modeling \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{aich07:STARBAC}, \cite{chen08:spatio-temporal}, \cite{joshi05:generalized}
-\item Example scenarios of the model in action \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{cholewka00:acontext-sensitive}, \cite{huang06:pervasive}, \cite{bao08:role}, \cite{jian2008extended}, \cite{yamazaki04:designing}, \cite{zou2009crbac}, \cite{ray07:spatio}, \cite{samuel07:spatio-temporal}, \cite{joshi05:generalized}, \cite{yao2008task}, \cite{zhou2007team}, \cite{oh2003task}
-%\item Experimental analysis of the model
-\item Case study of the model in practice \cite{motta03:contextual}
-\end{itemize}
+\subsection{Results}
 
@@
 
+\begin{table}
+\centering
+\caption{Evaluation types by primary source}
+\vspace{0.1 in}
+\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
+\cline{1-3}
+
+\textbf{Evaluation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
+Time-Based Performance
+&
+\cite{ni2010privacy}
+\cite{aich09:role}
+&
+2 \\ \cline{1-3}
+Complexity Analysis
+&
+\cite{bao08:role}
+\cite{zhang06:collaborative}
+\cite{chen08:spatio-temporal}
+\cite{aich09:role}
+&
+4 \\ \cline{1-3}
+Comparison to Standard RBAC
+& 
+\cite{bao08:role}
+\cite{zou2009crbac}
+\cite{zhang06:collaborative}
+\cite{zhao2008flexible}
+\cite{ray07:spatio}
+&
+5 \\ \cline{1-3}
+Mathematical Modeling
+& 
+\cite{damiani2007geo}
+\cite{hansen2003spatial}
+\cite{aich07:STARBAC}
+\cite{chen08:spatio-temporal}
+\cite{joshi05:generalized}
+&
+5 \\ \cline{1-3}
+Example Scenarios of the Model in Action
+& 
+\cite{alam06:constraint}
+\cite{tzelepi01:flexible}
+\cite{cholewka00:acontext-sensitive}
+\cite{huang06:pervasive}
+\cite{bao08:role}
+\cite{jian2008extended}
+\cite{yamazaki04:designing}
+\cite{zou2009crbac}
+\cite{ray07:spatio}
+\cite{samuel07:spatio-temporal}
+\cite{joshi05:generalized}
+\cite{yao2008task}
+\cite{zhou2007team}
+\cite{oh2003task}
+&
+14 \\ \cline{1-3}
+Case Study of the Model in Practice
+&
+\cite{motta03:contextual}
+&
+1 \\ \cline{1-3}
+
+\end{tabular*}
+\label{tab:implementations}
+\end{table}
+
 \section{Domains} \label{sec:domains}
@@
 scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
+
+\subsection{Results}
+
 We found that the domains mentioned and their associated sources are:
 
-\begin{itemize}
-\setlength{\itemsep}{0.25pt}
-\item Medical domain \cite{alam06:constraint}, \cite{tzelepi01:flexible}, \cite{motta03:contextual}, \cite{ni2010privacy}, \cite{damiani2007geo}, \cite{hansen2003spatial}, \cite{samuel07:spatio-temporal}, \cite{aich09:role}, \cite{zhou2007team}
-\item Pervasive computing environments \cite{huang06:pervasive}, \cite{chen08:spatio-temporal}, \cite{ray07:spatio}
-\item Web applications \cite{masoumzadeh2008purbac}
-\item Mobile computing \cite{thein2011leveraging}, \cite{zou2009crbac}, \cite{chandran05:llt}, \cite{ray07:spatio}, \cite{aich09:role}
-\item Large-scale organizations with many sub-departments \cite{yamazaki04:designing}, \cite{han08:extended}, \cite{yao2008task}
-\item Enterprise, organization workflows \cite{cholewka00:acontext-sensitive}, \cite{bao08:role}, \cite{zhang06:collaborative}, \cite{oh2003task}, \cite{joshi05:generalized}
-\end{itemize}
+\begin{table}
+\centering
+\caption{Domains by primary source}
+\vspace{0.1 in}
+\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
+\cline{1-3}
+
+\textbf{Domain} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
+Medical Domain
+&
+\cite{alam06:constraint}
+\cite{tzelepi01:flexible}
+\cite{motta03:contextual}
+\cite{ni2010privacy}
+\cite{damiani2007geo}
+\cite{hansen2003spatial}
+\cite{samuel07:spatio-temporal}
+\cite{aich09:role}
+\cite{zhou2007team}
+&
+9 \\ \cline{1-3}
+Pervasive Computing Environments
+& 
+\cite{huang06:pervasive}
+\cite{chen08:spatio-temporal}
+\cite{ray07:spatio}
+&
+3 \\ \cline{1-3}
+Web Applications
+& 
+\cite{masoumzadeh2008purbac}
+&
+1 \\ \cline{1-3}
+Mobile Computing
+& 
+\cite{thein2011leveraging}
+\cite{zou2009crbac}
+\cite{chandran05:llt}
+\cite{ray07:spatio}
+\cite{aich09:role}
+&
+5 \\ \cline{1-3}
+Organizations with Many Sub-departments
+& 
+\cite{yamazaki04:designing}
+\cite{jian2008extended}
+\cite{yao2008task}
+&
+3 \\ \cline{1-3}
+Enterprise Workflows
+&
+\cite{cholewka00:acontext-sensitive}
+\cite{bao08:role}
+\cite{zhang06:collaborative}
+\cite{oh2003task}
+\cite{joshi05:generalized}
+&
+5 \\ \cline{1-3}
+
+\end{tabular*}
+\label{tab:implementations}
+\end{table}
 
@@
 
+During the paper reading phase, we identified commonalities or generalizations within the primary sources 
+by looking for formal modeling of extending RBAC extension models.
+
 \subsection{Results}

### 2014-02-08 paper/sections/appendix.tex
Committing updates based on comments from reading group.
Adding greybox. Adding results subsections for each RQ. Fixing last table captions to be on top. Edits to categories.
Removing duplicate paper and updating total counts.
Updating data display tables.

@@
 \centering
+\caption{Primary sources grouped by categorization}
 \vline
@@
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{han08:extended}} \\ \cline{2-4}
-
-\multicolumn{1}{ c }{}  &
 \multicolumn{3}{ |l| }{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging}} \\ \cline{2-4}
@@
 \multicolumn{2}{ |l| }{PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac}} \\[1ex] \cline{2-4}
-
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Res. }}}} &
-\multicolumn{2}{ |l| }{A Flexible Role- and Resource-Based Access Control Model, 2008 \cite{zhao2008flexible}} \\[2.5ex]  \cline{2-4}
 
@@
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
-\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray07:spatio}} \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray2006lrbac}} \\[0.25ex] \cline{4-4}
 
@@
 \vline
-\caption{Primary sources grouped by categorization}
 \label{tab:categorization}

### 2014-02-09 paper/sections/categorization.tex
updates: RQ formats and combine of RQ3-4 fixing typos and some sentences.

@@
-\section{Categorization} \label{sec:categorization}
+\section{RQ1: Classification} \label{sec:categorization}
 
-RQ1: What categories exist within the RBAC extension models?
+\textit{How can RBAC extension models be classified?}
 
-During the paper reading phase, we identify categorical themes, which are
-central topic or context of the RBAC extension models. 
+During the paper reading phase, we identify classification themes, which are
+central topic or context of the RBAC extension models. Our classification process
+yields a set of categories based upon observations during the paper reading phase.
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model.
+We classify this paper into the Privacy category.
 Some papers presented this direct pronouncement of theme of the RBAC extension models, whereas others were less obvious.
-Thus, we developed a set of guidelines to aide in determining a set of eight categories based upon observations during the paper reading phase. 
+Thus, we developed a set of guidelines to aide in determining a set of eight categories. 
 In developing these guidelines, we defined each category by a single noun-phrase descriptor.
@@
 	control policies. For example, in organizations, locations are enforced whereas a
-	specific role is permitted to conduct an action. Consider that a part-time employee works only in a specific location.
-	In such cases, the part-time employee role should access required resources only when the user is in that location. 
+	specific role is permitted to conduct an action. Consider that a temporal employee works only in a specific location.
+	In such cases, the temporal employee role should access required resources only when the user is in that location. 
 	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.
@@
 	control policies. For example, in organizations, periodic temporal durations are enforced whereas a
-	specific role is permitted to conduct an action. Consider that a part-time employee works only from 9:00 a.m. to 3:00 p.m.
-	In such cases, the part-time employee role should be allowed to access required resources during the interval. 
+	specific role is permitted to conduct an action. Consider that a temporal employee works only from 9:00 a.m. to 3:00 p.m.
+	In such cases, the temporal employee role should be allowed to access required resources during the interval. 
 	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.

### 2014-02-09 paper/sections/analysis.tex
Appendix table edits and cleaning up bibliography.

@@
-\section{Motivations} \label{sec:motivations}
+\section{RQ2: Motivations} \label{sec:motivations}
 
-RQ2: What are the motivations behind the RBAC extension models?
+\textit{What are motivations behind the RBAC extension models?}
 
@@
 
-\section{Implementations} \label{sec:implementations}
+\section{RQ3: Implementations and Evaluations} \label{sec:implementations}
 
-RQ3: Do the RBAC extension models have corresponding implementations?
+\textit{How are RBAC extension models implemented and evaluated by their authors?}
 
 When designing and proposing a model targeted at a feature that is rooted in practical
-usage by real software systems, bringing the model to life is strong evidence that the
-proposed model can work in practice. The concept of authorization and access control
-is rooted in a business need. Thus, any access control model needs to be feasible
-in the real world not just on paper. We analyzed the primary sources to see how many
-proposed models actually had implementations associated with them.  Then, we quantified
-types of implementations. Whether the implementation was for a real system, for a prototype
-and/or used in a production environment.
+usage by real software systems, bringing the model to life is evidence that the
+proposed model can work in practice. The presence of an actual implementation substantiates the robustness of the design and/or the need for the extension.
+%The concept of authorization and access control
+%is rooted in a business need. Thus, any access control model needs to be feasible
+%in the real world not just on paper. 
+We analyzed the primary sources to see how many
+proposed models actually had implementations associated with them.
+We quantified whether the implementation was enterprise (i.e., an implementation used in a production environment) was for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).
+
+Furthermore, the 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+We identified a set of evaluation types found within  each of the primary sources and provide a list of the evaluation types and which primary sources provided which type.
+In some cases a single primary source provided multiple evaluation types.
 
@@
 Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
-Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
+Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
 system. The remaining 20 papers provide no
-mention of an implementation or prototype.
+mention of an implementation.
+
+Table \ref{tab:evaluations} shows evaluation types.
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
+and how application of their model would apply and resolve the situation. Six models provided some form of performance
+or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
+as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
+Five models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
+would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
+or show how the Reference Model is deficient in tackling said scenarios.
 
@@
 
-The RBAC standard is designed with enterprises in mind such that when practitioners implemented RBAC into their systems,
-there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC Reference Mode
+The RBAC standard is designed in mind such that when practitioners implemented RBAC into their systems, The RBAC standard demonstrates reasonable assurance being based off a well thought out model.
+%there would be a reasonable assurance being based off a well thought out model.  
+As extensions to the RBAC Reference Mode
 come along, thought and time should be given to how features and nuances of their models may impact implementation
-in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
+to achieve the same goals as the original standard.  The primary sources showed a lack of implementation
 with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as four papers did, should be
@@
 
-\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for.}
+When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
+Further, in the case of extensions to the RBAC Reference Model, the model should be accompanied by validation of the model as a stand-alone
+access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
+extension models are lacking.
+
+The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
+scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
+authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario. In rare cases, 
+one paper~\cite{motta03:contextual} from our results, a case study is performed to examine how the proposed model works in practice. Discussions
+of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
+for what types of situations the proposed model could handle.
+Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
+have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
+case studies of a model in action provide developers with evidence that the model works as intended when applied.
+
+When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
+Two of the primary sources examined provided time-based performance analysis of their extension model compared
+to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
+requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
+some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
+over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
+need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
+entities.
+
+%\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for.}
+\greybox{Research on RBAC extension models shows a lack of implementations in the real world scenarios that the models are ultimately designed for.
+RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
 
@@
 
-\section{Evaluations} \label{sec:evaluations}
-
-RQ4: How are extensions to RBAC evaluated theoretically and in practice?
-
-The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
-Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
-evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
-types.
-
-\subsection{Results}
-
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
-and how application of their model would apply and resolve the situation.  6 models provided some form of performance
-or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
-as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
-The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
-would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
-or show how the Reference Model is deficient in tackling said scenarios.
-
-\subsection{Analysis and Discussion}
-
-When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
-Further, in the case of extensions to the RBAC Reference Model, the model should be accompanied by validation of the model as a stand-alone
-access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
-extension models are lacking.
-
-The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
-scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario. In rare cases, 
-one paper~\cite{motta03:contextual} from our results, a case study is performed to examine how the proposed model works in practice. Discussions
-of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
-for what types of situations the proposed model could handle.
+%\section{Evaluations} \label{sec:evaluations}
+%
+%RQ4: How are extensions to RBAC evaluated theoretically and in practice?
+%
+%The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+%Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
+%evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
+%types.
+%
+%\subsection{Results}
+%
+%Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
+%and how application of their model would apply and resolve the situation.  6 models provided some form of performance
+%or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
+%as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
+%5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+%The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
+%would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
+%or show how the Reference Model is deficient in tackling said scenarios.
+%
+%\subsection{Analysis and Discussion}
+%
+%When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
+%Further, in the case of extensions to the RBAC Reference Model, the model should be accompanied by validation of the model as a stand-alone
+%access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
+%extension models are lacking. 
+%
+%The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
+%scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
+%authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario. In rare cases, 
+%one paper~\cite{motta03:contextual} from our results, a case study is performed to examine how the proposed model works in practice. Discussions
+%of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
+%for what types of situations the proposed model could handle.
 Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
-have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
-case studies of a model in action provide developers with evidence that the model works as intended when applied.
-
-When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
-Two of the primary sources examined provided time-based performance analysis of their extension model compared
-to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
-requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
-some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
-over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
-need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
-entities.
-
-\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
+%have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
+%case studies of a model in action provide developers with evidence that the model works as intended when applied.
+%
+%When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
+%Two of the primary sources examined provided time-based performance analysis of their extension model compared
+%to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
+%requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
+%some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
+%over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
+%need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
+%entities.
+%
+%\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
 
@@
 \end{tabular*}
-\label{tab:implementations}
+\label{tab:evaluations}
 \end{table}
 
-\section{Domains} \label{sec:domains}
+\section{RQ5: Domains} \label{sec:domains}
 
-RQ5: What domains have RBAC extensions been created for?
+\textit{What domains have been targeted by RBAC extension models?}
 
@@
 
-We found that the domains mentioned and their associated sources are:
+Table \ref{tab:domains} shows domains mentioned and their associated sources.
 
@@
 \cite{masoumzadeh2008purbac}
+\cite{haibo05:context}
 &
@@
 \end{tabular*}
-\label{tab:implementations}
+\label{tab:domains}
 \end{table}
@@
 of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.
-There were three papers~\cite{jian2008extended, aich07:STARBAC, haibo2005context} without any direct mention of a domain since Aich et al.
+There were three papers~\cite{jian2008extended, aich07:STARBAC} without any direct mention of a domain since Aich et al.
 \cite{aich09:role} fall under both the medical domain and mobile computing.
@@
 each of the eight categories except for the Organizational category. 
-The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
+The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and requirements.
 Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
@@
 
-\section{Generalizations} \label{sec:generalizations}
+\section{RQ6: Generalizations} \label{sec:generalizations}
 
-RQ6: What commonalities or generalizations exist across all categories?
+\textit{What commonalities exist across RBAC extension models?}
 
-During the paper reading phase, we identified commonalities or generalizations within the primary sources 
+During the paper reading phase, we identified commonalities within the primary sources
 by looking for formal modeling of extending RBAC extension models.

### 2014-02-09 paper/sections/appendix.tex
Appendix table edits and cleaning up bibliography.

@@
 
-\begin{sidewaystable}
+\begin{table}
 \centering
-\caption{Primary sources grouped by categorization}
+\caption{Primary sources by classification and properties (Implementation Type / Domain / Evaluation Type(s))}
 \vline
-\begin{tabularx}{\linewidth}{ | X | X | X | X | }
+\begin{tabularx}{\textwidth}{ | c | c | c | c | l | l | }
 \cline{1-4}
@@
 \multicolumn{1}{ c }{\multirow{2}{*} {\rotatebox{90}{ Context } }} & 
-\multicolumn{3}{ |l| }{A flexible content and context-based access control model for multimedia medical image database systems, 2001 \cite{tzelepi01:flexible}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.9\textwidth} | }{\cite{tzelepi01:flexible} A flexible content and context-based access control model for multimedia medical image database systems, 2001 (None / Medical / Scenario)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{ A Context-Aware Role-Based Access Control Model for Web Services, 2005 \cite{haibo2005context}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{haibo05:context} A Context-Aware Role-Based Access Control Model for Web Services, 2005 (None / Web / None)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A context-sensitive access control model and prototype implementation, 2000 \cite{cholewka00:acontext-sensitive}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{cholewka00:acontext-sensitive} A context-sensitive access control model and prototype implementation, 2000 (Prototype / Enterprise Workflow / Scenarios) } \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 \cite{huang06:pervasive}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{huang06:pervasive} A Context, Rule and Role-Based Access Control Model In Enterprise Pervasive Computing Environment, 2006 (Prototype / Pervasive Computing / Scenarios) } \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A contextual role-based access control authorization model for electronic patient record Motta, 2003 \cite{motta03:contextual}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{motta03:contextual} A contextual role-based access control authorization model for electronic patient record Motta, 2003 (Enterprise / Medical / Case Study) } \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{A Role and Context Based Access Control Model with UML, 2008 \cite{bao08:role}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{bao08:role} A Role and Context Based Access Control Model with UML, 2008 (Prototype / Enterprise Workflow / Complexity, Comparison to Standard, Scenarios)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{An extended RBAC model based on granular logic, 2008 \cite{jian2008extended}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{jian2008extended} An extended RBAC model based on granular logic, 2008 (None / Organizations / Scenarios)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{Designing an agent-based RBAC system for dynamic security policy, 2004 \cite{yamazaki04:designing}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{yamazaki04:designing} Designing an agent-based RBAC system for dynamic security policy, 2004 (None / Organizations / Scenarios)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 \cite{thein2011leveraging}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{thein2011leveraging} Leveraging Access Control Mechanism of Android Smartphone Using Context-Related Role-Based Access Control Model, 2011 (None / Mobile Computing / Scenarios)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ |l| }{CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 \cite{zou2009crbac}} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{zou2009crbac} CRBAC: Imposing multi-grained constraints on the RBAC model in the multi-application environment, 2009 (None / Mobile Computing / None)} \\ \cline{2-4}
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Const. }}}} &
-\multicolumn{2}{ |l| }{A constraint based role based access control in the SECTET a model-driven approach, 2006 \cite{alam06:constraint}} \\[2.75ex] \cline{2-4}
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{Const. }}}} &
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{alam06:constraint} A constraint based role based access control in the SECTET a model-driven approach, 2006 } \\[4ex] \cline{2-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org. }}}} &
-\multicolumn{2}{ |l| }{ROBAC: Scalable Role and Organization Based Access Control Models, 2006 \cite{zhang06:collaborative}} \\[2.75ex]  \cline{2-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhang06:collaborative} ROBAC: Scalable Role and Organization Based Access Control Models, 2006 } \\[2.75ex]  \cline{2-4}
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{ Priv. }}}} &
-\multicolumn{2}{ |l| }{Privacy-aware role-based access control, 2007 \cite{ni2010privacy}} \\[1ex] \cline{3-4}
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{Priv. }}}} &
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ni2010privacy} Privacy-aware role-based access control, 2007 } \\[1ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{PuRBAC: Purpose-Aware Role-Based Access Control, 2008 \cite{masoumzadeh2008purbac}} \\[1ex] \cline{2-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{masoumzadeh2008purbac} PuRBAC: Purpose-Aware Role-Based Access Control, 2008 } \\[1ex] \cline{2-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Task }}}} &
-\multicolumn{2}{ |l| }{A Task-Role Based Access Control Model with Multi-Constraints, 2008 \cite{yao2008task}} \\ \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{yao2008task} A Task-Role Based Access Control Model with Multi-Constraints, 2008 } \\ \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{Team and Task Based RBAC Access Control Model, 2009 \cite{zhou2007team}} \\ \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhou2007team} Team and Task Based RBAC Access Control Model, 2009 } \\ \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{Task-role-based access control model, 2003 \cite{oh2003task}} \\ \cline{2-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{oh2003task} Task-role-based access control model, 2003 } \\ \cline{2-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{10}{*}{\rotatebox{90}{ Spatio-Temporal }}}} &
-\multicolumn{2}{ |l| }{STARBAC: Spatiotemporal role based access control, 2007 \cite{aich07:STARBAC}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich07:STARBAC} STARBAC: Spatiotemporal role based access control, 2007 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{On spatio-temporal constraints and inheritance in role-based access control, 2008 \cite{chen08:spatio-temporal}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chen08:spatio-temporal} On spatio-temporal constraints and inheritance in role-based access control, 2008 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 \cite{samuel07:spatio-temporal}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{samuel07:spatio-temporal} A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{LoT-RBAC: A Location and Time-Based RBAC Model, 2005 \cite{chandran05:llt}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chandran05:llt} LoT-RBAC: A Location and Time-Based RBAC Model, 2005 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{A Spatio-temporal Role-Based Access Control Model, 2007 \cite{ray07:spatio}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ray07:spatio} A Spatio-temporal Role-Based Access Control Model, 2007 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ |l| }{Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 \cite{aich09:role}} \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich09:role} Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
-\multicolumn{1}{ |l| }{GEO-RBAC: a spatially aware RBAC, 2005 \cite{damiani2007geo}} \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ | p{.6\textwidth} | }{\cite{damiani2007geo} GEO-RBAC: a spatially aware RBAC, 2005 } \\[0.25ex] \cline{4-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
-\multicolumn{1}{ |l| }{LRBAC: A location-aware role-based access control model, 2006 \cite{ray2006lrbac}} \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ | p{.6\textwidth} | }{\cite{ray2006lrbac} LRBAC: A location-aware role-based access control model, 2006 } \\[0.25ex] \cline{4-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
-\multicolumn{1}{ |l| }{Spatial role-based access control model for wireless networks, 2003 \cite{hansen2003spatial}} \\[0.25ex] \cline{3-4}
+\multicolumn{1}{ | p{.6\textwidth} | }{\cite{hansen2003spatial} Spatial role-based access control model for wireless networks, 2003 } \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp. }}}} &
-\multicolumn{1}{ |l| }{A generalized temporal role-based access control model, 2005 \cite{joshi05:generalized}} \\[5ex]
+\multicolumn{1}{ | p{.6\textwidth} | }{\cite{joshi05:generalized} A generalized temporal role-based access control model, 2005 } \\[4ex]
 
@@
 \label{tab:categorization}
-\end{sidewaystable}
+\end{table}

### 2014-02-09 paper/sections/analysis.tex
Updating sections.

--- 2014-02-08 paper/sections/analysis.tex
@@
-\section{Motivations} \label{sec:motivations}
+\section{RQ2: Motivations} \label{sec:motivations}
 
-RQ2: What are the motivations behind the RBAC extension models?
-
-%\subsection{Results}
+\textit{What are the motivations behind the RBAC extension models?}
+\\
 
@@
 
-\section{Implementations} \label{sec:implementations}
+\section{RQ3: Evaluations} \label{sec:evaluations}
 
-RQ3: Do the RBAC extension models have corresponding implementations?
+\textit{How are RBAC extension models implemented and evaluated by their authors?}
+\\
 
-When designing and proposing a model targeted at a feature that is rooted in practical
-usage by real software systems, bringing the model to life is strong evidence that the
-proposed model can work in practice. The concept of authorization and access control
-is rooted in a business need. Thus, any access control model needs to be feasible
-in the real world not just on paper. We analyzed the primary sources to see how many
-proposed models actually had implementations associated with them.  Then, we quantified
-types of implementations. Whether the implementation was for a real system, for a prototype
-and/or used in a production environment.
+When designing and proposing a model targeted at a feature that is rooted in practical usage by real software systems, bringing the model to life is strong evidence that the proposed model can work in practice.
+The concept of authorization and access control is rooted in a business need. 
+Thus, any access control model needs to be feasible in the real world not just on paper. 
+We analyzed the primary sources to see how many proposed models actually had implementations associated with them. 
+Then, we quantified types of implementations. 
+Whether the implementation was for a real system, for a prototype and/or used in a production environment.
 
@@
 
-\section{Domains} \label{sec:domains}
+\section{RQ4: Domains} \label{sec:domains}
 
-RQ5: What domains have RBAC extensions been created for?
-
-%\subsection{Results}
+\textit{What domains have RBAC extensions been created for?}
+\\
 
@@
 
-\section{Generalizations} \label{sec:generalizations}
+\section{RQ5: Generalizations} \label{sec:generalizations}
 
-RQ6: What commonalities or generalizations exist across all categories?
+\textit{What commonalities or generalizations exist across all categories?}
+\\
 
-During the paper reading phase, we identified commonalities or generalizations within the primary sources 
-by looking for formal modeling of extending RBAC extension models.
+During the paper reading phase, we identified commonalities or generalizations within the primary sources by looking for formal modeling of extending RBAC extension models.
 

### 2014-02-09 paper/sections/intro.tex
Updating data display tables.
update intro
Add space before citation []

administration[? ] -> administration [? ]
the 1990s[? ], again in 2002[? ] -> the 1990s [? ], again in 2002 [? ]
and later confirmed in 2010[? ]. -> and later confirmed in 2010 [? ].

@@
 The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows and the number of permissions and users scales upward.
-As permissions can be managed by role instead of by user, RBAC has been shown to significantly reduce complexity of security administration\cite{impact2002}.
+As permissions can be managed by role instead of by user, RBAC has been shown to significantly reduce complexity of security administration \cite{impact2002}.
 For example, if a user requires access to resources associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead of assigning a set of individual permissions.
@@
 In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute\footnote{http://www.ansi.org/} (ANSI) and the InterNational Committee for Information Technology Standards\footnote{http://www.incits.org/} (INCITS).
-The development of a standard was inspired by an economic impact study done during the 1990s\cite{impact1996}, again in 2002\cite{impact2002} and later confirmed in 2010\cite{o20102010}. 
+The development of a standard was inspired by an economic impact study done during the 1990s \cite{impact1996}, again in 2002 \cite{impact2002} and later confirmed in 2010 \cite{o20102010}. 
 The study showed the cost savings of RBAC implementation and maintenance. 
@@
 \item A summary of current evaluation methods used in research on RBAC extension models.
-\item A starting direction for a new standard based on conclusions drawn from the RBAC extension models.
+\item A starting point for a new standard based on conclusions drawn from the RBAC extension models.
 \end{itemize}

### 2014-02-09 paper/sections/conclusion.tex
update category names
change updated categories naming.
-> categories with Const, Org, Priv, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context.

@@
 
-The RBAC extension models were revealed to fall into a number of categorizations with Organization, Privacy, Resource, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context.
+The RBAC extension models were revealed to fall into a number of categories with Const, Org, Priv, Task, Spatio-Temporal, Spatial, and Temporal falling under the general category of context. 
 The categories each had properties specific to their implementation, but were seen to generalize to being specialized instances of context tailored to the entities or actions the categories covered.

### 2014-02-09 paper/sections/analysis.tex
Update analysis.tex
combining..
Updating formatting of RQ3/4 by combining both sections entirely.

@@
 
+The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
+evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
+types.
+
 \subsection{Results}
@@
 mention of an implementation or prototype.
+
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
+and how application of their model would apply and resolve the situation.  6 models provided some form of performance
+or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
+as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
+5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
+would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
+or show how the Reference Model is deficient in tackling said scenarios.
 
@@
 implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
-
-\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for.}
 
@@
 
-\section{Evaluations} \label{sec:evaluations}
-
-RQ4: How are extensions to RBAC evaluated theoretically and in practice?
-
-The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
-Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
-evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
-types.
-
-\subsection{Results}
-
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
-and how application of their model would apply and resolve the situation.  6 models provided some form of performance
-or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
-as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
-The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
-would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
-or show how the Reference Model is deficient in tackling said scenarios.
-
-\subsection{Analysis and Discussion}
-
 When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
@@
 
-\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
+\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for. 
+Further, comprehensive evaluation of the models were found to be significantly lacking. }
 

### 2014-02-09 paper/sections/categorization.tex
Updating sections.
Updating formatting of RQ3/4 by combining both sections entirely.

@@
 \textit{How can RBAC extension models be classified?}
+\\
 
-During the paper reading phase, we identify classification themes, which are
-central topic or context of the RBAC extension models. Our classification process
-yields a set of categories based upon observations during the paper reading phase.
+During the paper reading phase, we identify classification themes, which are central topic or context of the RBAC extension models. 
+Our classification process yields a set of categories based upon observations during the paper reading phase.
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
@@
 
-The previous example paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in
-the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
+The previous example paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} contained ``privacy'' in the title and in the name of the model leading to the creation of the Privacy category and the subsequent placement of the paper under that category.
 By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name. 
 However, in reading the body of the paper, we determined that this paper discussed RBAC extension based on context. 
-We offer definitions for each of the eight observed categories based on the data extracted from the primary sources and the English definitions for each noun-phrase. 
-Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Table \ref{tab:categorization}.
+In Section 4.1, we offer definitions for each of the eight observed categories.
 
 \subsection{Results}
+
+Based on data extracted from the primary sources and the English definitions for each noun-phrase we provide a definition for each observed model category. 
+Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Appendix A \ref{tab:categorization} which presents each primary source within it's designated category.
 
@@
   \item \textbf{Context}: The extension model integrates contextual information into the RBAC standard model.
-Context is defined as a user's current state and environment (e.g., location, time, system resources, network state, network security configuration, etc) which may affect the user's access privileges.
+Context is defined as a user's current state and environment (e.g., location, time, system resources, network state, network security configuration, etc).
+The user's access privileges are dependent upon the values of the current state and environment at any given time.
 
@@
 For example, a doctor may modify any medical record for which the doctor is assigned as the designated primary care physician.
-This example describes a doctor's permission with a conditional restriction - ``only when the doctor is assigned as the designated primary care physician may modify a particular medical record''.
+This example describes a doctor's permission with a conditional restriction - ``only when the doctor is assigned as the designated primary care physician may the doctor modify a particular medical record''.
 
-  \item \textbf{Organizational (Org)}: The RBAC extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
-Typically, users may have the same role name in different organizations, but may have different access privileges due to different local variations.
+  \item \textbf{Organizational (Org)}: The extension model is concerned with providing mechanisms and entities that allow for RBAC across multiple organizations.
+Typically, users may have the same role name in different organizations, but may have different access privileges due to departmental variations.
 
-  \item \textbf{Privacy (Priv)}: The RBAC extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or date of birth.
+  \item \textbf{Privacy (Priv)}: The extension model provides entities and mechanisms to describe privacy policies, which are legal statements or documents about disclosure or management of personally identifiable information such as name, address, or date of birth.
 
-  \item \textbf{Task}: The RBAC extension model provides task entities which are associated with permissions and roles.
+  \item \textbf{Task}: The extension model provides task entities which are associated with permissions and roles.
 A task is a fundamental unit of a business activity.
@@
 Roles are associated with tasks that are associated with permissions.
-For example, the employee role is associated with a task, for example, to write a report.
-Then, this task is associated with a permission.
+For example, the employee role is associated with a task to write a report.
+This task is then associated with a permission.
 
-  \item \textbf{Spatio-Temporal}: The RBAC extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
 For example, specific locations permit roles to conduct actions from 8:00 a.m.
@@
 
-  \item \textbf{Spatial}: The RBAC extension model provides spatial constraints, location-based constraints in specifying access
+  \item \textbf{Spatial}: The extension model provides spatial (location-based) constraints in specifying access
 	control policies. For example, in organizations, locations are enforced whereas a
-	specific role is permitted to conduct an action. Consider that a temporal employee works only in a specific location.
-	In such cases, the temporal employee role should access required resources only when the user is in that location. 
+	specific role is permitted to conduct an action. Consider that an employee works only at a specific location.
+	In such cases, a role should allow access to required resources only when the user is in that location. 
 	Spatial constraints can integrate with roles, user-role assignments, or role-permission assignments.
 
-  \item \textbf{Temporal (Temp)}:  The RBAC extension model provides temporal constraints, time-based constraints in specifying access
+  \item \textbf{Temporal (Temp)}:  The extension model provides temporal (time-based) constraints in specifying access
 	control policies. For example, in organizations, periodic temporal durations are enforced whereas a
 	specific role is permitted to conduct an action. Consider that a temporal employee works only from 9:00 a.m. to 3:00 p.m.
-	In such cases, the temporal employee role should be allowed to access required resources during the interval. 
+	In such cases, the temporal employee role should only be allowed to access required resources during the interval. 
 	Temporal constraints can integrate with roles, user-role assignments, or role-permission assignments.

### 2014-02-09 paper/sections/categorization.tex
Adding back analysis section and cleaning up.

@@
 
-Based on data extracted from the primary sources and the English definitions for each noun-phrase we provide a definition for each observed model category. 
-Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Appendix A \ref{tab:categorization} which presents each primary source within it's designated category.
+We provide a definition for each observed model category based on data extracted from the primary sources and the English definitions for each noun-phrase.
+Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Appendix 1 \ref{tab:categorization} which presents each primary source within it's designated category.
 
@@
 The Constraint, Organizational, Privacy, Task, and Spatial and Temporal categories can be special cases of the Context category. 
-The Spatio-Temporal category subsumes Spatial and Temporal categories.
-The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and the Spatio-Temporal category contained more primary sources that the Spatial or Temporal categories alone.
+The Spatial and Temporal categories were treated as subsets of the broader category of Spatio-Temporal since this category encompasses them individually and the Spatio-Temporal category was derived from primary sources directly.
 
-When looking across all categories, we noted that each category added domain specific features on top of the RBAC Reference Model.
-These domain specific features were under the surface these features adding contextual relationships between the core user, permission and role entities.  Thus, we
-concluded that all categories stemmed from the context category, of which some primary sources were already deemed direct members.
+When looking across all categories, we noted that each category added specific features on top of the RBAC Reference Model.
+These features were under the surface adding contextual relationships between the core user, permission and role entities. 
+Thus, we concluded that all categories stemmed from the context category, of which some primary sources were already deemed direct members.
 

### 2014-02-09 paper/sections/process.tex
Updating data display tables.
Adding back analysis section and cleaning up.

@@
 \caption{Paper counts after applying search strategy}
-\begin{tabular}{|p{3.5cm}|p{1.25cm}|p{4.25cm}|p{4.25cm}|p{1cm}|}
+\vspace{0.1 in}
+\begin{tabular}{ | l | r | r | r | r | }
 
@@
 The title elimination round resulted in Reviewer 1 selecting 305 papers, and Reviewer 2 selecting 176 papers with 149 papers of overlap between the two. 
-Between the two reviewers selections, there were 332 papers not in common, of which, 149 were rejected and 141 retained after a second review.
+There were 332 papers found not to be in common, of which, 149 were rejected and 141 retained after a second review.
 
@@
 The abstract elimination round resulted in Reviewer 1 selecting 86 papers, and Reviewer 2 selecting 102 papers with 51 papers of overlap between the two. 
-Between the two reviewers selections, there were 137 papers not in common, of which, 116 were rejected and 21 retained after a second review.
+There were 137 papers found not to be in common, of which, 116 were rejected and 21 retained after a second review.
 
@@
 \caption{Elimination Rounds}
-\begin{tabular}{ c c | c | c | c | }
+\vspace{0.1 in}
+\begin{tabular}{ l l | r | r | r | }
 \cline{3-5}

### 2014-02-09 paper/sections/categorization.tex
Adding note about Figure 2 and cleaning up mentions of Table 6.

@@
 We provide a definition for each observed model category based on data extracted from the primary sources and the English definitions for each noun-phrase.
-Some category descriptors contain abbreviations in parenthesis that match the shortened name found within Appendix 1 \ref{tab:categorization} which presents each primary source within it's designated category.
+Some category descriptors contain abbreviations in parenthesis that match the shortened name found in Table 6 (within Appendix 1) \ref{tab:categorization} which presents each primary source within it's designated category.
 
@@
 Table \ref{tab:categorization} summarizes each primary source under their designated category and furthermore, displays the perceived hierarchy of the categories. 
-Figure~\ref{fig:category_structrure} shows the hierarchy structure among categories. 
+Figure~\ref{fig:category_structrure} shows the hierarchy structure among categories and the counts of primary sources for each.
+For categories that have sub-categories, the totals of the category and sub-categories is provided as the second number in the figure. 
 The Constraint, Organizational, Privacy, Task, and Spatial and Temporal categories can be special cases of the Context category. 

### 2014-02-11 paper/sections/analysis.tex
Adding back analysis section and cleaning up.
Fixing broken reference.
Adding note about Figure 2 and cleaning up mentions of Table 6.
Updating counts in appendix and other data tables.

--- 2014-02-09 paper/sections/analysis.tex
@@
 
-\textit{What are the motivations behind the RBAC extension models?}
+\textit{What are motivations behind the RBAC extension models?}
 \\
@@
 
-  \item \textbf{Const}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.
+  \item \textbf{Constraint}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.
 
-  \item \textbf{Org}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations.
+  \item \textbf{Organizational}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations.
 The model needs to reduce the administrative complexity of RBAC across multiple organizations.
 
-  \item \textbf{Priv}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.
+  \item \textbf{Privacy}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.
 
@@
 
-  \item \textbf{Temp}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.
+  \item \textbf{Temporal}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.
 
@@
 While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitations on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
+For example, in cases where the RBAC extension model does not support the notion of specific constraints, the RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
+The RBAC extension model provides general agreement on the definition for specific features.
+The RBAC extension model helps simplify theoretical modeling and practical implementation of features in the model.
 
-The Context, Org, Priv, Spatio-Temporal, Spatial, and Temp categories claim that the RBAC extension model does not support the notion of specific constraints. 
+%The Const category claims that the RBAC extension model has limitations on its features.
+The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model.
+For models in this category, changes occur within the RBAC Reference Model.
+%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles.
+Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations.
+For this combination, new associations between attributes of two models should be introduced.
+Models in this category require new associations between attributes of different access control models to work together.
+%
+%
+%The Context, Org, Priv, Spatio-Temporal, Spatial, and Temp categories claim that the RBAC extension model does not support the notion of specific constraints.
 The RBAC extension model in these categories incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
 The extension model does not change the RBAC Reference Model.
-
-The Const category claims that the RBAC extension model has limitations on its features. 
+%
+%The Const category claims that the RBAC extension model has limitations on its features.
 The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model.
 For models in this category, changes occur within the RBAC Reference Model.
-
-The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles. 
+%
+%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles.
 Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations.
@@
 
-\section{RQ3: Evaluations} \label{sec:evaluations}
+\section{RQ3: Implementations} \label{sec:implementations}
 
-\textit{How are RBAC extension models implemented and evaluated by their authors?}
+\textit{How are RBAC extension models implemented by their authors?}
 \\
 
-When designing and proposing a model targeted at a feature that is rooted in practical usage by real software systems, bringing the model to life is strong evidence that the proposed model can work in practice.
-The concept of authorization and access control is rooted in a business need. 
-Thus, any access control model needs to be feasible in the real world not just on paper. 
-We analyzed the primary sources to see how many proposed models actually had implementations associated with them. 
-Then, we quantified types of implementations. 
-Whether the implementation was for a real system, for a prototype and/or used in a production environment.
+When designing and proposing a model targeted at a feature that is rooted in practical
+usage by real software systems, implementing the model is evidence that the
+proposed model can work in practice. The presence of an actual implementation substantiates the robustness of the design and/or the need for the extension.
+We analyzed the primary sources to see how many
+proposed models actually had implementations associated with them.
+We quantified whether the implementation was enterprise (i.e., an implementation used in a production environment) was for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).
 
@@
 Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
-Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
+Of the 27 papers surveyed, RBAC extension models in four papers were prototypes developed by the authors whereas RBAC extension models in three papers were claimed to be implemented within a real
 system. The remaining 20 papers provide no
-mention of an implementation or prototype.
-
-\subsection{Analysis and Discussion}
-
-The RBAC standard is designed with enterprises in mind such that when practitioners implemented RBAC into their systems,
-there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC Reference Mode
-come along, thought and time should be given to how features and nuances of their models may impact implementation
-in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
-with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as four papers did, should be
-a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
-implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
-
-\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for.}
+mention of an implementation.
 
@@
 \cite{bao08:role}
-\cite{zhou07:network}
+\cite{zhou2007team}
 &
@@
 
-\section{Evaluations} \label{sec:evaluations}
+\subsection{Analysis and Discussion}
 
-RQ4: How are extensions to RBAC evaluated theoretically and in practice?
+The RBAC standard is designed such that when practitioners implemented RBAC into their systems, the RBAC standard demonstrates reasonable assurance being based off a well thought out model.
+As extensions to the RBAC Reference Mode come along, thought and time would be given to how features and nuances of their models may impact implementation to achieve the same goals as the original standard.
+The primary sources showed a lack of implementation
+with over 70\% of the models having no notion of attempting to implement them.
+Prototype implementation shows the feasibility of the RBAC extension models in the four papers.
+Of the models that produced an implementation within the enterprise world, two were from within the medical domain and one was implemented using web application technologies.
+
+\greybox{Research on RBAC extension models shows a lack of implementations in the real world scenarios that the models are ultimately designed for.}
+
+\section{RQ4: Evaluations} \label{sec:evaluations}
+
+\textit{How are RBAC extension models evaluated by their authors?}
+\\
 
 The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
-Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
-evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
-types.
+We identified a set of evaluation types found within  each of the primary sources and provide a list of the evaluation types and which primary sources provided which type.
+A primary source may use multiple evaluation types.
+For example, Aich et al.
+conducted performance and complexity analyses to show the effectiveness of their proposed RBAC extension model \cite{aich09:role}.
 
@@
 
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
-and how application of their model would apply and resolve the situation.  6 models provided some form of performance
+Table \ref{tab:evaluations} shows evaluation types.
+Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Fourteen models presented example scenarios
+and how application of their model would apply and resolve the situation. Six models provided some form of performance
 or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
 as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+Five models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
 The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
-would tackle those scenarios. Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
-or show how the Reference Model is deficient in tackling said scenarios.
-
-\subsection{Analysis and Discussion}
-
-When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
-Further, in the case of extensions to the RBAC Reference Model, the model should be accompanied by validation of the model as a stand-alone
-access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
-extension models are lacking.
-
-The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
-scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario. In rare cases, 
-one paper~\cite{motta03:contextual} from our results, a case study is performed to examine how the proposed model works in practice. Discussions
-of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
-for what types of situations the proposed model could handle.
-Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
-have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
-case studies of a model in action provide developers with evidence that the model works as intended when applied.
-
-When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
-Two of the primary sources examined provided time-based performance analysis of their extension model compared
-to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
-requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
-some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
-over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
-need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
-entities.
-
-\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
+would tackle those scenarios. 
+%Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
+%or show how the Reference Model is deficient in tackling said scenarios.
 
@@
 \cite{chen08:spatio-temporal}
-\cite{aich09:role}
+\cite{aich07:STARBAC}
 &
@@
 1 \\ \cline{1-3}
+None
+&
+\cite{masoumzadeh2008purbac}
+\cite{chandran05:llt}
+\cite{ray2006lrbac}
+&
+3 \\ \cline{1-3}
 
 \end{tabular*}
-\label{tab:implementations}
+\label{tab:evaluations}
 \end{table}
 
-\section{RQ4: Domains} \label{sec:domains}
+\subsection{Analysis and Discussion}
 
-\textit{What domains have RBAC extensions been created for?}
+When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
+Further, in the case of extensions to the RBAC Reference Model, the model is accompanied by validation of the model as a stand-alone
+access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
+extension models are lacking.
+
+The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
+scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
+authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
+In Table \ref{tab:evaluations}, we observe that one paper includes a case study to examine how the proposed model works in practice. Discussions
+of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
+for what types of situations the proposed model could handle.
+Further, by not providing a comparison to the RBAC Reference Model, developers may be left implementing a more complex model to address their requirements when the RBAC Reference Model would
+have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
+case studies of a model in action provide developers with evidence that the model works as intended when applied.
+
+When looking for an enterprise ready access control mechanism, developers must balance usability with security. 
+Two of the primary sources examined provided time-based performance analysis of their extension model compared
+to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
+requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
+some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
+over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
+need to ensure non-functional requirements are met that deal with the ability for a system administrator to effectively manage these
+entities.
+
+\greybox{RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
+
+\section{RQ5: Domains} \label{sec:domains}
+
+\textit{What domains have been targeted by RBAC extension models?}
 \\
@@
 \cite{chandran05:llt}
-\cite{ray07:spatio}
 \cite{aich09:role}
+\cite{ray2006lrbac}
 &
-5 \\ \cline{1-3}
+4 \\ \cline{1-3}
 Organizations with Many Sub-departments
@@
 \cite{yamazaki04:designing}
-\cite{jian2008extended}
 \cite{yao2008task}
 &
-3 \\ \cline{1-3}
+2 \\ \cline{1-3}
 Enterprise Workflows
@@
 5 \\ \cline{1-3}
+None
+&
+\cite{jian2008extended}
+\cite{aich2007:STARBAC}
+&
+2 \\ \cline{1-3}
 
 \end{tabular*}
-\label{tab:implementations}
+\label{tab:domains}
 \end{table}
@@
 of domains was pervasive computing environments and large-scale organizations with three each and web applications with one.
-There were three papers~\cite{jian2008extended, aich07:STARBAC} without any direct mention of a domain since Aich et al.
+Two papers~\cite{jian2008extended, aich07:STARBAC} do not mention domains explicitly since Aich et al.
 \cite{aich09:role} fall under both the medical domain and mobile computing.
@@
 \subsection{Analysis and Discussion}
+
+The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of RBAC extension models.  
+Moreover, we observed that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
+each of the eight categories except for the Organizational category. 
+The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and requirements.
+Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
+the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
+of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
+of both the RBAC standard and current regulations.
+
+The RBAC standard has been re-enforced by the economic impact that standardization has had on enterprises needing to apply access control.  The
+inclusion of extension models targeted at the enterprise workflow domain is indicative of the expansion of requirements for enterprises. Developers
+and researchers would take care when looking at extension models designed to address the newer requirements of enterprise workflows to
+achieve the same economic implementation and maintainability benefits the RBAC Reference Model presents.
+
+Figure~\ref{fig:dist_domains} shows domain distribution by 3-year period.
+We observed that medical and enterprise domain papers constantly appear for every period in Figure~\ref{fig:dist_domains}.
+For a domain that has roots in medical and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity. 
+Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone was introduced \cite{buck2013new}.
+Since 2004, the domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
+
+\greybox{RBAC extension models were found targeting domains such as medical, pervasive computing, web applications, enterprise workflows and mobile computing.}
 
@@
 
-The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of RBAC extension models.  
-Moreover, we observed that the categories associated with papers identifying the medical domain were not limited to one or two but cut across
-each of the eight categories except for the Organizational category. 
-The cross-category nature of the medical domain papers appears indicative of the complex nature of medical applications and the requirements therein.
-Given the growth of the research and development of medical applications over the past decade this result does not appear to be surprising. However,
-the RBAC standard was originally created to reduce cost and increase interoperability - two goals of current regulation around the standardization
-of electronic health record systems. The large number of proposed models, and the cross-category result stand in direct opposition of the goals
-of both the RBAC standard and current regulations.
+\section{RQ6: Generalizations} \label{sec:generalizations}
 
-The RBAC standard has been re-enforced by the economic impact that standardization has had on enterprises needing to apply access control.  The
-inclusion of extension models targeted at the enterprise workflow domain is indicative of the expansion of requirements for enterprises. Developers
-and researchers should take care when looking at extension models designed to address the newer requirements of enterprise workflows in order to
-achieve the same economic implementation and maintainability benefits the RBAC Reference Model presents.
-
-Figure~\ref{fig:dist_domains} shows domain distribution by 3-year period.
-We observed that medical and enterprise domain papers constantly appear for every period in Figure~\ref{fig:dist_domains}.
-For a domain that has roots in medical and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity. 
-Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone was introduced.
-Since 2004, the domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
-
-\greybox{RBAC extension models were found targeting domains such as medical, pervasive computing, web applications, enterprise workflows and mobile computing.}
-
-\section{RQ5: Generalizations} \label{sec:generalizations}
-
-\textit{What commonalities or generalizations exist across all categories?}
+\textit{What commonalities exist across RBAC extension models?}
 \\
 
-During the paper reading phase, we identified commonalities or generalizations within the primary sources by looking for formal modeling of extending RBAC extension models.
+During the paper reading phase, we identified commonalities within the primary sources
+by looking for formal modeling of extending RBAC extension models.
 
@@
 The RBAC standard is used in various aspects of computer systems.
-In order to reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
+To reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
 The authors use propositional logic to describe access control model across all categorizations.
@@
 In propositional logic, simple (i.e., atomic) or compound condition at given context is evaluated to true or false based on specified rules and access control logic.
-Researchers are concerned to extend limited set of propositions specific to core RBAC to meet real-world scenarios such as dynamic constraints, temporal, or spatial constraints.
-However, semantic meanings of such propositions are various based on researchers' intention.
+Researchers and practitioners are concerned to extend limited set of propositions specific to core RBAC to meet real-world scenarios such as dynamic constraints, temporal, or spatial constraints.
 
@@
 As access control is typically evaluated to either true or false based on predicates, propositional logic is sufficient to describe key ideas and definitions.
+Propositional logic specifies what combination of attributes values a request must satisfy to be authorized to roles.
+Combination of attribute values may handle constraints such as temporal and spatial constraints.
+For example, we describe user-role authorization as $ae$ $\Rightarrow$ $r$ using propositional logic where $ae$ is an attribute expression and $r$ is a role.
+Propositional logic for RBAC has two parts.
+The left hand side (LHS) is an attribute expression $ae$.
+The right hand side (RHS) is a role.
+If a request satisfies the attribute expression, a user $u$ in the request is authorized to the role specified in RHS.
+
 We found that 27 papers of RBAC extensions use propositional logic to describe its extended model.
@@
 reduce the administrative complexity of associations such as user- role associations.
-In order to support for reasoning of access control, one may describe the RBAC extended models in first-order logic. 
+To support for reasoning of access control, one may describe the RBAC extended models in first-order logic. 
 Given an RBAC extension model, Samuel et al.~\cite{samuel07:spatio-temporal} proposed verification of the model using a specification language, which is based first-order logic.

### 2014-02-11 paper/sections/intro.tex
Updating counts in appendix and other data tables.

@@
 
-Since the introduction of the RBAC standard, researchers have proposed domain-targeted extensions that add one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
+Since the introduction of the RBAC standard, researchers and practitioners have proposed domain-targeted extensions that add one or more features on top of components in the RBAC Reference Model \cite{kuhn2010adding}.
 For example, extensions that target the medical and mobile domains provide dynamic context or privacy around the access control policies.
@@
 \item A summary of current evaluation methods used in research on RBAC extension models.
-\item A starting point for a new standard based on conclusions drawn from the RBAC extension models.
 \end{itemize}

### 2014-02-11 paper/sections/abstract.tex
Updates for introduction references and research question refining.
Updating counts in appendix and other data tables.

@@
 {\bf Context}: Since the United States National Institute of Standards and Technology proposed Role-Based Access Control (RBAC) as documented in the RBAC Reference Model in the late 1990's, domain-targeted extensions have been proposed.
-For example, the mobile domain has surfaced a need for permission granting and role activation.
+For example, the mobile domain has identified a need for permission granting and role activation.
 \\
@@
 \noindent
-{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding
-how RBAC extension models are evaluated by providing an assessment of the state of RBAC extension models.
+{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding how RBAC extension models are evaluated by providing an assessment of the state of RBAC extension models.
 \\
@@
 \noindent
-{\bf Results}: We identified and classified the RBAC extension models into eight extension categories: Constraint, Context, 
-Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal.
+{\bf Results}: We identified and classified the RBAC extension models into eight extension categories: Constraint, Context, Organization, Privacy, Task, Spatio-Temporal, Spatial, and Temporal.
 Only 8 of the 27 papers provided an implementation of their model in the form of an enterprise application or prototype.
-The primary domains
-that inspired extensions were the medical domain with 9 of the 27 models, enterprise workflows with 5 of 27, and mobile computing with 5 of 27.
+The primary domains that inspired extensions were the medical domain with 9 of the 27 models, enterprise workflows with 5 of 27, and mobile computing with 5 of 27.
+The state of evaluation of RBAC model extensions is inconsistent with 4 papers comparing themselves to the RBAC Reference Model, 14 presenting example scenarios, and 12 lacking any evaluation that we could discern.
 \\
@@
 \noindent
-{\bf Conclusions}: Our literature review shows that all eight of the RBAC extension categories we identified deal with context.
-The state of evaluation of RBAC model extensions is limited with 1 paper doing an evaluation comparing itself to the RBAC Reference Model, 8 presenting example
-scenarios of their model in action, and 12 of the 27 models lacking an evaluation that we could discern.
+{\bf Conclusions}: Our literature review shows that all eight of the RBAC extension categories we identified deal with context whereby the privileges provided to a role are environmental depending upon factors such as location or time.
 The magnitude and scope of extensions to the RBAC standard suggests a revised standard may be beneficial.

### 2014-02-11 paper/sections/analysis.tex
Updating counts in appendix and other data tables.

--- 2014-02-09 paper/sections/analysis.tex
@@
 
-\section{RQ3: Evaluations} \label{sec:evaluations}
+\section{RQ3: Implementations and Evaluations} \label{sec:implementations}
 
@@
 
-When designing and proposing a model targeted at a feature that is rooted in practical usage by real software systems, bringing the model to life is strong evidence that the proposed model can work in practice.
-The concept of authorization and access control is rooted in a business need. 
-Thus, any access control model needs to be feasible in the real world not just on paper. 
-We analyzed the primary sources to see how many proposed models actually had implementations associated with them. 
-Then, we quantified types of implementations. 
-Whether the implementation was for a real system, for a prototype and/or used in a production environment.
+When designing and proposing a model targeted at a feature that is rooted in practical
+usage by real software systems, bringing the model to life is evidence that the
+proposed model can work in practice. The presence of an actual implementation substantiates the robustness of the design and/or the need for the extension.
+We analyzed the primary sources to see how many
+proposed models actually had implementations associated with them.
+We quantified whether the implementation was enterprise (i.e., an implementation used in a production environment) was for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).
 
-The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
-Further, we identified a set of evaluation types found within  each of the primary sources and provide below a list of the 
-evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
-types.
+Furthermore, the 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
+We identified a set of evaluation types found within  each of the primary sources and provide a list of the evaluation types and which primary sources provided which type.
+In some cases a single primary source provided multiple evaluation types.
 
@@
 Table \ref{tab:implementations} shows the breakdown of implementations found within the primary sources.
-Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were simply prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
+Of the 27 papers surveyed, 4 papers~\cite{zhou2007team,bao08:role, cholewka00:acontext-sensitive,huang06:pervasive} were prototypes developed by the authors whereas 4 papers~\cite{jian2008extended, aich09:role, yao2008task, motta03:contextual} were claimed to be implemented within a real
 system. The remaining 20 papers provide no
-mention of an implementation or prototype.
+mention of an implementation.
 
+Table \ref{tab:evaluations} shows evaluation types.
 Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. 14 models presented example scenarios
-and how application of their model would apply and resolve the situation.  6 models provided some form of performance
+and how application of their model would apply and resolve the situation. Six models provided some form of performance
 or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
 as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-5 models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+Five models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
 The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
@@
 or show how the Reference Model is deficient in tackling said scenarios.
-
-\subsection{Analysis and Discussion}
-
-The RBAC standard is designed with enterprises in mind such that when practitioners implemented RBAC into their systems,
-there would be a reasonable assurance being based off a well thought out model.  As extensions to the RBAC Reference Mode
-come along, thought and time should be given to how features and nuances of their models may impact implementation
-in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
-with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as four papers did, should be
-a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
-implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
 
@@
 
+\subsection{Analysis and Discussion}
+
+The RBAC standard is designed in mind such that when practitioners implemented RBAC into their systems, The RBAC standard demonstrates reasonable assurance being based off a well thought out model.
+As extensions to the RBAC Reference Mode come along, thought and time should be given to how features and nuances of their models may impact implementation to achieve the same goals as the original standard.
+The primary sources showed a lack of implementation
+with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as four papers did, should be
+a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
+implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.
+
 When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
@@
 
-\greybox{Research on RBAC extension models shows a lack of application and testing in the real world scenarios that the models are ultimately designed for. 
-Further, comprehensive evaluation of the models were found to be significantly lacking. }
+\greybox{Research on RBAC extension models shows a lack of implementations in the real world scenarios that the models are ultimately designed for.
+RBAC extension model research lacks comprehensive evaluation of the models both in theory and in practice.}
 
@@
 \cite{chen08:spatio-temporal}
-\cite{aich09:role}
+\citep{aich07:STARBAC}09:role}
 &
@@
 1 \\ \cline{1-3}
+None
+&
+\cite{masoumzadeh2008purbac}
+\cite{chandran05:llt}
+\cite{ray2006lrbac}
+&
+3 \\ \cline{1-3}
 
 \end{tabular*}
-\label{tab:implementations}
+\label{tab:evaluations}
 \end{table}
 
-\section{RQ4: Domains} \label{sec:domains}
+\section{RQ5: Domains} \label{sec:domains}
 
-\textit{What domains have RBAC extensions been created for?}
+\textit{What domains have been targeted by RBAC extension models?}
 \\
@@
 
-\begin{figure}[ht]
-    \centering
-        \includegraphics[width=4.0in]{sections/dist_domains_byYear.png}
-\vspace{-0.2 in}
-    \caption{\label{fig:dist_domains}Domain distribution by 3-year period. Y-axis represents the number of corresponding domain papers.
-    X-axis represents 3-year period from 2000 to 2011.}
-\end{figure}
-
 The medical domain produced the largest selection of papers when analyzing the domains influencing the proposals of RBAC extension models.  
@@
 
-\section{RQ5: Generalizations} \label{sec:generalizations}
+\begin{figure}[ht]
+    \centering
+        \includegraphics[width=4.0in]{sections/dist_domains_byYear.png}
+\vspace{-0.2 in}
+    \caption{\label{fig:dist_domains}Domain distribution by 3-year period. Y-axis represents the number of corresponding domain papers.
+    X-axis represents 3-year period from 2000 to 2011.}
+\end{figure}
 
-\textit{What commonalities or generalizations exist across all categories?}
+\section{RQ6: Generalizations} \label{sec:generalizations}
+
+\textit{What commonalities exist across RBAC extension models?}
 \\
 
-During the paper reading phase, we identified commonalities or generalizations within the primary sources by looking for formal modeling of extending RBAC extension models.
+During the paper reading phase, we identified commonalities within the primary sources
+by looking for formal modeling of extending RBAC extension models.
 

### 2014-02-11 paper/sections/appendix.tex
Adding back analysis section and cleaning up.
Updating counts in appendix and other data tables.

@@
 
-\begin{table}
+\begin{sidewaystable}
+
+\footnotesize
 \centering
@@
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ | p{.95\textwidth} | }{\cite{haibo05:context} A Context-Aware Role-Based Access Control Model for Web Services, 2005 (None / Web / None)} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{haibo05:context} A Context-Aware Role-Based Access Control Model for Web Services, 2005 (None / Web Applications / None)} \\ \cline{2-4}
 
@@
 \multicolumn{1}{ c }{}  &
-\multicolumn{3}{ | p{.95\textwidth} | }{\cite{jian2008extended} An extended RBAC model based on granular logic, 2008 (None / Organizations / Scenarios)} \\ \cline{2-4}
+\multicolumn{3}{ | p{.95\textwidth} | }{\cite{jian2008extended} An extended RBAC model based on granular logic, 2008 (None / None / Scenarios)} \\ \cline{2-4}
 
@@
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{Const. }}}} &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{alam06:constraint} A constraint based role based access control in the SECTET a model-driven approach, 2006 } \\[4ex] \cline{2-4}
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{Const}}}} &
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{alam06:constraint} A constraint based role based access control in the SECTET a model-driven approach, 2006 (None / Medical / Scenarios)} \\[4ex] \cline{2-4}
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org. }}}} &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhang06:collaborative} ROBAC: Scalable Role and Organization Based Access Control Models, 2006 } \\[2.75ex]  \cline{2-4}
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{1}{*}{\rotatebox{90}{ Org }}}} &
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{zhang06:collaborative} ROBAC: Scalable Role and Organization Based Access Control Models, 2006 (None / Enterprise Workflows / Complexity Analysis)} \\[2.75ex]  \cline{2-4}
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{Priv. }}}} &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ni2010privacy} Privacy-aware role-based access control, 2007 } \\[1ex] \cline{3-4}
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{2}{*}{\rotatebox{90}{Priv }}}} &
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{ni2010privacy} Privacy-aware role-based access control, 2007 (None / Medical / Time-based Performance)} \\[1ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{masoumzadeh2008purbac} PuRBAC: Purpose-Aware Role-Based Access Control, 2008 } \\[1ex] \cline{2-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{masoumzadeh2008purbac} PuRBAC: Purpose-Aware Role-Based Access Control, 2008 (None / Web Applications / None)} \\[1ex] \cline{2-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Task }}}} &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{yao2008task} A Task-Role Based Access Control Model with Multi-Constraints, 2008 } \\ \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{yao2008task} A Task-Role Based Access Control Model with Multi-Constraints, 2008 (Enterprise / Organizations / Scenarios)} \\ \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{zhou2007team} Team and Task Based RBAC Access Control Model, 2009 } \\ \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{zhou2007team} Team and Task Based RBAC Access Control Model, 2009 (Prototype / Medical / Scenarios)} \\ \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{oh2003task} Task-role-based access control model, 2003 } \\ \cline{2-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{oh2003task} Task-role-based access control model, 2003 (None / Enterprise Workflows / Scenarios)} \\ \cline{2-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{{\multirow{10}{*}{\rotatebox{90}{ Spatio-Temporal }}}} &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich07:STARBAC} STARBAC: Spatiotemporal role based access control, 2007 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{aich07:STARBAC} STARBAC: Spatiotemporal role based access control, 2007 (None / None / Mathematical Modeling)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chen08:spatio-temporal} On spatio-temporal constraints and inheritance in role-based access control, 2008 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{chen08:spatio-temporal} On spatio-temporal constraints and inheritance in role-based access control, 2008 (None / Pervasive Computing / Mathematical Modeling)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{samuel07:spatio-temporal} A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{samuel07:spatio-temporal} A framework for specification and verification of generalized spatio-temporal role based access control model, 2007 (None / Medical / Scenarios)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{chandran05:llt} LoT-RBAC: A Location and Time-Based RBAC Model, 2005 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{chandran05:llt} LoT-RBAC: A Location and Time-Based RBAC Model, 2005 (None / Mobile / None)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{ray07:spatio} A Spatio-temporal Role-Based Access Control Model, 2007 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{ray07:spatio} A Spatio-temporal Role-Based Access Control Model, 2007 (None / Pervasive / Scenarios, Comparison to Standard)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{}  &
-\multicolumn{2}{ | p{.7\textwidth} | }{\cite{aich09:role} Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 } \\[0.25ex] \cline{3-4}
+\multicolumn{2}{ | p{.9\textwidth} | }{\cite{aich09:role} Role Based Access Control with Spatiotemporal Context for Mobile Applications, 2009 (Enterprise / Medical, Mobile / Complexity Analysis, Time-based performance)} \\[0.25ex] \cline{3-4}
 
 \multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Spatial }}}} &
-\multicolumn{1}{ | p{.6\textwidth} | }{\cite{damiani2007geo} GEO-RBAC: a spatially aware RBAC, 2005 } \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ | p{.9\textwidth} | }{\cite{damiani2007geo} GEO-RBAC: a spatially aware RBAC, 2005 (None / Medical / Mathematical Modeling)} \\[0.25ex] \cline{4-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
-\multicolumn{1}{ | p{.6\textwidth} | }{\cite{ray2006lrbac} LRBAC: A location-aware role-based access control model, 2006 } \\[0.25ex] \cline{4-4}
+\multicolumn{1}{ | p{.9\textwidth} | }{\cite{ray2006lrbac} LRBAC: A location-aware role-based access control model, 2006 (None / Mobile / None)} \\[0.25ex] \cline{4-4}
 
 \multicolumn{1}{ c }{}  & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{}  &
-\multicolumn{1}{ | p{.6\textwidth} | }{\cite{hansen2003spatial} Spatial role-based access control model for wireless networks, 2003 } \\[0.25ex] \cline{3-4}
+\multicolumn{1}{ | p{.9\textwidth} | }{\cite{hansen2003spatial} Spatial role-based access control model for wireless networks, 2003 (None / Medical / Mathematical Modeling)} \\[0.25ex] \cline{3-4}
 
-\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp. }}}} &
-\multicolumn{1}{ | p{.6\textwidth} | }{\cite{joshi05:generalized} A generalized temporal role-based access control model, 2005 } \\[4ex]
+\multicolumn{1}{ c }{} & \multicolumn{1}{ |c }{} & \multicolumn{1}{ |c }{{\multirow{3}{*}{\rotatebox{90}{ Temp }}}} &
+\multicolumn{1}{ | p{.9\textwidth} | }{\cite{joshi05:generalized} A generalized temporal role-based access control model, 2005 (None / Enterprise Workflows / Scenarios)} \\[4ex]
 
@@
 \label{tab:categorization}
-\end{table}
+\end{sidewaystable}

### 2014-02-14 paper/sections/discussion.tex
Address comments given by Dr. Williams

@@
 
-A small percentage of primary sources presented substantial evaluations or implementations of their model in practice.
-This in turn leaves researchers and practitioners little data from which to compare one model to the next.
-And given the wide swath of popular and critical domains being targeted by RBAC extension models, Hu and Scarfone proposed metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
+A small percentage of primary sources presented evaluations or implementations of their model.
+This result leaves researchers and practitioners little data from which to compare one model to the next.
+Given the wide swath of popular and critical domains being targeted by RBAC extension models, Hu and Scarfone proposed metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
 They collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing.
-For example, ``Ease of Privilege Assignments'' metric measures number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC.
-If a model requires more steps for such a given task, there is a potential increase in human or system error due to increased complexity.
+For example, ``Ease of Privilege Assignments'' metric measures the number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC.
 This metric can be applicable for not only the RBAC Reference Model but also RBAC extension models.
@@
 
-In order to adopt a model in practice, software developers should implement the model in real system environments based on intended use of the models.
-Typically, a model provides abstract formal representation of extended RBAC and its operation.
+To adopt a model in practice, software developers would implement the model in real system environments based on intended use of the models.
+Typically, a model provides abstract formal representation of RBAC extension and its operation.
 To configure intended use (i.e., which subjects can access which resources), we define an access control policy, which specifies high-level rules such as which subjects can access to which resources.
@@
 While the concept of the Reference Model is clear and can be applicable to any system, it is challenging to implement the security mechanism of RBAC because the Reference Model is can be interpreted in more than one interpretation due to its complexity.
-In order to bridge gap between RBAC standard and its implementation, NIST published RBAC Implementation and Interoperability standard (RIIS/ ANSI INCITS-459) in 2011.
+To bridge gap between RBAC standard and its implementation, NIST published RBAC Implementation and Interoperability standard (RIIS/ ANSI INCITS-459) in 2011.
 This standard specifies how to implement RBAC system, which is consistent with NIST RBAC standard.
@@
 
-This standard does not provide a specific guideline for implementation of various extended RBAC models.
-Moreover, when software developers implemented extended RBAC models, they require clear security requirements of extended role-based access control, which identifies security objectives, intended environments, and assumed threats.
-Moreover, software developers understand how this extended RBAC overlaps with other access control models when more than two access control models are integrated into a single system.
+This standard does not provide a specific guideline for implementation of various RBAC extension models.
+Moreover, when software developers implemented RBAC extension models, they require clear security requirements of extended role-based access control, which identifies security objectives, intended environments, and assumed threats.
+Moreover, software developers understand how this RBAC extension model overlaps with other access control models when more than two access control models are integrated into a single system.
 A survey~\cite{o20102010} shows that over 50\% of users at organizations with more than 500 employees are given some of their permissions to access resources based on RBAC.
@@
 Therefore, when RBAC is extended and combined with other access control models, software developers often require to provide a formal representation, which can support of proof of the model.
-Then, software developers implement extended models based on NIST standard of RBAC implementation.
-Especially, software developers should understand intended use of extended access control models to meet a new system requirement.
+Then, software developers implement RBAC extension models based on NIST standard of RBAC implementation.
+Especially, software developers understand intended use of RBAC extension models to meet a new system requirement.
 For example, for a spatial RBAC model, software developers incorporate additional spatial constraints, which can be either static or dynamic, in practice into the RBAC model.

### 2014-02-15 paper/sections/categorization.tex
Updates.

@@
 
-During the paper reading phase, we identify classification themes, which are central topic or context of the RBAC extension models. 
-Our classification process yields a set of categories based upon observations during the paper reading phase.
+During the paper reading phase, we identified that a variety of common terms were emerging to describe the RBAC extension models. 
+We developed a process by which to build a classification of the RBAC extension models based upon observations during the paper reading phase.
 For example, the paper ``Privacy-aware role-based access control'' \cite{ni2010privacy} brings in the notion of privacy explicitly within the title of the paper and the name of their model. 
-We classify this paper into the Privacy category. 
-Some papers presented this direct pronouncement of theme of the RBAC extension models, whereas others were less obvious.
+Some papers appeared to present a direct pronouncement of their classification, whereas others were less obvious.
 Thus, we developed a set of guidelines to aide in determining a set of eight categories. 
@@
 By comparison, the paper ``An extended RBAC model based on granular logic'' \cite{jian2008extended} does not contain a direct categorization in the title or model name. 
-However, in reading the body of the paper, we determined that this paper discussed RBAC extension based on context. 
+However, in reading the body of the paper, we determined that the paper discussed RBAC extension based on context. 
 In Section 4.1, we offer definitions for each of the eight observed categories.
@@
 We provide a definition for each observed model category based on data extracted from the primary sources and the English definitions for each noun-phrase.
-Some category descriptors contain abbreviations in parenthesis that match the shortened name found in Table 6 (within Appendix 1) \ref{tab:categorization} which presents each primary source within it's designated category.
+Some category descriptors contain abbreviations in parenthesis that match the shortened name found in Table 6 (within Appendix 1) which presents each primary source within it's designated category.
 
@@
 
-  \item \textbf{Spatio-Temporal}: The extension model combines spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
+  \item \textbf{Spatio-Temporal}: The extension model combines the use of either or both spatial (location-based) and temporal (time-based) constraints in specifying access control policies.
 For example, specific locations permit roles to conduct actions from 8:00 a.m.
@@
 
-\greybox{We found eight categories that exist within the RBAC extension models: Constraint, Context, 
-Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal. We further found that the other
-seven categories fall under the category of Context.}
+\greybox{We found eight categories that exist within the RBAC extension models: Constraint, Context, Organization, Privacy, Task, Spatio-Temporal, Spatial and Temporal.
+The context category was a superset of all other categories since all categories were found to consider context.}

### 2014-02-15 paper/sections/core_rbac.tex
Updates.

@@
 Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
-This lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
+The lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
 Moreover, in cases where organizations were concerned with adopting RBAC,
@@
 Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
-The analysts estimate that RBAC technology has generated \$6.1 billion in net economic benefits to industry where NIST RBAC standard saved \$1.1 billion.
+The analysts estimate that the use of RBAC technology has saved industry \$6.1 billion from 1992 to 2009 and the research and development work by NIST to produce an RBAC standard has saved an additional \$1.1 billion that wold have otherwise been spent across the industry doing research and development on RBAC.
 
@@
 $PA$ represents assignment of permissions to roles.
-
-\begin{figure}[ht]
-    \centering
-        \includegraphics[width=4.0in]{sections/core-model.png}
-    \caption{\label{fig:overview}Diagram in Core RBAC reference model\cite{ferraiolokuhn}.}
-\end{figure}
-
-Figure~\ref{fig:overview} presents an overview of the core RBAC reference model where entities and their relations are described.
-Let "USERS", "ROLES", "OBS", "OPS", "PRM", and "SESSIONS" denote users, roles, objects, operations, permissions, and sessions, respectively.
 Permissions are associated with possible users' pre-defined operation on an object (e.g., execute a file).
@@
 The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. 
-Separation of duty relations enforce conflicts of interest among roles. This model defines two types of separation of duty relations: static and dynamic.
+Separation of duty relations enforce conflicts of interest among roles. 
+The Constrained RBAC model defines two types of separation of duty relations: static and dynamic.
 

### 2014-02-15 paper/sections/process.tex
Updates.

@@
 
-We adopted and applied a systematic literature review process following recommendations from Kitchenham's suggested processes~\cite{kitchenham2007guidelines}.
+We adopted and applied a systematic literature review process following recommendations from Kitchenham and Charter's suggested processes~\cite{kitchenham2007guidelines}. 
 The systematic literature review process was broken down into four stages and the rest of this section is broken down by each stage. 
@@
 Table \ref{tab:search_results} lists the four search engines along the left most column with the three search terms across the top row along with the papers for each criterion and engine combination. 
+The total column represents the total for each individual search engine with a grand total across all search engines. 
+The combined total represents the net paper total after removing duplicate entries. 
 The search algorithm was performed as follows:
@@
 After gathering all 12 data sets, we combined the papers into a master list, which includes only distinct papers by systematically comparing the bibliographic information for each.
-Out of the master list of 1,716 papers, two reviewers, denoted by Reviewer 1 and Reviewer 2, conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
+Out of the master list of 1,716 papers, two reviewers conducted a series of elimination rounds to narrow the list of papers and identify primary sources.
+The two reviewers are denoted by Author 1 and Author 2 where the former is the first author on the paper (Eric) and the latter is the second author on the paper (JeeHyun).
 Table \ref{tab:eliminations} shows the total number of papers selected by each reviewer for each round and how many papers from the disjoint set for each round survived to the next round.
@@
 \centering
-\caption{Paper counts after applying search strategy}
+\caption{Paper counts after applying Step 1}
 \vspace{0.1 in}
@@
 
-The title elimination round resulted in Reviewer 1 selecting 305 papers, and Reviewer 2 selecting 176 papers with 149 papers of overlap between the two. 
-There were 332 papers found not to be in common, of which, 149 were rejected and 141 retained after a second review.
+The title elimination round resulted in Author 1 selecting 305 papers, and Author 2 selecting 176 papers with 149 papers of overlap between the two. 
+All 149 papers found to be in both reviewers lists were kept. 
+The 332 papers found not to be in common were then slated for a second round of review. 
+The second round of review consisted of the reviewers discussing their perception of each paper title as the title related to the elimination criteria and making a joint decision to keep or reject. 
+The second review resulted in 149 rejections and 141 being retained.
 
@@
 
-The second round of elimination was based on reading of the abstracts of papers that survived title elimination.  Researchers read each abstract and evaluated relevancy based off:
+The second round of elimination was based on reading of the abstracts of papers that survived title elimination. 
+The reviewers read each abstract and evaluated relevancy based off:
 
@@
 
-The abstract elimination round resulted in Reviewer 1 selecting 86 papers, and Reviewer 2 selecting 102 papers with 51 papers of overlap between the two. 
-There were 137 papers found not to be in common, of which, 116 were rejected and 21 retained after a second review.
+The abstract elimination round resulted in Author 1 selecting 86 papers, and Author 2 selecting 102 papers with 51 papers of overlap between the two. 
+All 51 papers found to be in both reviewers lists were kept. 
+The 137 papers found not to be in common were then slated for a second round of review. 
+The second round of review consisted of the reviewers discussing their perception of each the paper abstract as each related to the elimination criteria and making a joint decision to keep or reject based on the joint outcome of the discussion.
+The second review resulted in 116 rejections and 21 being retained.
 
@@
 
-The content elimination round resulted in Reviewer 1 selecting 46 papers, and Reviewer 2 selecting 42 papers with 24 papers of overlap between the two. 
+The content elimination round resulted in Author 1 selecting 46 papers, and Author 2 selecting 42 papers with 24 papers of overlap between the two. 
 Between the two reviewers selections, there were 64 papers not in common, of which, 59 were rejected and 5 retained after a second review.
@@
 
-After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions in order to look for trends. 
+After selection of primary sources, the next step was to extract data from each paper that pertained to our research questions to look for trends. 
 The first step was to take the individual data generated from the final elimination round and organize this information around the research questions. 

### 2014-02-23 paper/sections/abstract.tex
Formatting updates.

@@
 \noindent
-{\bf Aim}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding how RBAC extension models are evaluated by providing an assessment of the state of RBAC extension models.
+{\bf Objective}: The goal of our work is to aid practitioners and researchers in choosing an RBAC extension model, and in understanding how RBAC extension models are evaluated by providing an assessment of the state of RBAC extension models.
 \\
@@
 The primary domains that inspired extensions were the medical domain with 9 of the 27 models, enterprise workflows with 5 of 27, and mobile computing with 5 of 27.
-The state of evaluation of RBAC model extensions is inconsistent with 4 papers comparing themselves to the RBAC Reference Model, 14 presenting example scenarios, and 12 lacking any evaluation that we could discern.
 \\
@@
 {\bf Conclusions}: Our literature review shows that all eight of the RBAC extension categories we identified deal with context whereby the privileges provided to a role are environmental depending upon factors such as location or time.
-The magnitude and scope of extensions to the RBAC standard suggests a revised standard may be beneficial.
+RBAC extension model evaluation lacks a consistent set of metrics and evaluation of current models was found to range from providing scenario examples of the model in action to comparison to the RBAC standard.
+The magnitude and scope of extensions to the RBAC standard suggests a revised standard may be beneficial but in the meantime this work can serve as a starting place for researchers and practitioners.
 

### 2014-02-23 paper/sections/analysis.tex
Address comments given by Dr. Williams
Updates.
Formatting updates.

--- 2014-02-11 paper/sections/analysis.tex
@@
 The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
-scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
-authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
+scenarios and how the proposed model can tackle those situations. 
+For five of the primary sources \cite{bao08:role} \cite{zou2009crbac} \cite{zhang06:collaborative} \cite{zhao2008flexible} \cite{ray07:spatio} the model authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
 In Table \ref{tab:evaluations}, we observe that one paper includes a case study to examine how the proposed model works in practice. Discussions
@@
 Moreover, this logic supports for concise and elegant formulation of the Reference Model and its relation. 
-First-Order logic is expressive enough to concisely represent access control systems.
-First-Order logic uses relations, variables, and quantifiers.
+First-Order logic is expressive enough to concisely represent access control systems. First-Order logic uses relations, variables, and quantifiers.

### 2014-02-23 paper/sections/discussion.tex
Formatting updates.

@@
 
-The research questions we identified, presented results for and analyzed provide specific views of portions of the RBAC extension model landscape. 
-By stepping back, and looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
+The research questions we identified, presented results for and analyzed provide a view of the RBAC extension model landscape. 
+By looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
 We break the discussion into two issues: recommendations for the enhancement of the evaluation of RBAC extension models to role-based access control models and guidelines for developers needing to choose an RBAC extension model to meet their requirements.
@@
 
-A small percentage of primary sources presented evaluations or implementations of their model.
+A small percentage of primary sources presented robust evaluations or implementations of their model. 
 This result leaves researchers and practitioners little data from which to compare one model to the next. 
@@
 This metric can be applicable for not only the RBAC Reference Model but also RBAC extension models.
-%Future work could include providing metric measurements for the primary sources presented in this paper. 
-We plan to use these metrics proposed by Hu and Scarfone for any extension models presented in this paper. 
+We propose the metrics introduced by Hu and Scarfone should be applied to the extension models presented in this paper as well as any future models.
 A developer could attempt to apply these metrics to an RBAC extension model not already evaluated when attempting to choose a model. 
-These metrics focus on comparing specific features support of access control models.
+These metrics focus on comparing feature support of access control models.
 
@@
 To adopt a model in practice, software developers would implement the model in real system environments based on intended use of the models. 
-Typically, a model provides abstract formal representation of RBAC extension and its operation.
-To configure intended use (i.e., which subjects can access which resources), we define an access control policy, which specifies high-level rules such as which subjects can access to which resources.
-Given an access control model and policy, security mechanism provides low-level implementation in a system, which controls access of roles.
+As shown by RQ6, all primary sources were found to provide an abstract formal representations of their extension model and the model's operation.
+The presence of an abstract formal representation of the model stands as a starting place for developers to configure intended use (i.e.
+which subjects can access which resources), and define an access control policy (i.e.
+which specifies high-level rules such as which subjects can access to which resources).
 
@@
 To adopt RBAC into a system, organizations often use hybrid approaches, which combine RBAC and access control lists because specific user types, systems, and workflow may not be effective to manage access based on roles.
-Therefore, when RBAC is extended and combined with other access control models, software developers often require to provide a formal representation, which can support of proof of the model.
+Therefore, when RBAC is extended and combined with other access control models, software developers can provide a formal representation which can serve as support for proof of the model. 
 Then, software developers implement RBAC extension models based on NIST standard of RBAC implementation. 

### 2014-02-23 paper/sections/intro.tex
Formatting updates.

@@
 \item RQ2: What are motivations behind the RBAC extension models?
-\item RQ3: How are RBAC extension models implemented and evaluated by their authors?
-\item RQ4: What domains have been targeted by RBAC extension models?
-\item RQ5: What commonalities exist across RBAC extension models?
+\item RQ3: How are RBAC extension models implemented by their authors?
+\item RQ4: How are RBAC extension models evaluated by their authors?
+\item RQ5: What domains have been targeted by RBAC extension models?
+\item RQ6: What commonalities exist across RBAC extension models?
 \end{itemize}
@@
 \setlength{\itemsep}{0.25pt}
-\item A starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
-the current state of the RBAC extension models. 
+\item A starting place for researchers in the realm of authorization and as a reference guide for discovering the current state of the RBAC extension models. 
 \item A basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs.

### 2014-04-15 paper/reviewer_responses.tex
Adding reviewer responses.

Reviewer #1:

The paper aims to present a systematic literature review (SLR) on RBAC
extensions.  However, a couple of serious issues raise doubts if the
literature review was actually "systematic". In particular, the
process described in the paper is intransparent and hides important
information that is needed to actually assess the paper.

Section 3 ("Methodology and Process") does not describe a systematic
literature review process, as claimed by the authors.  Instead, this
section merely gives an overview of the search and screening part of
the SLR.  However, even this part is roughly sketched only and omits
significant information. For example, it is never explained why
certain search engines resp. certain digital libraries were included
in the review while other very important libraries are ignored. Why
were Springer (SpringerLink) and Elsevier (ScienceDirect), two of the
most important publishers of scientific research, excluded from the
review ?

Moreover, Section 3 does neither reveal the search terms that were
used in the review, nor does it describe the actual search procedure.
Are the 3 terms from Table 1 the final search terms ? If so, these
terms are way too generic. Why didn't the authors focus the search on
their research questions (defined in Section 1) ?

Furthermore, every search engine works differently - how did the
search strings look like ?  Were the search terms applied to paper
titles and abstracts only or did the authors conduct a full-text
search ?  Which type of papers did the authors target (research
papers, opinion papers, white papers, conference papers, journal
papers)? Was the search restricted to a certain publication period ?

In steps 2-4 of the search and screening process (Section 3), the
paper reports on pairs of authors serving as raters on publications
(elimination steps). The paper also refers to (substantial)
negotiation rounds to establish agreement on whether to
include/exclude a publication item. However, the paper does not report
on the levels of agreement/disagreement between the raters over all
items. The elimination quantities from Table 2 are too coarse
grained. However, such IRR analyses are recommended by Kitchenham's
guidelines (allegedly adopted by the authors).

In addition, the stopping criteria are strange . Why was the search
stopped if the result set was greater than 1000 ? How could you
guarantee that the most relevant papers are amongst the first 1000
publications found in the search and not among the papers that were
simply ignored? The paper says "the result set size stopping criteria
was chosen due to a technical limitation of some search
engines". However, the alleged "technical limitations" are never
discussed in more detail. Everybody is well aware of certain "quota"
for searches per hour or searches per day that are enforced by
different scientific search engines. This, however, is certainly no
valid reason for truncating a systematic (!) literature review.

Moreover, stopping "if the last ten results did not contain the search
term phrase within the title" is certainly not the best option: a
potentially large number of relevant publications will not be found.
An alternative would be to establish a quasi-gold standard and search
until a threshold is reached.

Section 3.2.1.: Why was it a requirement that the paper titles must
include certain terms, such as "model"?  I fear that this requirement
is excessively limiting and potentially excluding relevant papers.
Moreover, why is a reference to the term "model" important in the
first place? This is never motivated, no rationale for this
requirement is provided to the reader.

Section 3.2.3.: The use of "conditional inclusion criteria" is not
clear to me (given that they are also not covered by the SLR
guidelines allegedly adopted by the authors). Moreover, the statement
that "without this conditional include, the paper list size" would
have been too "small" indicates that the authors deviated from the
initial procedure (protocol) without justifying or reporting this
deviation.  Arbitrarily changing the inclusion/exclusion criteria to
expand the paper base is not a systematic approach.  With respect to
the content elimination criteria: what does "inspired by real world
example" mean exactly ?

The search strategy results in a relatively small corpus of 27 papers,
searching the papers' references via snowballing could help here.

Moreover, the paper includes a number of inconsistencies that are
never explained. For example, the data from Table 3 and Table 4 are
inconsistent -- in Table 4, papers [21] and [23] are missing while [36]
is newly introduced.

Section 4/RQ1: It is not clear how the authors established the common
terms and how they systematically tagged the included papers in terms
of a content analysis. Also, the wording indicates that all of this
happened during paper reading (ad hoc), rather than as a preparatory
step.

Section 5/RQ2: I do not see how the answers to this RQ are related to
the SLR. The section rather reads like a discussion based on the
authors' intuitions or conjectures.

Section 7/RQ4: Why do the authors only distinguish between
enterprise/prototype "implementations" of RBAC approaches? Wouldn't
it be more interesting to distinguish different "implementation
techniques"?  Moreover, often an implementation alone doesn't say much
about the actual applicability of a certain approach. What about other
evaluation techniques such as case studies, static and dynamic
analysis, formal proofs, or controlled experiments for example ?

Section 9/RQ6: The authors' suggestion of using "propositional
logic" is a pure discussion item and not even remotely backed by the
data extracted from the SLR.

In chapter 10.1. (Discussion) the paper proposes the use of metrics
developed by Hu & Scarfone when choosing an RBAC extension model. Why
this metric in particular? Did the authors perform a prior comparative
analysis of different metrics?

From the results presented in the paper, i have major doubts that the
search terms were adequate and that the search was conducted
thoroughly.  For example, the authors did not find a single
contribution published in the proceedings of the the premier access
control conference -- the ACM Symposium on Access Control Models and
Technologies (SACMAT) which was established in 2001 as a successor of
the ACM RBAC workshop. Moreover, although the paper references some
papers published in ACM TISSEC, many other papers from ACM TISSEC that
do also present extensions to the RBAC reference model were not found
in the SLR.

In its current state, the paper has to be classified as work in
progress. It does neither increase our knowledge of RBAC nor of RBAC
extensions.  The literature review reported in this paper may serve as
a pre-study for an actual SLR. However, the paper needs a lot of work
before it can be accepted for publication in a journal. In particular,
the paper is intransparent and includes a number of serious
methodological flaws.

Minor:

- Section 2 should be shortened to at most one page

- The definitions of static and dynamic separation of duty in Section
2.3 are not correct.

- why are the paper references from Tables 3, 4, and 5 sorted randomly
and not in an ascending order ?

- Table 5 includes a broken literature reference (probably to [29])

- additional figures/visualizations would help to improve the paper

- a number of entries in the References section are incomplete

- the paper includes a number of spelling errors, grammatical errors,
and punctuation flaws.

- Table 6 (Appendix A) should be revised, esp. the structure needs to
be improved

Reviewer #2:

The topic analysed in this paper is interesting and the literature review has been carried out by following a formal approach.

Nevertheles, the authors should justify some decisions made in the systematic review.

Their search strategy should be justified.
        - Why have they selected that search chain?
        - What is the criteria for selecting those sources (Google Shoolar, etc.) and no other?
        - Why were they using as a title elimination criteria that title contains a reference to 'model' and not to "policy", etc. ?

On the other hand, the authors have defined the classification criteria by observation when they were reading the papers. And also, they have defined the comparison items by themself. I consider that these criteria should be defined based on existing frameworks for comparing access control policies and if they do not exist, justify why they propose these criteria.

### 2014-04-19 paper/review/breakdown.tex
Adding bulleted breakdown of reviewer comments and strategies to tackle the methodological flaws.

\pagebreak

\section{General Breakdown}

\begin{itemize}

\item Broken literature references in Table 5

\item Tables 3,4,5 are not sorted in any order

\item Additional figures/visualizations would help

\item Reference section has incomplete entries

\item Spelling, grammar and punctuation flaws

\item Table 6 should be revised to have better structure

\end{itemize}

\section{Background Breakdown}

\begin{itemize}

\item This section should be shortened to at most 1 page.

\item Definitions of static and dynamic separation of duty are incorrect.

\end{itemize}

\section{Methodology Breakdown}

\begin{itemize}

\item Calls into question the "systematic" nature of the paper.

We can address this by adjusting our search method to follow an accepted systematic approach as noted in the SLR literature. Notably the Zhang \cite{zhang2011identifying} approach mentioned in \cite{kitchenham2013systematic}. 
In this approach we do a two phase manual + automatic search. 
The manual phase identifies premier journals and conferences specifically related to our research questions and pulls out the "quasi-gold standard" set of papers.
We derive our search term from this set of papers and use them as an oracle for the precision and sensitivity of our automatic search

\item Why the search we chose? Why not others like Springer? Elsevier?

Looking at the recent work by Kitchenham to review SLRs from 2013 \cite{kitchenham2013systematic}, the literature justifies the use of:

  \begin{itemize}
    \item ACM Digital Library (already included)
    \item IEEE Explore (already included)
  \end{itemize}

Two of the following indexes:
  \begin{itemize}
    \item SCOPUS
    \item EI Compndix
    \item Web of Science
    \item Google Scholar (already included)
  \end{itemize}

\item Where are the search terms?

We will explicitly enumerate and mention the search term.

\item Search process not described

This will be beefed up with added information from the adjusted search strategy derived from the Zhang method.

\item Search terms too generic. Why isn't search focused on research questions?

Zhang \cite{zhang2011identifying} proposes (based on empirical testing) two methods for selecting search terms.

Subjective - researchers derive search terms based on expert knowledge and the results of the "quasi-gold standard (QGS)" paper set or the use of Kitchenhams suggested population, intervention, comparison, outcomes, context, study designs
Objective - researchers use a textual analysis tool aimed at the QGS paper set to derive most commonly used words and word relationships to then build the search term

\item No mention of using different syntax for different search engines based on how different engines parse search strings.

We will note any differences in how the search term is applied for different search engines.

\item Type of search: full-text? title? abstract?

We can address this by only specifically mentioning it for the manual search process since the automatic search will use common search engines.

\item Whyat type of papers did the authors target? research? white paper? journal? conference?

We can address this by only specifically mentioning it for the manual search process since the automatic search will use common search engines.

\item Was search restricted to a publication period?

We can address this by only specifically mentioning it for the manual search process since the automatic search will use common search engines.

\item Reports of reviewers agreement/disagreement too coarse. IRR analyses recommended by Kitchenham.

We'll calculate and use the Cohen Kappa value for this.

\item Reasoning for stopping criteria never explained, especially technical limitations and what they are.

This one may be trickier as a lot of SLRs don't require a stopping criteria due to how the search term is constructed.
This item may take care of itself with the adjustments to the process and if not we'll specify the technical limitations for any search engine as the reason for why we "stopped".

\item Why not use quasi-gold standard for search threshold?

Zhang method will achieve this.

\item No motivation for why "model" is important as a search term.

Choice of subjective or objective will handle this justification.

\item "conditional inclusion criteria" not allowed by SLR process

We can pair down our inclusion/exclusion criteria and move some of the items into the data collection aspect.
I noted that a number of SLRs use only a few inclusion/exclusion criteria.

\item Corpus too small

We can add a threat to validity if the corpus is too small and note that the search process followed a systematic approach.

\item Question about changing the inclusion/exclusion criteria as not being systematic

We can justify this more in the text as Kitchenham even mentions being able to iterate some elements as the search proceeds.

\item No jutification for deviating from initial procedure

Changing our methodology to Zhang should take care of this.

\item "inspired by real world example" - what does this mean?

\item Table 3/4 inconsistent

\item No papers from joournals/conferences related to access control such as SACMAT, missing papers from ACM TISSEC

We can include these journals as part of the manual QGS approach prior to the automatic search.

\item Why the search chain?

Handled by Zhang.

\item Criteria for selecting sources?

Handled by Zhang method manual and automatic engines listed from Kitchenham.

\item Why "model" in the title is important and not something like "policy"?

This can be handled by the derivation of search terms from the subjective/objective method.

\end{itemize}

\section{RQ1 Breakdown}

\begin{itemize}

\item Not clear how the authors established the common terms.

\item How did the authors systematically tag the included papers in terms of a content analysis?

\item Was this done ad hoc (during paper reading) or as a preparatory step?

\item Authors defined the classification criteria by observation during paper reading.

\item The authors defined comparison items by themselves. Are there existing frameworks for comparing access control policies?

\end{itemize}

\section{RQ2 Breakdown}

\begin{itemize}

\item How do the answers to this RQ relate to the SLR?

\item This section reads like a discussion based on authors intuitions or conjectures

\end{itemize}

\section{RQ4 Breakdown}

\begin{itemize}

\item Why do the authors only distinguish between enterprise/prototype implementations?

\item What about distinguishing different "implementation techniques" ?

\item What about other evaluation techniques such as case studies, static and dynamic analysis, formal proofs or controller experiments for example?

\end{itemize}

\section{RQ6 Breakdown}

\begin{itemize}

\item Authors suggestion of "propositional logic" is purely discussion and not backed by any data from the SLR

\end{itemize}

\section{Discussion Breakdown}

\begin{itemize}

\item Why the "Hu \& Scarfone" metrics?

\item Did the authors perform a prior comparative analysis of different metrics?

\end{itemize}

### 2014-04-20 paper/review/breakdown.tex
Adding list of manual search journals and conferences.

@@
     \item Google Scholar (already included)
+  \end{itemize}
+
+Journals/Conferences of Access Control:
+
+  \begin{itemize}
+    \item ACM Transactions on Information and System Security (TISSEC)
+    \item International Conference on Emerging Security Information, Systems and Technologies, SECURWARE
+    \item International Conference on Information Systems Security
+    \item International Conference on Network and System Security
+    \item ACM Symposium on Access Control Models and Technologies (SACMAT)
+    \item ACM Symposium on Information, Computer and Communications Security
+    \item Information Security Conference
+    \item International Conference on Information Security and Assurance
+    \item International Conference on Information Systems Security
+    \item International Conference on Information Assurance and Security
   \end{itemize}

### 2014-04-22 paper/review/breakdown.tex
My comments on review comments (breakdown) Addressed some of comments.

@@
 \item Broken literature references in Table 5
+We address this issue.
 
 \item Tables 3,4,5 are not sorted in any order
+We manually sort by the order of the paper.
 
 \item Additional figures/visualizations would help
+[TBD] Eric and I would discuss this issue because the reviewer did not clearly which kinds of figures/visualizations would be needed.
 
 \item Reference section has incomplete entries
+We address this comment by looking at the each reference item and find missing entry. We check each entry with ACM or IEEE Bibtex. Consistent formatting will be done as well.
 
 \item Spelling, grammar and punctuation flaws
+We address this by spell check and grammar check (in Latex compiler).
+In addition, we convert it to MS word for additional spell check and grammar check.
+After this step, we will conduct manual proof reading.
 
 \item Table 6 should be revised to have better structure
+[TBD]  Eric and I would discuss this issue because the reviewer did not clearly mention what's the issue or how to revise Table 6.
 
@@
 \item This section should be shortened to at most 1 page.
+We address this comment by revising the background section.
 
 \item Definitions of static and dynamic separation of duty are incorrect.
+We find reference of static and dynamic separation of duty from NIST RBAC standard. We refer to "The NIST Model for Role-Based Access Control: Towards A Unified Standard."
 
@@
     \item International Conference on Information Assurance and Security
+    \item International Symposium on Policies for Distributed Systems and Networks
+    \item ACM Conference on Data and Application Security and Privacy
   \end{itemize}
@@
 \item Not clear how the authors established the common terms.
+We find the terms by bottom-up approach by looking at the paper and finding common terms that can characterize the paper.
 
@@
 \item Was this done ad hoc (during paper reading) or as a preparatory step?
+We address this issue by describing our approach.
+We do it by bottom-up search to find terms and construct it.
+Need to find such a method for classification (I don't think that our approach is ad-hoc if we find it during paper reading).
 
 \item Authors defined the classification criteria by observation during paper reading.
+Same as above.
 
 \item The authors defined comparison items by themselves. Are there existing frameworks for comparing access control policies?
+
+In our SLR, we measure not only access control policies, but also measure approaches of papers.
+There are some such as "Hu \& Scarfone" metrics to measure access control policies themselves.
+We may find some papers that describe metrics.
+However, their metrics are limited to measure a specific property (such as complexity) of a given access control policy.
 
@@
 \item How do the answers to this RQ relate to the SLR?
+We address this issue by checking SLR process to understand what kinds of answers would be expected. Then, we adjust our answers to fit to SLR.
 
 \item This section reads like a discussion based on authors intuitions or conjectures
+We would move this section to discussion section if this one may not fit well to the RQs for SLR.
 
@@
 \item Why do the authors only distinguish between enterprise/prototype implementations?
+In the development process, modeling -> research prototype -> implementations in the real world systems.
+Our selected papers dealt with modeling of access control policies. Then, we would like to check whether this model is provided with a prototype and implementation.
 
@@
 \item What about other evaluation techniques such as case studies, static and dynamic analysis, formal proofs or controller experiments for example?
+We find from bottom-up approach.
 
@@
 \item Authors suggestion of "propositional logic" is purely discussion and not backed by any data from the SLR
+We find reference paper about this propositional logic, which is used for papers collected for us.
 
@@
 \item Why the "Hu \& Scarfone" metrics?
+We use this metric to show additional metrics.
 
 \item Did the authors perform a prior comparative analysis of different metrics?
+[TBD]
 

### 2014-04-22 paper/sections/analysis.tex
My comments on review comments (breakdown) Addressed some of comments.

@@
 &
+\cite{ni2010privacy}
+\cite{jian2008extended}
 \cite{alam06:constraint}
@@
 \cite{yamazaki04:designing}
-\cite{jian2008extended}
 \cite{thein2011leveraging}
@@
 \cite{zhang06:collaborative}
-\cite{ni2010privacy}
 \cite{masoumzadeh2008purbac}
@@
 \cite{zhang06:collaborative}
+\cite{aich07:STARBAC}
 \cite{chen08:spatio-temporal}
-\cite{aich07:STARBAC}
 &
@@
 \cite{zhang06:collaborative}
+\cite{ray07:spatio}
 \cite{zhao2008flexible}
-\cite{ray07:spatio}
 &
@@
 &
-\cite{alam06:constraint}
-\cite{tzelepi01:flexible}
+\cite{jian2008extended}
+\cite{yao2008task}
 \cite{cholewka00:acontext-sensitive}
@@
 \cite{bao08:role}
-\cite{jian2008extended}
+\cite{zhou2007team} 
+\cite{alam06:constraint}
+\cite{tzelepi01:flexible}
 \cite{yamazaki04:designing}
 \cite{zou2009crbac}
+\cite{samuel07:spatio-temporal}
 \cite{ray07:spatio}
-\cite{samuel07:spatio-temporal}
 \cite{joshi05:generalized}
-\cite{yao2008task}
-\cite{zhou2007team}
 \cite{oh2003task}
@@
 \cite{masoumzadeh2008purbac}
+\cite{ray2006lrbac}
 \cite{chandran05:llt}
-\cite{ray2006lrbac}
 &
@@
 &
+\cite{ni2010privacy}
+\cite{motta03:contextual}
+\cite{aich09:role}
+\cite{zhou2007team}
 \cite{alam06:constraint}
 \cite{tzelepi01:flexible}
-\cite{motta03:contextual}
-\cite{ni2010privacy}
 \cite{damiani2007geo}
@@
 \cite{samuel07:spatio-temporal}
-\cite{aich09:role}
-\cite{zhou2007team}
+
 &
@@
 \cite{huang06:pervasive}
+\cite{ray07:spatio}
 \cite{chen08:spatio-temporal}
-\cite{ray07:spatio}
 &
@@
 &
+\cite{haibo05:context} 
 \cite{masoumzadeh2008purbac}
-\cite{haibo05:context}
 &
@@
 &
+\cite{aich09:role} 
 \cite{thein2011leveraging}
 \cite{zou2009crbac}
+\cite{ray2006lrbac}
 \cite{chandran05:llt}
-\cite{aich09:role}
-\cite{ray2006lrbac}
 &
@@
 &
+\cite{yao2008task} 
 \cite{yamazaki04:designing}
-\cite{yao2008task}
 &
@@
 \cite{jian2008extended}
-\cite{aich2007:STARBAC}
+\cite{aich07:STARBAC}
 &

### 2014-06-17 paper/sections/analysis.tex
Revised version
 - Address comments in background, RQ2-RQ5, and discussion section.
 - In addition, I sort paper list in an increasing order (e.g., [4][5][10] instead of [10][4][5]) in the table.

@@
-\section{RQ2: Motivations} \label{sec:motivations}
-
-\textit{What are motivations behind the RBAC extension models?}
-\\
-
-During the paper reading phase, we identify motivations of the RBAC extension models by category. 
-We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues. 
-For example, the papers in the Spatial Category argue that the RBAC reference model does not support spatial constraints, which are used in determining users' access decisions based on spatial dimensions (i.e., location).
-To address this issue, the authors developed the RBAC extension models by incorporating additional spatial entities and their relations with existing entities of the RBAC Reference Model.
-
-\subsection{Results}
-
-We describe motivations behind the RBAC extension models by category as follows:
-
-\begin{itemize}
-
-  \item \textbf{Context}: The RBAC Reference Model does not support the notion of context constraints related to changes in environments.
-Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
-
-  \item \textbf{Constraint}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.
-
-  \item \textbf{Organizational}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations.
-The model needs to reduce the administrative complexity of RBAC across multiple organizations.
-
-  \item \textbf{Privacy}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.
-
-  \item \textbf{Task}: The RBAC extension model does not support the notion of a task, team, purpose, and organizational roles, which help specify a business activity in enterprises.
-
-  \item \textbf{Spatio-Temporal}: The RBAC extension model does not support spatial (location-based) and temporal (time-based) constraints, which specify role-assignment, role-activation and permissions based on location and time.
-
-  \item \textbf{Spatial}: The RBAC extension model does not support the notion of spatial (location-based)  constraints, which specify role-assignment and permissions based on location.
-
-  \item \textbf{Temporal}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.
-
-\end{itemize}
-
-\subsection{Analysis and Discussion}
-
-The RBAC standard provides the RBAC Reference Model, which is an abstraction on top of authorization based on $UA$ and $PA$.
-RBAC Reference Model presents four entities: roles, sessions, users, and permissions, and their relations.
-While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitations on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
-For example, in cases where the RBAC extension model does not support the notion of specific constraints, the RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
-The RBAC extension model provides general agreement on the definition for specific features.
-The RBAC extension model helps simplify theoretical modeling and practical implementation of features in the model.
-
-%The Const category claims that the RBAC extension model has limitations on its features.
-The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model.
-For models in this category, changes occur within the RBAC Reference Model.
-%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles.
-Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations.
-For this combination, new associations between attributes of two models should be introduced.
-Models in this category require new associations between attributes of different access control models to work together.
-%
-%
-%The Context, Org, Priv, Spatio-Temporal, Spatial, and Temp categories claim that the RBAC extension model does not support the notion of specific constraints.
-The RBAC extension model in these categories incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
-The extension model does not change the RBAC Reference Model.
-%
-%The Const category claims that the RBAC extension model has limitations on its features.
-The RBAC extension model in this category incorporates modification of existing features in the RBAC Reference Model.
-For models in this category, changes occur within the RBAC Reference Model.
-%
-%The Task category claims that the RBAC extension model does not support the notion of task, team, purpose, and organizational roles.
-Because task-based access control has been developed, the RBAC extension model in this category is a combination of the RBAC Reference Model and another access control model, which uses model-specific attributes and their relations.
-For this combination, new associations between attributes of two models should be introduced.
-Models in this category require new associations between attributes of different access control models to work together.
-
-\greybox{Motivations for RBAC extension models vary across category but are rooted in adding new entities or relationships to allow authorization flexibility in meeting the demands of emerging requirements.}
-
-\section{RQ3: Implementations} \label{sec:implementations}
+\section{RQ2: Implementations} \label{sec:implementations}
 
@@
 proposed models actually had implementations associated with them.
-We quantified whether the implementation was enterprise (i.e., an implementation used in a production environment) was for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).
+We quantified whether the implementation was an enterprise implementation (i.e., an implementation used in a production environment for a real system) or a prototype implementation (i.e., proof of concept to demonstrate the feasibility).
+In scientific research software life cycle, NuNamaker et al.~\cite{nunamaker1990systems} illustrated three implementation stages: prototype, enterprise (i.e., product), and technology transfer.
+Researchers often conduct their research by building a prototype implementation.
+A prototype implementation serves as a proof of concept to verify that certain concepts of proposed RBAC extension models.
+The prototype implementation demonstrates feasibility.
+However, this prototype implementation does not represent a deliverable real-world application.
+Majority of implementations in RBAC research stops with prototyping.
+After prototypes are successfully demonstrated, researchers make an effort to further develop a prototype implementation into an enterprise implementation.
+An enterprise implementation is a deliverable real-world application and allows to conduct realistic evaluations of the RBAC extension models.
+Technology transfer means that the technology of the proposed RBAC extension models are eventually transferred to the public marketplace.
+We found that none of implementations of the primary sources reach to the stage of the transfer of technology.
 
@@
 
-\section{RQ4: Evaluations} \label{sec:evaluations}
+\section{RQ3: Evaluations} \label{sec:evaluations}
 
@@
 The 27 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors. 
-We identified a set of evaluation types found within  each of the primary sources and provide a list of the evaluation types and which primary sources provided which type.
-A primary source may use multiple evaluation types.
+Zelkowitz and Wllace~\cite{Zelkowitz1998experimentalmodels} proposed evaluation types and methods that are used to validate the claims in the paper. 
+Table~\ref{tab:evaluationtypes} illustrates 12 evaluation methods with their corresponding general evaluation types and descriptions.
+We classified each paper according to the evaluation methods shown in Table~\ref{tab:evaluationtypes}.
 For example, Aich et al.
-conducted performance and complexity analyses to show the effectiveness of their proposed RBAC extension model \cite{aich09:role}.
+conducted project monitoring to show the performance and complexity of their proposed RBAC extension model \cite{aich09:role}.
 
@@
 
-Table \ref{tab:evaluations} shows evaluation types.
-Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation. Fourteen models presented example scenarios
-and how application of their model would apply and resolve the situation. Six models provided some form of performance
-or complexity analysis of their model.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
+Table~\ref{tab:evaluations} shows evaluation methods and criteria.
+Based on the diverse evaluation methods, 12 models presented no evidence of an evaluation. Fourteen models presented 
+example scenarios and how application of their model would apply and resolve the situation. These methods refer to assertions.
+Six models provided some form of performance
+or complexity analysis of their model through project monitoring.  The performance and complexity analysis included graphs of the model's time to determine calculate authorization 
 as the number of entities grew, and the size of the role space for the extension model compared to the RBAC Reference Model. 
-Five models provided mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
-The most widely used evaluation approach was providing sample scenarios with accompanying workflows of how the extension model
+Five models provided simulations based on mathematical descriptions and analysis as a way to provide evaluation in the form of completeness. 
+The most widely used evaluation method was assertion, which provides sample scenarios with accompanying workflows of how the extension model
 would tackle those scenarios.
@@
 
+% Table generated by Excel2LaTeX from sheet 'Sheet1'
+\begin{table}[t]
+  \centering
+  \caption{Evaluation Types and methods classified by Zelkowitz and Wllace~\cite{Zelkowitz1998experimentalmodels}}
+  %  \begin{tabular}{|l|l|l|}
+   \begin{tabular*}{.9\linewidth}{| p{.25\linewidth}| p{.2\linewidth} | p{.35\linewidth} |}
+		\hline
+
+    \textbf{Evaluation Type} & \textbf{Method} & \textbf{Description} \\\hline
+
+    \multirow{4}[8]{*}{Observational methods} & Project monitoring & Monitor project and collect development data \\\cline{2-3}
+          & Case study & Monitor project in depth and collect data with a specific goal for the project \\\cline{2-3}
+          & Assertion & This evaluation uses examples for validation\\\cline{2-3}
+          & Field study & Monitor multiple projects and collect relevant information to compare the projects \\\hline
+    \multirow{3}[6]{*}{Historical methods} & Literature research & Examine previously published projects \\\cline{2-3}
+          & Legacy & Examine data from completed projects \\\cline{2-3}
+          & Lessons learned & Examine qualitative data from projects \\\hline
+    \multirow{5}[10]{*}{Controlled methods} & Static analysis & Examine structure of dveloped product \\\cline{2-3}
+          & Replicated experiment & Develop mulitple versions of product \\\cline{2-3}
+          & Synthetic environment & Replicate one factor in laboratory setting \\\cline{2-3}
+          & Dynamic analysis & Execute developed product for perormance \\\cline{2-3}
+          & Simulation & Execute product using a model of real enviroment \\\hline
+
+    \end{tabular*}%
+  \label{tab:evaluationtypes}%
+\end{table}%
+
 \begin{table}
 \centering
-\caption{Evaluation types by primary source}
+\caption{Evaluation methods by primary source}
 \vspace{0.1 in}
-\begin{tabular*}{.9\linewidth}{| p{.45\linewidth} | p{.3\linewidth} | r | }
-\cline{1-3}
-
-\textbf{Evaluation Type} & \textbf{Papers} & \textbf{Count} \\ \cline{1-3}
+\begin{tabular*}{.9\linewidth}{| p{.2\linewidth}| p{.35\linewidth} | p{.2\linewidth} | r | }
+\cline{1-4}
+\textbf{Methods} & \textbf{Criteria} & \textbf{Papers} & \textbf{Count} \\ \cline{1-4}
+\multirow{2}[0]{*}{Project Monitoring}
+&
 Time-Based Performance
@@
 &
-2 \\ \cline{1-3}
+2 \\ \cline{2-4}
+
+&
 Complexity Analysis
@@
 &
-4 \\ \cline{1-3}
-Comparison to Standard RBAC
+4 \\ 
+\cline{1-4}
+Case Study
 &
-\cite{bao08:role}
-\cite{zou2009crbac}
-\cite{zhang06:collaborative}
-\cite{ray07:spatio}
-\cite{zhao2008flexible}
+RBAC Model in Practice
 &
-5 \\ \cline{1-3}
-Mathematical Modeling
+\cite{motta03:contextual}
 &
-\cite{damiani2007geo}
-\cite{hansen2003spatial}
-\cite{aich07:STARBAC}
-\cite{chen08:spatio-temporal}
-\cite{joshi05:generalized}
+1 \\ \cline{1-4}
+Assertion
 &
-5 \\ \cline{1-3}
 Example Scenarios of the Model in Action
@@
 &
-14 \\ \cline{1-3}
-Case Study of the Model in Practice
+14 \\ \cline{1-4}
+Field Study
 &
-\cite{motta03:contextual}
+Comparison to Standard RBAC
 &  
-1 \\ \cline{1-3}
-None
+\cite{bao08:role}
+\cite{zou2009crbac}
+\cite{zhang06:collaborative}
+\cite{ray07:spatio}
+\cite{zhao2008flexible}
+&
+5 \\ \cline{1-4}
+Simulation
+&
+Mathematical Modeling
+&  
+\cite{damiani2007geo}
+\cite{hansen2003spatial}
+\cite{aich07:STARBAC}
+\cite{chen08:spatio-temporal}
+\cite{joshi05:generalized}
+&
+5 \\ \cline{1-4}
+No Evaluations
+&
+Not Applicable
 &
@@
 &
-3 \\ \cline{1-3}
+3 \\ \cline{1-4}
 
@@
 scenarios and how the proposed model can tackle those situations. 
-For five of the primary sources \cite{bao08:role} \cite{zou2009crbac} \cite{zhang06:collaborative} \cite{zhao2008flexible} \cite{ray07:spatio} the model authors provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
+For five of the primary sources \cite{bao08:role} \cite{zou2009crbac} \cite{zhang06:collaborative} \cite{ray07:spatio} \cite{zhao2008flexible}, the model authors conduct field study to provide some discussion of how the RBAC Reference Model is deficient in tackling the scenario.
 In Table \ref{tab:evaluations}, we observe that one paper includes a case study to examine how the proposed model works in practice. Discussions
@@
 Two of the primary sources examined provided time-based performance analysis of their extension model compared
-to the RBAC Reference Model. This inclusion of time analysis provides some assurances to developers that any non-functional
+to the RBAC Reference Model through project monitoring. This inclusion of time analysis provides some assurances to developers that any non-functional
 requirements surrounding time to compute authorizations compete or beat the RBAC Reference Model. Further, four of the models provided
-some form of complexity of their model. This complexity analysis plays a key role in the management of the access control mechanism
+some form of complexity of their model through project monitoring. This complexity analysis plays a key role in the management of the access control mechanism
 over the course of the models implementation lifetime. As the number of roles, users and additional entities grows, developers will 
@@
 
-\section{RQ5: Domains} \label{sec:domains}
+\section{RQ4: Domains} \label{sec:domains}
 
@@
 
-\section{RQ6: Generalizations} \label{sec:generalizations}
+\section{RQ5: Generalizations} \label{sec:generalizations}
 
@@
 To reduce efforts for modeling access control used in various applications, researchers often focus on developing generalized core concepts of access control.
-The authors use propositional logic to describe access control model across all categorizations.
-Propositional logic is concerned with propositions and their logical relationships.
-In propositional logic, simple (i.e., atomic) or compound condition at given context is evaluated to true or false based on specified rules and access control logic.
-Researchers and practitioners are concerned to extend limited set of propositions specific to core RBAC to meet real-world scenarios such as dynamic constraints, temporal, or spatial constraints.
+The authors use formal model to provide high level description of access control model across all categorizations.
+Formal model of access control provides precise understanding, specification, and analsysi of access control.
+Formal model~\cite{abadi2008variations} of access control can be specified in one of logic expressions such as propositional logic.
+For example, in propositional logic, simple (i.e., atomic) or compound condition at given context is evaluated to true or false based on specified rules and access control logic.
+Aich et al.~\cite{aich09:role} proposed STARBAC model,  which can be expressed in propositional logic.
+While propositional logic does not support for quantifiers, first-order logic (i.e., predicate logic) extends propositional logic by the use of quantifiers.
+For example, ESTARBAC model~\cite{aich09:role} can be expressed in first-order logic.
 
@@
 
-Since NIST proposed RBAC standard using propositional logic, researchers describe extended RBAC models using propositional logic.
+We found that 27 papers of RBAC extensions provide formal model based on logic expressions to describe its extended model.
+%Since NIST proposed RBAC standard using logic expression, we observed that researchers describe extended RBAC models using logic expressions.
+We discuss two widely used logic expressions for access control: propositional logic and first-order logic. 
 Given context, propositional logic is used to evaluate true or false based on specified rules and access control logic.
 As access control is typically evaluated to either true or false based on predicates, propositional logic is sufficient to describe key ideas and definitions.
-Propositional logic specifies what combination of attributes values a request must satisfy to be authorized to roles.
-Combination of attribute values may handle constraints such as temporal and spatial constraints.
+Propositional logic is suitable for specifying what combination of attributes values a request must satisfy to be authorized to roles.
+However, propositional logic has limitations. While this logic is simple, this logic does not support for quantifiers and reasoning about RBAC, which helps
+reduce the administrative complexity of associations such as user- role associations.
+To support for reasoning of access control, one may describe the RBAC extended models in first-order logic. First-order logic is expressive enough to concisely represent access control systems.
+Given an RBAC extension model, Samuel et al.~\cite{samuel07:spatio-temporal} proposed verification of the model using a specification language, which is based first-order logic. 
+This logic is sufficient to model RBAC and extended RBAC for reasoning. 
+Moreover, this logic supports for concise and elegant formulation of the Reference Model and its relation.
+
+%First-order logic uses relations, variables, and quantifiers.
+
+%Combination of attribute values may handle constraints such as temporal and spatial constraints.
 For example, we describe user-role authorization as $ae$ $\Rightarrow$ $r$ using propositional logic where $ae$ is an attribute expression and $r$ is a role.
@@
 The left hand side (LHS) is an attribute expression $ae$.
-The right hand side (RHS) is a role. 
-If a request satisfies the attribute expression, a user $u$ in the request is authorized to the role specified in RHS.
+%The right hand side (RHS) is a role. 
+%If a request satisfies the attribute expression, a user $u$ in the request is authorized to the role specified in RHS.
 
-We found that 27 papers of RBAC extensions use propositional logic to describe its extended model.
-However, propositional logic has limitations. While this logic is simple, this logic does not support for reasoning about RBAC, which helps
-reduce the administrative complexity of associations such as user- role associations.
-To support for reasoning of access control, one may describe the RBAC extended models in first-order logic. 
-Given an RBAC extension model, Samuel et al.~\cite{samuel07:spatio-temporal} proposed verification of the model using a specification language, which is based first-order logic. 
-This logic is sufficient to model RBAC and extended RBAC for reasoning. 
-Moreover, this logic supports for concise and elegant formulation of the Reference Model and its relation. 
-First-Order logic is expressive enough to concisely represent access control systems. First-Order logic uses relations, variables, and quantifiers.
+.

### 2014-06-17 paper/sections/core_rbac.tex
Revised version
 - Address comments in background, RQ2-RQ5, and discussion section.
 - In addition, I sort paper list in an increasing order (e.g., [4][5][10] instead of [10][4][5]) in the table.

@@
 
-RBAC provides effective and efficient permissions management for operations, especially when sharing resources within an organization.
-Prior to the creation of the NIST RBAC standard, no general agreement on the definition 
-of RBAC existed among practitioners or within the research community. 
-Without a unified definition of RBAC, software developers described similar concepts and features of RBAC models using different terminology. 
-The lack of consistent terminology was shown to slow the implementation of RBAC~\cite{o20102010}.  
-Moreover, in cases where organizations were concerned with adopting RBAC,
-evaluation and comparison of RBAC technologies developed by different vendors was difficult.
-NIST, in collaboration with industry and academics, worked on defining a set of consensus RBAC concepts and terminology and proposed a standard for 
-RBAC that addressed these cost and interoperability issues by developing a common definition that can be used across different vendors.
+RBAC provides a high level abstraction of permissions management for operations, especially when sharing resources among roles within an organization. 
+In cases where organizations were concerned with adopting RBAC, evaluation and comparison of RBAC technologies developed by different vendors was difficult.
+To address this issue, NIST proposed a standard for RBAC.
+NIST's RBAC standard can benefit organizations by lowering the cost of RBAC adoption~\cite{o20102010}.
+The RBAC standard includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC. 
+Hierarchical RBAC and constrained RBAC are developed by incorporating new features into core RBAC.
+Each component includes a corresponding RBAC Reference model.
 
-NIST's work can directly benefit organizations by lowering the cost of early phase research and development, and implementation of RBAC.
-Since the RBAC standard was first introduced, a 2010 report by RTI International showed that the rate of RBAC adoption has rapidly grown~\cite{o20102010}. 
-The analysts estimate that the use of RBAC technology has saved industry \$6.1 billion from 1992 to 2009 and the research and development work by NIST to produce an RBAC standard has saved an additional \$1.1 billion that wold have otherwise been spent across the industry doing research and development on RBAC.
-
-The RBAC standard includes three components of RBAC: core RBAC, hierarchical RBAC, and constrained RBAC.
-Each component includes a corresponding RBAC Reference model.
-We first describe the core RBAC, associated entities and other terminology.
-We next describe hierarchical RBAC and constrained RBAC, which are developed by incorporating new features into core RBAC.
-
-\subsection{Core RBAC}
-
-The four entities of the core RBAC Reference Model are:
+We describe the four entities of these reference models.
 
@@
 
+%\subsection{Core RBAC} 
+%The four entities of the core RBAC Reference Model are:
+
 In RBAC, a user can exercise a permission only if the user is assigned to a role that contains the permission.
@@
 
-\subsection{Hierarchical RBAC}
-
-The hierarchical RBAC Reference Model adds role hierarchies ($RH$) as a feature to the core RBAC Reference Model.
-$RH$ incorporates a structure of roles in an organization using inheritance relationships among attributes such as roles.
+On the top of the core RBAC Reference Model, the hierarchical RBAC Reference Model adds role hierarchies ($RH$) as a feature. 
 The role structure in an organization may use
 a role $r_1$, which inherits all permissions of another role $r_2$.
-For example, a manager role may inherit all permissions of an employee role.
-Role hierarchies help simplify access control policy creation and maintenance by reducing the number of
-individual role assignments for a user. 
 Concept of the role inheritance describes the many-to-many mapping role inheritance relations among roles.
-Therefore, more than one role (e.g., two roles $r_1$ and $r_1'$) can inherit all permissions of $r_2$ 
-General role hierarchies can be extended to use the concept of multiple inheritances where
-$r_1$ inherits all permissions from more than one role (e.g., two roles $r_2$ and $r_2'$) .
+Therefore, more than one role (e.g., two roles $r_1$ and $r_1'$) can inherit all permissions of $r_2$.
 
-\subsection{Constrained RBAC}
-
-The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. 
-Separation of duty relations enforce conflicts of interest among roles. 
-The Constrained RBAC model defines two types of separation of duty relations: static and dynamic.
+The constrained RBAC Reference Model adds separation of duty relations to the core and hierarchical RBAC Reference Models. Separation of duty relations enforce conflicts of interest among roles. 
+The Constrained RBAC model defines two types of constraints placed on the user-role assignments:
+Static Separation of Duty (SSoD) and Dynamic Separation of Duty (DSD).
 
 \begin{itemize}
-\item Static Separation of Duty (SSoD): SSoD restricts the conflicting-roles, which can be assigned to a single user statically. 
-Suppose roles $Role_A$ and $Role_B$ conflict with each other.
-On situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ statically.
-SSoD is known to be too rigid for practical use in cases where a user should have permissions when a user is assigned to both $Role_A$ and $Role_B$. 
-For example, an accounting clerk role requests a check and an accounting manager role approves the check. 
-These two roles must be mutually exclusive to avoid a situation where one approves the check that the other requested.
+\item SSoD relations define constraints as a pair ($role set$, $n$) statically where no user is assigned to more than $n$ roles from the $role set$.
+Suppose roles $Role_A$ and $Role_B$ in $role set$ conflict with each other and $n$ is 2. A user who is assinged to $Role_A$ cannot be assigned to $Role_B$.
+%For example, an accounting clerk role requests a check and an accounting manager role approves the check. 
+%These two roles must be mutually exclusive to avoid a situation where one approves the check that the other requested.
 
-\item Dynamic Separation of Duty (DSD): DSD is known to be more flexible than SSoD. 
-DSD restricts the conflicting-role assignments dynamically that are associated with a user. 
-Suppose roles $Role_A$ and $Role_B$ conflict with each other. 
-For situations where multiple roles can be associated with a single user, no permission is given to a user who is assigned to both $Role_A$ and $Role_B$ dynamically.	
-Consider that a user can be assigned to the accounting clerk and accounting manager role at the same time.
-In such a situation, DSD can enforce that the accounting manager role does not approve her/his requested checks but can only approve checks that others have requested.
+\item DSD relations define constraints as a pair ($role set$, $n$) dynamically where $n$ is a number, with the property that
+user session may not activate more than $n$ roles from the $role set$.
+For situations where multiple roles can be associated with a single user and $n$ is 2, a user session cannot be assigned to both $Role_A$ and $Role_B$ at the same time. 
+%Consider that a user can be assigned to the accounting clerk and accounting manager role at the same time.
+%In such a situation, DSD can enforce that the accounting manager role does not approve her/his requested checks but can only approve checks that others have requested.
 \end{itemize}

### 2014-06-17 paper/sections/discussion.tex
Revised version
 - Address comments in background, RQ2-RQ5, and discussion section.
 - In addition, I sort paper list in an increasing order (e.g., [4][5][10] instead of [10][4][5]) in the table.

@@
 By looking across all the research questions we can arrive at cross-cutting concerns and identify areas that may benefit from future work.
-We break the discussion into two issues: recommendations for the enhancement of the evaluation of RBAC extension models to role-based access control models and guidelines for developers needing to choose an RBAC extension model to meet their requirements.
+We discuss motivations behind the RBAC extension models and guidelines for choosing an RBAC extension model.
 
-\subsection{Metrics for the Evaluation of RBAC Extension Models}
+\subsection{Motivations behind the RBAC extension models}
 
-A small percentage of primary sources presented robust evaluations or implementations of their model. 
-This result leaves researchers and practitioners little data from which to compare one model to the next. 
-Given the wide swath of popular and critical domains being targeted by RBAC extension models, Hu and Scarfone proposed metrics for access control system evaluation or a security system in general~\cite{hu2012:NIST7874}.
-They collect metrics, which can be applicable for not only access control models, but also requirements, system implementation, and extended application such as verification and testing. 
-For example, ``Ease of Privilege Assignments'' metric measures the number of steps for assigning, changing, or deleting a privilege for users and roles when a user or administrator manages RBAC. 
-This metric can be applicable for not only the RBAC Reference Model but also RBAC extension models.
-We propose the metrics introduced by Hu and Scarfone should be applied to the extension models presented in this paper as well as any future models.
-A developer could attempt to apply these metrics to an RBAC extension model not already evaluated when attempting to choose a model. 
-These metrics focus on comparing feature support of access control models.
+While the RBAC Reference Model is considered fundamental in any RBAC systems, the RBAC Reference Model has limitations on providing features such as dealing with context for emerging applications such as healthcare and mobile devices.
+For example, in cases where the RBAC extension model does not support the notion of specific constraints, the RBAC extension model incorporates additional entities with regards to context/constraints and their relations with existing entities of the RBAC Reference Model.
+The RBAC extension model provides general agreement on the definition for specific features.
+The RBAC extension model helps simplify theoretical modeling and practical implementation of features in the model.
+
+During the paper reading phase, we identify motivations of the RBAC extension models by category.
+We identify motivations by looking for issues regarding why the RBAC reference model was inadequate and how the authors addressed the issues.
+We describe motivations behind the RBAC extension models by category as follows:
+
+\begin{itemize}
+
+  \item \textbf{Context}: The RBAC Reference Model does not support the notion of context constraints related to changes in environments.
+Therefore, RBAC belongs to the static access control model, which may not capture changes in environments.
+
+  \item \textbf{Constraint}: The RBAC Reference Model has limitations on its features such as delegation and role hierarchy. For example, partial inheritance in role hierarchy needs to be developed.
+
+  \item \textbf{Organizational}: The RBAC extension model does not handle RBAC administrative tasks efficiently across multiple organizations.
+The model needs to reduce the administrative complexity of RBAC across multiple organizations.
+
+  \item \textbf{Privacy}: The RBAC extension model does not support the notion of privacy. For example, the model lacks components, constraints, and obligations to handle privacy in RBAC.
+
+  \item \textbf{Task}: The RBAC extension model does not support the notion of a task, team, purpose, and organizational roles, which help specify a business activity in enterprises.
+
+  \item \textbf{Spatio-Temporal}: The RBAC extension model does not support spatial (location-based) and temporal (time-based) constraints, which specify role-assignment, role-activation and permissions based on location and time.
+
+  \item \textbf{Spatial}: The RBAC extension model does not support the notion of spatial (location-based)  constraints, which specify role-assignment and permissions based on location.
+
+  \item \textbf{Temporal}: The RBAC extension model does not support the notion of temporal (time-based)  constraints, which specify role-assignments and permissions based on time.
+
+\end{itemize}
+
+We found that motivations for RBAC extension models vary across category.
+In general, motivations are rooted in adding new entities or relationships to allow authorization flexibility in meeting the demands of emerging requirements.
 
@@
 To adopt a model in practice, software developers would implement the model in real system environments based on intended use of the models. 
-As shown by RQ6, all primary sources were found to provide an abstract formal representations of their extension model and the model's operation.
+As shown by RQ5, all primary sources were found to provide an abstract formal representations of their extension model and the model's operation.
 The presence of an abstract formal representation of the model stands as a starting place for developers to configure intended use (i.e.

### 2014-06-17 paper/sections/intro.tex
Revised version
 - Address comments in background, RQ2-RQ5, and discussion section.
 - In addition, I sort paper list in an increasing order (e.g., [4][5][10] instead of [10][4][5]) in the table.

@@
 \item RQ1: How can RBAC extension models be classified?
-\item RQ2: What are motivations behind the RBAC extension models?
-\item RQ3: How are RBAC extension models implemented by their authors?
-\item RQ4: How are RBAC extension models evaluated by their authors?
-\item RQ5: What domains have been targeted by RBAC extension models?
-\item RQ6: What commonalities exist across RBAC extension models?
+%\item RQ2: What are motivations behind the RBAC extension models?
+\item RQ2: How are RBAC extension models implemented by their authors?
+\item RQ3: How are RBAC extension models evaluated by their authors?
+\item RQ4: What domains have been targeted by RBAC extension models?
+\item RQ5: What commonalities exist across RBAC extension models?
 \end{itemize}
